### 2024-10-20 paper/paper.qmd
Prepare Appendix

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Pollster

# Idealized Methodology

# References

### 2024-10-20 scripts/03-clean_data.R
Created a sript for cleaning data. Cleaned the raw data and saved for analysis.

@@ -1,44 +1,61 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw poll data from FiveThirtyEight, which contains observations on the number of polls for the U.S. election.
+# Author: Shamayla Durrin Islam
+# Date: 19 October 2024
+# Contact: shamayla.islam@mil.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Requires raw polling data from FiveThirtyEight. Ensure tidyverse, janitor and arrow packages are installed for data saving and cleaning.
+# Any other information: This script assumes that the raw data is in CSV format and structured according to FiveThirtyEight's poll dataset.
 
-#### Workspace setup ####
+#### Workspace setup and loading data  ####
+
 library(tidyverse)
+library(janitor)
+library(arrow)
+raw_data <- read_csv("data/01-raw_data/raw_data.csv")
 
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+#### Data Cleaning ####
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
[…]

### 2024-10-20 scripts/03-clean_data.R
Saved the data as parquet file.

@@ -58,4 +58,5 @@ cutoffscore <- 0.22
 cleaned_data <- cleaned_data %>% filter(pollscore >= cutoffscore)
 
 #### Save data ####
-write_parquet(cleaned_data, "data/02-analysis_data/analysis_data.csv")
+
+write_parquet(cleaned_data, "data/02-analysis_data/analysis_data.parquet")

### 2024-10-20 scripts/03-clean_data.R
keeping only relevant candidates.

@@ -57,6 +57,11 @@ cutoffscore <- 0.22
 
 cleaned_data <- cleaned_data %>% filter(pollscore >= cutoffscore)
 
+#Filter the dataset to keep only rows with Kamala Harris or Donald Trump
+
+cleaned_data <- cleaned_data %>%
+  filter(candidate_name %in% c("Kamala Harris", "Donald Trump"))
+
 #### Save data ####
 
 write_parquet(cleaned_data, "data/02-analysis_data/analysis_data.parquet")

### 2024-10-20 scripts/03-clean_data.R
Fixing a mistake in data cleaning.

@@ -32,7 +32,7 @@ cleaned_data <- cleaned_data %>%
 
 # Keep only the relevant columns
 cleaned_data <- cleaned_data %>%
-  select(poll_id, pollster_id, numeric_grade, pollscore, state, start_date, end_date, sample_size, population, candidate_name, pct)
+  select(poll_id, pollster_id, pollster, numeric_grade, pollscore, state, start_date, end_date, sample_size, population, candidate_name, pct)
 
 # Remove any rows that have missing values in any column
 cleaned_data <- cleaned_data %>% drop_na()

### 2024-10-21 scripts/03-clean_data.R
Calculating recensyof polls based on start date,

@@ -32,7 +32,7 @@ cleaned_data <- cleaned_data %>%
 
 # Keep only the relevant columns
 cleaned_data <- cleaned_data %>%
-  select(poll_id, pollster_id, pollster, numeric_grade, pollscore, state, start_date, end_date, sample_size, population, candidate_name, pct)
+  select(poll_id, pollster_id, pollster, numeric_grade, pollscore, state, start_date, poll_scope, end_date, sample_size, population, candidate_name, pct)
 
 # Remove any rows that have missing values in any column
 cleaned_data <- cleaned_data %>% drop_na()

### 2024-10-21 scripts/03-clean_data.R
calculated recency of polls.

@@ -12,6 +12,7 @@
 library(tidyverse)
 library(janitor)
 library(arrow)
+library(lubridate)
 raw_data <- read_csv("data/01-raw_data/raw_data.csv")
 
 #### Data Cleaning ####
@@ -61,6 +62,15 @@ cleaned_data <- cleaned_data %>% filter(pollscore >= cutoffscore)
 
 cleaned_data <- cleaned_data %>%
   filter(candidate_name %in% c("Kamala Harris", "Donald Trump"))
+
+# Ensure the start_date and end_date columns are in date format
+cleaned_data$start_date <- as.Date(cleaned_data$start_date, format = "%m/%d/%y")
+cleaned_data$end_date <- as.Date(cleaned_data$end_date, format = "%m/%d/%y")
+
+# Create a new column to calculate the recency of a poll 
+
+cleaned_data$recency <- as.numeric(Sys.Date() - cleaned_data$end_date)
+
 
 #### Save data ####
 

### 2024-10-21 scripts/03-clean_data.R
Fixed a code.

@@ -64,8 +64,15 @@ cleaned_data <- cleaned_data %>%
   filter(candidate_name %in% c("Kamala Harris", "Donald Trump"))
 
 # Ensure the start_date and end_date columns are in date format
-cleaned_data$start_date <- as.Date(cleaned_data$start_date, format = "%m/%d/%y")
-cleaned_data$end_date <- as.Date(cleaned_data$end_date, format = "%m/%d/%y")
+cleaned_data <- cleaned_data %>%
+  mutate(
+    start_date = parse_date_time(start_date, orders = c("mdy", "ymd")),
+    end_date = parse_date_time(end_date, orders = c("mdy", "ymd"))
+  )
+
+# Ensure that end_date is a Date object
+
+cleaned_data$end_date <- as.Date(cleaned_data$end_date)
 
 # Create a new column to calculate the recency of a poll 
 

### 2024-10-21 scripts/03-clean_data.R
Fixed an error.

@@ -48,15 +48,6 @@ cutoffgrade <- 1.5 # this cutoff was selected  to retain mid to high quality pol
 
 cleaned_data <- cleaned_data %>% filter(numeric_grade >= cutoffgrade)
 
-# Look at the range of pollscore to select a cutoff 
-
-pollscore_range <- range(cleaned_data$pollscore)
-
-cutoffscore <- 0.22 
-
-# remove observations that has high pollsore (lower reflects less error and bias)
-
-cleaned_data <- cleaned_data %>% filter(pollscore >= cutoffscore)
 
 #Filter the dataset to keep only rows with Kamala Harris or Donald Trump
 

### 2024-10-21 scripts/03-clean_data.R
fixed an error.

@@ -42,7 +42,7 @@ cleaned_data <- cleaned_data %>% drop_na()
 
 numeric_grade_range <- range(cleaned_data$numeric_grade)
 
-cutoffgrade <- 1.5 # this cutoff was selected  to retain mid to high quality pollster 
+cutoffgrade <- 1 # this cutoff was selected  to retain mid to high quality pollster 
 
 # Remove pollster that has low numeric grade 
 

### 2024-10-21 paper/paper.qmd
Made some changes.

@@
 ---
-title: "My title"
+title: Forcasting US Elections using PollsofPolls
 subtitle: "My subtitle if needed"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+  - Shamayla Durrin
+  - Denise Chang 
+  - Krishna Kumar
+thanks: "Code and data are available at: [https://github.com/krishnak30/US_elections](https://github.com/krishnak30/US_elections)."
 date: today
@@
 library(tidyverse)
-library(palmerpenguins)
+library(arrow)
+library(here)
+library(ggplot2)
+analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
 ```
@@
 
-## Predictor variables
+# Polls of Polls to Forcast Winning Candidate
 
-Add graphs, tables and text.
+## Forecasting Approach
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+The polls of polls methodology is widely used in election prediction as it aggregates multiple polls to provide a more reliable estimate of voter support, rather than relying on any single poll.
+The goal is to reduce errors and biases present in individual polls by using a weighted average of many different polls.
+In our approach, we calculate the weights based on several important variables.
+Sample size is included because larger polls are generally more reliable and have smaller margins of error.
+Pollscore (or pollster rating) reflects the empirical quality of a pollster, balancing both accuracy and bias, allowing us to give more weight to those with a proven track record.
+We also account for the pollscope, differentiating between national and state-level polls, as state polls may offer more localized insights into voter sentiment.
+We also use recency as more recent polls give us a more accurate estimate of voter sentiment.
+
+However, we acknowledge that even with these adjustments, there may still be underlying biases from individual pollsters toward the candidates.
+To account for this, we will select a baseline pollster with a strong rating, high transparency, and minimal historical error, and regress the support percentage (pct) on pollster, including the same variables as controls whcih was used in the weighting method, such as sample size, pollscore and pollscope.Additionally, we will control for the state of the poll, as support levels can vary regionally.
+Therefore each pollster's coefficient in the regression model will represent the bias of that pollster relative to the baseline pollster.
+Controlling for these factors is crucial because it ensures that the differences between pollsters are not driven by other factors, such as differences in sample size or the types of polls they conduct, but by their inherent bias in support estimates.
+Once we have calculated the weights, we will use them to generate a weighted average of support for Kamala Harris and Donald Trump, ultimately comparing their weighted support to forecast the likely winner of the election.
+
+## **Exploration of Control Variables**
+
+### Investigating the Effect of Pollscore on Candidate Support
+
+```{r}
+
+# Define pollscore ranges (you can adjust the binwidth if needed)
+cleaned_data <- cleaned_data %>%
+  mutate(pollscore_range = cut(pollscore, breaks = seq(-1.5, 1, by = 0.3), include.lowest = TRUE))
+
+# Calculate the average support for each candidate within each pollscore range
+average_support <- cleaned_data %>%
+  group_by(pollscore_range, candidate_name) %>%
+  summarise(avg_support = mean(pct, na.rm = TRUE)) %>%
+  ungroup()
+
+# Create a bar plot
+ggplot(average_support, aes(x = pollscore_range, y = avg_support, fill = candidate_name)) +
+  geom_bar(stat = "identity", position = "dodge") +  # Dodge to separate bars for each candidate
+  scale_fill_manual(values = c("Donald Trump" = "indianred", "Kamala Harris" = "steelblue")) +  # Custom colors
+  labs(title = "Average Support for Candidates by Pollscore Range",
+       x = "Pollscore Range",
+       y = "Average Support (%)",
+       fill = "Candidate") +
+  theme_minimal() +  # Minimal theme
+  theme(
+    legend.position = "bottom",
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
+
+### Investigating the Effect of Poll Recency on Candidate Support
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+# Custom colors for candidates
+trump_color <- "indianred"
+harris_color <- "steelblue"
+
+# Create the plot with the requested adjustments (recency vs support)
+ggplot(analysis_data, aes(x = recency, y = pct, color = candidate_name)) +
+  geom_point(size = 0.5, alpha = 0.7) +
+  geom_smooth(method = "lm", se = FALSE) + # Smaller dots with slight transparency
+  scale_color_manual(values = c("Donald Trump" = trump_color, "Kamala Harris" = harris_color)) +  # Custom colors
+  theme_minimal() +  # Remove grey background
+  theme(
+    panel.grid.major = element_blank(),  # Remove major gridlines
+    panel.grid.minor = element_blank(),  # Remove minor gridlines
+    legend.title = element_blank(),  # Remove "candidate_name" as title
+    legend.position = "bottom",  # Move legend to the bottom
+    plot.title = element_text(hjust = 0.5),  # Center the title
+    axis.line = element_line(color = "black"),  # Add axis lines
+    panel.border = element_blank(),  # Remove panel border
+    axis.line.y.right = element_line(color = "black"),  # Add right spine for Kamala Harris
+    axis.line.x.top = element_blank()  # Remove top spine
+  ) +
+  labs(title = "Recency vs Support for Kamala Harris and Donald Trump",
+       x = "Poll Recency (days since poll end date)", 
+       y = "Support (%)") +
+  facet_wrap(~ candidate_name, scales = "free_y")  # Separate plots for each candidate
+
+```
+
+### Investigate the effect of sample size on support estimates
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+
+# Custom colors for candidates
+trump_color <- "indianred"
+harris_color <- "steelblue"
+
+# Create the plot with the requested adjustments
+ggplot(analysis_data, aes(x = sample_size, y = pct, color = candidate_name)) +
+  geom_point(size = 0.5, alpha = 0.9) + # Smaller dots with slight transparency
+  geom_smooth(method = "lm", se = FALSE)+
+
+  scale_color_manual(values = c("Donald Trump" = trump_color, "Kamala Harris" = harris_color)) +  # Custom colors
+  theme_minimal() +  # Remove grey background
+  theme(
+    panel.grid.major = element_blank(),  # Remove major gridlines
+    panel.grid.minor = element_blank(),  # Remove minor gridlines
+    legend.title = element_blank(),  # Remove "candidate_name" as title
+    legend.position = "bottom",  # Move legend to the bottom
+    plot.title = element_text(hjust = 0.5),  # Center the title
+    axis.line = element_line(color = "black"),  # Add axis lines
+    panel.border = element_blank(),  # Remove panel border
+    axis.line.y.right = element_line(color = "black"),  # Add right spine for Kamala Harris
+    axis.line.x.top = element_blank()  # Remove top spine
+  ) +
+  labs(title = "Sample Size vs Support for Kamala Harris and Donald Trump",
+       x = "Sample Size", 
+       y = "Support (%)") +
+  facet_wrap(~ candidate_name, scales = "free_y")
+```
+
+### Distribution of support for Candiates by State
+
+```{r}
+# Calculate mean support for each candidate in each state
+state_support <- analysis_data %>%
+  group_by(state, candidate_name) %>%
+  summarise(mean_support = mean(pct)) %>%
+  ungroup()
+
+# Pivot the data to have one row per state with separate columns for Harris and Trump
+state_support_wide <- state_support %>%
+  pivot_wider(names_from = candidate_name, values_from = mean_support)
+```
+
+```{r}
+# Calculate mean support for each candidate in each state
+state_support <- analysis_data %>%
+  group_by(state, candidate_name) %>%
+  summarise(mean_support = mean(pct))
+
+# Calculate the proportion of support for Kamala Harris
+state_support <- state_support %>%
+  mutate(harris_proportion = Kamala_Harris / (Kamala_Harris + Donald_Trump))
+
+```
 
@@
 
-```{=tex}
 \begin{align} 
@@
 \end{align}
-```
+
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
@@
 
-# Pollster
+# Additional data details
 
-# Idealized Methodology
+# Model details {#sec-model-details}
+
+## Posterior predictive check
+
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-ppcheckandposteriorvsprior
+#| layout-ncol: 2
+#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+
+pp_check(first_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+posterior_vs_prior(first_model) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
+```
+
+## Diagnostics
+
+@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+
+@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-stanareyouokay
+#| fig-cap: "Checking the convergence of the MCMC algorithm"
+#| fig-subcap: ["Trace plot", "Rhat plot"]
+#| layout-ncol: 2
+
+plot(first_model, "trace")
+
+plot(first_model, "rhat")
+```
+
+\newpage
 

### 2024-10-21 paper/paper.qmd
intial draft of the data section

--- 2024-10-20 paper/paper.qmd
@@
 author: 
-  - First author
+  - Shamalya D
   - Another author
@@
 
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
-library(tidyverse)
-library(palmerpenguins)
-```
-
 # Introduction
 
-Overview paragraph
+# Data
 
-Estimand paragraph
+## Data Set Overview
 
-Results paragraph
+add a summary statistics table here maybe?
 
-Why it matters paragraph
+## Variables
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+Mention any variables that were constructed here, as well
 
-# Data {#sec-data}
+## High Level Data Cleaning
 
-## Overview
+## Summary Statistics
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
+## Graphs for Imp. Variables (relationships)
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+Public support for political candidates is an essential part of understanding electoral dynamics for a variety of reasons.
+Public polls provide insights into the preferences and beliefs of certain groups of people, usually voters in specific states or districts, and sometimes to also correlate preferences to demographic information (race, age, neighbourhood, etc).
 
-## Outcome variables
+The results help inform policy decisions, help track the likely outcomes of elections, help government officials and candidates decide where to invest time, funds, and resources to get the desired results.
+Polling organizations aim to capture data by surveying a representative sample of voters, free from government intervention.
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+Polling organizations conduct these surveys through various methods. For our particular data set, the following methods have been used to gauge public sentiment:
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+-   App panel
+
+-   Online panel
+
+-   Email
+
+-   Online ad
+
+-   IVR (interactive voice response)
+
+-   Live phone
+
+-   Text
+
+-   Text-to-web
+
+-   Probability panel
+
+-   Mail-to-phone
+
+Each pollster chooses a method(s) based on what they find to be most convenient, efficient, and providing the best results.
+Moreover, there are some variables which affect the validity and reliability of the polling results, such as sample size, methodology, population type, etc.
+Larger samples generally lead to more reliable estimates of public opinion.
+Some methodologies might be considered more reliable due to certain factors.
+The pollsters also use different sampling methods (random, stratified, etc) which can affect the validity of the polling results as well.
+
+After collecting the required data, the polling organizations transform these results into more easily comprehensible variables such as numeric grade, poll score, transparency score, and a percentage.
+These scores are calculated through an extensive process of calculating excess error and bias, adjusting for race difficulty, calculating predictive error and bias, calculating pollster transparency, and combining everything into one single ‘POLLSCORE’, while accounting for luck as well.
+
+Variables such as pollster rating and transparency score help validate the reliability of the pollster and the polling results they publish.
+
+Each entry in the dataset represents a glimpse of public opinion at a specific time and location, capturing measured support percentages alongside contextual variables such as pollster, methodology, and sample size.
+This extensive approach allows for an enhanced understanding of public support dynamics, facilitating informed analyses of electoral outcomes.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
+library(arrow)
+analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+head(analysis_data)
 ```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-```{=tex}
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-```
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {.unnumbered}
-
-# Pollster
-
-# Idealized Methodology
-
-# References

### 2024-10-22 paper/paper.qmd
updated data section with two new graphs

--- 2024-10-21 paper/paper.qmd
@@
 ---
-title: Forcasting US Elections using PollsofPolls
+title: Forcasting US Elections using Polls-of-Polls
 subtitle: "My subtitle if needed"
@@
 thanks: "Code and data are available at: [https://github.com/krishnak30/US_elections](https://github.com/krishnak30/US_elections)."
+
 date: today
@@
 
-Overview paragraph
+# Data
 
-Estimand paragraph
+## Data Set Overview
 
-Results paragraph
+add a summary statistics table here maybe?
 
-Why it matters paragraph
+```{r}
+library(arrow)
+analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+head(analysis_data, 20)
 
-# Data {#sec-data}
+```
 
-## Overview
+## Variables
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+The data set contains a wide range of variables that are essential for understanding polling outcomes. The following are key variables:
 
-Overview text
+-   **`poll_id`**: Unique identifier for each poll conducted.
+
+-   **`pollster`**: The polling organization that conducted the poll (e.g., YouGov, TIPP).
+
+-   **`numeric_grade`**: A numeric rating indicating the pollster’s quality or reliability (e.g., 1.8).
+
+-   **`pollscore`**: A score representing the reliability of the pollster in terms of error and bias.
+
+-   **`state`**: The state where the poll was conducted (or "Not Applicable" for national polls).
+
+-   **`start_date` / `end_date`**: Dates during which the poll was conducted.
+
+-   **`sample_size`**: The number of respondents in the poll.
+
+-   **`population`**: The group surveyed (e.g., likely voters or registered voters).
+
+-   **`candidate_name`**: The name of the candidate being polled (e.g., Kamala Harris, Donald Trump).
+
+-   **`pct`**: The percentage of support for the candidate in the poll.
+
+-   **`methodology`**: The method used to conduct the poll (e.g. online panel)
+
+-   **`transparency_score`**: A grade for how transparent a pollster is, calculated based on how much information it discloses about its polls and weighted by recency.
+The highest Transparency Score is 10.
+
+## High Level Data Cleaning
+
+## Summary Statistics
+
+### 1) Average Pollster Quality (pollscore): mean, median, mode, min, max
+
+```{r}
+library(dplyr)
+
+# Generate summary statistics for national and state polls
+summary_stats <- analysis_data %>%
+  filter(!is.na(pollscore)) %>%  # Exclude rows with missing pollscore
+  group_by(poll_scope) %>%        # Group by poll_scope
+  summarize(
+    Count = n(),                   
+    Min = round(min(pollscore), 2),
+    `1st Qu.` = round(quantile(pollscore, 0.25), 2),
+    Median = round(median(pollscore), 2),
+    Mean = round(mean(pollscore), 2),
+    `3rd Qu.` = round(quantile(pollscore, 0.75), 2),
+    Max = round(max(pollscore), 2),
+    .groups = 'drop'  # Drop grouping structure for cleaner output
+  )
+
+# Print the summary statistics
+print(summary_stats)
+
+```
+
+As shown above, the summary statistics table **\[CROSS-REFERENCE HERE\]** offers an overview of poll scores for national and state polls, with 2,971 and 2,953 entries, respectively.
+Both categories feature a minimum poll score of -1.5, which a high level of validity and reliability.
+The first quartile shows national polls at -0.9, while state polls are slightly lower at -1.1.
+Median scores are -0.3 for national polls and -0.4 for state polls, suggesting that half of the polls in both categories yield moderately negative scores, reinforcing a trend of favorable sentiment.
+The mean score is -0.37 for national polls, indicating a less reliable perception compared to the -0.51 mean for state polls.
+The third quartile for national polls (0.0) indicates that the upper 25% of scores are neutral, while state polls remain just below neutral at 0.7.
+The maximum score of 1.7 for national polls reflects a notable negative perception for at least one candidate, while the maximum for state polls is 0.7.
+Overall, these statistics illustrate a strong reliability of pollsters considering their errors and bias.
+
+### 2) Distribution of sample sizes: same as above
+
+```{r}
+library(ggplot2)
+library(dplyr)
+library(gridExtra)  # Load the gridExtra package
+
+# Create a new variable for sample size ranges
+analysis_data <- analysis_data %>%
+  mutate(sample_size_range = case_when(
+    sample_size <= 10000 ~ "0-10,000",
+    sample_size > 10000 & sample_size <= 20000 ~ "10,001-20,000",
+    sample_size > 20000 & sample_size <= 30000 ~ "20,001-30,000",
+    TRUE ~ "Above 30,000"
+  ))
+
+# Plot for 0-10,000
+p1 <- ggplot(analysis_data %>% filter(sample_size <= 10000), aes(x = sample_size)) +
+  geom_histogram(binwidth = 500, fill = "lightgreen", color = NA, alpha = 0.7) +
+  labs(title = "Sample Sizes 
+       (0-10,000)",
+       x = "Sample Size",
+       y = "Count") +
+  scale_x_continuous(limits = c(0, 10000)) +  # Set x limits for the first plot
+  scale_y_continuous(limits = c(0, 2300)) +  # Set y limits for the first plot
+  theme_minimal() +
+  theme(plot.title = element_text(size = 10),  # **Decreased font size of the title**
+        axis.title.y = element_text(margin = margin(r = 15)))  # **Increased gap for y-axis label**
+
+# Plot for 10,001 - 20,000
+p2 <- ggplot(analysis_data %>% filter(sample_size > 10000 & sample_size <= 20000), aes(x = sample_size)) +
+  geom_histogram(binwidth = 500, fill = "indianred", color = NA, alpha = 0.7) +
+  labs(title = "Sample Sizes 
+       (10,001-20,000)",
+       x = "Sample Size",
+       y = "Count") +
+  scale_x_continuous(limits = c(10001, 20000)) +  # Set x limits for the second plot
+  scale_y_continuous(limits = c(0, 50)) +  # Set y limits for the second plot
+  theme_minimal() +
+  theme(plot.title = element_text(size = 10),  # **Decreased font size of the title**
+        axis.title.y = element_text(margin = margin(r = 10)))  # **Increased gap for y-axis label**
+
+# Plot for 20,001 - 30,000
+p3 <- ggplot(analysis_data %>% filter(sample_size > 20000 & sample_size <= 30000), aes(x = sample_size)) +
+  geom_histogram(binwidth = 500, fill = "lightblue", color = NA, alpha = 0.7) +
+  labs(title = "Sample Sizes 
+       (20,001-30,000)",
+       x = "Sample Size",
+       y = "Count") +
+  scale_x_continuous(limits = c(20001, 30000)) +  # Set x limits for the third plot
+  scale_y_continuous(limits = c(0, 10)) +  # Set y limits for the third plot
+  theme_minimal() +
+  theme(plot.title = element_text(size = 10),  # **Decreased font size of the title**
+        axis.title.y = element_text(margin = margin(r = 5)))  # **Increased gap for y-axis label**
+
+# Arrange the plots side by side
+grid.arrange(p1, p2, p3, ncol = 3)  # ncol=3 arranges the plots in one row
+
+```
+
+The histograms **\[CROSS-REFERENCE HERE\]** reveal significant trends in sample sizes across the different ranges, indicating a general tendency towards smaller samples in the polling data.
+Most values within the 0-10,000 range cluster at the lower end, suggesting that many polls are conducted with limited respondent numbers, which may compromise the reliability and representativeness of the results.
+**However, this conclusion is very much dependent on what we determine to be an acceptable sample size for the results to be considered reliable**.
+In the 10,001-20,000 range, while there is some increase in sample sizes, the majority still falls within the 10,000-12,500 bracket, with few reaching the upper limits.
+
+The findings from the 20,001-30,000 range further illustrate the scarcity of larger sample sizes, with only a couple of values noted at each of the significant points.
+This distribution emphasizes a preference for smaller samples in the polling data analyzed.
+
+In the context of polling results, these findings suggest potential limitations in the accuracy and validity of the insights generated.
+Larger sample sizes are typically preferred in polling as they tend to yield more reliable estimates and reduce the margin of error, leading to a better representation of the population's views.
+Consequently, the predominance of smaller sample sizes in this data set may raise concerns about the robustness of the polling findings, potentially affecting their applicability in broader analyses or decision-making processes.
+**However, we must determine an acceptable and ideal sample size that would allow us to determine the reliability of these results.**
+
+### 3) Average percentage of support for top candidates (pct): how do we display this, which kind of graph, plot?
+
+### 4) Number of polls conducted by State: map showing different colours for the different states and the number of polls conducted for them.
+this will only include the 'state' polls then.
+not national polls.
+
+### 5) population, type fo voters simple graph, chart: how many likely vs registered voters.
+
+### 6) methodology: count for each type fo methodology and how many times it was used
+
+## Graphs for Imp. Variables (relationships)
 
@@
 
-## Outcome variables
+Public support for political candidates is an essential part of understanding electoral dynamics for a variety of reasons.
+Public polls provide insights into the preferences and beliefs of certain groups of people, usually voters in specific states or districts, and sometimes to also correlate preferences to demographic information (race, age, neighbourhood, etc).
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+The results help inform policy decisions, help track the likely outcomes of elections, help government officials and candidates decide where to invest time, funds, and resources to get the desired results.
+Polling organizations aim to capture data by surveying a representative sample of voters, free from government intervention.
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+Polling organizations conduct these surveys through various methods. For our particular data set, the following methods have been used to gauge public sentiment:
 
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
+-   App panel
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
+-   Online panel
 
-Talk more about it.
+-   Email
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+-   Online ad
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
+-   IVR (interactive voice response)
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+-   Live phone
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
+-   Text
 
-Talk way more about it.
+-   Text-to-web
+
+-   Probability panel
+
+-   Mail-to-phone
+
+Each pollster chooses a method(s) based on what they find to be most convenient, efficient, and providing the best results.
+Moreover, there are some variables which affect the validity and reliability of the polling results, such as sample size, methodology, population type, etc.
+Larger samples generally lead to more reliable estimates of public opinion.
+Some methodologies might be considered more reliable due to certain factors.
+The pollsters also use different sampling methods (random, stratified, etc) which can affect the validity of the polling results as well.
+
+After collecting the required data, the polling organizations transform these results into more easily comprehensible variables such as numeric grade, poll score, transparency score, and a percentage.
+These scores are calculated through an extensive process of calculating excess error and bias, adjusting for race difficulty, calculating predictive error and bias, calculating pollster transparency, and combining everything into one single ‘POLLSCORE’, while accounting for luck as well.
+
+Variables such as pollster rating and transparency score help validate the reliability of the pollster and the polling results they publish.
+
+Each entry in the data set represents a glimpse of public opinion at a specific time and location, capturing measured support percentages alongside contextual variables such as pollster, methodology, and sample size.
+This extensive approach allows for an enhanced understanding of public support dynamics, facilitating informed analyses of electoral outcomes.
 
@@
 # Define pollscore ranges (you can adjust the binwidth if needed)
-cleaned_data <- cleaned_data %>%
+cleaned_data <- analysis_data %>%
   mutate(pollscore_range = cut(pollscore, breaks = seq(-1.5, 1, by = 0.3), include.lowest = TRUE))

### 2024-10-22 scripts/03-clean_data.R
Added some new things to data cleaning.

@@ -21,19 +21,19 @@ raw_data <- read_csv("data/01-raw_data/raw_data.csv")
 
 cleaned_data <- raw_data %>% distinct()
 
-# Create a new column 'poll_scope' based on whether the 'state' column is empty or not
+# Create a new column 'National' based on whether the 'state' column is empty or not
 
 cleaned_data <- cleaned_data %>%
-  mutate(poll_scope = ifelse( is.na(state), "National", "State"))
+  mutate(national = ifelse(is.na(state), 1, 0))
 
-# Replace missing values (NA) in the 'state' column with NA
+# Replace missing values (NA) in the 'state' column with Not Applicable
 
 cleaned_data <- cleaned_data %>%
   mutate(state = ifelse(is.na(state), "Not Applicable", state))
 
 # Keep only the relevant columns
 cleaned_data <- cleaned_data %>%
-  select(poll_id, pollster_id, pollster, numeric_grade, pollscore, state, start_date, poll_scope, end_date, sample_size, population, candidate_name, pct)
+  select(poll_id, pollster_id, pollster, numeric_grade, pollscore, state, start_date, national, end_date, sample_size, population, candidate_name, pct)
 
 # Remove any rows that have missing values in any column
 cleaned_data <- cleaned_data %>% drop_na()
@@ -52,7 +52,7 @@ cleaned_data <- cleaned_data %>% filter(numeric_grade >= cutoffgrade)
 #Filter the dataset to keep only rows with Kamala Harris or Donald Trump
 
 cleaned_data <- cleaned_data %>%
-  filter(candidate_name %in% c("Kamala Harris", "Donald Trump"))
+  filter(candidate_name %in% c("Kamala Harris"))
 
 # Ensure the start_date and end_date columns are in date format
 cleaned_data <- cleaned_data %>%
@@ -61,14 +61,33 @@ cleaned_data <- cleaned_data %>%
     end_date = parse_date_time(end_date, orders = c("mdy", "ymd"))
   )
 
+# Apply clean_names to clean the column names
+cleaned_data <- cleaned_data %>%
+  clean_names()
+
 # Ensure that end_date is a Date object
 
 cleaned_data$end_date <- as.Date(cleaned_data$end_date)
 
-# Create a new column to calculate the recency of a poll 
 
-cleaned_data$recency <- as.numeric(Sys.Date() - cleaned_data$end_date)
[…]

### 2024-10-22 paper/paper.qmd
Update pollster appendix

--- 2024-10-21 paper/paper.qmd
@@
 author: 
-  - Shamalya D
+  - First author
   - Another author
@@
 
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+library(tidyverse)
+library(palmerpenguins)
+```
+
 # Introduction
 
-# Data
+Overview paragraph
 
-## Data Set Overview
+Estimand paragraph
 
-add a summary statistics table here maybe?
+Results paragraph
 
-## Variables
+Why it matters paragraph
 
-Mention any variables that were constructed here, as well
+Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
 
-## High Level Data Cleaning
+# Data {#sec-data}
 
-## Summary Statistics
+## Overview
 
-## Graphs for Imp. Variables (relationships)
+We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+
+Overview text
 
@@
 
-Public support for political candidates is an essential part of understanding electoral dynamics for a variety of reasons.
-Public polls provide insights into the preferences and beliefs of certain groups of people, usually voters in specific states or districts, and sometimes to also correlate preferences to demographic information (race, age, neighbourhood, etc).
+Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
 
-The results help inform policy decisions, help track the likely outcomes of elections, help government officials and candidates decide where to invest time, funds, and resources to get the desired results.
-Polling organizations aim to capture data by surveying a representative sample of voters, free from government intervention.
+## Outcome variables
 
-Polling organizations conduct these surveys through various methods. For our particular data set, the following methods have been used to gauge public sentiment:
+Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
 
--   App panel
-
--   Online panel
-
--   Email
-
--   Online ad
-
--   IVR (interactive voice response)
-
--   Live phone
-
--   Text
-
--   Text-to-web
-
--   Probability panel
-
--   Mail-to-phone
-
-Each pollster chooses a method(s) based on what they find to be most convenient, efficient, and providing the best results.
-Moreover, there are some variables which affect the validity and reliability of the polling results, such as sample size, methodology, population type, etc.
-Larger samples generally lead to more reliable estimates of public opinion.
-Some methodologies might be considered more reliable due to certain factors.
-The pollsters also use different sampling methods (random, stratified, etc) which can affect the validity of the polling results as well.
-
-After collecting the required data, the polling organizations transform these results into more easily comprehensible variables such as numeric grade, poll score, transparency score, and a percentage.
-These scores are calculated through an extensive process of calculating excess error and bias, adjusting for race difficulty, calculating predictive error and bias, calculating pollster transparency, and combining everything into one single ‘POLLSCORE’, while accounting for luck as well.
-
-Variables such as pollster rating and transparency score help validate the reliability of the pollster and the polling results they publish.
-
-Each entry in the dataset represents a glimpse of public opinion at a specific time and location, capturing measured support percentages alongside contextual variables such as pollster, methodology, and sample size.
-This extensive approach allows for an enhanced understanding of public support dynamics, facilitating informed analyses of electoral outcomes.
+Some of our data is of penguins (@fig-bills), from @palmerpenguins.
 
 ```{r}
-library(arrow)
-analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
+#| label: fig-bills
+#| fig-cap: Bills of penguins
+#| echo: false
 
-head(analysis_data)
+ggplot(penguins, aes(x = island, fill = species)) +
+  geom_bar(alpha = 0.8) +
+  scale_fill_manual(values = c("darkorange","purple","cyan4"),
+                    guide = "none") +
+  theme_minimal() +
+  facet_wrap(~species, ncol = 1) +
+  coord_flip()
 ```
+
+Talk more about it.
+
+And also planes (@fig-planes).
+(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+
+```{r}
+#| label: fig-planes
+#| fig-cap: Relationship between wing length and width
+#| echo: false
+#| warning: false
+#| message: false
+
+analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+
+analysis_data |> 
+  ggplot(aes(x = width, y = length)) +
+  geom_point(alpha = 0.8) +
+  theme_minimal() +
+  labs(x = "Wing width (mm)",
+       y = "Wing length (mm)")
+```
+
+Talk way more about it.
+
+## Predictor variables
+
+Add graphs, tables and text.
+
+Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+
+# Model
+
+The goal of our modelling strategy is twofold. Firstly,...
+
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+
+## Model set-up
+
+Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+
+```{=tex}
+\begin{align} 
+y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_i + \gamma_i\\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta &\sim \mbox{Normal}(0, 2.5) \\
+\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\sigma &\sim \mbox{Exponential}(1)
+\end{align}
+```
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
+### Model justification
+
+We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+
+We can use maths by including latex between dollar signs, for instance $\theta$.
+
+# Results
+
+Our results are summarized in @tbl-modelresults.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+library(rstanarm)
+
+first_model <-
+  readRDS(file = here::here("models/first_model.rds"))
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| warning: false
+
+modelsummary::modelsummary(
+  list(
+    "First model" = first_model
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
+
+# Discussion
+
+## First discussion point {#sec-first-point}
+
+If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+
+## Second discussion point
+
+Please don't use these as sub-heading labels - change them to be what your point actually is.
+
+## Third discussion point
+
+## Weaknesses and next steps
+
+Weaknesses and next steps should also be included.
+
+\newpage
+
+\appendix
+
+# Appendix {.unnumbered}
+
+# Emerson College Polling Methodology (October 14-16, 2024)
+
+## Overview
+Emerson College Polling, known for its extensive online polling methods, conducted a survey from October 14 to 16, 2024, targeting 1,000 likely voters.
+The poll measured voter preferences between Kamala Harris and Donald Trump, revealing a near tie with Harris at 50% and Trump at 49%.
+
+## Population, Frame, and Sample
+The target population is likely voters in the U.S.
+elections.
+These voters are determined by the voters' likeliness to vote in the upcoming elections and the voter's voting history.
+Both of these factors are self-reported.
+
+The sampling frame is the voters that are on the Aristotle list and on the online panel provided by CINT.
+Aristotle is an online supplier of voter files and CINT is a software company that connects researchers, such as Emerson, to online panels.
+This approach helps ensure a representative sample by accurately reflecting the electorate.
+
+The sample size is the 1,000 likely voters chosen arbitrarily from the sampling frame.
+Likely voters are determined by a combination of voter history, registration status, and demographic data.
+These factors are self-reported.
+
+## Sample Recruitment and Approach
+Emerson College used a mixed-mode sampling approach for its polling. Specifically, three primary modes were used to collect data for its October 14-16, 2024 poll:
+
+- MMS-to-web text surveys: Respondents were contacted via text messages sent to cell phones.
+These messages directed participants to complete the survey online.
+This method was conducted using voter lists provided by Aristotle.
+
+- Interactive Voice Response (IVR): Landlines were targeted using automated phone calls where respondents could answer questions using their keypads.
+The contact information for this was also sourced from Aristotle’s voter lists.
+
+- Online Panel from CINT: Emerson utilized a pre-screened, opt-in online panel of voters provided by CINT.
+
+## Advantages and Trade-offs in Mixed-Mode Sampling
+The multi-mode sampling approach used in this poll has several advantages:
+
+- Wider demographic reach: Having a combination of data collection methods allows Emerson to gather responses from a broader demographic group.
+In this polling, MMS-to-web surveys mostly gather responses from younger voters and those who primarily use mobile devices.
+IVR target older and rural voters, who are more likely to respond via landlines.
+Online panels is more efficient in reaching tech-savvy individuals.
+
+- Cost-efficient: Automated technologies reduce the cost compared to live interviews. In a polling context, this allows for a larger sample size.
+
+However, the trade-offs include:
+
+- Coverage bias: IVR only reaches individuals with landlines, which may skew results towards older demographics.
+
+- Self-selection bias: Respondents in online panels tend to be more engaged, which can lead to over-representation of politically active individuals.
+
+## Non-response Handling
+Emerson addresses non-response using a weighting system.
+The data is adjusted based on key demographic variables like age, gender, race, education, and party affiliation, making the sample more representative of the electorate.
+Weights are applied to compensate for under- or over-represented groups, ensuring balanced results.
+
+## Questionnaire Design
+The Emerson survey features a straightforward, efficient design focused on key issues and voter preferences. Strengths of the questionnaire include:
+
+- Simplicity: Clear, concise questions makes the survey easy for respondents to understand and to provide accurate answers.
+
+- Topical relevance: The poll zeroes in on vote preference and demographic splits, delivering timely insights during election cycles.
+
+However, weaknesses include:
+
+- Lack of nuance: While the survey captures basic preferences, it does not delve deeply into voters' motivations or the factors driving their choices.
+
+- Mode limitations: Different survey modes can affect how respondents engage with the questions. This can potentially influence the depth and accuracy of voter responses.
+
+## Conclusion
+Emerson College Polling's mixed-mode methodology is effective in balancing cost with broad demographic reach.
+Its weighting techniques contribute to reliable and accurate results despite non-responses.
+While the use of multiple survey modes increases population representation, possible biases—such as lack of nuance and mode effects need to be taken into consideration when interpreting results.
+
+In sum, this method is a strong predictor of election outcomes and continues to adapt to the evolving challenges of modern polling.
+
+# Idealized Methodology
+
+# References

### 2024-10-22 paper/paper.qmd
complete overview and sampling approach for Appendix B

@@
 
+## Overview
+This methodology outlines an election forecasting plan with a budget of $100,000.
+Heavily inspired by TIPP Insights and Emerson College's hybrid approaches, the survey methodology balances cost-efficiency and data accuracy by using different modes of data collection and wide demographic reach.
+
+This forecasting plan will predict the popular vote and Electoral College outcomes, emphasizing key swing states and underrepresented voter groups.
+
+## Sampling Approach
+This idealized methodology uses a multi-mode hybrid sampling approach. This allows the use of both probability-based and non-probability-based methods:
+
+- Probability-based methods:
+
+  - Live phone interviews (landline and cell): As used by TIPP, live phone interviews is a direct and personal method of collecting responses.
+Voter lists would be sourced from Aristotle to ensure we contact likely voters.
+This mode reaches older voters and voters with less internet connection.
+
+  - Interactive Voice Response (IVR): Similar to Emerson’s approach, IVR for landline users balances costs and ensures demographic representation, particularly among older or less tech-savvy populations.
+This method captures quick responses and is less expensive than live interviews.
+
+- Non-probability methods:
+
+  - MMS-to-web (mobile): Following Emerson’s MMS-to-web strategy, we would send multimedia messages to likely voters’ cell phones, encouraging them to complete an online survey.
+This method works well for younger, mobile-first voters, helping fill in gaps where traditional methods might miss.
+
+  - Online panel recruitment via CINT: To mirror TIPP’s online panel approach, a portion of the survey would be conducted via an opt-in panel of pre-screened voters from CINT, capturing tech-savvy respondents.
+CINT’s platform ensures that different voter groups, including hard-to-reach groups, are represented in the polling.
+
 # References

### 2024-10-22 paper/paper.qmd
include oversampling of key states in recruitment

@@
 
+## Recruitment Strategy
+By using Aristotle’s voter lists and CINT's platform as a sampling frame, the sample consists of likely voters based on registration status and past voting behavior.
+Quota sampling would be used to match the demographic profile of the U.S.
+electorate.
+This ensures that respondents in the sample reflect key demographic variables such as age, gender, race, and education level.
+
+We would oversample key swing states like Pennsylvania, Wisconsin, and Florida.
+Key swing states are states where there is no overwhelming support for a particular candidate or political party.
+The outcome in these states can often determine the overall result of the election, making them critical in securing a win in the Electoral College.
+Oversampling these states allows a more accurate forecasting in these battlegrounds, which are critical for predicting the Electoral College outcome.
+
 # References

### 2024-10-22 paper/paper.qmd
include data validation section and conclusion

@@
 Key swing states are states where there is no overwhelming support for a particular candidate or political party.
-The outcome in these states can often determine the overall result of the election, making them critical in securing a win in the Electoral College.
+The outcome in these states often determine the overall result of the election, making them critical in securing a win in the Electoral College.
 Oversampling these states allows a more accurate forecasting in these battlegrounds, which are critical for predicting the Electoral College outcome.
 
+## Data Validation
+
+Given that the data is collected in a hybrid format, we would use duplicate detection methods via IP addresses or phone numbers to eliminate multiple responses from the same individual.
+Also, we would implement attention checks to identify respondents who are not fully engaged.
+Both duplicate respondents and unengaged respondents biases results, and should be removed from the data set.
+
+Following TIPP’s iterative weighting process, the data will be weighted to ensure that the sample aligns with key demographics (race, gender, education) based on census data.
+
+## Conclusion
+By combining TIPP’s live phone/IVR hybrid and Emerson College’s mixed-mode survey methods, this idealized methodology captures diverse voter demographics.
+It uses both probability-based and non-probability-based sampling methods to forecast the 2024 U.S.
+Presidential Election.
+
 # References

### 2024-10-22 paper/paper.qmd
Couple of changes, deleted unnecessary files, made an interactive map.

--- 2024-10-22 paper/paper.qmd
@@
 ---
-title: "My title"
+title: Forcasting US Elections using Polls-of-Polls
 subtitle: "My subtitle if needed"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+  - Shamayla Durrin
+  - Denise Chang 
+  - Krishna Kumar
+thanks: "Code and data are available at: [https://github.com/krishnak30/US_elections](https://github.com/krishnak30/US_elections)."
+
 date: today
@@
 #| message: false
-
+library(psych)
 library(tidyverse)
-library(palmerpenguins)
+library(arrow)
+library(here)
+library(ggplot2)
+library(kableExtra)
+library(sf)
+library(tidyverse)
+library(plotly)
+library(RColorBrewer)
+library(stringr)
+library(modelsummary)
+analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
+mapping_data <- st_read(here("data/03-shapefiles_data/US"))
 ```
@@
 
-Overview paragraph
+# Data
 
-Estimand paragraph
+### Measurement
 
-Results paragraph
+In the dataset of our analysis, the process of measurement begins by capturing real-world public opinion through surveys.
+Polling organizations transform this observed phenomenon into structured data by recording support percentages, pollster details, and methodological choices, thus turning public sentiment into quantifiable entries in the dataset.
 
-Why it matters paragraph
+Public opinion polls are essential for understanding voter preferences and electoral dynamics, providing data that informs policy decisions and campaign strategies.
+Our data set captures public support for candidates at specific times, along with information on methodologies, sample sizes, and pollster ratings.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+Polling organizations use various methods, such as App panels, IVR (Interactive Voice Response), Live phone calls, Text-to-web systems etc.
+These methods can influence the validity and reliability of results, with larger sample sizes generally yielding more reliable estimates.
+Some methods, like live phone interviews, may be seen as more trustworthy than others.
 
-# Data {#sec-data}
+Our dataset also includes important measures of pollster performance, such as numeric grades, poll scores, and transparency scores.
+These metrics are calculated by considering bias, race difficulty, predictive error, and transparency, allowing us to assess the accuracy of each poll.
+The "pollscore," for instance, reflects a pollster’s past performance and reliability, adjusted for potential biases.
 
-## Overview
+### Data Cleaning
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+The raw data for this project, sourced from FiveThirtyEight, [@FiveThirtyEight] underwent a series of cleaning steps to prepare it for analysis.
+Initially, duplicate rows were removed to ensure that only unique observations remained, facilitated by the janitor package \[\@janitor\].
+A new binary variable, 'national', was created to indicate whether a poll was conducted at the national or state level.
+Missing values in the 'state' column were replaced with "Not Applicable," and numeric grades were evaluated to filter out low-quality pollsters, keeping only those with a numeric grade above 1.
+This cutoff was selected to retain mid to high-level pollsters for more reliable results.
+These steps were performed using functions from the dpylr package \[\@dplyr\].
+Furthermore, dates were standardized and converted into a proper format for analysis using the lubridate package \[\@lubridate\].
+Polls related to Kamala Harris were retained for further analysis, and percentage support values were transformed into actual numbers of supporters based on sample size.
+Additionally, pollster counts below five were excluded to focus on more reliable data sources.
+Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the arrow package \[\@arrow\].
 
-Overview text
-
-## Measurement
-
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
-
-## Outcome variables
-
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+### Summary of Variables of Interest
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
+# Assuming you have the dataset numeric_data already loaded
+#| include: false
 #| warning: false
@@
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+numeric_data <- analysis_data %>% select(numeric_grade, pollscore, sample_size, pct, num_support)
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+# Create a summary table
+summary_table <- numeric_data %>%
+  reframe(
+    `Variable` = c("numeric_grade", "pollscore", "sample_size", "pct", "num_support"),
+    `Mean` = c(mean(numeric_grade, na.rm = TRUE), 
+               mean(pollscore, na.rm = TRUE), 
+               mean(sample_size, na.rm = TRUE), 
+               mean(pct, na.rm = TRUE), 
+               mean(num_support, na.rm = TRUE)),
+    `Max` = c(max(numeric_grade, na.rm = TRUE), 
+              max(pollscore, na.rm = TRUE), 
+              max(sample_size, na.rm = TRUE), 
+              max(pct, na.rm = TRUE), 
+              max(num_support, na.rm = TRUE)),
+    `Min` = c(min(numeric_grade, na.rm = TRUE), 
+              min(pollscore, na.rm = TRUE), 
+              min(sample_size, na.rm = TRUE), 
+              min(pct, na.rm = TRUE), 
+              min(num_support, na.rm = TRUE)),
+    `Standard Deviation` = c(sd(numeric_grade, na.rm = TRUE), 
+                             sd(pollscore, na.rm = TRUE), 
+                             sd(sample_size, na.rm = TRUE), 
+                             sd(pct, na.rm = TRUE), 
+                             sd(num_support, na.rm = TRUE))
+  )
+
+kable(summary_table)
 ```
 
-Talk way more about it.
+## Variables
 
-## Predictor variables
+The data set contains a wide range of variables that are essential for understanding polling outcomes. The following are key variables:
 
-Add graphs, tables and text.
+-   **`poll_id`**: Unique identifier for each poll conducted.
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+-   **`pollster`**: The polling organization that conducted the poll (e.g., YouGov, TIPP).
+
+-   **`numeric_grade`**: A numeric rating indicating the pollster’s quality or reliability (e.g., 1.8).
+
+-   **`pollscore`**: A score representing the reliability of the pollster in terms of error and bias.
+
+-   **`state`**: The state where the poll was conducted (or "Not Applicable" for national polls).
+
+-   **`start_date` / `end_date`**: Dates during which the poll was conducted.
+
+-   **`sample_size`**: The number of respondents in the poll.
+
+-   **`population`**: The group surveyed (e.g., likely voters or registered voters).
+
+-   **`candidate_name`**: The name of the candidate being polled (e.g., Kamala Harris, Donald Trump).
+
+-   **`pct`**: The percentage of support for the candidate in the poll.
+
+-   **`methodology`**: The method used to conduct the poll (e.g. online panel)
+
+-   **`transparency_score`**: A grade for how transparent a pollster is, calculated based on how much information it discloses about its polls and weighted by recency.
+The highest Transparency Score is 10.
+
+##
+
+## Summary Statistics
+
+### 1) Average Pollster Quality (pollscore): mean, median, mode, min, max
+
+```{r}
+library(dplyr)
+
+# Generate summary statistics for national and state polls
+summary_stats <- analysis_data %>%
+  filter(!is.na(pollscore)) %>%  # Exclude rows with missing pollscore
+  group_by(national) %>%        # Group by poll_scope
+  summarize(
+    Count = n(),                   
+    Min = round(min(pollscore), 2),
+    `1st Qu.` = round(quantile(pollscore, 0.25), 2),
+    Median = round(median(pollscore), 2),
+    Mean = round(mean(pollscore), 2),
+    `3rd Qu.` = round(quantile(pollscore, 0.75), 2),
+    Max = round(max(pollscore), 2),
+    .groups = 'drop'  # Drop grouping structure for cleaner output
+  )
+
+# Print the summary statistics
+kable(summary_stats)
+
+```
+
+As shown above, the summary statistics table **\[CROSS-REFERENCE HERE\]** offers an overview of poll scores for national and state polls, with 2,971 and 2,953 entries, respectively.
+Both categories feature a minimum poll score of -1.5, which a high level of validity and reliability.
+The first quartile shows national polls at -0.9, while state polls are slightly lower at -1.1.
+Median scores are -0.3 for national polls and -0.4 for state polls, suggesting that half of the polls in both categories yield moderately negative scores, reinforcing a trend of favorable sentiment.
+The mean score is -0.37 for national polls, indicating a less reliable perception compared to the -0.51 mean for state polls.
+The third quartile for national polls (0.0) indicates that the upper 25% of scores are neutral, while state polls remain just below neutral at 0.7.
+The maximum score of 1.7 for national polls reflects a notable negative perception for at least one candidate, while the maximum for state polls is 0.7.
+Overall, these statistics illustrate a strong reliability of pollsters considering their errors and bias.
+
+### 2) Distribution of sample sizes: same as above
+
+```{r}
+library(ggplot2)
+library(dplyr)
+library(gridExtra)  # Load the gridExtra package
+
+# Create a new variable for sample size ranges
+analysis_data <- analysis_data %>%
+  mutate(sample_size_range = case_when(
+    sample_size <= 10000 ~ "0-10,000",
+    sample_size > 10000 & sample_size <= 20000 ~ "10,001-20,000",
+    sample_size > 20000 & sample_size <= 30000 ~ "20,001-30,000",
+    TRUE ~ "Above 30,000"
+  ))
+
+# Plot for 0-10,000
+p1 <- ggplot(analysis_data %>% filter(sample_size <= 10000), aes(x = sample_size)) +
+  geom_histogram(binwidth = 500, fill = "lightgreen", color = NA, alpha = 0.7) +
+  labs(title = "Sample Sizes 
+       (0-10,000)",
+       x = "Sample Size",
+       y = "Count") +
+  scale_x_continuous(limits = c(0, 10000)) +  # Set x limits for the first plot
+  scale_y_continuous(limits = c(0, 2300)) +  # Set y limits for the first plot
+  theme_minimal() +
+  theme(plot.title = element_text(size = 10),  # **Decreased font size of the title**
+        axis.title.y = element_text(margin = margin(r = 15)))  # **Increased gap for y-axis label**
+
+# Plot for 10,001 - 20,000
+p2 <- ggplot(analysis_data %>% filter(sample_size > 10000 & sample_size <= 20000), aes(x = sample_size)) +
+  geom_histogram(binwidth = 500, fill = "indianred", color = NA, alpha = 0.7) +
+  labs(title = "Sample Sizes 
+       (10,001-20,000)",
+       x = "Sample Size",
+       y = "Count") +
+  scale_x_continuous(limits = c(10001, 20000)) +  # Set x limits for the second plot
+  scale_y_continuous(limits = c(0, 50)) +  # Set y limits for the second plot
+  theme_minimal() +
+  theme(plot.title = element_text(size = 10),  # **Decreased font size of the title**
+        axis.title.y = element_text(margin = margin(r = 10)))  # **Increased gap for y-axis label**
+
+# Plot for 20,001 - 30,000
+p3 <- ggplot(analysis_data %>% filter(sample_size > 20000 & sample_size <= 30000), aes(x = sample_size)) +
+  geom_histogram(binwidth = 500, fill = "lightblue", color = NA, alpha = 0.7) +
+  labs(title = "Sample Sizes 
+       (20,001-30,000)",
+       x = "Sample Size",
+       y = "Count") +
+  scale_x_continuous(limits = c(20001, 30000)) +  # Set x limits for the third plot
+  scale_y_continuous(limits = c(0, 10)) +  # Set y limits for the third plot
+  theme_minimal() +
+  theme(plot.title = element_text(size = 10),  # **Decreased font size of the title**
+        axis.title.y = element_text(margin = margin(r = 5)))  # **Increased gap for y-axis label**
+
+# Arrange the plots side by side
+grid.arrange(p1, p2, p3, ncol = 3)  # ncol=3 arranges the plots in one row
+
+```
+
+The histograms **\[CROSS-REFERENCE HERE\]** reveal significant trends in sample sizes across the different ranges, indicating a general tendency towards smaller samples in the polling data.
+Most values within the 0-10,000 range cluster at the lower end, suggesting that many polls are conducted with limited respondent numbers, which may compromise the reliability and representativeness of the results.
+**However, this conclusion is very much dependent on what we determine to be an acceptable sample size for the results to be considered reliable**.
+In the 10,001-20,000 range, while there is some increase in sample sizes, the majority still falls within the 10,000-12,500 bracket, with few reaching the upper limits.
+
+The findings from the 20,001-30,000 range further illustrate the scarcity of larger sample sizes, with only a couple of values noted at each of the significant points.
+This distribution emphasizes a preference for smaller samples in the polling data analyzed.
+
+In the context of polling results, these findings suggest potential limitations in the accuracy and validity of the insights generated.
+Larger sample sizes are typically preferred in polling as they tend to yield more reliable estimates and reduce the margin of error, leading to a better representation of the population's views.
+Consequently, the predominance of smaller sample sizes in this data set may raise concerns about the robustness of the polling findings, potentially affecting their applicability in broader analyses or decision-making processes.
+**However, we must determine an acceptable and ideal sample size that would allow us to determine the reliability of these results.**
+
+# Polls of Polls to Forcast Winning Candidate
+
+## Forecasting Approach
+
+The polls of polls methodology is widely used in election prediction as it aggregates multiple polls to provide a more reliable estimate of voter support, rather than relying on any single poll.
+The goal is to reduce errors and biases present in individual polls by using a weighted average of many different polls.
+
+In our approach, we will employ linear modeling of voter support percentage (pct) on pollster and other independent variables such as sample size, poll recency, and poll scope (state vs.
+national).
+This will allow us to smooth out the inherent noise, biases, and variability across different pollsters.
+Once we obtain the predicted values from our model, we will weight these predictions based on the numeric grade (quality score) of each pollster to calculate an overall national estimate of the outcome.
+Additionally, we will separately compute estimates for key battleground states, where voter behavior can be more volatile and pivotal in deciding the final outcome of the election.
+This approach helps us capture both national trends and critical state-level dynamics.
+
+## **Exploration of Independent Variables**
+
+### Support for Kamala Harris by Sample Size and Polling Level
+
+\[\@fig-sample\] suggests that the relationship between support for Kamala Harris and sample size shows a slight positive trend, indicating a small increase in support as sample sizes grow.
+Both national (blue) and state (green) polls reflect this upward trend, although the effect is minimal.
+Since voting outcomes often depend on small margins, this subtle increase in support with larger sample sizes could still be relevant in predicting overall support.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-sample
+#| fig-cap: Sample Size vs. Support for Kamala Harris, Colored by Pollscope (National vs. State)
+
+# Custom colors for pollscope
+national_color <- "chartreuse3"
+state_color <- "steelblue"
+
+# Create the plot for Kamala Harris with the requested adjustments
+ggplot(analysis_data %>% filter(candidate_name == "Kamala Harris"), 
+       aes(x = sample_size, y = pct, color = factor(national))) +
+  geom_point(size = 0.5, alpha = 0.9) +  # Smaller dots with slight transparency
+  geom_smooth(aes(group = 1), method = "lm", se = FALSE, color = "azure4") +  # Linear model smoothing
+
+  # Custom colors for national (1 for National, 0 for State)
+  scale_color_manual(values = c("1" = national_color, "0" = state_color), 
+                     labels = c("National", "State")) +  # Pollscope categories with custom colors
+
+  # Same minimal styling
+  theme_minimal() +
+  theme(
+    panel.grid.major = element_blank(),  # Remove major gridlines
+    panel.grid.minor = element_blank(),  # Remove minor gridlines
+    legend.title = element_blank(),  # No title for the legend
+    legend.position = "bottom",  # Move legend to the bottom
+    plot.title = element_text(hjust = 0.5),  # Center the title
+    axis.line = element_line(color = "black"),  # Add axis lines
+    panel.border = element_blank(),  # Remove panel border
+    axis.line.y.right = element_line(color = "black"),  # Add right spine
+    axis.line.x.top = element_blank()  # Remove top spine
+  ) +
+  labs(x = "Sample Size", y = "Support (%)")  # No title, just axis labels
+
+```
+
+### Variation in Support for Kamala Harris Among Most Frequent Pollsters
+
+[@fig-top5pollsters] above shows the variation in support for Kamala Harris across the five most frequent pollsters.
+Notably, the median support levels differ between pollsters, indicating variability in central estimates of Harris’s support.
+For instance, Siena/NYT reports a higher median compared to Redfield & Wilton Strategies.
+Additionally, the differing ranges of support estimates highlight potential discrepancies in the methodologies, sample sizes, and biases employed by each pollster.
+This suggests that different pollster have different interpretations of candidate support, underlining the importance of aggregating data from multiple sources to avoid over-reliance on any single pollster.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-top5pollsters
+#| fig-cap: "Support for Kamala Harris across the top 5 pollsters by poll count."
+
+# Get top 5 pollsters by poll count
+top5_pollsters <- analysis_data %>%
+  filter(candidate_name == "Kamala Harris") %>%
+  count(pollster, sort = TRUE) %>%
+  top_n(5, n) %>%
+  pull(pollster)
+
+# Filter the data for only these top 5 pollsters
+top5_data <- analysis_data %>%
+  filter(pollster %in% top5_pollsters & candidate_name == "Kamala Harris")
+
+# Custom colors for boxplot fill
+boxplot_colors <- c("chartreuse3", "steelblue", "indianred", "purple", "darkorange")
+
+# Create the boxplots with thinner borders, smaller points, and rotated x-axis labels
+ggplot(top5_data, aes(x = pollster, y = pct, fill = pollster)) +
+  geom_boxplot(size = 0.2, outlier.size = 0.5) +  # Thinner box lines and smaller outliers
+  scale_fill_manual(values = boxplot_colors) +
+  theme_minimal() +  # Minimal theme
+  theme(
+    panel.grid.major = element_blank(),  # Remove major gridlines
+    panel.grid.minor = element_blank(),  # Remove minor gridlines
+    legend.position = "none",  # No legend needed for boxplots
+    axis.line = element_line(color = "black"),  # Add axis lines
+    plot.title = element_text(hjust = 0.5),  # Center title
+    axis.text.x = element_text(angle = 20, hjust = 1),  # Rotate x-axis labels for better readability
+    axis.line.y.right = element_line(color = "black"),  # Add right spine for clean look
+    axis.line.x.top = element_blank(),  # Remove top spine
+    panel.border = element_blank()  # Remove panel border
+  ) +
+  labs(x = "Pollster", y = "Support (%)")  # Labels
+
+```
+
+### Support for Kamala Harris Based on Recency of Polls and Pollster Quality
+
+In \[\@fig-enddate\], we observe a slight increasing trend in support for Kamala Harris as the end date of polls progresses towards October.
+This may indicate a gradual improvement in her polling performance over time.
+The color gradient, reflecting the pollster's error and bias (Pollscore), reveals variations in bias: greener points represent more negative biases, while redder points indicate more positive biases.
+
+While the recency of the polls shows a stable trend in Harris's support, the wide distribution of bias (Pollscore) values suggests that pollster quality can vary substantially.
+These variations could influence the model's predictions if left unaddressed.
+Hence, including both recency and pollster quality as predictor variables in the model might be valuable to account for such variations.
+Further model diagnostics and validation will help decide if these variables are significant.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-enddate
+#| fig-cap: "Support for Kamala Harris based on End Date and colored by Pollster's Error and Bias."
+
+# Define the color gradient based on pollscore
+ggplot(analysis_data %>% filter(candidate_name == "Kamala Harris"), aes(x = end_date, y = pct, color = pollscore)) +
+  geom_point(size = 0.8, alpha = 1) +  # Adjust point size and transparency
+  geom_smooth(method = "lm", se = FALSE, color = "azure4") +  # Add a smooth trend line for overall support
+  scale_color_gradient2(low = "blue", mid = "green", high = "red", midpoint = 0) +  # Color gradient
+  theme_minimal() +  # Minimal theme as usual
+  theme(
+    panel.grid.major = element_blank(),  # Remove major gridlines
+    panel.grid.minor = element_blank(),  # Remove minor gridlines
+    legend.position = "bottom",  # Legend at the bottom
+    axis.line = element_line(color = "black"),  # Add axis lines
+    panel.border = element_blank(),  # Remove panel border
+    axis.line.y.right = element_line(color = "black"),  # Add right spine for clean look
+    axis.line.x.top = element_blank()  # Remove top spine
+  ) +
+  labs(x = "Date Poll Ended",
+       y = "Support (%)",
+       color = "Error and Bias of Pollster (Low to High)")  # Updated label for color gradient
+
+```
+
+### Political Leanings of Different State
+
+[@fig-mapinterac] provides a state-level representation of Kamala Harris's average support across the U.S., with the darker blue shades indicating states where her support is higher, and the redder tones showing where her support is lower.
+Gray areas represent states where we have no polling data available.
+
+The gradient highlights notable geographic trends.
+Harris has the strongest support in states like Vermont, Massachusetts, and New York, while states like Utah and West Virginia show considerably lower levels of support.
+A distinct pattern emerges in the West, where coastal states like California favor Harris more strongly, contrasting with inland states like Wyoming and Utah.
+
+This reveals that political leanings differ greatly by state, and this regional variation should be controlled for in any regression model analyzing poll data.
+The map serves as a useful tool for visualizing regional variations in voter support for Kamala Harris as reflected in the available polling data.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-mapinterac
+#| fig-cap: "Average Support for Kamala Harris in Different States estimated by Polls"
+  # For string manipulation
+
+# Filter out non-continental states (Hawaii and Alaska)
+us_states_filtered <- us_states %>% 
+  filter(!NAME %in% c("Alaska", "Hawaii"))
+
+# Ensure correct merge between map data and Kamala's support data
+map_data_full <- left_join(us_states_filtered, mapping_data, by = c("NAME" = "state"))
+
+# Create centroids for state labels
+state_centroids <- st_centroid(us_states_filtered)
+
+# Plot using geom_sf() for sf objects
+p <- ggplot(map_data_full) +
+  geom_sf(aes(fill = mean_support_kamala, text = paste("State:", str_to_title(NAME), "<br>Average Support:", round(mean_support_kamala, 2))), color = "white") +  # Capitalize state names
+  scale_fill_distiller(palette = "RdBu", direction = 1, na.value = "gray", name = "Kamala's Support") +  # Reverse RdBu color gradient
+  geom_text(data = state_centroids, aes(geometry = geometry, label = STUSPS), stat = "sf_coordinates", size = 2) +  # State labels
+  labs(title = "Support for Kamala Harris by State", 
+       caption = "States with no polling data are shown in gray") +
+  theme_minimal() +
+  theme(
+    legend.position = "right",
+    plot.title = element_text(hjust = 0.5),
+    panel.grid = element_blank(),  # Remove grid lines
+    axis.title = element_blank(),  # No axis titles for maps
+    axis.text = element_blank(),   # No axis labels
+    axis.ticks = element_blank()   # No axis ticks
+  )
+
+# Turn ggplot to plotly for interactivity, add hover template for better control over display
+ggplotly(p, tooltip = "text")
+
+```
 
@@
 
-# Emerson College Polling Methodology (October 14-16, 2024)
+# Additional data details
 
-## Overview
-Emerson College Polling, known for its extensive online polling methods, conducted a survey from October 14 to 16, 2024, targeting 1,000 likely voters.
-The poll measured voter preferences between Kamala Harris and Donald Trump, revealing a near tie with Harris at 50% and Trump at 49%.
+# Model details {#sec-model-details}
 
-## Population, Frame, and Sample
-The target population is likely voters in the U.S.
-elections.
-These voters are determined by the voters' likeliness to vote in the upcoming elections and the voter's voting history.
-Both of these factors are self-reported.
+## Posterior predictive check
 
-The sampling frame is the voters that are on the Aristotle list and on the online panel provided by CINT.
-Aristotle is an online supplier of voter files and CINT is a software company that connects researchers, such as Emerson, to online panels.
-This approach helps ensure a representative sample by accurately reflecting the electorate.
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
 
-The sample size is the 1,000 likely voters chosen arbitrarily from the sampling frame.
-Likely voters are determined by a combination of voter history, registration status, and demographic data.
-These factors are self-reported.
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
 
-## Sample Recruitment and Approach
-Emerson College used a mixed-mode sampling approach for its polling. Specifically, three primary modes were used to collect data for its October 14-16, 2024 poll:
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-ppcheckandposteriorvsprior
+#| layout-ncol: 2
+#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
 
-- MMS-to-web text surveys: Respondents were contacted via text messages sent to cell phones.
-These messages directed participants to complete the survey online.
-This method was conducted using voter lists provided by Aristotle.
+pp_check(first_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
 
-- Interactive Voice Response (IVR): Landlines were targeted using automated phone calls where respondents could answer questions using their keypads.
-The contact information for this was also sourced from Aristotle’s voter lists.
+posterior_vs_prior(first_model) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
+```
 
-- Online Panel from CINT: Emerson utilized a pre-screened, opt-in online panel of voters provided by CINT.
+## Diagnostics
 
-## Advantages and Trade-offs in Mixed-Mode Sampling
-The multi-mode sampling approach used in this poll has several advantages:
+@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
 
-- Wider demographic reach: Having a combination of data collection methods allows Emerson to gather responses from a broader demographic group.
-In this polling, MMS-to-web surveys mostly gather responses from younger voters and those who primarily use mobile devices.
-IVR target older and rural voters, who are more likely to respond via landlines.
-Online panels is more efficient in reaching tech-savvy individuals.
+@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 
-- Cost-efficient: Automated technologies reduce the cost compared to live interviews. In a polling context, this allows for a larger sample size.
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-stanareyouokay
+#| fig-cap: "Checking the convergence of the MCMC algorithm"
+#| fig-subcap: ["Trace plot", "Rhat plot"]
+#| layout-ncol: 2
 
-However, the trade-offs include:
+plot(first_model, "trace")
 
-- Coverage bias: IVR only reaches individuals with landlines, which may skew results towards older demographics.
+plot(first_model, "rhat")
+```
 
-- Self-selection bias: Respondents in online panels tend to be more engaged, which can lead to over-representation of politically active individuals.
-
-## Non-response Handling
-Emerson addresses non-response using a weighting system.
-The data is adjusted based on key demographic variables like age, gender, race, education, and party affiliation, making the sample more representative of the electorate.
-Weights are applied to compensate for under- or over-represented groups, ensuring balanced results.
-
-## Questionnaire Design
-The Emerson survey features a straightforward, efficient design focused on key issues and voter preferences. Strengths of the questionnaire include:
-
-- Simplicity: Clear, concise questions makes the survey easy for respondents to understand and to provide accurate answers.
-
-- Topical relevance: The poll zeroes in on vote preference and demographic splits, delivering timely insights during election cycles.
-
-However, weaknesses include:
-
-- Lack of nuance: While the survey captures basic preferences, it does not delve deeply into voters' motivations or the factors driving their choices.
-
-- Mode limitations: Different survey modes can affect how respondents engage with the questions. This can potentially influence the depth and accuracy of voter responses.
-
-## Conclusion
-Emerson College Polling's mixed-mode methodology is effective in balancing cost with broad demographic reach.
-Its weighting techniques contribute to reliable and accurate results despite non-responses.
-While the use of multiple survey modes increases population representation, possible biases—such as lack of nuance and mode effects need to be taken into consideration when interpreting results.
-
-In sum, this method is a strong predictor of election outcomes and continues to adapt to the evolving challenges of modern polling.
-
-# Idealized Methodology
-
-## Overview
-This methodology outlines an election forecasting plan with a budget of $100,000.
-Heavily inspired by TIPP Insights and Emerson College's hybrid approaches, the survey methodology balances cost-efficiency and data accuracy by using different modes of data collection and wide demographic reach.
-
-This forecasting plan will predict the popular vote and Electoral College outcomes, emphasizing key swing states and underrepresented voter groups.
-
-## Sampling Approach
-This idealized methodology uses a multi-mode hybrid sampling approach. This allows the use of both probability-based and non-probability-based methods:
-
-- Probability-based methods:
-
-  - Live phone interviews (landline and cell): As used by TIPP, live phone interviews is a direct and personal method of collecting responses.
-Voter lists would be sourced from Aristotle to ensure we contact likely voters.
-This mode reaches older voters and voters with less internet connection.
-
-  - Interactive Voice Response (IVR): Similar to Emerson’s approach, IVR for landline users balances costs and ensures demographic representation, particularly among older or less tech-savvy populations.
-This method captures quick responses and is less expensive than live interviews.
-
-- Non-probability methods:
-
-  - MMS-to-web (mobile): Following Emerson’s MMS-to-web strategy, we would send multimedia messages to likely voters’ cell phones, encouraging them to complete an online survey.
-This method works well for younger, mobile-first voters, helping fill in gaps where traditional methods might miss.
-
-  - Online panel recruitment via CINT: To mirror TIPP’s online panel approach, a portion of the survey would be conducted via an opt-in panel of pre-screened voters from CINT, capturing tech-savvy respondents.
-CINT’s platform ensures that different voter groups, including hard-to-reach groups, are represented in the polling.
+\newpage
 

### 2024-10-22 scripts/03-clean_data.R

@@ -92,3 +92,21 @@ cleaned_data <- cleaned_data %>%
 #### Save data ####
 
 write_parquet(cleaned_data, "data/02-analysis_data/analysis_data.parquet")
+
+#### Creating Data For Mapping ####
+
+deduped_data <- raw_data %>%
+  distinct()  # Remove duplicate rows
+
+# Step 2: Filter for Kamala Harris and non-empty states
+harris_data <- deduped_data %>%
+  filter(candidate_name == "Kamala Harris" & !is.na(state))
+
+# Step 3: Group by state and calculate mean support for Kamala Harris
+mapping_data <- harris_data %>%
+  group_by(state) %>%
+  summarise(mean_support_kamala = mean(pct, na.rm = TRUE))
+
+# Step 4: Save the aggregated data to a parquet file
+write_parquet(mapping_data, "data/02-analysis_data/mapping_data.parquet")
+

### 2024-10-22 scripts/03-clean_data.R
Calculated recency of polls.

@@ -89,6 +89,14 @@ cleaned_data <- cleaned_data %>%
 cleaned_data <- cleaned_data %>%
   filter(end_date >= as.Date("2024-07-21"))
 
+# Calculate recency of every poll and make a column 
+
+reference_date <- max(cleaned_data$end_date, na.rm = TRUE)  # Most recent poll date
+
+analysis_data <- cleaned_data %>%
+  mutate(recency = as.numeric(reference_date - end_date))
+
+
 #### Save data ####
 
 write_parquet(cleaned_data, "data/02-analysis_data/analysis_data.parquet")

### 2024-10-22 paper/paper.qmd
fixed typo and add link to survey
downloaded some other packages.
finished summary statistics and data section + added all of the LLM usage instances and documentation

--- 2024-10-22 paper/paper.qmd
@@
 
-Overview paragraph
+# Data
 
-Estimand paragraph
+## Data Set Overview
 
-Results paragraph
+add a summary statistics table here maybe?
 
-Why it matters paragraph
+```{r}
+library(arrow)
+analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+#head(analysis_data, 20)
+colnames(analysis_data)
+#print(nrow(analysis_data))
+```
 
-# Data {#sec-data}
+## Variables
 
-## Overview
+The data set contains a wide range of variables that are essential for understanding polling outcomes. The following are key variables:
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+-   **`poll_id`**: Unique identifier for each poll conducted.
 
-Overview text
+-   **`pollster`**: The polling organization that conducted the poll (e.g., YouGov, TIPP).
+
+-   **`numeric_grade`**: A numeric rating indicating the pollster’s quality or reliability (e.g., 1.8).
+
+-   **`pollscore`**: A score representing the reliability of the pollster in terms of error and bias.
+
+-   **`state`**: The state where the poll was conducted (or "Not Applicable" for national polls).
+
+-   **`start_date` / `end_date`**: Dates during which the poll was conducted.
+
+-   **`national`**: A binary variable which determines where a poll was national (1) or state (0)
+
+-   **`sample_size`**: The number of respondents in the poll.
+
+-   **`population`**: The group surveyed (e.g., likely voters or registered voters).
+
+-   **`candidate_name`**: The name of the candidate being polled (e.g., Kamala Harris, Donald Trump).
+
+-   **`pct`**: The percentage of support for the candidate in the poll.
+
+-   **`methodology`**: The method used to conduct the poll (e.g. online panel)
+
+-   **`num_support`**: The number of people supporting a candidate in a poll
+
+## High Level Data Cleaning
+
+blah blah blah
+
+## Summary Statistics
+
+### 1) Average Pollster Quality by Poll Type (pollscore vs state/national): mean, median, mode, min, max
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+
+library(dplyr)
+
+# Generate summary statistics for national and state polls
+summary_stats <- analysis_data %>%
+  filter(!is.na(pollscore)) %>%  # Exclude rows with missing pollscore
+  group_by(national) %>%         # Group by the binary 'national' column (1 for national, 0 for state)
+  summarize(
+    Count = n(),
+    Min = round(min(pollscore), 2),
+    `1st Qu.` = round(quantile(pollscore, 0.25), 2),
+    Median = round(median(pollscore), 2),
+    Mean = round(mean(pollscore), 2),
+    `3rd Qu.` = round(quantile(pollscore, 0.75), 2),
+    Max = round(max(pollscore), 2),
+    .groups = 'drop'  # Drop grouping structure for cleaner output
+  ) %>%
+  # Rename 'national' to 'Poll Type' and replace values
+  mutate(`Poll Type` = ifelse(national == 1, "National", "State")) %>%
+  select(`Poll Type`, Count, Min, `1st Qu.`, Median, Mean, `3rd Qu.`, Max)  # Reorder columns
+
+# Print the summary statistics
+print(summary_stats)
+
+```
+
+As shown above, the summary statistics table **\[CROSS-REFERENCE HERE\]** offers an overview of poll scores for national and state polls, with 427 and 696 entries, respectively.
+Both categories feature a minimum poll score of -1.5, which a high level of validity and reliability.
+The first quartile shows national polls at -0.9, while state polls are slightly lower at -1.1.
+Median scores are -0.3 for national polls and -0.4 for state polls, suggesting that half of the polls in both categories yield moderately negative scores, reinforcing a trend of favorable sentiment.
+The mean score is -0.39 for national polls, indicating a less reliable perception compared to the -0.54 mean for state polls.
+The third quartile for national and state polls (-0.3) indicates that the upper 25% of scores are slightly above neutral.
+The maximum score of 1.7 for national polls reflects a notable negative perception for at least one candidate, while the maximum for state polls is 0.6.
+Overall, these statistics illustrate a strong reliability of pollsters considering their errors and bias.
+
+### 2) Distribution of Sample Sizes
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+
+library(ggplot2)
+library(dplyr)
+library(gridExtra)
+
+# Adjust the sample size range categories
+analysis_data <- analysis_data %>%
+  mutate(sample_size_range = case_when(
+    sample_size <= 6000 ~ "0-6,000",
+    sample_size > 6000 & sample_size <= 13000 ~ "6,001-13,000",
+    TRUE ~ "Above 13,000"  # Catch any values above 13,000
+  ))
+
+# Plot for 0-6,000
+p1 <- ggplot(analysis_data %>% filter(sample_size <= 6000), aes(x = sample_size)) +
+  geom_histogram(binwidth = 500, fill = "lightgreen", color = NA, alpha = 0.7) +
+  labs(title = "Sample Sizes (0-6,000)",
+       x = "Sample Size",
+       y = "Count") +
+  scale_x_continuous(limits = c(0, 6000)) +  # Set x limits for the first plot
+  scale_y_continuous(limits = c(0, 500)) +  # Adjust y limits based on data
+  theme_minimal() +
+  theme(plot.title = element_text(size = 10, hjust = 0.5),  # **Center the title**
+        axis.title.y = element_text(margin = margin(r = 15)))  # **Increased gap for y-axis label**
+
+# Plot for 6,001 - 13,000
+p2 <- ggplot(analysis_data %>% filter(sample_size > 6000 & sample_size <= 13000), aes(x = sample_size)) +
+  geom_histogram(binwidth = 500, fill = "indianred", color = NA, alpha = 0.7) +
+  labs(title = "Sample Sizes (6,001-13,000)",
+       x = "Sample Size",
+       y = "Count") +
+  scale_x_continuous(limits = c(6001, 13000)) +  # Set x limits for the second plot
+  scale_y_continuous(limits = c(0, 20)) +  # Adjust y limits based on data
+  theme_minimal() +
+  theme(plot.title = element_text(size = 10, hjust = 0.5),  # **Center the title**
+        axis.title.y = element_text(margin = margin(r = 10)))  # **Increased gap for y-axis label**
+
+# Arrange the two plots side by side
+grid.arrange(p1, p2, ncol = 2)  # ncol=2 arranges the plots in one row
+
+```
+
+The histograms **\[CROSS-REFERENCE HERE\]** reveal significant trends in sample sizes across the different ranges, indicating a general tendency towards smaller samples in the polling data.
+**In Plot 1**, most of the values are concentrated on the left side, with very few values extending beyond the 4,500 mark.
+This clustering suggests that many polls are conducted with limited respondent numbers, which may compromise the reliability and representativeness of the results.
+
+In **Plot 2**, the values are somewhat evenly divided between 6,000 and the maximum sample size of 12,084.
+This distribution shows a broader representation of sample sizes, indicating that while there are still smaller samples, there are also a notable number of polls that feature larger sample sizes.
+This contrast with Plot 1 highlights a shift towards more robust polling efforts.
+
+The findings from these histograms suggest potential limitations and advantages in the accuracy and validity of the insights generated from the polling data.
+Smaller sample sizes are typically less desirable in polling, as they can yield less reliable estimates and increase the margin of error.
+Consequently, the predominance of smaller sample sizes in Plot 1 raises concerns about the robustness of the polling findings.
+However, the more even distribution in Plot 2 suggests that some polls conducted with larger sample sizes could enhance the reliability of the results.
+
+Ultimately, the analysis emphasizes the need to consider sample sizes critically when interpreting polling data, as larger sample sizes tend to provide a better representation of the population's views.
+Thus, while many polls may fall short of ideal sample sizes, the existence of larger samples in the dataset can offer valuable insights into public opinion.
+**However, we need to determine an ideal and acceptable sample size that would be enough for the polling results to be considered reliable, before we conclude smaller samples as being unreliable.**
+
+### 3) Summary Statistics Table for Numeric Variables
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+
+library(dplyr)
+library(modelsummary)
+library(tools)  # Make sure to load the tools package for toTitleCase()
+
+# Set options to use kableExtra for table formatting
+options(modelsummary_factory_default = 'kableExtra')
+
+# Create the summary statistics table for the specified numerical variables
+summary_table <- analysis_data %>%
+  select(numeric_grade, pollscore, sample_size, num_support) %>%
+  # Remove underscores and capitalize the first letter of each word
+  rename_with(~ toTitleCase(gsub("_", " ", .))) %>%
+  datasummary_skim(
+    fun_numeric = list(
+      Mean = mean,
+      SD = sd,
+      Min. = min,
+      Median = median,
+      Max. = max
+    )
+  )
+
+# Display the summary statistics table
+print(summary_table)
+
+```
+
+The summary statistics **\[CROSS REFERENCE HERE\]** for the numeric variables provide significant insights into the polling data.
+The **numeric grade** has a mean of 2.3 and a standard deviation of 0.6, suggesting that most polls are graded around a middle value, though there is moderate variability.
+The range, with a minimum of 1.1 and a maximum of 3.0, indicates the presence of both low-quality and higher-quality polls within the dataset.
+
+The **pollscore** averages at −0.5, with a standard deviation of 0.6.
+Since lower poll scores indicate more favorable outcomes for certain candidates or parties, this negative mean suggests that the polls are, on average, skewing positively towards their targets.
+The minimum of −1.5 and maximum of 1.7 further illustrate the wide distribution of poll results, highlighting the variability in outcomes across different polls.
+
+Looking at the **sample size**, the mean of 1962.5, accompanied by a substantial standard deviation of 2608.0, reflects significant variability in respondent numbers.
+With a minimum sample size of 382 and a maximum of 12,084, it is evident that while many polls are conducted with relatively small sample sizes, a few achieve much larger respondent pools.
+This variability can enhance the reliability of findings, particularly for those polls with larger sample sizes.
+
+Finally, the **number of supporters** reported has a mean of 949.2 and a standard deviation of 1290.0, indicating that while the average number of supporters is relatively high, there is considerable variability.
+The minimum value of 172 and maximum of 6042 underscores the diverse contexts in which these polls were conducted.
+
+Overall, these summary statistics emphasize the variability and range present in the polling data, which may impact the reliability and interpretation of the findings.
+The mixture of low and high values in numeric grade, pollscore, and sample size points to the importance of careful analysis when drawing conclusions from these polling efforts.
+
+### 4) Top 10 Pollsters and their Average Pollscores and Numeric Grades
+
+```{r}
+library(dplyr)
+
+# Get the top 10 pollsters by frequency along with their average pollscore and numeric grade
+top_pollsters <- analysis_data %>%
+  group_by(pollster) %>%           # Group by the 'pollster' variable
+  summarize(
+    Count = n(),                   # Count the number of occurrences
+    Average_Pollscore = round(mean(pollscore, na.rm = TRUE), 2),  # Average pollscore
+    Average_Numeric_Grade = round(mean(numeric_grade, na.rm = TRUE), 2),  # Average numeric grade
+    .groups = 'drop'               # Drop grouping structure for cleaner output
+  ) %>%
+  arrange(desc(Count)) %>%         # Arrange in descending order of count
+  slice_head(n = 10)                # Select the top 10 pollsters
+
+# Rename the columns to remove underscores and capitalize 'Pollster'
+colnames(top_pollsters) <- c("Pollster", "Count", "Average Pollscore", "Average Numeric Grade")
+
+# Print the top pollsters table
+print(top_pollsters)
+
+```
+
+The examination of the top 10 pollsters in **\[CROSS REFERENCE HERE\]** highlights their significance in electoral forecasting.
+**Morning Consult**, with the highest volume of polls (235), shows a slightly optimistic average pollscore of −0.3, indicating some reliability and validity.
+Conversely, **Siena/NYT** stands out with the lowest average pollscore of −1.5 and the highest numeric grade of 3.0, suggesting a rigorous methodology that enhances its credibility in predicting electoral outcomes.
+
+In contrast, **Redfield & Wilton Strategies** displays an average pollscore of 0.4, which may create a misleadingly pessimistic view of voter sentiment.
+This underscores the necessity of considering multiple polling sources for a nuanced understanding of electoral dynamics.
+
+The presence of several pollsters with negative scores indicates a general trend towards specific candidates, which is crucial for shaping campaign strategies.
+Overall, these findings reinforce the critical role of trustworthy polling in influencing public perception and political strategy during elections.
+
+### 5) Histogram of End Dates
+
+## Graphs for Imp. Variables (relationships)
 
@@
 
-## Outcome variables
+Public support for political candidates is an essential part of understanding electoral dynamics for a variety of reasons.
+Public polls provide insights into the preferences and beliefs of certain groups of people, usually voters in specific states or districts, and sometimes to also correlate preferences to demographic information (race, age, neighborhood, etc).
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+The results help inform policy decisions, help track the likely outcomes of elections, help government officials and candidates decide where to invest time, funds, and resources to get the desired results.
+Polling organizations aim to capture data by surveying a representative sample of voters, free from government intervention.
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+Polling organizations conduct these surveys through various methods. For our particular data set, the following methods have been used to gauge public sentiment:
+
+-   App panel
+
+-   Online panel
+
+-   Email
+
+-   Online ad
+
+-   IVR (interactive voice response)
+
+-   Live phone
+
+-   Text
+
+-   Text-to-web
+
+-   Probability panel
+
+-   Mail-to-phone
+
+Each pollster chooses a method(s) based on what they find to be most convenient, efficient, and providing the best results.
+Moreover, there are some variables which affect the validity and reliability of the polling results, such as sample size, methodology, population type, etc.
+Larger samples generally lead to more reliable estimates of public opinion.
+Some methodologies might be considered more reliable due to certain factors.
+The pollsters also use different sampling methods (random, stratified, etc) which can affect the validity of the polling results as well.
+
+After collecting the required data, the polling organizations transform these results into more easily comprehensible variables such as numeric grade, poll score, transparency score, and a percentage.
+These scores are calculated through an extensive process of calculating excess error and bias, adjusting for race difficulty, calculating predictive error and bias, calculating pollster transparency, and combining everything into one single ‘POLLSCORE’, while accounting for luck as well.
+
+Variables such as pollster rating and transparency score help validate the reliability of the pollster and the polling results they publish.
+
+Each entry in the data set represents a glimpse of public opinion at a specific time and location, capturing measured support percentages alongside contextual variables such as pollster, methodology, and sample size.
+This extensive approach allows for an enhanced understanding of public support dynamics, facilitating informed analyses of electoral outcomes.
+
+# Polls of Polls to Forecast Winning Candidate
+
+## Forecasting Approach
+
+The polls of polls methodology is widely used in election prediction as it aggregates multiple polls to provide a more reliable estimate of voter support, rather than relying on any single poll.
+The goal is to reduce errors and biases present in individual polls by using a weighted average of many different polls.
+In our approach, we calculate the weights based on several important variables.
+Sample size is included because larger polls are generally more reliable and have smaller margins of error.
+Pollscore (or pollster rating) reflects the empirical quality of a pollster, balancing both accuracy and bias, allowing us to give more weight to those with a proven track record.
+We also account for the pollscope, differentiating between national and state-level polls, as state polls may offer more localized insights into voter sentiment.
+We also use recency as more recent polls give us a more accurate estimate of voter sentiment.
+
+However, we acknowledge that even with these adjustments, there may still be underlying biases from individual pollsters toward the candidates.
+To account for this, we will select a baseline pollster with a strong rating, high transparency, and minimal historical error, and regress the support percentage (pct) on pollster, including the same variables as controls whcih was used in the weighting method, such as sample size, pollscore and pollscope.Additionally, we will control for the state of the poll, as support levels can vary regionally.
+Therefore each pollster's coefficient in the regression model will represent the bias of that pollster relative to the baseline pollster.
+Controlling for these factors is crucial because it ensures that the differences between pollsters are not driven by other factors, such as differences in sample size or the types of polls they conduct, but by their inherent bias in support estimates.
+Once we have calculated the weights, we will use them to generate a weighted average of support for Kamala Harris and Donald Trump, ultimately comparing their weighted support to forecast the likely winner of the election.
+
+## **Exploration of Control Variables**
+
+### Investigating the Effect of Pollscore on Candidate Support
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+
+# Define pollscore ranges (you can adjust the binwidth if needed)
+cleaned_data <- analysis_data %>%
+  mutate(pollscore_range = cut(pollscore, breaks = seq(-1.5, 1, by = 0.3), include.lowest = TRUE))
+
+# Calculate the average support for each candidate within each pollscore range
+average_support <- cleaned_data %>%
+  group_by(pollscore_range, candidate_name) %>%
+  summarise(avg_support = mean(pct, na.rm = TRUE)) %>%
+  ungroup()
+
+# Create a bar plot
+ggplot(average_support, aes(x = pollscore_range, y = avg_support, fill = candidate_name)) +
+  geom_bar(stat = "identity", position = "dodge") +  # Dodge to separate bars for each candidate
+  scale_fill_manual(values = c("Donald Trump" = "indianred", "Kamala Harris" = "steelblue")) +  # Custom colors
+  labs(title = "Average Support for Candidates by Pollscore Range",
+       x = "Pollscore Range",
+       y = "Average Support (%)",
+       fill = "Candidate") +
+  theme_minimal() +  # Minimal theme
+  theme(
+    legend.position = "bottom",
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
+
+### Investigating the Effect of Poll Recency on Candidate Support
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+# Custom colors for candidates
+trump_color <- "indianred"
+harris_color <- "steelblue"
+
+# Create the plot with the requested adjustments (recency vs support)
+ggplot(analysis_data, aes(x = recency, y = pct, color = candidate_name)) +
+  geom_point(size = 0.5, alpha = 0.7) +
+  geom_smooth(method = "lm", se = FALSE) + # Smaller dots with slight transparency
+  scale_color_manual(values = c("Donald Trump" = trump_color, "Kamala Harris" = harris_color)) +  # Custom colors
+  theme_minimal() +  # Remove grey background
+  theme(
+    panel.grid.major = element_blank(),  # Remove major gridlines
+    panel.grid.minor = element_blank(),  # Remove minor gridlines
+    legend.title = element_blank(),  # Remove "candidate_name" as title
+    legend.position = "bottom",  # Move legend to the bottom
+    plot.title = element_text(hjust = 0.5),  # Center the title
+    axis.line = element_line(color = "black"),  # Add axis lines
+    panel.border = element_blank(),  # Remove panel border
+    axis.line.y.right = element_line(color = "black"),  # Add right spine for Kamala Harris
+    axis.line.x.top = element_blank()  # Remove top spine
+  ) +
+  labs(title = "Recency vs Support for Kamala Harris and Donald Trump",
+       x = "Poll Recency (days since poll end date)", 
+       y = "Support (%)") +
+  facet_wrap(~ candidate_name, scales = "free_y")  # Separate plots for each candidate
+
+```
+
+### Investigate the effect of sample size on support estimates
+
+```{r}
+#| warning: false
+#| message: false
 #| echo: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+# Custom colors for candidates
+trump_color <- "indianred"
+harris_color <- "steelblue"
+
+# Create the plot with the requested adjustments
+ggplot(analysis_data, aes(x = sample_size, y = pct, color = candidate_name)) +
+  geom_point(size = 0.5, alpha = 0.9) + # Smaller dots with slight transparency
+  geom_smooth(method = "lm", se = FALSE)+
+
+  scale_color_manual(values = c("Donald Trump" = trump_color, "Kamala Harris" = harris_color)) +  # Custom colors
+  theme_minimal() +  # Remove grey background
+  theme(
+    panel.grid.major = element_blank(),  # Remove major gridlines
+    panel.grid.minor = element_blank(),  # Remove minor gridlines
+    legend.title = element_blank(),  # Remove "candidate_name" as title
+    legend.position = "bottom",  # Move legend to the bottom
+    plot.title = element_text(hjust = 0.5),  # Center the title
+    axis.line = element_line(color = "black"),  # Add axis lines
+    panel.border = element_blank(),  # Remove panel border
+    axis.line.y.right = element_line(color = "black"),  # Add right spine for Kamala Harris
+    axis.line.x.top = element_blank()  # Remove top spine
+  ) +
+  labs(title = "Sample Size vs Support for Kamala Harris and Donald Trump",
+       x = "Sample Size", 
+       y = "Support (%)") +
+  facet_wrap(~ candidate_name, scales = "free_y")
 ```
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+### Distribution of support for Candidates by State
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
+# Calculate mean support for each candidate in each state
+state_support <- analysis_data %>%
+  group_by(state, candidate_name) %>%
+  summarise(mean_support = mean(pct)) %>%
+  ungroup()
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+# Pivot the data to have one row per state with separate columns for Harris and Trump
+state_support_wide <- state_support %>%
+  pivot_wider(names_from = candidate_name, values_from = mean_support)
 ```
 
-Talk way more about it.
+```{r}
+# Calculate mean support for each candidate in each state
+state_support <- analysis_data %>%
+  group_by(state, candidate_name) %>%
+  summarise(mean_support = mean(pct))
 
-## Predictor variables
+# Calculate the proportion of support for Kamala Harris
+state_support <- state_support %>%
+  mutate(harris_proportion = Kamala_Harris / (Kamala_Harris + Donald_Trump))
 
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+```
 
@@
 
-# Emerson College Polling Methodology (October 14-16, 2024)
+# Additional data details
 
-## Overview
-Emerson College Polling, known for its extensive online polling methods, conducted a survey from October 14 to 16, 2024, targeting 1,000 likely voters.
-The poll measured voter preferences between Kamala Harris and Donald Trump, revealing a near tie with Harris at 50% and Trump at 49%.
+# Model details {#sec-model-details}
 
-## Population, Frame, and Sample
-The target population is likely voters in the U.S.
-elections.
-These voters are determined by the voters' likeliness to vote in the upcoming elections and the voter's voting history.
-Both of these factors are self-reported.
+## Posterior predictive check
 
-The sampling frame is the voters that are on the Aristotle list and on the online panel provided by CINT.
-Aristotle is an online supplier of voter files and CINT is a software company that connects researchers, such as Emerson, to online panels.
-This approach helps ensure a representative sample by accurately reflecting the electorate.
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
 
-The sample size is the 1,000 likely voters chosen arbitrarily from the sampling frame.
-Likely voters are determined by a combination of voter history, registration status, and demographic data.
-These factors are self-reported.
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
 
-## Sample Recruitment and Approach
-Emerson College used a mixed-mode sampling approach for its polling. Specifically, three primary modes were used to collect data for its October 14-16, 2024 poll:
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-ppcheckandposteriorvsprior
+#| layout-ncol: 2
+#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
 
-- MMS-to-web text surveys: Respondents were contacted via text messages sent to cell phones.
-These messages directed participants to complete the survey online.
-This method was conducted using voter lists provided by Aristotle.
+pp_check(first_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
 
-- Interactive Voice Response (IVR): Landlines were targeted using automated phone calls where respondents could answer questions using their keypads.
-The contact information for this was also sourced from Aristotle’s voter lists.
+posterior_vs_prior(first_model) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
+```
 
-- Online Panel from CINT: Emerson utilized a pre-screened, opt-in online panel of voters provided by CINT.
+## Diagnostics
 
-## Advantages and Trade-offs in Mixed-Mode Sampling
-The multi-mode sampling approach used in this poll has several advantages:
+@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
 
-- Wider demographic reach: Having a combination of data collection methods allows Emerson to gather responses from a broader demographic group.
-In this polling, MMS-to-web surveys mostly gather responses from younger voters and those who primarily use mobile devices.
-IVR target older and rural voters, who are more likely to respond via landlines.
-Online panels is more efficient in reaching tech-savvy individuals.
+@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 
-- Cost-efficient: Automated technologies reduce the cost compared to live interviews. In a polling context, this allows for a larger sample size.
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-stanareyouokay
+#| fig-cap: "Checking the convergence of the MCMC algorithm"
+#| fig-subcap: ["Trace plot", "Rhat plot"]
+#| layout-ncol: 2
 
-However, the trade-offs include:
+plot(first_model, "trace")
 
-- Coverage bias: IVR only reaches individuals with landlines, which may skew results towards older demographics.
+plot(first_model, "rhat")
+```
 
-- Self-selection bias: Respondents in online panels tend to be more engaged, which can lead to over-representation of politically active individuals.
-
-## Non-response Handling
-Emerson addresses non-response using a weighting system.
-The data is adjusted based on key demographic variables like age, gender, race, education, and party affiliation, making the sample more representative of the electorate.
-Weights are applied to compensate for under- or over-represented groups, ensuring balanced results.
-
-## Questionnaire Design
-The Emerson survey features a straightforward, efficient design focused on key issues and voter preferences. Strengths of the questionnaire include:
-
-- Simplicity: Clear, concise questions makes the survey easy for respondents to understand and to provide accurate answers.
-
-- Topical relevance: The poll zeroes in on vote preference and demographic splits, delivering timely insights during election cycles.
-
-However, weaknesses include:
-
-- Lack of nuance: While the survey captures basic preferences, it does not delve deeply into voters' motivations or the factors driving their choices.
-
-- Mode limitations: Different survey modes can affect how respondents engage with the questions. This can potentially influence the depth and accuracy of voter responses.
-
-## Conclusion
-Emerson College Polling's mixed-mode methodology is effective in balancing cost with broad demographic reach.
-Its weighting techniques contribute to reliable and accurate results despite non-responses.
-While the use of multiple survey modes increases population representation, possible biases—such as lack of nuance and mode effects need to be taken into consideration when interpreting results.
-
-In sum, this method is a strong predictor of election outcomes and continues to adapt to the evolving challenges of modern polling.
-
-# Idealized Methodology
-
-## Overview
-This methodology outlines an election forecasting plan with a budget of $100,000.
-Heavily inspired by TIPP Insights and Emerson College's hybrid approaches, the survey methodology balances cost-efficiency and data accuracy by using different modes of data collection and wide demographic reach.
-
-This forecasting plan will predict the popular vote and Electoral College outcomes, emphasizing key swing states and underrepresented voter groups.
-
-## Sampling Approach
-This idealized methodology uses a multi-mode hybrid sampling approach. This allows the use of both probability-based and non-probability-based methods:
-
-- Probability-based methods:
-
-  - Live phone interviews (landline and cell): As used by TIPP, live phone interviews is a direct and personal method of collecting responses.
-Voter lists would be sourced from Aristotle to ensure we contact likely voters.
-This mode reaches older voters and voters with less internet connection.
-
-  - Interactive Voice Response (IVR): Similar to Emerson’s approach, IVR for landline users balances costs and ensures demographic representation, particularly among older or less tech-savvy populations.
-This method captures quick responses and is less expensive than live interviews.
-
-- Non-probability methods:
-
-  - MMS-to-web (mobile): Following Emerson’s MMS-to-web strategy, we would send multimedia messages to likely voters’ cell phones, encouraging them to complete an online survey.
-This method works well for younger, mobile-first voters, helping fill in gaps where traditional methods might miss.
-
-  - Online panel recruitment via CINT: To mirror TIPP’s online panel approach, a portion of the survey would be conducted via an opt-in panel of pre-screened voters from CINT, capturing tech-savvy respondents.
-CINT’s platform ensures that different voter groups, including hard-to-reach groups, are represented in the polling.
-
-## Recruitment Strategy
-By using Aristotle’s voter lists and CINT's platform as a sampling frame, the sample consists of likely voters based on registration status and past voting behavior.
-Quota sampling would be used to match the demographic profile of the U.S.
-electorate.
-This ensures that respondents in the sample reflect key demographic variables such as age, gender, race, and education level.
-
-We would oversample key swing states like Pennsylvania, Wisconsin, and Florida.
-Key swing states are states where there is no overwhelming support for a particular candidate or political party.
-The outcome in these states can often determine the overall result of the election, making them critical in securing a win in the Electoral College.
-Oversampling these states allows a more accurate forecasting in these battlegrounds, which are critical for predicting the Electoral College outcome.
+\newpage
 

### 2024-10-22 paper/paper.qmd

@@
 
+#install.packages("psych")
+#install.packages("plotly")
+install.packages("stringr")
+library(psych)
 library(tidyverse)
@@
 library(ggplot2)
+library(kableExtra)
+library(sf)
+library(tidyverse)
+library(plotly)
+library(RColorBrewer)
+library(stringr)
+library(modelsummary)
 analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
+mapping_data <- st_read(here("data/03-shapefiles_data/US"))
 ```
@@
 
-## Data Set Overview
+### Measurement
 
-add a summary statistics table here maybe?
+In the dataset of our analysis, the process of measurement begins by capturing real-world public opinion through surveys.
+Polling organizations transform this observed phenomenon into structured data by recording support percentages, pollster details, and methodological choices, thus turning public sentiment into quantifiable entries in the dataset.
+
+Public opinion polls are essential for understanding voter preferences and electoral dynamics, providing data that informs policy decisions and campaign strategies.
+Our data set captures public support for candidates at specific times, along with information on methodologies, sample sizes, and pollster ratings.
+
+Polling organizations use various methods, such as App panels, IVR (Interactive Voice Response), Live phone calls, Text-to-web systems etc.
+These methods can influence the validity and reliability of results, with larger sample sizes generally yielding more reliable estimates.
+Some methods, like live phone interviews, may be seen as more trustworthy than others.
+
+Our data set also includes important measures of pollster performance, such as numeric grades, poll scores, and transparency scores.
+These metrics are calculated by considering bias, race difficulty, predictive error, and transparency, allowing us to assess the accuracy of each poll.
+The "pollscore," for instance, reflects a pollster’s past performance and reliability, adjusted for potential biases.
+
+### Data Cleaning
+
+The raw data for this project, sourced from FiveThirtyEight, [@FiveThirtyEight] underwent a series of cleaning steps to prepare it for analysis.
+Initially, duplicate rows were removed to ensure that only unique observations remained, facilitated by the janitor package \[\@janitor\].
+A new binary variable, 'national', was created to indicate whether a poll was conducted at the national or state level.
+Missing values in the 'state' column were replaced with "Not Applicable," and numeric grades were evaluated to filter out low-quality pollsters, keeping only those with a numeric grade above 1.
+This cutoff was selected to retain mid to high-level pollsters for more reliable results.
+These steps were performed using functions from the dpylr package \[\@dplyr\].
+Furthermore, dates were standardized and converted into a proper format for analysis using the lubridate package \[\@lubridate\].
+Polls related to Kamala Harris were retained for further analysis, and percentage support values were transformed into actual numbers of supporters based on sample size.
+Additionally, pollster counts below five were excluded to focus on more reliable data sources.
+Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the arrow package \[\@arrow\].
+
+### Summary of Variables of Interest
 
 ```{r}
-library(arrow)
-analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
 
-#head(analysis_data, 20)
-colnames(analysis_data)
-#print(nrow(analysis_data))
+#| include: false
+#| warning: false
+#| message: false
+
+library(dplyr)
+library(modelsummary)
+library(tools)  
+library(kableExtra)
+
+# Set options to use kableExtra for table formatting
+options(modelsummary_factory_default = 'kableExtra')
+
+# Create the summary statistics table for the specified numerical variables
+summary_table <- analysis_data %>%
+  select(numeric_grade, pollscore, sample_size, num_support) %>%
+  # Remove underscores and capitalize the first letter of each word
+  rename_with(~ toTitleCase(gsub("_", " ", .))) %>%
+  datasummary_skim(
+    fun_numeric = list(
+      Mean = mean,
+      SD = sd,
+      Min. = min,
+      Median = median,
+      Max. = max
+    )
+  )
+
+# Display the summary statistics table
+print(summary_table)
+
+#kable(summary_table, caption = "Summary of Variables of Interest") %>%
+  kable_styling(full_width = FALSE)
+
 ```
+
+The summary statistics **\[CROSS REFERENCE HERE\]** for the numeric variables provide significant insights into the polling data.
+The **numeric grade** has a mean of 2.3 and a standard deviation of 0.6, suggesting that most polls are graded around a middle value, though there is moderate variability.
+The range, with a minimum of 1.1 and a maximum of 3.0, indicates the presence of both low-quality and higher-quality polls within the dataset.
+
+The **pollscore** averages at −0.5, with a standard deviation of 0.6.
+Since lower poll scores indicate more favorable outcomes for certain candidates or parties, this negative mean suggests that the polls are, on average, skewing positively towards their targets.
+The minimum of −1.5 and maximum of 1.7 further illustrate the wide distribution of poll results, highlighting the variability in outcomes across different polls.
+
+Looking at the **sample size**, the mean of 1962.5, accompanied by a substantial standard deviation of 2608.0, reflects significant variability in respondent numbers.
+With a minimum sample size of 382 and a maximum of 12,084, it is evident that while many polls are conducted with relatively small sample sizes, a few achieve much larger respondent pools.
+This variability can enhance the reliability of findings, particularly for those polls with larger sample sizes.
+
+Finally, the **number of supporters** reported has a mean of 949.2 and a standard deviation of 1290.0, indicating that while the average number of supporters is relatively high, there is considerable variability.
+The minimum value of 172 and maximum of 6042 underscores the diverse contexts in which these polls were conducted.
+
+Overall, these summary statistics emphasize the variability and range present in the polling data, which may impact the reliability and interpretation of the findings.
+The mixture of low and high values in numeric grade, pollscore, and sample size points to the importance of careful analysis when drawing conclusions from these polling efforts.
 
@@
   filter(!is.na(pollscore)) %>%  # Exclude rows with missing pollscore
+
   group_by(national) %>%         # Group by the binary 'national' column (1 for national, 0 for state)
+
+  group_by(national) %>%        # Group by poll_scope
+
   summarize(
@@
 # Print the summary statistics
-print(summary_stats)
+kable(summary_stats)
 
@@
 Ultimately, the analysis emphasizes the need to consider sample sizes critically when interpreting polling data, as larger sample sizes tend to provide a better representation of the population's views.
-Thus, while many polls may fall short of ideal sample sizes, the existence of larger samples in the dataset can offer valuable insights into public opinion.
+Thus, while many polls may fall short of ideal sample sizes, the existence of larger samples in the data set can offer valuable insights into public opinion.
 **However, we need to determine an ideal and acceptable sample size that would be enough for the polling results to be considered reliable, before we conclude smaller samples as being unreliable.**
 
-### 3) Summary Statistics Table for Numeric Variables
-
-```{r}
-#| warning: false
-#| message: false
-#| echo: false
-
-library(dplyr)
-library(modelsummary)
-library(tools)  # Make sure to load the tools package for toTitleCase()
-
-# Set options to use kableExtra for table formatting
-options(modelsummary_factory_default = 'kableExtra')
-
-# Create the summary statistics table for the specified numerical variables
-summary_table <- analysis_data %>%
-  select(numeric_grade, pollscore, sample_size, num_support) %>%
-  # Remove underscores and capitalize the first letter of each word
-  rename_with(~ toTitleCase(gsub("_", " ", .))) %>%
-  datasummary_skim(
-    fun_numeric = list(
-      Mean = mean,
-      SD = sd,
-      Min. = min,
-      Median = median,
-      Max. = max
-    )
-  )
-
-# Display the summary statistics table
-print(summary_table)
-
-```
-
-The summary statistics **\[CROSS REFERENCE HERE\]** for the numeric variables provide significant insights into the polling data.
-The **numeric grade** has a mean of 2.3 and a standard deviation of 0.6, suggesting that most polls are graded around a middle value, though there is moderate variability.
-The range, with a minimum of 1.1 and a maximum of 3.0, indicates the presence of both low-quality and higher-quality polls within the dataset.
-
-The **pollscore** averages at −0.5, with a standard deviation of 0.6.
-Since lower poll scores indicate more favorable outcomes for certain candidates or parties, this negative mean suggests that the polls are, on average, skewing positively towards their targets.
-The minimum of −1.5 and maximum of 1.7 further illustrate the wide distribution of poll results, highlighting the variability in outcomes across different polls.
-
-Looking at the **sample size**, the mean of 1962.5, accompanied by a substantial standard deviation of 2608.0, reflects significant variability in respondent numbers.
-With a minimum sample size of 382 and a maximum of 12,084, it is evident that while many polls are conducted with relatively small sample sizes, a few achieve much larger respondent pools.
-This variability can enhance the reliability of findings, particularly for those polls with larger sample sizes.
-
-Finally, the **number of supporters** reported has a mean of 949.2 and a standard deviation of 1290.0, indicating that while the average number of supporters is relatively high, there is considerable variability.
-The minimum value of 172 and maximum of 6042 underscores the diverse contexts in which these polls were conducted.
-
-Overall, these summary statistics emphasize the variability and range present in the polling data, which may impact the reliability and interpretation of the findings.
-The mixture of low and high values in numeric grade, pollscore, and sample size points to the importance of careful analysis when drawing conclusions from these polling efforts.
-
-### 4) Top 10 Pollsters and their Average Pollscores and Numeric Grades
+### 3) Top 10 Pollsters and their Average Pollscores and Numeric Grades
 
@@
 
-# Print the top pollsters table
-print(top_pollsters)
+kable(top_pollsters, format = "html", caption = "Top 10 Pollsters by Frequency")
 
@@
 
-## Graphs for Imp. Variables (relationships)
+yet to do.
 
@@
 The goal is to reduce errors and biases present in individual polls by using a weighted average of many different polls.
-In our approach, we calculate the weights based on several important variables.
-Sample size is included because larger polls are generally more reliable and have smaller margins of error.
-Pollscore (or pollster rating) reflects the empirical quality of a pollster, balancing both accuracy and bias, allowing us to give more weight to those with a proven track record.
-We also account for the pollscope, differentiating between national and state-level polls, as state polls may offer more localized insights into voter sentiment.
-We also use recency as more recent polls give us a more accurate estimate of voter sentiment.
 
-However, we acknowledge that even with these adjustments, there may still be underlying biases from individual pollsters toward the candidates.
-To account for this, we will select a baseline pollster with a strong rating, high transparency, and minimal historical error, and regress the support percentage (pct) on pollster, including the same variables as controls whcih was used in the weighting method, such as sample size, pollscore and pollscope.Additionally, we will control for the state of the poll, as support levels can vary regionally.
-Therefore each pollster's coefficient in the regression model will represent the bias of that pollster relative to the baseline pollster.
-Controlling for these factors is crucial because it ensures that the differences between pollsters are not driven by other factors, such as differences in sample size or the types of polls they conduct, but by their inherent bias in support estimates.
-Once we have calculated the weights, we will use them to generate a weighted average of support for Kamala Harris and Donald Trump, ultimately comparing their weighted support to forecast the likely winner of the election.
+In our approach, we will employ linear modeling of voter support percentage (pct) on pollster and other independent variables such as sample size, poll recency, and poll scope (state vs.
+national).
+This will allow us to smooth out the inherent noise, biases, and variability across different pollsters.
+Once we obtain the predicted values from our model, we will weight these predictions based on the numeric grade (quality score) of each pollster to calculate an overall national estimate of the outcome.
+Additionally, we will separately compute estimates for key battleground states, where voter behavior can be more volatile and pivotal in deciding the final outcome of the election.
+This approach helps us capture both national trends and critical state-level dynamics.
 
-## **Exploration of Control Variables**
+## **Exploration of Independent Variables**
 
-### Investigating the Effect of Pollscore on Candidate Support
+### Support for Kamala Harris by Sample Size and Polling Level
+
+\[\@fig-sample\] suggests that the relationship between support for Kamala Harris and sample size shows a slight positive trend, indicating a small increase in support as sample sizes grow.
+Both national (blue) and state (green) polls reflect this upward trend, although the effect is minimal.
+Since voting outcomes often depend on small margins, this subtle increase in support with larger sample sizes could still be relevant in predicting overall support.
 
 ```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-sample
+#| fig-cap: Sample Size vs. Support for Kamala Harris, Colored by Pollscope (National vs. State)
 
-# Define pollscore ranges (you can adjust the binwidth if needed)
-cleaned_data <- analysis_data %>%
-  mutate(pollscore_range = cut(pollscore, breaks = seq(-1.5, 1, by = 0.3), include.lowest = TRUE))
+# Custom colors for pollscope
+national_color <- "chartreuse3"
+state_color <- "steelblue"
 
-# Calculate the average support for each candidate within each pollscore range
-average_support <- cleaned_data %>%
-  group_by(pollscore_range, candidate_name) %>%
-  summarise(avg_support = mean(pct, na.rm = TRUE)) %>%
-  ungroup()
+# Create the plot for Kamala Harris with the requested adjustments
+ggplot(analysis_data %>% filter(candidate_name == "Kamala Harris"), 
+       aes(x = sample_size, y = pct, color = factor(national))) +
+  geom_point(size = 0.5, alpha = 0.9) +  # Smaller dots with slight transparency
+  geom_smooth(aes(group = 1), method = "lm", se = FALSE, color = "azure4") +  # Linear model smoothing
 
-# Create a bar plot
-ggplot(average_support, aes(x = pollscore_range, y = avg_support, fill = candidate_name)) +
-  geom_bar(stat = "identity", position = "dodge") +  # Dodge to separate bars for each candidate
-  scale_fill_manual(values = c("Donald Trump" = "indianred", "Kamala Harris" = "steelblue")) +  # Custom colors
-  labs(title = "Average Support for Candidates by Pollscore Range",
-       x = "Pollscore Range",
-       y = "Average Support (%)",
-       fill = "Candidate") +
+  # Custom colors for national (1 for National, 0 for State)
+  scale_color_manual(values = c("1" = national_color, "0" = state_color), 
+                     labels = c("National", "State")) +  # Pollscope categories with custom colors
+
+  # Same minimal styling
+  theme_minimal() +
+  theme(
+    panel.grid.major = element_blank(),  # Remove major gridlines
+    panel.grid.minor = element_blank(),  # Remove minor gridlines
+    legend.title = element_blank(),  # No title for the legend
+    legend.position = "bottom",  # Move legend to the bottom
+    plot.title = element_text(hjust = 0.5),  # Center the title
+    axis.line = element_line(color = "black"),  # Add axis lines
+    panel.border = element_blank(),  # Remove panel border
+    axis.line.y.right = element_line(color = "black"),  # Add right spine
+    axis.line.x.top = element_blank()  # Remove top spine
+  ) +
+  labs(x = "Sample Size", y = "Support (%)")  # No title, just axis labels
+
+```
+
+### Variation in Support for Kamala Harris Among Most Frequent Pollsters
+
+[@fig-top5pollsters] above shows the variation in support for Kamala Harris across the five most frequent pollsters.
+Notably, the median support levels differ between pollsters, indicating variability in central estimates of Harris’s support.
+For instance, Siena/NYT reports a higher median compared to Redfield & Wilton Strategies.
+Additionally, the differing ranges of support estimates highlight potential discrepancies in the methodologies, sample sizes, and biases employed by each pollster.
+This suggests that different pollster have different interpretations of candidate support, underlining the importance of aggregating data from multiple sources to avoid over-reliance on any single pollster.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-top5pollsters
+#| fig-cap: "Support for Kamala Harris across the top 5 pollsters by poll count."
+
+# Get top 5 pollsters by poll count
+top5_pollsters <- analysis_data %>%
+  filter(candidate_name == "Kamala Harris") %>%
+  count(pollster, sort = TRUE) %>%
+  top_n(5, n) %>%
+  pull(pollster)
+
+# Filter the data for only these top 5 pollsters
+top5_data <- analysis_data %>%
+  filter(pollster %in% top5_pollsters & candidate_name == "Kamala Harris")
+
+# Custom colors for boxplot fill
+boxplot_colors <- c("chartreuse3", "steelblue", "indianred", "purple", "darkorange")
+
+# Create the boxplots with thinner borders, smaller points, and rotated x-axis labels
+ggplot(top5_data, aes(x = pollster, y = pct, fill = pollster)) +
+  geom_boxplot(size = 0.2, outlier.size = 0.5) +  # Thinner box lines and smaller outliers
+  scale_fill_manual(values = boxplot_colors) +
   theme_minimal() +  # Minimal theme
   theme(
-    legend.position = "bottom",
-    plot.title = element_text(hjust = 0.5)
+    panel.grid.major = element_blank(),  # Remove major gridlines
+    panel.grid.minor = element_blank(),  # Remove minor gridlines
+    legend.position = "none",  # No legend needed for boxplots
+    axis.line = element_line(color = "black"),  # Add axis lines
+    plot.title = element_text(hjust = 0.5),  # Center title
+    axis.text.x = element_text(angle = 20, hjust = 1),  # Rotate x-axis labels for better readability
+    axis.line.y.right = element_line(color = "black"),  # Add right spine for clean look
+    axis.line.x.top = element_blank(),  # Remove top spine
+    panel.border = element_blank()  # Remove panel border
+  ) +
+  labs(x = "Pollster", y = "Support (%)")  # Labels
+
+```
+
+### Support for Kamala Harris Based on Recency of Polls and Pollster Quality
+
+In \[\@fig-enddate\], we observe a slight increasing trend in support for Kamala Harris as the end date of polls progresses towards October.
+This may indicate a gradual improvement in her polling performance over time.
+The color gradient, reflecting the pollster's error and bias (Pollscore), reveals variations in bias: greener points represent more negative biases, while redder points indicate more positive biases.
+
+While the recency of the polls shows a stable trend in Harris's support, the wide distribution of bias (Pollscore) values suggests that pollster quality can vary substantially.
+These variations could influence the model's predictions if left unaddressed.
+Hence, including both recency and pollster quality as predictor variables in the model might be valuable to account for such variations.
+Further model diagnostics and validation will help decide if these variables are significant.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-enddate
+#| fig-cap: "Support for Kamala Harris based on End Date and colored by Pollster's Error and Bias."
+
+# Define the color gradient based on pollscore
+ggplot(analysis_data %>% filter(candidate_name == "Kamala Harris"), aes(x = end_date, y = pct, color = pollscore)) +
+  geom_point(size = 0.8, alpha = 1) +  # Adjust point size and transparency
+  geom_smooth(method = "lm", se = FALSE, color = "azure4") +  # Add a smooth trend line for overall support
+  scale_color_gradient2(low = "blue", mid = "green", high = "red", midpoint = 0) +  # Color gradient
+  theme_minimal() +  # Minimal theme as usual
+  theme(
+    panel.grid.major = element_blank(),  # Remove major gridlines
+    panel.grid.minor = element_blank(),  # Remove minor gridlines
+    legend.position = "bottom",  # Legend at the bottom
+    axis.line = element_line(color = "black"),  # Add axis lines
+    panel.border = element_blank(),  # Remove panel border
+    axis.line.y.right = element_line(color = "black"),  # Add right spine for clean look
+    axis.line.x.top = element_blank()  # Remove top spine
+  ) +
+  labs(x = "Date Poll Ended",
+       y = "Support (%)",
+       color = "Error and Bias of Pollster (Low to High)")  # Updated label for color gradient
+
+```
+
+### Political Leanings of Different State
+
+[@fig-mapinterac] provides a state-level representation of Kamala Harris's average support across the U.S., with the darker blue shades indicating states where her support is higher, and the redder tones showing where her support is lower.
+Gray areas represent states where we have no polling data available.
+
+The gradient highlights notable geographic trends.
+Harris has the strongest support in states like Vermont, Massachusetts, and New York, while states like Utah and West Virginia show considerably lower levels of support.
+A distinct pattern emerges in the West, where coastal states like California favor Harris more strongly, contrasting with inland states like Wyoming and Utah.
+
+This reveals that political leanings differ greatly by state, and this regional variation should be controlled for in any regression model analyzing poll data.
+The map serves as a useful tool for visualizing regional variations in voter support for Kamala Harris as reflected in the available polling data.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-mapinterac
+#| fig-cap: "Average Support for Kamala Harris in Different States estimated by Polls"
+  # For string manipulation
+
+# Filter out non-continental states (Hawaii and Alaska)
+us_states_filtered <- us_states %>% 
+  filter(!NAME %in% c("Alaska", "Hawaii"))
+
+# Ensure correct merge between map data and Kamala's support data
+map_data_full <- left_join(us_states_filtered, mapping_data, by = c("NAME" = "state"))
+
+# Create centroids for state labels
+state_centroids <- st_centroid(us_states_filtered)
+
+# Plot using geom_sf() for sf objects
+p <- ggplot(map_data_full) +
+  geom_sf(aes(fill = mean_support_kamala, text = paste("State:", str_to_title(NAME), "<br>Average Support:", round(mean_support_kamala, 2))), color = "white") +  # Capitalize state names
+  scale_fill_distiller(palette = "RdBu", direction = 1, na.value = "gray", name = "Kamala's Support") +  # Reverse RdBu color gradient
+  geom_text(data = state_centroids, aes(geometry = geometry, label = STUSPS), stat = "sf_coordinates", size = 2) +  # State labels
+  labs(title = "Support for Kamala Harris by State", 
+       caption = "States with no polling data are shown in gray") +
+  theme_minimal() +
+  theme(
+    legend.position = "right",
+    plot.title = element_text(hjust = 0.5),
+    panel.grid = element_blank(),  # Remove grid lines
+    axis.title = element_blank(),  # No axis titles for maps
+    axis.text = element_blank(),   # No axis labels
+    axis.ticks = element_blank()   # No axis ticks
   )
@@
   mutate(harris_proportion = Kamala_Harris / (Kamala_Harris + Donald_Trump))
+
+# Turn ggplot to plotly for interactivity, add hover template for better control over display
+ggplotly(p, tooltip = "text")
 

### 2024-10-22 paper/paper.qmd
Some modelling done.

--- 2024-10-22 paper/paper.qmd
@@
 library(modelsummary)
+library(broom)
+library(sf)
 analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
 mapping_data <- st_read(here("data/03-shapefiles_data/US"))
+mean_support_data <- read_parquet(here("data/02-analysis_data/mapping_data.parquet"))
 ```
@@
 
-Our dataset also includes important measures of pollster performance, such as numeric grades, poll scores, and transparency scores.
+The dataset also includes important measures of pollster performance, such as numeric grades, poll scores, and transparency scores.
 These metrics are calculated by considering bias, race difficulty, predictive error, and transparency, allowing us to assess the accuracy of each poll.
@@
 
-##
-
 ## Summary Statistics
@@
 Overall, these statistics illustrate a strong reliability of pollsters considering their errors and bias.
+
+```         
+```
 
@@
 Consequently, the predominance of smaller sample sizes in this data set may raise concerns about the robustness of the polling findings, potentially affecting their applicability in broader analyses or decision-making processes.
-**However, we must determine an acceptable and ideal sample size that would allow us to determine the reliability of these results.**
+However, we must determine an acceptable and ideal sample size that would allow us to determine the reliability of these results.
 
-# Polls of Polls to Forcast Winning Candidate
-
-## Forecasting Approach
+# Forecasting Approach Using Polls of Polls Method
 
@@
 ```{r}
-#| warning: false
-#| message: false
-#| echo: false
-#| label: fig-mapinterac
-#| fig-cap: "Average Support for Kamala Harris in Different States estimated by Polls"
-  # For string manipulation
-
 # Filter out non-continental states (Hawaii and Alaska)
-us_states_filtered <- us_states %>% 
+us_states_filtered <- mapping_data %>% 
   filter(!NAME %in% c("Alaska", "Hawaii"))
@@
 # Ensure correct merge between map data and Kamala's support data
-map_data_full <- left_join(us_states_filtered, mapping_data, by = c("NAME" = "state"))
+map_data_full <- left_join(us_states_filtered, mean_support_data, by = c("NAME" = "state"))
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+In this section, we aim to address the inherent biases and differences present in various polling data to arrive at a robust prediction model.
+The core challenge lies in selecting a model with an optimal balance between complexity and fit, ensuring it accurately captures the dynamics of polling data while avoiding overfitting.
+To this end, we carefully evaluated different model specifications to determine the most appropriate one for our forecasting purpose.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+Given that variables like numeric grade and pollscore are perfectly collinear with pollster, they were excluded from the regression analysis to avoid multicollinearity issues.
+These variables, however, remain integral to our weighting strategy, where they will be used to adjust for differences in polling accuracy and reliability.
+Instead, we focus on key features such as pollster, sample size, state, and recency, gradually adding complexity to the model.
 
-## Model set-up
+By systematically comparing model specifications that incorporate these variables, we aim to select the model with the right balance between predictive accuracy and generalizability, ultimately providing the best possible forecast.
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+## Model Set Up
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+We aim to model the percentage of support for Kamala Harris in each poll as a function of the pollster the sample size, the state, and the recency of the poll.
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+$$ y_i = \alpha + \beta_1 \cdot \texttt{pollster}_i + \beta_2 \cdot \texttt{sample_size}_i + \beta_3 \cdot \texttt{recency}_i +  \beta_4 \cdot \texttt{state}_i + \epsilon_i $$
 
-### Model justification
+Where
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+-   $y_i$​ is the percentage of support for Kamala Harris in poll iii,
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+-   $α$ is the intercept,
 
-# Results
+-   $β_1$ captures the effect of the polling organization,
+
+-   $β_2​$ captures the effect of the sample size,
+
+-   $β_3$​ captures the effect of recency (how recent the poll is),
+
+-   $β_4$ capture the effects of the different states
+
+-   ​ $\epsilon_i$ represents the error term, assumed to follow a normal distribution with mean 0.
+
+## Model Justification
+
+We compared three models with different sets of predictors, gradually increasing the model complexity.
+Model 1 only included **pollster** and **sample size** as predictors, while Model 2 introduced **state** as an additional factor.
+Finally, Model 3 incorporated **recency** as another key predictor.
+Among these models, Model 3 demonstrated the highest $R^2$ value (0.731), indicating the best fit and ability to explain the variation in the data.
+Additionally, Model 3 had the lowest AIC (4786.815) and BIC (5203.787) values, suggesting that it provided the best balance between model complexity and predictive performance.
+Therefore, Model 3 was selected as the most appropriate model for smoothing out polling biases and differences, offering the most accurate prediction of support for Kamala Harris.
+
+```{r}
+# Load necessary libraries
+library(broom)
+library(dplyr)
+library(knitr)
+
+# Load the models from RDS files
+model_1 <- readRDS(here("models/model_1.rds"))
+model_2 <- readRDS(here("models/model_2.rds"))
+model_3 <- readRDS(here("models/model_3.rds"))
+
+# Function to get model details
+get_model_details <- function(model, model_name, include_recency = FALSE) {
+  r2 <- summary(model)$r.squared
+  adj_r2 <- summary(model)$adj.r.squared
+  f_stat <- summary(model)$fstatistic[1]
+
+  # Extract p-values for Sample Size and Recency
+  sample_size_pvalue <- tidy(model) %>% filter(term == "sample_size") %>% pull(p.value)
+  sample_size_significance <- ifelse(sample_size_pvalue < 0.05, paste0("Significant (p = ", round(sample_size_pvalue, 4), ")"), paste0("Not significant (p = ", round(sample_size_pvalue, 4), ")"))
+
+  recency_significance <- "Not included"
+  if (include_recency) {
+    recency_pvalue <- tidy(model) %>% filter(term == "recency") %>% pull(p.value)
+    recency_significance <- ifelse(recency_pvalue < 0.05, paste0("Highly significant (p < 2e-16)"), paste0("Not significant (p = ", round(recency_pvalue, 4), ")"))
+  }
+
+  return(data.frame(
+    Model = model_name,
+    Variables_Included = ifelse(include_recency, "Pollster + Sample Size + State + Recency", "Pollster + Sample Size + State"),
+    R2 = round(r2, 3),
+    Adjusted_R2 = round(adj_r2, 3),
+    F_statistic = round(f_stat, 2),
+    Sample_Size_Significance = sample_size_significance,
+    Recency_Significance = recency_significance,
+    States_Significant = ifelse(model_name == "Model 1", "Not included", ifelse(model_name == "Model 2", "Several states", "More states"))
+  ))
+}
+
+# Generate the summaries for all models
+model_1_summary <- get_model_details(model_1, "Model 1")
+model_2_summary <- get_model_details(model_2, "Model 2")
+model_3_summary <- get_model_details(model_3, "Model 3", include_recency = TRUE)
+
+# Combine all summaries into one data frame
+all_summaries <- bind_rows(model_1_summary, model_2_summary, model_3_summary)
+
+# Print the table using kable for better formatting
+
+kable(all_summaries, format = "markdown",row.names = FALSE, col.names = c("Model", "Variables Included", "R²", "Adjusted R²", "F-statistic", "Sample Size Significance", "Recency Significance", "States Significant"))
+
+```
+
+```{r}
+# Create a data frame for AIC and BIC comparison
+aic_bic_table <- data.frame(
+  Model = c("Model 1", "Model 2", "Model 3"),
+  AIC = c(AIC(model_1), AIC(model_2), AIC(model_3)),
+  BIC = c(BIC(model_1), BIC(model_2), BIC(model_3))
+)
+
+# Print the table using kable for LaTeX format
+kable(aic_bic_table, format = "markdown", col.names = c("Model", "AIC", "BIC"), booktabs = TRUE)
+
+```
+
+## Model Results
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Load necessary libraries
+library(modelsummary)
+library(broom)
+
+# Load Model 3 from RDS file
+model_3 <- readRDS(file = here::here("models/model_3.rds"))
+
+# Generate LaTeX output for the model
+latex_output <- modelsummary(model_3, output = "latex")
+
+# Ensure the output is a character string
+latex_output_string <- capture.output(cat(latex_output))
+
+# Write the LaTeX string to a .tex file
+writeLines(latex_output_string, here::here("model_3_summary.tex"))
+
+```
+
+```{r}
+# Save LaTeX output to a string
+latex_output <- modelsummary(model_3, output = "latex")
+
+# Write the LaTeX string to a .tex file
+writeLines(latex_output, here::here("output/model_3_summary.tex"))
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Load necessary libraries
+library(modelsummary)
+library(broom)
+
+# Load Model 3 from RDS file
+model_3 <- readRDS(file = here::here("models/model_3.rds"))
+
+modelsummary(model_3, output = "latex", file = "model_3_summary.tex")
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Load necessary libraries
+library(modelsummary)
+library(broom)
+
+# Load Model 3 from RDS file
+model_3 <- readRDS(file = here::here("models/model_3.rds"))
+
+modelsummary(model_3)
+
+# Display the summary of Model 3 using modelsummary
+#| echo: false
+#| eval: true
+#| label: tbl-model3results
+#| tbl-cap: "Regression results for Model 3 including Pollster, Sample Size, State, and Recency"
+
+modelsummary::modelsummary(
+  list(
+    "Model 3" = model_3
+  ),
+  statistic = "std.error",
+  fmt = 2
+)
+
+```
 

### 2024-10-22 paper/paper.qmd

--- 2024-10-22 paper/paper.qmd
@@
 
-Our data set also includes important measures of pollster performance, such as numeric grades, poll scores, and transparency scores.
+The dataset also includes important measures of pollster performance, such as numeric grades, poll scores, and transparency scores.
 These metrics are calculated by considering bias, race difficulty, predictive error, and transparency, allowing us to assess the accuracy of each poll.
@@
 
-### 2) Distribution of Sample Sizes
+### 2) Distribution of sample sizes: same as above
 
@@
 This contrast with Plot 1 highlights a shift towards more robust polling efforts.
+
+In the context of polling results, these findings suggest potential limitations in the accuracy and validity of the insights generated.
+Larger sample sizes are typically preferred in polling as they tend to yield more reliable estimates and reduce the margin of error, leading to a better representation of the population's views.
+Consequently, the predominance of smaller sample sizes in this data set may raise concerns about the robustness of the polling findings, potentially affecting their applicability in broader analyses or decision-making processes.
+However, we must determine an acceptable and ideal sample size that would allow us to determine the reliability of these results.
+
+# Forecasting Approach Using Polls of Polls Method
 

### 2024-10-22 scripts/03-clean_data.R
fixed merging conflict.

@@ -89,32 +89,6 @@ cleaned_data <- cleaned_data %>%
 cleaned_data <- cleaned_data %>%
   filter(end_date >= as.Date("2024-07-21"))
 
-# Calculate recency of every poll and make a column 
-
-reference_date <- max(cleaned_data$end_date, na.rm = TRUE)  # Most recent poll date
-
-analysis_data <- cleaned_data %>%
-  mutate(recency = as.numeric(reference_date - end_date))
-
-
 #### Save data ####
 
 write_parquet(cleaned_data, "data/02-analysis_data/analysis_data.parquet")
-
-#### Creating Data For Mapping ####
-
-deduped_data <- raw_data %>%
-  distinct()  # Remove duplicate rows
-
-# Step 2: Filter for Kamala Harris and non-empty states
-harris_data <- deduped_data %>%
-  filter(candidate_name == "Kamala Harris" & !is.na(state))
-
-# Step 3: Group by state and calculate mean support for Kamala Harris
-mapping_data <- harris_data %>%
-  group_by(state) %>%
-  summarise(mean_support_kamala = mean(pct, na.rm = TRUE))
-
-# Step 4: Save the aggregated data to a parquet file
-write_parquet(mapping_data, "data/02-analysis_data/mapping_data.parquet")
-

### 2024-10-22 scripts/06-model_data.R

@@ -1,22 +1,19 @@
 #### Preamble ####
-# Purpose: Models the percentage of support for Kamala Harris (pct) based on various factors such as pollster, poll type (state vs. national), 
-#          sample size, and other relevant variables.
-# Author: Shamayla Durrin
-# Date: 19 October 2024
-# Contact: shamayla.islam@mil.utoronto.ca
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
 # License: MIT
-# Pre-requisites: Requires the cleaned dataset of polls, including variables like pollster, state, sample size, and pct for Kamala Harris.
-# Any other information needed: Ensure the necessary packages (e.g., tidyverse, broom, and modeling packages) are installed before running the model.
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
-
 library(tidyverse)
 library(rstanarm)
-library(arrow)
 
 #### Read data ####
-analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
+analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
 ### Model data ####
 first_model <-
@@ -34,7 +31,7 @@ first_model <-
 #### Save model ####
 saveRDS(
   first_model,
-  file = "models/first_m\del.rds"
+  file = "models/first_model.rds"
 )
 
 

### 2024-10-22 paper/paper.qmd
Made some changes.

--- 2024-10-22 paper/paper.qmd
@@
 thanks: "Code and data are available at: [https://github.com/krishnak30/US_elections](https://github.com/krishnak30/US_elections)."
-
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
-format: pdf
+abstract: "This paper uses a polls of polls approach to predict voter support for Kamala Harris in the 2024 U.S.
+election, aggregating multiple polls to reduce individual bias and improve accuracy.
+Through a regression model incorporating factors such as pollster, sample size, state, and recency, we predict her overall support to be 47.78%, just below the 50% threshold.
+Our analysis highlights the importance of poll quality and recency in understanding trends in public opinion.
+These findings are significant as they provide a more reliable estimate of electoral outcomes, helping to inform political strategies and public discourse."
+format: html
 number-sections: true
@@
 #| message: false
+
 library(psych)
@@
 analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
-mapping_data <- st_read(here("data/03-shapefiles_data/US"))
-mean_support_data <- read_parquet(here("data/02-analysis_data/mapping_data.parquet"))
+mapping_data <- st_read(here("data/03-mapping_data/US"))
+mean_support_data <- read_parquet(here("data/02-analysis_data/map_data.parquet"))
 ```
@@
 # Introduction
+
+While individual polls provide snapshots of public opinion, they are often subject to biases and methodological differences.
+To address these limitations, this paper uses a "polls of polls" approach to predict support for Kamala Harris, aggregating multiple polls to smooth out inconsistencies.
+By combining data from various polling organizations, our aim is to create a more accurate forecast of her voter support leading up to the 2024 U.S.
+election.
+
+We employ a linear regression model to estimate Kamala Harris' support percentage, incorporating key variables such as pollster, sample size, state, and recency.
+To account for the varying reliability of different polls, we weight predictions using each pollster’s numeric grade.
+Our model also explores the role of recency, with recent polls showing a decline in support for Harris.
+The final weighted prediction for her support stands at 47.78%, indicating that her overall backing falls just below the 50% threshold, suggesting potential challenges in gaining majority support.
+
+This paper contributes to the field of election forecasting by emphasizing the importance of poll quality and recency in prediction models.
+The structure of the paper proceeds as follows: we first introduce the dataset and methodology, followed by model selection and results.
+We then discuss the implications of our findings, and conclude with limitations and potential directions for future research.
 
@@
 
-### Summary of Variables of Interest
+### Summary Statistics of Variables of Interest
 
 ```{r}
-# Assuming you have the dataset numeric_data already loaded
 #| include: false
@@
 #| message: false
+#| echo: false
 
-numeric_data <- analysis_data %>% select(numeric_grade, pollscore, sample_size, pct, num_support)
+library(dplyr)
+library(modelsummary)
+library(tools)  
+library(kableExtra)
 
-# Create a summary table
-summary_table <- numeric_data %>%
-  reframe(
-    `Variable` = c("numeric_grade", "pollscore", "sample_size", "pct", "num_support"),
-    `Mean` = c(mean(numeric_grade, na.rm = TRUE), 
-               mean(pollscore, na.rm = TRUE), 
-               mean(sample_size, na.rm = TRUE), 
-               mean(pct, na.rm = TRUE), 
-               mean(num_support, na.rm = TRUE)),
-    `Max` = c(max(numeric_grade, na.rm = TRUE), 
-              max(pollscore, na.rm = TRUE), 
-              max(sample_size, na.rm = TRUE), 
-              max(pct, na.rm = TRUE), 
-              max(num_support, na.rm = TRUE)),
-    `Min` = c(min(numeric_grade, na.rm = TRUE), 
-              min(pollscore, na.rm = TRUE), 
-              min(sample_size, na.rm = TRUE), 
-              min(pct, na.rm = TRUE), 
-              min(num_support, na.rm = TRUE)),
-    `Standard Deviation` = c(sd(numeric_grade, na.rm = TRUE), 
-                             sd(pollscore, na.rm = TRUE), 
-                             sd(sample_size, na.rm = TRUE), 
-                             sd(pct, na.rm = TRUE), 
-                             sd(num_support, na.rm = TRUE))
+# Set options to use kableExtra for table formatting
+options(modelsummary_factory_default = 'kableExtra')
+
+# Create the summary statistics table for the specified numerical variables
+summary_table <- analysis_data %>%
+  select(numeric_grade, pollscore, sample_size, num_support) %>%
+  # Remove underscores and capitalize the first letter of each word
+  rename_with(~ toTitleCase(gsub("_", " ", .))) %>%
+  datasummary_skim(
+    fun_numeric = list(
+      Mean = mean,
+      SD = sd,
+      Min. = min,
+      Median = median,
+      Max. = max
+    )
   )
 
-kable(summary_table)
-```
+print(summary_table)
 
-## Variables
+#(summary_table)
 
-The data set contains a wide range of variables that are essential for understanding polling outcomes. The following are key variables:
-
--   **`poll_id`**: Unique identifier for each poll conducted.
-
--   **`pollster`**: The polling organization that conducted the poll (e.g., YouGov, TIPP).
-
--   **`numeric_grade`**: A numeric rating indicating the pollster’s quality or reliability (e.g., 1.8).
-
--   **`pollscore`**: A score representing the reliability of the pollster in terms of error and bias.
-
--   **`state`**: The state where the poll was conducted (or "Not Applicable" for national polls).
-
--   **`start_date` / `end_date`**: Dates during which the poll was conducted.
-
--   **`sample_size`**: The number of respondents in the poll.
-
--   **`population`**: The group surveyed (e.g., likely voters or registered voters).
-
--   **`candidate_name`**: The name of the candidate being polled (e.g., Kamala Harris, Donald Trump).
-
--   **`pct`**: The percentage of support for the candidate in the poll.
-
--   **`methodology`**: The method used to conduct the poll (e.g. online panel)
-
--   **`transparency_score`**: A grade for how transparent a pollster is, calculated based on how much information it discloses about its polls and weighted by recency.
-The highest Transparency Score is 10.
-
-## Summary Statistics
-
-### 1) Average Pollster Quality (pollscore): mean, median, mode, min, max
-
-```{r}
-library(dplyr)
-
-# Generate summary statistics for national and state polls
-summary_stats <- analysis_data %>%
-  filter(!is.na(pollscore)) %>%  # Exclude rows with missing pollscore
-  group_by(national) %>%        # Group by poll_scope
-  summarize(
-    Count = n(),                   
-    Min = round(min(pollscore), 2),
-    `1st Qu.` = round(quantile(pollscore, 0.25), 2),
-    Median = round(median(pollscore), 2),
-    Mean = round(mean(pollscore), 2),
-    `3rd Qu.` = round(quantile(pollscore, 0.75), 2),
-    Max = round(max(pollscore), 2),
-    .groups = 'drop'  # Drop grouping structure for cleaner output
-  )
-
-# Print the summary statistics
-kable(summary_stats)
+#kable(summary_table, caption = "Summary of Variables of Interest") %>% kable_styling(full_width = FALSE)
 
@@
 
-As shown above, the summary statistics table **\[CROSS-REFERENCE HERE\]** offers an overview of poll scores for national and state polls, with 2,971 and 2,953 entries, respectively.
-Both categories feature a minimum poll score of -1.5, which a high level of validity and reliability.
-The first quartile shows national polls at -0.9, while state polls are slightly lower at -1.1.
-Median scores are -0.3 for national polls and -0.4 for state polls, suggesting that half of the polls in both categories yield moderately negative scores, reinforcing a trend of favorable sentiment.
-The mean score is -0.37 for national polls, indicating a less reliable perception compared to the -0.51 mean for state polls.
-The third quartile for national polls (0.0) indicates that the upper 25% of scores are neutral, while state polls remain just below neutral at 0.7.
-The maximum score of 1.7 for national polls reflects a notable negative perception for at least one candidate, while the maximum for state polls is 0.7.
-Overall, these statistics illustrate a strong reliability of pollsters considering their errors and bias.
+The summary statistics table presents key descriptive measures for the numeric variables of interest.
+The Numeric Grade variable has a mean of 2.3, with values ranging from a minimum of 1.1 to a maximum of 3.0.
+It has a standard deviation of 0.6, indicating relatively low variability.
+The Pollscore variable averages -0.5, with a standard deviation of 0.6 and a range between -1.5 and 1.7.
+The Sample Size has a wider range, with an average of 1962.5 and a standard deviation of 2608.0, indicating significant variation across polls, with values spanning from 382 to 12084.
+Lastly, Num Support, calculated as the number of respondents supporting Kamala Harris, shows a mean of 949.2, with values ranging from 172 to 6042 and a standard deviation of 1290.0, highlighting substantial differences across the polls.
 
-```         
-```
-
-### 2) Distribution of sample sizes: same as above
+### Distribution of Sample Size of Polls
 
 ```{r}
+#| warning: false
+#| message: false
+#| echo: false
+
 library(ggplot2)
-library(dplyr)
-library(gridExtra)  # Load the gridExtra package
 
-# Create a new variable for sample size ranges
-analysis_data <- analysis_data %>%
-  mutate(sample_size_range = case_when(
-    sample_size <= 10000 ~ "0-10,000",
-    sample_size > 10000 & sample_size <= 20000 ~ "10,001-20,000",
-    sample_size > 20000 & sample_size <= 30000 ~ "20,001-30,000",
-    TRUE ~ "Above 30,000"
-  ))
-
-# Plot for 0-10,000
-p1 <- ggplot(analysis_data %>% filter(sample_size <= 10000), aes(x = sample_size)) +
-  geom_histogram(binwidth = 500, fill = "lightgreen", color = NA, alpha = 0.7) +
-  labs(title = "Sample Sizes 
-       (0-10,000)",
+# Plot histogram for all sample sizes
+ggplot(analysis_data, aes(x = sample_size)) +
+  geom_histogram(binwidth = 500, fill = "lightblue", color = "black", alpha = 0.7) +
+  labs(title = "Distribution of Sample Sizes",
        x = "Sample Size",
        y = "Count") +
-  scale_x_continuous(limits = c(0, 10000)) +  # Set x limits for the first plot
-  scale_y_continuous(limits = c(0, 2300)) +  # Set y limits for the first plot
   theme_minimal() +
-  theme(plot.title = element_text(size = 10),  # **Decreased font size of the title**
-        axis.title.y = element_text(margin = margin(r = 15)))  # **Increased gap for y-axis label**
-
-# Plot for 10,001 - 20,000
-p2 <- ggplot(analysis_data %>% filter(sample_size > 10000 & sample_size <= 20000), aes(x = sample_size)) +
-  geom_histogram(binwidth = 500, fill = "indianred", color = NA, alpha = 0.7) +
-  labs(title = "Sample Sizes 
-       (10,001-20,000)",
-       x = "Sample Size",
-       y = "Count") +
-  scale_x_continuous(limits = c(10001, 20000)) +  # Set x limits for the second plot
-  scale_y_continuous(limits = c(0, 50)) +  # Set y limits for the second plot
-  theme_minimal() +
-  theme(plot.title = element_text(size = 10),  # **Decreased font size of the title**
-        axis.title.y = element_text(margin = margin(r = 10)))  # **Increased gap for y-axis label**
-
-# Plot for 20,001 - 30,000
-p3 <- ggplot(analysis_data %>% filter(sample_size > 20000 & sample_size <= 30000), aes(x = sample_size)) +
-  geom_histogram(binwidth = 500, fill = "lightblue", color = NA, alpha = 0.7) +
-  labs(title = "Sample Sizes 
-       (20,001-30,000)",
-       x = "Sample Size",
-       y = "Count") +
-  scale_x_continuous(limits = c(20001, 30000)) +  # Set x limits for the third plot
-  scale_y_continuous(limits = c(0, 10)) +  # Set y limits for the third plot
-  theme_minimal() +
-  theme(plot.title = element_text(size = 10),  # **Decreased font size of the title**
-        axis.title.y = element_text(margin = margin(r = 5)))  # **Increased gap for y-axis label**
-
-# Arrange the plots side by side
-grid.arrange(p1, p2, p3, ncol = 3)  # ncol=3 arranges the plots in one row
+  theme(
+    panel.grid = element_blank(),  # Remove grid lines
+    plot.title = element_text(size = 14, hjust = 0.5),  # Center the title
+    axis.title.y = element_text(margin = margin(r = 10))  # Increased gap for y-axis label
+  )
 
@@
 
-The histograms **\[CROSS-REFERENCE HERE\]** reveal significant trends in sample sizes across the different ranges, indicating a general tendency towards smaller samples in the polling data.
-Most values within the 0-10,000 range cluster at the lower end, suggesting that many polls are conducted with limited respondent numbers, which may compromise the reliability and representativeness of the results.
-**However, this conclusion is very much dependent on what we determine to be an acceptable sample size for the results to be considered reliable**.
-In the 10,001-20,000 range, while there is some increase in sample sizes, the majority still falls within the 10,000-12,500 bracket, with few reaching the upper limits.
+The histogram of sample sizes reveals a clear right-skewed distribution.
+Most of the sample sizes are clustered between 0 and 3000, with a sharp peak around 1000-1500.
+This indicates that the majority of polls are conducted with smaller sample sizes.
+As sample size increases, the frequency significantly drops, with very few polls conducted with sample sizes larger than 5000, though there are a few outliers with sizes approaching 10,000 or more.
+This wide range in sample sizes might affect the precision of estimates in different polls.
 
-The findings from the 20,001-30,000 range further illustrate the scarcity of larger sample sizes, with only a couple of values noted at each of the significant points.
-This distribution emphasizes a preference for smaller samples in the polling data analyzed.
+### Most Frequent Pollsters
 
-In the context of polling results, these findings suggest potential limitations in the accuracy and validity of the insights generated.
-Larger sample sizes are typically preferred in polling as they tend to yield more reliable estimates and reduce the margin of error, leading to a better representation of the population's views.
-Consequently, the predominance of smaller sample sizes in this data set may raise concerns about the robustness of the polling findings, potentially affecting their applicability in broader analyses or decision-making processes.
-However, we must determine an acceptable and ideal sample size that would allow us to determine the reliability of these results.
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+# Get the top 10 pollsters by frequency along with their average pollscore and numeric grade
+top_pollsters <- analysis_data %>%
+  group_by(pollster) %>%           # Group by the 'pollster' variable
+  summarize(
+    Count = n(),                   # Count the number of occurrences
+    Average_Pollscore = round(mean(pollscore, na.rm = TRUE), 2),  # Average pollscore
+    Average_Numeric_Grade = round(mean(numeric_grade, na.rm = TRUE), 2),  # Average numeric grade
+    .groups = 'drop'               # Drop grouping structure for cleaner output
+  ) %>%
+  arrange(desc(Count)) %>%         # Arrange in descending order of count
+  slice_head(n = 10)                # Select the top 10 pollsters
 
-# Forecasting Approach Using Polls of Polls Method
+# Rename the columns to remove underscores and capitalize 'Pollster'
+colnames(top_pollsters) <- c("Pollster", "Count", "Average Pollscore", "Average Numeric Grade")
+
+kable(top_pollsters, format = "markdown", caption = "Top 10 Pollsters by Frequency")
+
+```
+
+This table lists the top 10 pollsters based on the frequency of polls they conducted.
+Morning Consult leads with 235 polls, followed by Siena/NYT and Redfield & Wilton Strategies with 94 and 88 polls, respectively.
+The table also shows each pollster's average poll score and average numeric grade.
+Most pollsters have negative poll scores, indicating a lower perceived reliability, except for Redfield & Wilton Strategies.
+However, the numeric grades vary, with Siena/NYT, YouGov, and Emerson receiving relatively high grades around 3, while others like TIPP and Redfield & Wilton Strategies are rated slightly lower.
+
+# Forcasting Election Outcome through Pooling Polls
+
+## Forecasting Approach
 
@@
 #| echo: false
-#| label: fig-enddate
-#| fig-cap: "Support for Kamala Harris based on End Date and colored by Pollster's Error and Bias."
+#| label: fig-mapinterac
+#| fig-cap: "Average Support for Kamala Harris in Different States estimated by Polls"
 
@@
 ```{r}
+#| warning: false
+#| message: false
+#| echo: false
 # Filter out non-continental states (Hawaii and Alaska)
@@
 p <- ggplot(map_data_full) +
-  geom_sf(aes(fill = mean_support_kamala, text = paste("State:", str_to_title(NAME), "<br>Average Support:", round(mean_support_kamala, 2))), color = "white") +  # Capitalize state names
+  geom_sf(aes(fill = mean_support, text = paste("State:", str_to_title(NAME), "<br>Average Support:", round(mean_support, 2))), color = "white") +  # Capitalize state names
   scale_fill_distiller(palette = "RdBu", direction = 1, na.value = "gray", name = "Kamala's Support") +  # Reverse RdBu color gradient
@@
 
-$$ y_i = \alpha + \beta_1 \cdot \texttt{pollster}_i + \beta_2 \cdot \texttt{sample_size}_i + \beta_3 \cdot \texttt{recency}_i +  \beta_4 \cdot \texttt{state}_i + \epsilon_i $$
+$$
+y_i = \alpha + \beta_1 \cdot \mathrm{pollster}_i + \beta_2 \cdot \mathrm{sample\_size}_i + \beta_3 \cdot \mathrm{recency}_i +  \beta_4 \cdot \mathrm{state}_i + \epsilon_i
+$$
 
@@
 ```{r}
-# Load necessary libraries
-library(broom)
-library(dplyr)
-library(knitr)
-
+#| warning: false
+#| message: false
+#| echo: false
 # Load the models from RDS files
@@
 ```{r}
+#| warning: false
+#| message: false
+#| echo: false
+
 # Create a data frame for AIC and BIC comparison
@@
 
+The table below shows the model output, where only a few states and pollsters are displayed to keep the output concise, as the full model includes many variables.
+We included "State California," "State Texas," "Pollster Redfield & Wilton Strategies," and "Pollster Echelon Insights" to provide examples of how different states and pollsters influence support for Kamala Harris.
+
+The variable "Sample Size" has a positive coefficient, suggesting that polls with larger sample sizes are associated with slightly higher support percentages.
+"Recency" has a negative and highly significant coefficient, indicating that more recent polls tend to show lower support levels.
+Additionally, certain states like California and Texas show significant effects on the percentage of support, with California showing a notable increase.
+Similarly, different pollsters also display varying biases, as seen with the coefficients for "Pollster Redfield & Wilton Strategies" and "Pollster Echelon Insights." The model shows strong goodness-of-fit, with an $R^2$ of 0.731, indicating that 73.1% of the variability in support can be explained by the included variables.
+
 ```{r}
@@
 #| message: false
+# Load necessary package
+library(stargazer)
 
-# Load necessary libraries
-library(modelsummary)
-library(broom)
-
-# Load Model 3 from RDS file
-model_3 <- readRDS(file = here::here("models/model_3.rds"))
-
-# Generate LaTeX output for the model
-latex_output <- modelsummary(model_3, output = "latex")
-
-# Ensure the output is a character string
-latex_output_string <- capture.output(cat(latex_output))
-
-# Write the LaTeX string to a .tex file
-writeLines(latex_output_string, here::here("model_3_summary.tex"))
+# Use stargazer to display only the specified variables and custom labels
+stargazer(model_3, 
+          type = "text",  # Can be "text", "html", or "latex"
+          covariate.labels = c("Sample Size", "Recency", "State California", "State Texas",
+                               "Pollster Redfield & Wilton Strategies", "Pollster Echelon Insights"),  # Renaming the variables
+          keep = c("sample_size", "recency", "stateCalifornia", "stateTexas", 
+                   "pollsterRedfield & Wilton Strategies", "pollsterEchelon Insights"),  # Keep only these variables
+          omit.stat = c("f", "ser"),  # Optionally omit F-statistics or other GOF stats
+          star.cutoffs = c(0.05, 0.01, 0.001),  # Custom significance stars
+          single.row = TRUE)  # Display coefficients on one row for readability
 
@@
 
-```{r}
-# Save LaTeX output to a string
-latex_output <- modelsummary(model_3, output = "latex")
-
-# Write the LaTeX string to a .tex file
-writeLines(latex_output, here::here("output/model_3_summary.tex"))
-```
+# Prediction
 
@@
 
-# Load necessary libraries
-library(modelsummary)
-library(broom)
+# Create a new dataset called 'predict_data' with predicted values and weights
+predict_data <- analysis_data %>%
+  # Add the predicted values (fitted values from model_3)
+  mutate(fitted_values = fitted(model_3),
+         # Create weights based on numeric grade
+         weight = numeric_grade / sum(numeric_grade))
 
-# Load Model 3 from RDS file
-model_3 <- readRDS(file = here::here("models/model_3.rds"))
-
-modelsummary(model_3, output = "latex", file = "model_3_summary.tex")
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# Load necessary libraries
-library(modelsummary)
-library(broom)
-
-# Load Model 3 from RDS file
-model_3 <- readRDS(file = here::here("models/model_3.rds"))
-
-modelsummary(model_3)
-
-# Display the summary of Model 3 using modelsummary
-#| echo: false
-#| eval: true
-#| label: tbl-model3results
-#| tbl-cap: "Regression results for Model 3 including Pollster, Sample Size, State, and Recency"
-
-modelsummary::modelsummary(
-  list(
-    "Model 3" = model_3
-  ),
-  statistic = "std.error",
-  fmt = 2
-)
+# Calculate the weighted fitted values for Kamala Harris support
+weighted_support <- sum(predict_data$fitted_values * predict_data$weight)
 
@@
 
-Our results are summarized in @tbl-modelresults.
+### Prediction
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
+To predict Kamala Harris' overall support, we used a weighted average approach based on the quality of each poll.
+The weights are calculated using each poll's `numeric_grade`, which reflects the reliability and transparency of the polling methodology.
 
-library(rstanarm)
+We define the weight for each pollster $w_i$ as follows:
 
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
+$$
+w_i = \frac{\mathrm{numeric\_grade}_i}{\sum_{i=1}^{n} \mathrm{numeric\_grade}_i}
+$$
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
+where:
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
+-   $w_i$​ represents the weight assigned to poll i,
+
+-   $numericgrade_i$ is the numeric grade of poll i, and
+
+-   $n$ is the total number of polls used in the analysis.
+
+Using these weights, the overall weighted prediction of Kamala Harris' support is calculated by summing the weighted predicted values from our regression model:
+
+$$
+\text{Overall Weighted Support} = \sum_{i=1}^{n} w_i \cdot \hat{y}_i
+$$
+
+-   $\hat{y}_i$​ is the predicted percentage of support for Kamala Harris from poll i.
+
+Applying this approach, our model predicts Kamala Harris' overall support to be **47.78%**.
+This prediction indicates that, after adjusting for biases in the polls, Harris' support is slightly below the 50% threshold.
+The weighted approach accounts for the differences in poll reliability and thus provides a more accurate estimate than using raw averages of the polls.
+
+This result suggests that Kamala Harris is facing some challenges in gaining majority support based on the current polling data.
+However, the model reflects the weighted average from multiple sources, smoothing out individual poll biases to offer a comprehensive view of her support landscape.
 
@@
 
-## First discussion point {#sec-first-point}
+In this paper, we set out to predict Kamala Harris' support using a polls-of-polls approach.
+By aggregating multiple polls, we aimed to smooth out the inherent biases present in individual surveys and create a more accurate and reliable forecast.
+Our analysis was driven by a linear regression model that used key predictors, such as pollster, sample size, state, and recency of the polls.
+After obtaining predicted values from the model, we applied a weighting scheme based on the numeric grade of each pollster, reflecting the quality and reliability of the poll.
+By assigning higher weights to more reliable pollsters, we generated a final prediction for Kamala Harris' overall support, accounting for variations across different states and polling organizations.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+One key insight from our analysis is the significance of poll recency.
+The negative and highly significant coefficient for recency indicates that **older** polls show lower support for Kamala Harris, while more recent polls tend to show higher support.
+This suggests that her popularity may be increasing over time, or that certain factors have shifted voter sentiment in her favor in more recent months.
+Inncorporating recency into the model also greatly improved its overall performance.
+Adding recency as a predictor significantly increased the model's explanatory power, raising the R² value and lowering both AIC and BIC scores compared to models that excluded it.
+This indicates that accounting for the timing of polls is essential to accurately predicting voter sentiment, as public opinion is highly dynamic and evolves rapidly in response to events.
+Including recency strengthens the model's ability to reflect these temporal shifts, leading to more reliable forecasts.
 
-## Second discussion point
+Our analysis also highlights the variability in support for Kamala Harris across different states and pollsters.
+For example, certain states like California and Texas exhibit distinct effects, with California significantly boosting Harris’ predicted support.
+Additionally, different pollsters show varying levels of bias, as seen with Redfield & Wilton Strategies and Echelon Insights, which have divergent effects on support levels.
+This underscores the geographic and methodological complexities in predicting election outcomes, demonstrating that voter preferences are not homogenous across regions or polling firms.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+Despite the strengths of the polls-of-polls methodology, there are several limitations to our approach.
+First, our model assumes that the relationship between the predictors (pollster, sample size, state, recency) and support for Kamala Harris is linear, which may not always be the case.
+Non-linear trends or interactions between variables might exist that were not captured by the model.
+Additionally, while we weighted polls based on their numeric grade, this metric might not fully reflect the true accuracy or bias of each pollster.
+Moreover, we only included polls conducted after Harris' official candidacy announcement, potentially omitting relevant data from earlier in the election cycle that could provide a fuller picture.
 
-## Third discussion point
+Future research could delve into more sophisticated methods to capture non-linear relationships and interactions between variables.
+For instance, incorporating machine learning algorithms like random forests or neural networks could allow for more flexible modeling of polling data.
+Another area of improvement could be refining the pollster weighting mechanism by integrating other factors, such as the historical performance of the pollster or specific characteristics of the electorate surveyed.
+Additionally, conducting more in-depth analyses of key battleground states could reveal more granular insights into voter behavior, especially in regions where small shifts in support could have outsized impacts on election outcomes.
 
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+In conclusion, while our model provides valuable predictions of Kamala Harris' support, there is always room to enhance the predictive accuracy by refining methodologies and incorporating new data sources.
+The evolving nature of elections demands continuous adjustment of models to reflect changes in public opinion, candidate strategies, and electoral dynamics.
 
@@
 
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
 ## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
 

### 2024-10-22 scripts/03-clean_data.R

@@ -89,6 +89,30 @@ cleaned_data <- cleaned_data %>%
 cleaned_data <- cleaned_data %>%
   filter(end_date >= as.Date("2024-07-21"))
 
+# Calculate recency 
+
+reference_date <- max(analysis_data$end_date, na.rm = TRUE)  # Most recent poll date
+
+# Calculate recency in days
+analysis_data <- analysis_data %>%
+  mutate(recency = as.numeric(reference_date - end_date))
+
+# Check the dataset after adding recency
+head(analysis_data)
+
 #### Save data ####
 
 write_parquet(cleaned_data, "data/02-analysis_data/analysis_data.parquet")
+
+#### Creating Mapping Data ####
+
+map_data <- raw_data %>%
+  filter(candidate_name == "Kamala Harris") %>%  # Filter for Kamala Harris
+  distinct() %>%  # Remove duplicates
+  select(state, pct) %>%  # Keep only the 'state' and 'pct' columns
+  group_by(state) %>%  # Group by state
+  summarise(mean_support = mean(pct, na.rm = TRUE))  # Calculate the mean pct for each state
+
+#### Save data ####
+
+write_parquet(map_data, "data/02-analysis_data/map_data.parquet")

### 2024-10-22 paper/paper.qmd
include appendix

@@
 
-# Additional data details
+# Emerson College Polling Methodology (October 14-16, 2024)
 
-# Model details {#sec-model-details}
+## Overview
+Emerson College Polling, known for its extensive online polling methods, conducted a survey from October 14 to 16, 2024, targeting 1,000 likely voters.
+The poll measured voter preferences between Kamala Harris and Donald Trump, revealing a near tie with Harris at 50% and Trump at 49%.
 
-## Diagnostics
+## Population, Frame, and Sample
+The target population is likely voters in the U.S.
+elections.
+These voters are determined by the voters' likeliness to vote in the upcoming elections and the voter's voting history.
+Both of these factors are self-reported.
 
+The sampling frame is the voters that are on the Aristotle list and on the online panel provided by CINT.
+Aristotle is an online supplier of voter files and CINT is a software company that connects researchers, such as Emerson, to online panels.
+This approach helps ensure a representative sample by accurately reflecting the electorate.
+
+The sample size is the 1,000 likely voters chosen arbitrarily from the sampling frame.
+Likely voters are determined by a combination of voter history, registration status, and demographic data.
+These factors are self-reported.
+
+## Sample Recruitment and Approach
+Emerson College used a mixed-mode sampling approach for its polling. Specifically, three primary modes were used to collect data for its October 14-16, 2024 poll:
+
+- MMS-to-web text surveys: Respondents were contacted via text messages sent to cell phones.
+These messages directed participants to complete the survey online.
+This method was conducted using voter lists provided by Aristotle.
+
+- Interactive Voice Response (IVR): Landlines were targeted using automated phone calls where respondents could answer questions using their keypads.
+The contact information for this was also sourced from Aristotle’s voter lists.
+
+- Online Panel from CINT: Emerson utilized a pre-screened, opt-in online panel of voters provided by CINT.
+
+## Advantages and Trade-offs in Mixed-Mode Sampling
+The multi-mode sampling approach used in this poll has several advantages:
+
+- Wider demographic reach: Having a combination of data collection methods allows Emerson to gather responses from a broader demographic group.
+In this polling, MMS-to-web surveys mostly gather responses from younger voters and those who primarily use mobile devices.
+IVR target older and rural voters, who are more likely to respond via landlines.
+Online panels is more efficient in reaching tech-savvy individuals.
+
+- Cost-efficient: Automated technologies reduce the cost compared to live interviews. In a polling context, this allows for a larger sample size.
+
+However, the trade-offs include:
+
+- Coverage bias: IVR only reaches individuals with landlines, which may skew results towards older demographics.
+
+- Self-selection bias: Respondents in online panels tend to be more engaged, which can lead to over-representation of politically active individuals.
+
+## Non-response Handling
+Emerson addresses non-response using a weighting system.
+The data is adjusted based on key demographic variables like age, gender, race, education, and party affiliation, making the sample more representative of the electorate.
+Weights are applied to compensate for under- or over-represented groups, ensuring balanced results.
+
+## Questionnaire Design
+The Emerson survey features a straightforward, efficient design focused on key issues and voter preferences. Strengths of the questionnaire include:
+
+- Simplicity: Clear, concise questions makes the survey easy for respondents to understand and to provide accurate answers.
+
+- Topical relevance: The poll zeroes in on vote preference and demographic splits, delivering timely insights during election cycles.
+
+However, weaknesses include:
+
+- Lack of nuance: While the survey captures basic preferences, it does not delve deeply into voters' motivations or the factors driving their choices.
+
+- Mode limitations: Different survey modes can affect how respondents engage with the questions. This can potentially influence the depth and accuracy of voter responses.
+
+## Conclusion
+Emerson College Polling's mixed-mode methodology is effective in balancing cost with broad demographic reach.
+Its weighting techniques contribute to reliable and accurate results despite non-responses.
+While the use of multiple survey modes increases population representation, possible biases—such as lack of nuance and mode effects need to be taken into consideration when interpreting results.
+
+In sum, this method is a strong predictor of election outcomes and continues to adapt to the evolving challenges of modern polling.
+
+# Idealized Methodology
+
+## Overview
+This methodology outlines an election forecasting plan with a budget of $100,000.
+Heavily inspired by TIPP Insights and Emerson College's hybrid approaches, the survey methodology balances cost-efficiency and data accuracy by using different modes of data collection and wide demographic reach.
+
+This forecasting plan will predict the popular vote and Electoral College outcomes, emphasizing key swing states and underrepresented voter groups.
+
+## Sampling Approach
+This idealized methodology uses a multi-mode hybrid sampling approach. This allows the use of both probability-based and non-probability-based methods:
+
+- Probability-based methods:
+
+  - Live phone interviews (landline and cell): As used by TIPP, live phone interviews is a direct and personal method of collecting responses.
+Voter lists would be sourced from Aristotle to ensure we contact likely voters.
+This mode reaches older voters and voters with less internet connection.
+
+  - Interactive Voice Response (IVR): Similar to Emerson’s approach, IVR for landline users balances costs and ensures demographic representation, particularly among older or less tech-savvy populations.
+This method captures quick responses and is less expensive than live interviews.
+
+- Non-probability methods:
+
+  - MMS-to-web (mobile): Following Emerson’s MMS-to-web strategy, we would send multimedia messages to likely voters’ cell phones, encouraging them to complete an online survey.
+This method works well for younger, mobile-first voters, helping fill in gaps where traditional methods might miss.
+
+  - Online panel recruitment via CINT: To mirror TIPP’s online panel approach, a portion of the survey would be conducted via an opt-in panel of pre-screened voters from CINT, capturing tech-savvy respondents.
+CINT’s platform ensures that different voter groups, including hard-to-reach groups, are represented in the polling.
+
+## Recruitment Strategy
+By using Aristotle’s voter lists and CINT's platform as a sampling frame, the sample consists of likely voters based on registration status and past voting behavior.
+Quota sampling would be used to match the demographic profile of the U.S.
+electorate.
+This ensures that respondents in the sample reflect key demographic variables such as age, gender, race, and education level.
+
+We would oversample key swing states like Pennsylvania, Wisconsin, and Florida.
+Key swing states are states where there is no overwhelming support for a particular candidate or political party.
+The outcome in these states often determine the overall result of the election, making them critical in securing a win in the Electoral College.
+Oversampling these states allows a more accurate forecasting in these battlegrounds, which are critical for predicting the Electoral College outcome.
+
+## Data Validation
+
+Given that the data is collected in a hybrid format, we would use duplicate detection methods via IP addresses or phone numbers to eliminate multiple responses from the same individual.
+Also, we would implement attention checks to identify respondents who are not fully engaged.
+Both duplicate respondents and unengaged respondents bias results, and should be removed from the data set.
+
+Following TIPP’s iterative weighting process, the data will be weighted to ensure that the sample aligns with key demographics (race, gender, education) based on census data.
+
+## Conclusion
+By combining TIPP’s live phone/IVR hybrid and Emerson College’s mixed-mode survey methods, this idealized methodology captures diverse voter demographics.
+It uses both probability-based and non-probability-based sampling methods to forecast the 2024 U.S.
+Presidential Election.
+
+A link to the survey can be found at: https://forms.gle/haGLaQBPQhrXE6Xa7 
 \newpage

### 2024-10-22 paper/paper.qmd
Added a subtitle.

@@
 title: Forcasting US Elections using Polls-of-Polls
-subtitle: "My subtitle if needed"
+subtitle: Using a Linear Regression Model to Adjust Pollster Biases and Predict Kamala Harris' Vote Percentage at 47.78%
 author: 
@@
 These findings are significant as they provide a more reliable estimate of electoral outcomes, helping to inform political strategies and public discourse."
-format: html
+format: pdf
 number-sections: true
@@
 
-### Political Leanings of Different State
-
-[@fig-mapinterac] provides a state-level representation of Kamala Harris's average support across the U.S., with the darker blue shades indicating states where her support is higher, and the redder tones showing where her support is lower.
-Gray areas represent states where we have no polling data available.
-
-The gradient highlights notable geographic trends.
-Harris has the strongest support in states like Vermont, Massachusetts, and New York, while states like Utah and West Virginia show considerably lower levels of support.
-A distinct pattern emerges in the West, where coastal states like California favor Harris more strongly, contrasting with inland states like Wyoming and Utah.
-
-This reveals that political leanings differ greatly by state, and this regional variation should be controlled for in any regression model analyzing poll data.
-The map serves as a useful tool for visualizing regional variations in voter support for Kamala Harris as reflected in the available polling data.
-
-```{r}
-#| warning: false
-#| message: false
-#| echo: false
-# Filter out non-continental states (Hawaii and Alaska)
-us_states_filtered <- mapping_data %>% 
-  filter(!NAME %in% c("Alaska", "Hawaii"))
-
-# Ensure correct merge between map data and Kamala's support data
-map_data_full <- left_join(us_states_filtered, mean_support_data, by = c("NAME" = "state"))
-
-# Create centroids for state labels
-state_centroids <- st_centroid(us_states_filtered)
-
-# Plot using geom_sf() for sf objects
-p <- ggplot(map_data_full) +
-  geom_sf(aes(fill = mean_support, text = paste("State:", str_to_title(NAME), "<br>Average Support:", round(mean_support, 2))), color = "white") +  # Capitalize state names
-  scale_fill_distiller(palette = "RdBu", direction = 1, na.value = "gray", name = "Kamala's Support") +  # Reverse RdBu color gradient
-  geom_text(data = state_centroids, aes(geometry = geometry, label = STUSPS), stat = "sf_coordinates", size = 2) +  # State labels
-  labs(title = "Support for Kamala Harris by State", 
-       caption = "States with no polling data are shown in gray") +
-  theme_minimal() +
-  theme(
-    legend.position = "right",
-    plot.title = element_text(hjust = 0.5),
-    panel.grid = element_blank(),  # Remove grid lines
-    axis.title = element_blank(),  # No axis titles for maps
-    axis.text = element_blank(),   # No axis labels
-    axis.ticks = element_blank()   # No axis ticks
-  )
-
-# Turn ggplot to plotly for interactivity, add hover template for better control over display
-ggplotly(p, tooltip = "text")
-
-```
-
 # Model
@@
 ## Overview
+
 Emerson College Polling, known for its extensive online polling methods, conducted a survey from October 14 to 16, 2024, targeting 1,000 likely voters.
@@
 ## Population, Frame, and Sample
+
 The target population is likely voters in the U.S.
@@
 ## Sample Recruitment and Approach
+
 Emerson College used a mixed-mode sampling approach for its polling. Specifically, three primary modes were used to collect data for its October 14-16, 2024 poll:
@@
 ## Advantages and Trade-offs in Mixed-Mode Sampling
+
 The multi-mode sampling approach used in this poll has several advantages:
@@
 ## Non-response Handling
+
 Emerson addresses non-response using a weighting system.
@@
 ## Questionnaire Design
+
 The Emerson survey features a straightforward, efficient design focused on key issues and voter preferences. Strengths of the questionnaire include:
@@
 ## Conclusion
+
 Emerson College Polling's mixed-mode methodology is effective in balancing cost with broad demographic reach.
@@
 ## Overview
-This methodology outlines an election forecasting plan with a budget of $100,000.
+
+This methodology outlines an election forecasting plan with a budget of \$100,000.
 Heavily inspired by TIPP Insights and Emerson College's hybrid approaches, the survey methodology balances cost-efficiency and data accuracy by using different modes of data collection and wide demographic reach.
@@
 ## Sampling Approach
+
 This idealized methodology uses a multi-mode hybrid sampling approach. This allows the use of both probability-based and non-probability-based methods:
@@
 ## Recruitment Strategy
+
 By using Aristotle’s voter lists and CINT's platform as a sampling frame, the sample consists of likely voters based on registration status and past voting behavior.
@@
 ## Conclusion
+
 By combining TIPP’s live phone/IVR hybrid and Emerson College’s mixed-mode survey methods, this idealized methodology captures diverse voter demographics.
@@
 
-A link to the survey can be found at: https://forms.gle/haGLaQBPQhrXE6Xa7 
-\newpage
+A link to the survey can be found at: https://forms.gle/haGLaQBPQhrXE6Xa7 \newpage
 

### 2024-10-23 scripts/00-simulate_data.R
remove redundant library (dplyr and lubridate are in tidyverse)

@@ -1,8 +1,7 @@
 #### Preamble ####
-# Purpose: Simulates a data set of polling results for 
-# forecasting the upcoming US elections. 
+# Purpose: Simulates a data set of polling results for forecasting the upcoming US elections. 
 # Author: Krishna Kumar
-# Date: today
+# Date: October 19 2024
 # Contact: krishna.kumar@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
@@ -11,13 +10,10 @@
 #### Workspace setup ####
 
 library(tidyverse)
-library(dplyr)
-library(lubridate)
-
-set.seed(420)
-
 
 #### Simulate data ####
+
+set.seed(420)
 
 # Number of observations #
 n <- 100

### 2024-10-23 scripts/01-test_simulated_data.R

@@ -1,15 +1,12 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests the structure and validity of the simulated polling results data set
+# Author: Krishna Kumar
+# Date: October 19 2024
+# Contact: krishna.kumar@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
   # - The `tidyverse` package must be installed and loaded
   # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-10-23 scripts/02-clean_data.R

@@ -12,7 +12,7 @@
 library(tidyverse)
 library(janitor)
 library(arrow)
-library(lubridate)
+
 raw_data <- read_csv("data/01-raw_data/raw_data.csv")
 
 #### Data Cleaning ####

### 2024-10-23 paper/paper.qmd
fix sub-section format and a few typos

@@
 ---
-title: Forcasting US Elections using Polls-of-Polls
+title: Forcasting 2024 U.S. Election using Polls-of-Polls
 subtitle: Using a Linear Regression Model to Adjust Pollster Biases and Predict Kamala Harris' Vote Percentage at 47.78%
@@
 
+#### Workspace set up ####
 library(psych)
@@
 library(kableExtra)
-library(sf)
-library(tidyverse)
 library(plotly)
@@
 library(sf)
+
+# Load data
 analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
@@
 
-### Measurement
+## Measurement
 
@@
 
-### Data Cleaning
+## Data Cleaning
 
 The raw data for this project, sourced from FiveThirtyEight, [@FiveThirtyEight] underwent a series of cleaning steps to prepare it for analysis.
-Initially, duplicate rows were removed to ensure that only unique observations remained, facilitated by the janitor package \[\@janitor\].
+Initially, duplicate rows were removed to ensure that only unique observations remained, facilitated by the janitor package [@janitor].
 A new binary variable, 'national', was created to indicate whether a poll was conducted at the national or state level.

### 2024-10-23 paper/paper.qmd
fix reference and include survey copy in appendix

@@
 ---
-title: Forcasting 2024 U.S. Election using Polls-of-Polls
-subtitle: Using a Linear Regression Model to Adjust Pollster Biases and Predict Kamala Harris' Vote Percentage at 47.78%
+title: Forecasting 2024 U.S. Election using Polls-of-Polls
+subtitle: Using a Linear Regression Model to Adjust Pollster Biases and Predict Kamala Harris Vote Percentage at 47.78%
 author: 
@@
 These findings are significant as they provide a more reliable estimate of electoral outcomes, helping to inform political strategies and public discourse."
-format: pdf
+format: 
+  pdf:
+    documentclass: article
     number-sections: true
+    header-includes:
+      - \usepackage{pdfpages}
 bibliography: references.bib
@@
 The raw data for this project, sourced from FiveThirtyEight, [@FiveThirtyEight] underwent a series of cleaning steps to prepare it for analysis.
-Initially, duplicate rows were removed to ensure that only unique observations remained, facilitated by the janitor package [@janitor].
+Initially, duplicate rows were removed to ensure that only unique observations remained, facilitated by the `janitor` package [@janitor].
 A new binary variable, 'national', was created to indicate whether a poll was conducted at the national or state level.
@@
 This cutoff was selected to retain mid to high-level pollsters for more reliable results.
-These steps were performed using functions from the dpylr package \[\@dplyr\].
-Furthermore, dates were standardized and converted into a proper format for analysis using the lubridate package \[\@lubridate\].
+These steps were performed using functions from the `dpylr` package [@dplyr].
+Furthermore, dates were standardized and converted into a proper format for analysis using the `lubridate` package [@lubridate].
 Polls related to Kamala Harris were retained for further analysis, and percentage support values were transformed into actual numbers of supporters based on sample size.
 Additionally, pollster counts below five were excluded to focus on more reliable data sources.
-Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the arrow package \[\@arrow\].
+Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the `arrow` package [@arrow].
 
@@
 
-# Forcasting Election Outcome through Pooling Polls
+# Forecasting Election Outcome through Pooling Polls
 
@@
 
-\[\@fig-sample\] suggests that the relationship between support for Kamala Harris and sample size shows a slight positive trend, indicating a small increase in support as sample sizes grow.
+[@fig-sample] suggests that the relationship between support for Kamala Harris and sample size shows a slight positive trend, indicating a small increase in support as sample sizes grow.
 Both national (blue) and state (green) polls reflect this upward trend, although the effect is minimal.
@@
 
-In \[\@fig-enddate\], we observe a slight increasing trend in support for Kamala Harris as the end date of polls progresses towards October.
+In [@fig-enddate], we observe a slight increasing trend in support for Kamala Harris as the end date of polls progresses towards October.
 This may indicate a gradual improvement in her polling performance over time.
@@
 #| echo: false
-#| label: fig-mapinterac
+#| label: fig-enddate
 #| fig-cap: "Average Support for Kamala Harris in Different States estimated by Polls"
@@
 ```
-
-### Prediction
 
@@
 
-A link to the survey can be found at: https://forms.gle/haGLaQBPQhrXE6Xa7 \newpage
+A link to the survey can be found at: https://forms.gle/haGLaQBPQhrXE6Xa7
+
+# Copy of the Survey
+\includepdf[pages=-]{../other/surveyfiles/SurveyCopy.pdf}
+
+\newpage
 

### 2024-10-30 scripts/03-model_data.R
Fixed an error in the modelling script.

@@ -1,37 +1,42 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models the percentage of support for Kamala Harris in each poll as a function of various predictors.
+# Author: Shamayla Durrin
+# Date: 30-10-2024
+# Contact: shamayla.islam@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Ensure that the necessary libraries (tidyverse, arrow, etc.) are installed and data is available in the required format.
+# Any other information needed? Data should include variables for support, pollster, sample size, state, and recency.
 
 
 #### Workspace setup ####
+
 library(tidyverse)
-library(rstanarm)
+library(here)
+library(arrow)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# Load necessary libraries
+library(tidyverse)
 
+# Assuming your data is loaded in a dataframe called `data`, with variables:
+# - support: percentage of support for Kamala Harris
+# - pollster: categorical variable for the polling organization
+# - sample_size: numeric variable for sample size
+# - state: categorical variable for the state
+# - recency: numeric variable representing the recency of the poll
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
+# Model 1: Only pollster and sample size as predictors
+model1 <- lm(pct ~ pollster + sample_size, data = analysis_data)
+saveRDS(model1, "model_1.rds")
+
+# Model 2: Adding state as a predictor
[…]

### 2024-10-31 scripts/02-clean_data.R
Wrapped up predictions. Did model diagnostic. Fixed title and subtitle. Fixed abstract.

@@ -13,7 +13,7 @@ library(tidyverse)
 library(janitor)
 library(arrow)
 
-raw_data <- read_csv("data/01-raw_data/raw_data.csv")
+raw_data <- read_csv(here("data/01-raw_data/raw_data.csv"))
 
 #### Data Cleaning ####
 
@@ -33,7 +33,7 @@ cleaned_data <- cleaned_data %>%
 
 # Keep only the relevant columns
 cleaned_data <- cleaned_data %>%
-  select(poll_id, pollster_id, pollster, numeric_grade, pollscore, state, start_date, national, end_date, sample_size, population, candidate_name, pct)
+  select(poll_id, pollster, numeric_grade, pollscore, state, national, end_date, sample_size,candidate_name, pct)
 
 # Remove any rows that have missing values in any column
 cleaned_data <- cleaned_data %>% drop_na()
@@ -49,15 +49,9 @@ cutoffgrade <- 1 # this cutoff was selected  to retain mid to high quality polls
 cleaned_data <- cleaned_data %>% filter(numeric_grade >= cutoffgrade)
 
 
-#Filter the dataset to keep only rows with Kamala Harris or Donald Trump
-
-cleaned_data <- cleaned_data %>%
-  filter(candidate_name %in% c("Kamala Harris"))
-
 # Ensure the start_date and end_date columns are in date format
 cleaned_data <- cleaned_data %>%
   mutate(
-    start_date = parse_date_time(start_date, orders = c("mdy", "ymd")),
     end_date = parse_date_time(end_date, orders = c("mdy", "ymd"))
   )
 
@@ -77,12 +71,7 @@ cleaned_data <- cleaned_data %>%
   filter(n() >= 5) %>%
   ungroup()
 
-# Convert percentage to actual number for modelling
 
-cleaned_data <- cleaned_data %>%
-  mutate(
-    num_support = round((pct / 100) * sample_size, 0)  
-  )
 
 # Filter Observatons after Harris was declared 
 
@@ -91,28 +80,17 @@ cleaned_data <- cleaned_data %>%
 
 # Calculate recency 
 
-reference_date <- max(analysis_data$end_date, na.rm = TRUE)  # Most recent poll date
+reference_date <- max(cleaned_data$end_date, na.rm = TRUE)  # Most recent poll date
 
 # Calculate recency in days
-analysis_data <- analysis_data %>%
+cleaned_data <- cleaned_data %>%
[…]

### 2024-10-31 scripts/03-model_data.R

@@ -17,26 +17,40 @@ library(arrow)
 #### Read data ####
 analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
 
-# Load necessary libraries
-library(tidyverse)
 
-# Assuming your data is loaded in a dataframe called `data`, with variables:
-# - support: percentage of support for Kamala Harris
-# - pollster: categorical variable for the polling organization
-# - sample_size: numeric variable for sample size
-# - state: categorical variable for the state
-# - recency: numeric variable representing the recency of the poll
+# Filter all the observations regarding Harris. Model 1,2,3 will be about Harris. 
+
+harris_data <- analysis_data %>%
+  filter(candidate_name %in% c("Kamala Harris"))
 
 # Model 1: Only pollster and sample size as predictors
-model1 <- lm(pct ~ pollster + sample_size, data = analysis_data)
-saveRDS(model1, "model_1.rds")
+model1 <- lm(pct ~ pollster + sample_size, data = harris_data)
+
+saveRDS(model1, here("models", "model1.rds"))
 
 # Model 2: Adding state as a predictor
-model2 <- lm(support ~ pollster + sample_size + state, data = data)
-saveRDS(model2, "model_2.rds")
+model2 <- lm(pct ~ pollster + sample_size + state, data = harris_data)
+saveRDS(model2, here("models", "model2.rds"))
 
 # Model 3: Adding recency as a predictor
-model3 <- lm(support ~ pollster + sample_size + state + recency, data = data)
-saveRDS(model3, "model_3.rds")
+model3 <- lm(pct ~ pollster + sample_size + state + recency, data = harris_data)
+saveRDS(model3, here("models", "model3.rds"))
 
+# Filter all the observations regarding Trump. Model 4,5,6 will be about Trump. 
+
+trump_data <- analysis_data %>%
+  filter(candidate_name %in% c("Donald Trump"))
+
+# Model 4: Only pollster and sample size as predictors
+model4 <- lm(pct ~ pollster + sample_size, data = trump_data)
+
+saveRDS(model4, here("models", "model4.rds"))
+
+# Model 5: Adding state as a predictor
+model5 <- lm(pct ~ pollster + sample_size + state, data = trump_data)
+saveRDS(model5, here("models", "model5.rds"))
+
+# Model 6: Adding recency as a predictor
+model6 <- lm(pct ~ pollster + sample_size + state + recency, data = trump_data)
+saveRDS(model6, here("models", "model6.rds"))
 

### 2024-10-31 scripts/04-create_graph.R

@@ -1,0 +1,66 @@
+#### Preamble ####
+# Purpose: Creates and exports a DAG (Directed Acyclic Graph) illustrating causal relationships in the polls-of-polls model.
+# Author: Shamayla Durrin
+# Date: 30-10-2024
+# Contact: shamayla.islam@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Libraries DiagrammeR, DiagrammeRsvg, and rsvg must be installed.
+# Additional Notes: This chunk saves the DAG as a PNG file to the specified directory for use in the main report.
+
+# Load necessary libraries
+library(DiagrammeR)
+library(DiagrammeRsvg)
+library(rsvg)
+library(ggfortify)
+library(here)
+
+# Define the causal model
+causal_model <- "digraph {
+  # Define nodes
+  node [shape = plaintext]
+    A [label = 'Candidate Support']
+    B [label = 'Pollster']
+    C [label = 'Sample Size']
+    D [label = 'State']
+    E [label = 'Recency']
+    F [label = 'Unobserved Confounders']
+
+  # Define directed edges (causal relationships)
+  edge []
+    B -> A
+    C -> A
+    D -> A
+    E -> A
+
+  # Dashed edges for unobserved confounders
+  edge [style = dashed]
+    F -> B
+    F -> E
+    F -> A
+}"
+
+# Render and export the graph
+graph <- grViz(causal_model)
+
+# Save the graph as a PNG file
+export_svg(graph) |>
+  charToRaw() |>
+  rsvg::rsvg_png("other/graphs/polls_of_polls_dag.png", width = 1000)
+
+
+
+# Define the file paths using the here() function
+output_path_model3 <- here("other", "graphs", "diagnostic_plot_model3.png")
+output_path_model6 <- here("other", "graphs", "diagnostic_plot_model6.png")
+
+# Save diagnostic plot for model3
+png(output_path_model3, width = 800, height = 600)
+autoplot(model3, label.size = 3, smooth.colour = "blue", smooth.linetype = "dashed") + theme_classic()
+dev.off()
+
+# Save diagnostic plot for model6
+png(output_path_model6, width = 800, height = 600)
+autoplot(model6, label.size = 3, smooth.colour = "blue", smooth.linetype = "dashed") + theme_classic()
+dev.off()
+
+

### 2024-10-31 paper.qmd
Updated the readme.

---
title: "Forecasting the 2024 U.S. Presidential Election: An Analysis for Battleground States"
subtitle: "Trump Favored Nationally, While Harris Leads in 4 of 7 Battleground States"
author: 
  - Shamayla Durrin
  - Denise Chang 
  - Krishna Kumar
thanks: "Code and data are available at: [https://github.com/krishnak30/US_elections](https://github.com/krishnak30/US_elections)."
date: today
date-format: long
abstract: >
  This study employs a polls-of-polls approach to predict support for Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election, with a focus on key battleground states to forecast the likely winner of the electoral college. By aggregating multiple polls and applying a weighted linear regression model with predictors like pollster reliability, sample size, state, and recency, we estimate a higher national support for Trump relative to Harris. Our analysis also reveals a close competition in battleground states, with Trump holding a narrow lead in Arizona, Georgia, and North Carolina, while Harris leads in Michigan, Nevada, Pennsylvania, and Wisconsin. These findings underscore the value of aggregating polls over relying on individual surveys, offering a more comprehensive and robust forecast of electoral outcomes.
format: 
  pdf:
    number-sections: true
    header-includes:
      - \usepackage{pdfpages}
      - '\KOMAoption{captions}{tableheading}'
      - \usepackage[margin=1in]{geometry}

bibliography: references.bib

---

```{r}
#| include: false
#| warning: false
#| message: false

#### Workspace set up ####
library(psych)
library(tidyverse)
library(arrow)
library(here)
library(ggplot2)
library(kableExtra)
library(plotly)
library(RColorBrewer)
library(stringr)
library(modelsummary)
library(broom)
library(sf)
library(patchwork)
library(ggfortify)

# Load data
analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
us_states <- st_read(here("data/03-mapping_data/US"))
```

\vspace{0.7cm}

# Introduction

While individual polls provide snapshots of public opinion, they are often subject to biases and methodological differences. In this paper, we aim to forecast voter support for Kamala Harris and Donald Trump by aggregating data from multiple polling sources, reducing individual poll biases, and improving overall prediction accuracy. Our analysis focuses on estimating the levels of support for each candidate not only nationally but also in key swing states that are likely to be decisive in determining the electoral college outcome.

The estimand of this study is the level of voter support for each candidate, Kamala Harris and Donald Trump, as reported across multiple polls.To estimate this, we developed a regression model incorporating variables such as pollster, sample size, state, and recency of the poll, with an emphasis on accurately capturing state-level dynamics. Our findings indicate that, on a national level, support between Kamala Harris and Donald Trump is closely balanced. Harris’s estimated national support stands close to 48%, while Trump’s support is slightly higher, around 49%. We found that Trump holds a marginal lead in battleground states like Arizona and Georgia, with a support margin slightly above 1%. In contrast, Harris shows narrow leads in Michigan, Nevada, Pennsylvania, and Wisconsin, with her support margin in Wisconsin reaching over 2%, making it her strongest battleground. North Carolina emerges as one of the tightest races, with Trump leading by only 0.26%, underscoring the competitive nature of these regions.

This paper contributes to the field of election forecasting by emphasizing the importance of poll quality and recency in prediction models. By focusing on battleground states where the support margins are particularly close, our approach highlights the areas where campaign strategies and voter turnout efforts may have the greatest impact. This could provides insights for political strategists, media analysts, and the general public by offering a nuanced view of the electoral landscape.

This paper is organized as follows: In {#sec-data}, we clean the data, explore summary statistics, plot distributions of key variables, and examine relationships between variables. In {#sec-model}, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions. Finally, in {#sec-discussion}, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work.
\vspace{0.7cm}
# Data {#sec-data}
\vspace{0.7cm}
## Measurement

In the dataset of our analysis, the process of measurement begins by capturing real-world public opinion through surveys. Polling organizations transform this observed phenomenon into structured data by recording support percentages, pollster details, and methodological choices, thus turning public sentiment into quantifiable entries in the dataset.

Public opinion polls are essential for understanding voter preferences and electoral dynamics, providing data that informs policy decisions and campaign strategies. Our data set captures public support for candidates at specific times, along with information on methodologies, sample sizes, and pollster ratings.

Polling organizations use various methods, such as App panels, IVR (Interactive Voice Response), Live phone calls, Text-to-web systems etc. These methods can influence the validity and reliability of results, with larger sample sizes generally yielding more reliable estimates. Some methods, like live phone interviews, may be seen as more trustworthy than others.

The dataset also includes important measures of pollster performance, such as numeric grades, poll scores, and transparency scores. These metrics are calculated by considering bias, race difficulty, predictive error, and transparency, allowing us to assess the accuracy of each poll. The "pollscore," for instance, reflects a pollster’s past performance and reliability, adjusted for potential biases.

\vspace{0.7cm}

## Data Cleaning

The raw data for this project, sourced from FiveThirtyEight, [@FiveThirtyEight] underwent a series of cleaning steps to prepare it for analysis. Initially, duplicate rows were removed to ensure that only unique observations remained, facilitated by the `janitor` package [@janitor]. A new binary variable, 'national', was created to indicate whether a poll was conducted at the national or state level. Missing values in the 'state' column were replaced with "Not Applicable," and numeric grades were evaluated to filter out low-quality pollsters, keeping only those with a numeric grade above 1. This cutoff was selected to retain mid to high-level pollsters for more reliable results. These steps were performed using functions from the `Tidyverse` package [@tidy]. Furthermore, dates were standardized and converted into a proper format for analysis using the `lubridate` package [@lubridate]. Polls related to Kamala Harris were retained for further analysis, and percentage support values were transformed into actual numbers of supporters based on sample size. Additionally, pollster counts below five were excluded to focus on more reliable data sources. Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the `arrow` package [@arrow].

## Explorations of Variables of Interest

### Summary Statistics of Key Variables

\vspace{0.7cm}

For the purposes of analysis in this paper, only a subset of variables was selected. Two additional variable, national and state, was created using the existing 'state' and 'end date' variable. A short description of each variable of interest is given below.

-   *Pollster*: The name of the polling organization conducting the poll (e.g., YouGov, RMG Research). This variable helps adjust for poll-specific biases.

-   *Numeric Grade*: A numeric rating given to the pollster, representing the quality or reliability of the organization (e.g., 3.0), with higher grades indicating more reliable pollsters.

-   *Pollscore*: A score that reflects the reliability of each pollster, capturing their historical accuracy and any systematic bias. Negative values indicate better predictive accuracy.

-   *State*: The U.S. state where the poll was conducted, allowing for analysis of regional differences in candidate support.

-   *National*: A binary variable indicating whether the poll is national (1 for national polls, 0 for state polls).

-   *End Date*: The date the poll ended, reflecting the currency of the data.

-   *Sample Size*: The total number of respondents in the poll.

-   *Candidate Name*: The name of the candidate being polled (e.g., Kamala Harris or Donald Trump), identifying the focus of each poll result.

-   *Pct (Percentage)*: The percentage of respondents in the poll who support the specified candidate.

-   *Recency*: This variable measures how recent each poll is. It was calculated by subtracting the end date of the poll from the end date of the most recent poll.

\newpage

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-sumstat
#| tbl-cap: "Summary Statistics of Numerical Variable"
analysis_data |> 
  select(numeric_grade, pollscore, sample_size, pct, recency) |> 
  datasummary_skim(histogram = FALSE)
```

\vspace{0.7cm}

From [@tbl-sumstat] we see that the average poll in our dataset has a numeric grade of 2.2, suggesting that most polls are of moderate to high quality in terms of reliability. The mean pollscore of -0.5 indicates a general trend toward negative values, which suggests these polls are relatively accurate, as more negative values imply reduced bias. Additionally, the large average sample size of 1908 respondents across polls provides a stable basis for our model, reducing variability and enhancing the reliability of predictions for candidate support.

\newpage

### Variation of Poll Quality and Support for Candiates by Pollster

\vspace{0.7cm}

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-topster
#| tbl-cap: "Top 5 most frequent pollsters, with count of polls, average pollscore (lower scores indicate less bias), and average numeric grade (higher values indicate greater reliability)."
# Get the top 10 pollsters by frequency along with their average pollscore and numeric grade
top_pollsters <- analysis_data %>%
  group_by(pollster) %>%           # Group by the 'pollster' variable
  summarize(
    Count = n(),                   # Count the number of occurrences
    Average_Pollscore = round(mean(pollscore, na.rm = TRUE), 2),  # Average pollscore
    Average_Numeric_Grade = round(mean(numeric_grade, na.rm = TRUE), 2),  # Average numeric grade
    .groups = 'drop'               # Drop grouping structure for cleaner output
  ) %>%
  arrange(desc(Count)) %>%         # Arrange in descending order of count
  slice_head(n = 5)                # Select the top 10 pollsters

# Rename the columns to remove underscores and capitalize 'Pollster'
colnames(top_pollsters) <- c("Pollster", "Count", "Average Pollscore", "Average Numeric Grade")

kable(top_pollsters, format = "markdown")

```

[@tbl-topster] displays the top 5 most frequent pollsters in the dataset, with each pollster’s total poll count, average pollscore (measuring reliability and historical accuracy), and average numeric grade (indicating overall quality). We observe that the most frequent pollster, Morning Consult, has a moderate pollscore and a slightly above-average numeric grade, indicating it provides fairly reliable data. However, there is variability in pollscore and numeric grade across the top pollsters, reflecting differences in quality and potential biases among them, which the model needs to adjust for to achieve accurate predictions.

```{r, fig.width = 12, fig.height = 6}
#| warning: false
#| message: false
#| echo: false
#| label: fig-ster
#| fig-cap: 'Support distribution for Kamala Harris and Donald Trump by major pollsters, highlighting variability in reported support across organizations and the value of aggregating multiple polls for balanced insights.'

# Filter the data for top 5 pollsters, if needed
top_pollsters <- analysis_data %>%
  count(pollster) %>%
  top_n(4, n) %>%
  pull(pollster)

filtered_data <- analysis_data %>%
  filter(pollster %in% top_pollsters)

# Plot the density plots for each pollster with overlaid densities for each candidate
ggplot(filtered_data, aes(x = pct, fill = candidate_name)) +
  geom_density(alpha = 0.5, color = "black") +
  scale_fill_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
  facet_wrap(~ pollster, ncol = 1) +
  labs(x = "Support (%)", y = "Density") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1),
        axis.text.y = element_text(size = 10),
        legend.title = element_blank(),
        legend.position = "right",
        panel.spacing = unit(0.5, "lines"))

```

[@fig-ster] shows the distribution of support for Kamala Harris and Donald Trump by pollster, highlighting variability across different polling organizations. Morning Consult, for example, demonstrates a wide range in reported support, with more spread in Trump’s support. In contrast, Siena/NYT shows less variation, with Harris consistently leading. This variability across pollsters emphasizes the importance of aggregating multiple polls to account for organization-specific biases and ensure a more balanced view of candidate support.

### Sample Size of Polls

\vspace{0.7cm}

```{r, fig.width = 12, fig.height = 6}
#| warning: false
#| message: false
#| echo: false
#| label: fig-ss
#| fig-cap: "Distribution of Sample Sizes Across Polls: Majority of polls have sample sizes under 5,000, with a few outliers at larger sizes."

# Plot histogram for all sample sizes
ggplot(analysis_data, aes(x = sample_size)) +
  geom_histogram(binwidth = 500, fill = "lightblue", color = "black", alpha = 0.7) +
  labs(
       x = "Sample Size",
       y = "Count") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),  # Remove grid lines
    plot.title = element_text(size = 14, hjust = 0.5),  # Center the title
    axis.title.y = element_text(margin = margin(r = 10))  # Increased gap for y-axis label
  )

```

[@fig-ss] reveals a clear right-skewed distribution. Most of the sample sizes are clustered between 0 and 3000, with a sharp peak around 1000-1500. This indicates that the majority of polls are conducted with smaller sample sizes. As sample size increases, the frequency significantly drops, with very few polls conducted with sample sizes larger than 5000, though there are a few outliers with sizes approaching 10,000 or more. This wide range in sample sizes might affect the precision of estimates in different polls.

### Distribution of Numeric Grade and Pollscore of Polls

```{r, fig.width = 12, fig.height = 6}
#| warning: false
#| message: false
#| echo: false
#| label: fig-reli
#| fig-cap: "Distribution of Numeric Grade and Pollscore among polling organizations, highlighting variability in pollster reliability and potential bias across polls."
# Histogram with density line for numeric_grade
density_plot <- ggplot(analysis_data, aes(x = numeric_grade)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.1, fill = "lightblue", color = "black", alpha = 0.6) +
  geom_density(color = "black", size = 0.7) +
  labs(x = "Numeric Grade", y = "Density") +
  theme_minimal()

# Histogram with density line for pollscore with more x-axis labels
bar_plot <- ggplot(analysis_data, aes(x = pollscore)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.2, fill = "lightblue", color = "black", alpha = 0.6) +
  geom_density(color = "black", size = 0.7) +
  labs(x = "Pollscore", y = "Density") +
  scale_x_continuous(breaks = seq(-1.5, 1.5, by = 0.5)) +  # Adds more frequent x-axis labels
  theme_minimal() +
  theme(legend.position = "none")

# Combine the plots side by side with a wider layout
combined_plot <- density_plot + bar_plot + plot_layout(ncol = 2, widths = c(1.2, 1.2))

# Display the combined plot
combined_plot

```

[@fig-reli] displays the distribution of Numeric Grade (left) and Pollscore (right) across polling organizations. The Numeric Grade distribution shows that most pollsters are rated between 1.5 and 3, with peaks around 2.0 and 2.5, indicating a concentration of pollsters with moderate to high reliability scores. In contrast, the Pollscore distribution, where lower values indicate higher reliability, shows a range primarily between -1.5 and 0, with a notable peak around -0.5. This suggests that while many polls demonstrate relatively low bias, there is still variability in reliability across organizations. The distinction between these two metrics emphasizes the need to consider both quality (Numeric Grade) and potential systematic bias (Pollscore) when weighting polls in the model.

\vspace{0.7cm}

### Distribution of Polls by Poll Type and Candidate

[@fig-pie] charts illustrate the distribution of polls between candidates and poll types. The left chart shows that the majority of polls are conducted at the state level, with a smaller portion being national. The right chart shows that polling is almost evenly split between Kamala Harris and Donald Trump. This distribution underscores the model’s balanced approach to capturing state-level nuances as well as broader national trends, providing a comprehensive view of candidate support across different contexts

```{r, fig.width = 12, fig.height = 4}
#| warning: false
#| message: false
#| echo: false
#| label: fig-pie
#| fig-cap: "Poll distribution by type (state vs. national) and candidate (Trump vs. Harris), showing a majority of state polls and near-equal coverage for each candidate."
# Prepare data for National vs. State pie chart
national_state_data <- analysis_data %>%
  count(national) %>%
  mutate(label = ifelse(national == 1, "National", "State"))

# Pie chart for National vs. State
national_state_pie <- ggplot(national_state_data, aes(x = "", y = n, fill = label)) +
  geom_col(color = "darkgrey") +
  coord_polar(theta = "y") +
  scale_fill_manual(values = c("State" = "darkgrey", "National" = "white")) +
  labs(fill = "Poll Type") +
  theme_void() +
  theme(legend.position = "bottom")

# Prepare data for Kamala vs. Trump pie chart
candidate_data <- analysis_data %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  count(candidate_name)

# Pie chart for Kamala vs. Trump
candidate_pie <- ggplot(candidate_data, aes(x = "", y = n, fill = candidate_name)) +
  geom_col(color = "white") +
  coord_polar(theta = "y") +
  scale_fill_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
  labs( fill = "Candidate") +
  theme_void() +
  theme(legend.position = "bottom")

# Combine the two pie charts side by side
combined_plot <- national_state_pie + candidate_pie + plot_layout(ncol = 2)

# Display the combined plot
combined_plot

```

\newpage

### Support Trend For Candidates

```{r, fig.width = 12, fig.height = 6}
#| warning: false
#| message: false
#| echo: false
#| label: fig-trend
#| fig-cap: "Figure showing support trends for Kamala Harris and Donald Trump over time. Trend lines indicate slight shifts in support as the election nears, with consistent polling frequency throughout the period."

# Sample data creation (Replace this with your actual data)
# Assuming your dataset `analysis_data` has columns: date, pct, and candidate_name

# Filtering for only the necessary candidates if required
filtered_data <- analysis_data %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump"))

# Plotting
ggplot(filtered_data, aes(x = as.Date(end_date), y = pct, color = candidate_name)) +
  geom_point(alpha = 0.3, size = 1.5) +
  geom_smooth(method = "loess", se = TRUE, span = 0.2) +
  scale_color_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
  labs(x = "Date", y = "Support (%)", title = "Support Trends for Kamala Harris and Donald Trump") +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    axis.text.x = element_text(angle = 0),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  )

```

[@fig-trend] shows the support trends for Kamala Harris and Donald Trump from August to October. Each point represents a poll result, color-coded by candidate, with the trend lines highlighting the overall changes in support over time. Kamala Harris's support remains relatively steady but shows slight fluctuations around mid-September, while Donald Trump's support appears to have a small upward trend toward October. The distribution of points is dense throughout, reflecting consistent polling activity, though some dates show more concentrated polling. This visualization indicates that while both candidates maintain stable support levels, minor shifts occur as the election approaches, underscoring the importance of tracking trends over time rather than relying on individual polls.

\newpage

### Relationship Between Sample Size and Suppoert for Candidates

```{r, fig.width = 12, fig.height = 6}
#| echo: false
#| warning: false
#| message: false
#| label: fig-pct
#| fig-cap: "Relationship between Sample Size and Support Percentage for Donald Trump and Kamala Harris, showing slight downward and upward trends in support with increasing sample size for Trump and Harris, respectively."

# Filter data for both candidates and cap sample size at 12,000
filtered_data <- analysis_data %>%
  filter(sample_size <= 12000, candidate_name %in% c("Kamala Harris", "Donald Trump"))

# Create the plot for both Kamala Harris and Donald Trump
ggplot(filtered_data, aes(x = sample_size, y = pct, color = candidate_name)) +
  geom_point(size = 0.5, alpha = 0.7) +  # Smaller dots with slight transparency
  geom_smooth(aes(group = candidate_name), method = "lm", se = FALSE, color = "darkgrey", size = 0.8) +  # LOESS smoothing line

  # Custom colors for candidate
  scale_color_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +

  # Facet wrap by candidate name to get two plots
  facet_wrap(~ candidate_name, scales = "free_y") +

  # Minimal theme adjustments
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),  # Remove major gridlines
    panel.grid.minor = element_blank(),  # Remove minor gridlines
    legend.position = "none",  # Remove legend as colors are self-explanatory
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.line = element_line(color = "black"),  # Add axis lines
    panel.border = element_blank(),  # Remove panel border
    axis.line.y.right = element_line(color = "black"),  # Add right spine
    axis.line.x.top = element_blank()  # Remove top spine
  ) +
  labs(x = "Sample Size", y = "Support (%)")  # Title and axis labels

```

[@fig-pct] shows the relationship between sample size and support percentage for Donald Trump and Kamala Harris. For both candidates, the majority of polls have a sample size below 2,500, but there are some larger polls exceeding 10,000 respondents. In Trump’s chart, there’s a slight downward trend, suggesting that larger sample sizes may show marginally lower support. In contrast, Harris's chart shows a slight upward trend with larger sample sizes, indicating a marginal increase in support with larger poll samples. This highlights the variability in polling support depending on the sample size, emphasiszing the importance of including sampel size in our model.

### Variability of Candidate support Across US States

```{r}
#| echo: false
#| warning: false
#| message: false
#| inlcude: false

# Create the dataset for mapping support
support_by_state <- analysis_data %>%
  # Filter for only Kamala Harris and Donald Trump
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  # Group by state and candidate
  group_by(state, candidate_name) %>%
  # Calculate the mean support percentage for each candidate in each state
  summarize(mean_support = mean(pct, na.rm = TRUE)) %>%
  # Spread the data so that each state has a column for each candidate's mean support
  pivot_wider(names_from = candidate_name, values_from = mean_support) %>%
  # Rename columns for clarity
  rename(kamala_mean_support = `Kamala Harris`, trump_mean_support = `Donald Trump`) %>%
  # Calculate total mean support for both candidates in each state
  mutate(total_support = kamala_mean_support + trump_mean_support,
         # Calculate proportion of support for each candidate out of the total
         kamala_proportion = kamala_mean_support / total_support,
         trump_proportion = trump_mean_support / total_support) %>%
# Select relevant columns
  select(state, kamala_proportion, trump_proportion)

```

```{r}
#| warning: false
#| message: false
#| echo: false
#| include: false
# Load the shapefile (adjust path as needed)

map_data <- us_states %>%
  left_join(support_by_state, by = c("NAME" = "state")) 
```

```{r, fig.width = 12, fig.height = 6}
#| echo: false
#| warning: false
#| message: false
#| label: fig-map1
#| fig-cap: "Map showing the proportion of support for Kamala Harris relative to Donald Trump across U.S. states. Blue indicates states where Harris has higher support, while red indicates states where Trump leads. Color intensity reflects the magnitude of support difference, with gray indicating insufficient polling data."

# Filter out Alaska, Hawaii, and Puerto Rico using subset
map_data_continental <- subset(map_data, !NAME %in% c("Alaska", "Hawaii", "Puerto Rico"))

ggplot(map_data_continental) +
  geom_sf(aes(fill = kamala_proportion * 100), color = "white", size = 0.2) +
  geom_text(aes(label = STUSPS, geometry = geometry), 
            stat = "sf_coordinates", color = "black", size = 3) +  # Adjust color and size as needed
  scale_fill_distiller(palette = "RdBu", direction = 1, na.value = "grey", 
                       name = "Harris's Mean Support in Polls (%)") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    aspect.ratio = 0.5  # Make the plot wider
  )

```

[@fig-map1] shows the average support for Kamala Harris as a proportion relative to both Kamala Harris and Donald Trump across the continental U.S. states. States shaded in blue indicate a higher proportion of support for Harris, while red shades represent stronger support for Trump, with color intensity reflecting the support margin. Gray-shaded states lack sufficient polling data.

In key swing states, such as Pennsylvania, Michigan, and Arizona, we observe a balanced distribution of support, suggesting close competition in these battleground areas. Meanwhile, traditional Democratic strongholds, like California and New York, show a clear preference for Harris, with deeper blue shades. Conversely, traditionally Republican states, such as Texas and Tennessee, exhibit higher support for Trump, with intense red shades highlighting his advantage. This visualization emphasizes the varied regional dynamics of candidate support, with distinct patterns in both competitive and historically partisan states.

# Forecasting Election Outcome through Pooling Polls {#sec-model}

## Forecasting Approach

The polls of polls methodology is widely used in election prediction as it aggregates multiple polls to provide a more reliable estimate of voter support, rather than relying on any single poll. The goal is to reduce errors and biases present in individual polls by using a weighted average of many different polls.

In our approach, we will employ linear modeling of voter support percentage (pct) on pollster and other independent variables such as sample size, poll recency, and poll scope (state vs. national). This will allow us to smooth out the inherent noise, biases, and variability across different pollsters. Once we obtain the predicted values from our model, we will weight these predictions based on the numeric grade (quality score) of each pollster to calculate an overall national estimate of the outcome. Additionally, we will separately compute estimates for key battleground states, where voter behavior can be more volatile and pivotal in deciding the final outcome of the election. This approach helps us capture both national trends and critical state-level dynamics.

## Model

In this section, we aim to address the inherent biases and differences present in various polling data to arrive at a robust prediction model. The core challenge lies in selecting a model with an optimal balance between complexity and fit, ensuring it accurately captures the dynamics of polling data while avoiding overfitting. To this end, we carefully evaluated different model specifications to determine the most appropriate one for our forecasting purpose.

Given that variables like numeric grade and pollscore are perfectly collinear with pollster, they were excluded from the regression analysis to avoid multicollinearity issues. These variables, however, remain integral to our weighting strategy, where they will be used to adjust for differences in polling accuracy and reliability. Instead, we focus on key features such as pollster, sample size, state, and recency, gradually adding complexity to the model.

By systematically comparing model specifications that incorporate these variables, we aim to select the model with the right balance between predictive accuracy and generalizability, ultimately providing the best possible forecast.

## Model Set Up

We aim to model the percentage of support for Kamala Harris and Donald Trump in each poll as a function of the pollster the sample size, the state, and the recency of the poll.

$$
y_i = \alpha + \beta_1 \cdot \mathrm{pollster}_i + \beta_2 \cdot \mathrm{sample\_size}_i + \beta_3 \cdot \mathrm{recency}_i +  \beta_4 \cdot \mathrm{state}_i + \epsilon_i
$$

Where

-   $y_i$​ is the percentage of support for candidate in poll iii,

-   $α$ is the intercept,

-   $β_1$ captures the effect of the polling organization,

-   $β_2​$ captures the effect of the sample size,

-   $β_3$​ captures the effect of recency (how recent the poll is),

-   $β_4$ capture the effects of the different states

-   ​ $\epsilon_i$ represents the error term, assumed to follow a normal distribution with mean 0.

## Model Justification

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-causal-model
#| fig-cap: "Directed Acyclic Graph illustrating factors influencing candidate support, with poll attributes (Pollster, Sample Size, State, Recency) as direct contributors and unobserved confounders representing potential biases in polling results."
# Load and display the saved PNG
knitr::include_graphics(here("other", "graphs", "polls_of_polls_dag.png"))

```

In our model, we aim to smooth out discrepancies and biases across various polling organizations using a polls-of-polls approach. Given the potential for individual pollsters to introduce systematic differences—due to variations in sampling methods, question phrasing, and historical leanings—our model includes a pollster variable to adjust for these organization-specific biases. This allows us to capture an aggregated view of public support that is less susceptible to the idiosyncrasies of any single poll. Furthermore, we incorporate sample size as a predictor, as polls with larger samples tend to yield more stable and reliable results, reducing random fluctuations caused by smaller samples. The state variable accounts for regional political differences, ensuring that the model captures varying levels of support across geographic and demographic groups, which is crucial for understanding the nuanced political landscape. Additionally, recency is included to prioritize more recent polls, as public opinion can shift rapidly in response to political events, and recent data is generally more reflective of current sentiment. By integrating these factors, our model seeks to produce a more stable and comprehensive measure of candidate support, reducing noise from individual poll discrepancies and focusing on a balanced, up-to-date aggregation of polling data.

We opted for a linear model due to its capacity to quantify the marginal effects of each predictor (pollster, sample size, state, and recency) on candidate support in a straightforward manner. This structure is well-suited to our polls-of-polls approach, as it allows for the estimation of fixed effects that can control for systematic biases across pollsters, while accommodating the influence of sample size and recency as continuous variables. All modeling was conducted using the base R package [@citeR], specifically utilizing the lm() function from the stats package for linear regression analysis.

## Model Results

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-summary
#| tbl-cap: "Model performance summary showing improved fit and accuracy as State and Recency are added, with R² increasing from 0.375 in Model 1 to 0.774 in Model 3, and RMSE decreasing from 3.053 to 1.836." 
# Load your models
model1 <- readRDS(here("models/model1.rds"))
model2 <- readRDS(here("models/model2.rds"))
model3 <- readRDS(here("models/model3.rds"))

# Extract summary statistics and add variables included in each model
model_summary <- tibble::tibble(
  Model = c("Model 1", "Model 2", "Model 3"),
  Variables = c("Pollster, Sample Size", "Pollster, Sample Size, State", "Pollster, Sample Size, State, Recency"),
  `R²` = c(summary(model1)$r.squared, summary(model2)$r.squared, summary(model3)$r.squared),
  `Adjusted R²` = c(summary(model1)$adj.r.squared, summary(model2)$adj.r.squared, summary(model3)$adj.r.squared),
  `AIC` = c(AIC(model1), AIC(model2), AIC(model3)),
  `BIC` = c(BIC(model1), BIC(model2), BIC(model3)),
  `RMSE` = c(sqrt(mean(residuals(model1)^2)), sqrt(mean(residuals(model2)^2)), sqrt(mean(residuals(model3)^2)))
)

# Display the table using kable
model_summary %>%
  kable(caption = "Model Summary with Included Variables", digits = 3) %>%
  kable_styling(full_width = FALSE)

```

[@tbl-summary] summarizes the performance metrics for three models with progressively added variables. Model 1, which includes only Pollster and Sample Size, achieves an R² of 0.375, indicating that these variables alone explain about 37.5% of the variance in candidate support. Model 2 incorporates State as an additional predictor, resulting in a substantial improvement, with an R² of 0.720 and a reduction in both AIC and RMSE, showing better model fit and predictive accuracy. Model 3 further adds Recency, which increases the R² to 0.774 and decreases the RMSE to 1.836, indicating enhanced explanatory power and prediction accuracy. This progression highlights the benefit of adding contextual variables like State and Recency to better capture the complexities of polling data.

## Prediction

To predict Kamala Harris' overall support, we used a weighted average approach based on the quality of each poll. The weights are calculated using each poll's `numeric_grade`, which reflects the reliability and transparency of the polling methodology.

We define the weight for each pollster $w_i$ as follows:

$$
w_i = \frac{\mathrm{numeric\_grade}_i \times (\mathrm{maxPollscore} - \mathrm{pollscore}_i)}{\sum_{i=1}^{n} \mathrm{numeric\_grade}_i \times (\mathrm{maxPollscore} - \mathrm{pollscore}_i)}
$$

where:

-   $w_i$​ represents the weight assigned to poll i,

-   $numericgrade_i$ is the numeric grade of poll i, and

-   $n$ is the total number of polls used in the analysis.

-   $pollscore_i$ is the is the pollscore of poll i which reflects the estimated bias of the poll (with more negative values indicating less bias)

-   $maxPollscore$ is the maximum pollscore across all polls.

The weight assigned to each poll combines both its quality (as represented by the numeric grade) and its level of bias (as indicated by the pollscore). Since a more negative pollscore reflects a lower level of bias, the formula uses the difference between the maximum pollscore and each poll’s specific pollscore. This approach gives more weight to polls that are both highly graded (indicating higher reliability and transparency) and less biased.By combining these two factors, the weighting system emphasizes polls that are reliable and minimally biased, ensuring that they have a stronger influence in the overall calculation. Additionally, the total of all weights is normalized to sum to one, so each poll’s weight is proportionate to its quality and relative lack of bias, resulting in a more balanced and accurate average of public support.

Using these weights, the overall weighted prediction of candidate's support is calculated by summing the weighted predicted values from our regression model:

$$
\text{Overall Weighted Support} = \sum_{i=1}^{n} w_i \cdot \hat{y}_i
$$

-   $\hat{y}_i$​ is the predicted percentage of support for Kamala Harris from poll i.

### Comparing Overall Weighted Support Across All Polls

Aggregating support across all polls and applying our weighted approach, we estimate the overall support for each candidate. The weights, calculated based on both the poll's numeric grade and poll score, ensure that higher-quality polls with less bias contribute more to our estimates. Based on this approach, Kamala Harris has an estimated overall weighted support of around 47.77%, while Donald Trump stands at around 46.31%. This aggregation provides a comprehensive view, taking into account the varied methodologies and sampling qualities of different polling organizations.

```{r}
#| warning: false
#| message: false
#| echo: false
#| include: true

model6 <- readRDS(here("models/model6.rds"))

# Step 1: Subset data for Kamala Harris, calculate weights, join with fitted values from model3, and calculate weighted_pct
kamala_data <- analysis_data %>%
  filter(candidate_name == "Kamala Harris") %>%
  mutate(
    weight = numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore) /
             sum(numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore), na.rm = TRUE),
    fitted_values = fitted(model3),  # Replace this with actual fitted values for Kamala
    weighted_pct = fitted_values * weight  # Calculate weighted percentage for Kamala
  )

# Calculate the overall weighted prediction for Kamala Harris
kamala_weighted_support <- sum(kamala_data$weighted_pct, na.rm = TRUE)

# Step 2: Subset data for Donald Trump, calculate weights, join with fitted values from model6, and calculate weighted_pct
trump_data <- analysis_data %>%
  filter(candidate_name == "Donald Trump") %>%
  mutate(
    weight = numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore) /
             sum(numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore), na.rm = TRUE),
    fitted_values = fitted(model6),  # Replace this with actual fitted values for Trump
    weighted_pct = fitted_values * weight  # Calculate weighted percentage for Trump
  )

# Calculate the overall weighted prediction for Donald Trump
trump_weighted_support <- sum(trump_data$weighted_pct, na.rm = TRUE)

```

### State-Level Predictions

In this section, we examine the predicted support for Kamala Harris and Donald Trump within each state, based on our weighted approach. By aggregating poll results within each state and applying weights that adjust for poll quality and bias, we can estimate candidate support with a more localized perspective. This allows us to identify state-by-state competition and highlight key battleground states where the margins are close.

For each state, we compare the predicted support percentages and determine the projected winner based on which candidate has higher weighted support. Additionally, we calculate the margin of difference between the candidates, which can provide insights into how close each state’s race is.

```{r}
#| warning: false
#| message: false
#| echo: false
#| include: false
# Define the max pollscore across the dataset
max_pollscore <- max(analysis_data$pollscore, na.rm = TRUE)

# Step 1: Subset data for Kamala Harris, join fitted values, then calculate weights within each state
kamala_data_a2 <- analysis_data %>%
  filter(candidate_name == "Kamala Harris") %>%
  mutate(fitted_values = fitted(model3)) %>%  # Add fitted values from model 3 for Kamala Harris
  group_by(state) %>%
  mutate(
    weight = numeric_grade * (max_pollscore - pollscore) /
             sum(numeric_grade * (max_pollscore - pollscore), na.rm = TRUE)
  ) %>%
  # Calculate the overall weighted prediction for each state
  summarize(support_kamala = sum(fitted_values * weight, na.rm = TRUE))

# Step 2: Repeat for Donald Trump with model6
trump_data_a2 <- analysis_data %>%
  filter(candidate_name == "Donald Trump") %>%
  mutate(fitted_values = fitted(model6)) %>%  # Add fitted values from model 6 for Donald Trump
  group_by(state) %>%
  mutate(
    weight = numeric_grade * (max_pollscore - pollscore) /
             sum(numeric_grade * (max_pollscore - pollscore), na.rm = TRUE)
  ) %>%
  # Calculate the overall weighted prediction for each state
  summarize(support_trump = sum(fitted_values * weight, na.rm = TRUE))

# Step 3: Join the two datasets by state
state_support_data <- left_join(kamala_data_a2, trump_data_a2, by = "state")

# Add a new column 'winner' based on higher support values
state_support_data <- state_support_data %>%
  mutate(
    winner = ifelse(support_kamala > support_trump, "Kamala Harris", "Donald Trump"),
    margin = abs(support_kamala - support_trump)
  )
```

```{r, fig.width = 12, fig.height = 6}
#| echo: false
#| warning: false
#| message: false
#| label: fig-map2
#| fig-cap: "Predicted 2024 U.S. Presidential Election winner by state, with Kamala Harris-leaning states in blue, Donald Trump-leaning states in red, and states lacking sufficient polling data in gray. The map highlights traditional Democratic and Republican strongholds as well as key battleground states."

# Filter out Alaska, Hawaii, and Puerto Rico from state_support_data
map_data_continental <- subset(us_states, !NAME %in% c("Alaska", "Hawaii", "Puerto Rico"))

# Merge with state_support_data to add winner info
map_data_continental <- map_data_continental %>%
  left_join(state_support_data,  by = c("NAME" = "state"))

# Define the colors for each category
color_scale <- c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred", "No Polling Data" = "grey")

# Plotting the map with ggplot2
ggplot(map_data_continental) +
  geom_sf(aes(fill = winner), color = "white", size = 0.2) +
  geom_text(aes(label = STUSPS, geometry = geometry), 
            stat = "sf_coordinates", color = "black", size = 3) +  # Adjust text size as needed
  scale_fill_manual(
    values = color_scale, 
    na.value = "grey",
    name = "2024 Election Winner",
    labels = c("Kamala Harris", "Donald Trump", "No Polling Data")
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    aspect.ratio = 0.5  # Adjust the aspect ratio to make the map wider
  )

```

[@fig-map2] presents the predicted winner of the 2024 U.S. Presidential Election by state based on aggregated polling data, with red representing states projected to favor Kamala Harris and blue indicating those favoring Donald Trump. Notably, traditional Democratic strongholds in the Northeast and West Coast, such as California, New York, and Washington, show solid support for Harris. Conversely, traditional Republican states, including Texas, Florida, and much of the South, show strong support for Trump.

Swing states, like Pennsylvania, Michigan, and Wisconsin, are predominantly leaning toward Harris, indicating a possible advantage for her in these key battlegrounds, although other typical swing states like Arizona and Georgia lean toward Trump. Gray-shaded states lack sufficient polling data to make a prediction, reflecting areas of data scarcity in the model's projections. This visualization highlights the geographical and political divides across the U.S., as well as the critical role swing states play in determining the election outcome.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-battleground-states
#| tbl-cap: "Predicted support for Kamala Harris and Donald Trump across key battleground states, indicating the winner and the support margin (%) in each state."

# Define battleground states based on provided list
battleground_states <- c("Pennsylvania", "Michigan", "North Carolina", "Wisconsin", "Nevada", "Arizona", "Georgia")

# Display the table of battleground states with adjusted column names
battleground_table <- state_support_data %>%
  filter(state %in% battleground_states) %>%
  select(
    State = state,
    `Kamala Support (%)` = support_kamala,
    `Trump Support (%)` = support_trump,
    `Predicted Winner` = winner,
    `Support Margin (%)` = margin
  )

# Display the table using kable
knitr::kable(battleground_table)

```

USAFacts[@usafacts],FiveThirtyEight[@FiveThirtyEight] and many other polling websites are calling Arizona, Georgia, Michigan, Pennsylvania, Wisconsin, North Carolina, Nevada the 'swing states' of the 2024 presidential election.Polling in swing states is crucial for election predictions because these states, with their historically close voting margins, often determine the overall outcome by tipping the electoral balance toward one candidate. [@tbl-battleground-states] summarizes the predicted support percentages for Kamala Harris and Donald Trump in key battleground states, along with the predicted winner and the support margin. The data shows close margins in these states, with some like Arizona and Georgia leaning slightly towards Trump, while others such as Michigan and Pennsylvania favor Harris by narrow margins. The support margin column highlights the competitiveness of these states, with all margins below 2.2%, indicating highly contested races.

# Discussion {#sec-discussion}

In this paper, we set out to predict the support for both Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election using a “polls-of-polls” approach. By aggregating multiple polls, we sought to mitigate the inherent biases present in individual surveys, creating a more accurate and balanced forecast across candidates. Our analysis was driven by a linear regression model using key predictors, such as pollster, sample size, state, and recency of the polls. After calculating predicted values from the model, we applied a weighting scheme based on each pollster’s numeric grade and pollscore, incorporating both reliability and bias. This weighted approach enabled us to produce an overall prediction of candidate support, reflecting variations across different states and polling organizations.

One key insight from our analysis is the importance of poll recency. Including recency notably increased the model’s explanatory power, with improvements in R² and reductions in RMSE compared to models without it. This underscores the dynamic nature of public opinion, where recent events can influence voter perceptions and candidate support.

Our state-level analysis highlights close competition observed in key battleground states, emphasizing their pivotal role in the 2024 election. Our model reveals that in states like Pennsylvania, Michigan, and Wisconsin, support levels for Harris and Trump are nearly tied, reflecting the intense contest for these critical electoral votes. The predictions show that even minor shifts in support within these swing states could decisively impact the overall election outcome. This analysis shows the importance of regional polling in capturing nuanced voter sentiment and the heightened influence that battleground states hold in shaping the final results.

However, there are limitations to our approach. Our model assumes linear relationships between predictors and candidate support, which might not capture more complex or interaction-based trends. Additionally, while we weighted polls based on their numeric grades, these scores may not fully reflect each pollster’s accuracy, leaving room for improvement. Future research could explore non-linear methods, such as machine learning, to capture potential interactions between variables. Refining the weighting mechanism by considering pollster performance history or state-specific polling nuances could further enhance predictive accuracy.

In conclusion, this paper contributes to election forecasting by providing a comprehensive, weighted estimate of candidate support across both national and state levels. While our predictions provide valuable insights into potential election outcomes, the evolving nature of voter sentiment and poll accuracy reflects the need for adaptive models that account for ongoing changes in public opinion.

\newpage

\appendix

# Appendix {.unnumbered}

# Emerson College Polling Methodology (October 14-16, 2024)

## Overview

Emerson College Polling, known for its extensive online polling methods, conducted a survey from October 14 to 16, 2024, targeting 1,000 likely voters. The poll measured voter preferences between Kamala Harris and Donald Trump, revealing a near tie with Harris at 50% and Trump at 49%.

## Population, Frame, and Sample

The target population is likely voters in the U.S. elections. These voters are determined by the voters' likeliness to vote in the upcoming elections and the voter's voting history. Both of these factors are self-reported.

The sampling frame is the voters that are on the Aristotle list and on the online panel provided by CINT. Aristotle is an online supplier of voter files and CINT is a software company that connects researchers, such as Emerson, to online panels. This approach helps ensure a representative sample by accurately reflecting the electorate.

The sample size is the 1,000 likely voters chosen arbitrarily from the sampling frame. Likely voters are determined by a combination of voter history, registration status, and demographic data. These factors are self-reported.

## Sample Recruitment and Approach

Emerson College used a mixed-mode sampling approach for its polling. Specifically, three primary modes were used to collect data for its October 14-16, 2024 poll:

-   MMS-to-web text surveys: Respondents were contacted via text messages sent to cell phones. These messages directed participants to complete the survey online. This method was conducted using voter lists provided by Aristotle.

-   Interactive Voice Response (IVR): Landlines were targeted using automated phone calls where respondents could answer questions using their keypads. The contact information for this was also sourced from Aristotle’s voter lists.

-   Online Panel from CINT: Emerson utilized a pre-screened, opt-in online panel of voters provided by CINT.

## Advantages and Trade-offs in Mixed-Mode Sampling

The multi-mode sampling approach used in this poll has several advantages:

-   Wider demographic reach: Having a combination of data collection methods allows Emerson to gather responses from a broader demographic group. In this polling, MMS-to-web surveys mostly gather responses from younger voters and those who primarily use mobile devices. IVR target older and rural voters, who are more likely to respond via landlines. Online panels is more efficient in reaching tech-savvy individuals.

-   Cost-efficient: Automated technologies reduce the cost compared to live interviews. In a polling context, this allows for a larger sample size.

However, the trade-offs include:

-   Coverage bias: IVR only reaches individuals with landlines, which may skew results towards older demographics.

-   Self-selection bias: Respondents in online panels tend to be more engaged, which can lead to over-representation of politically active individuals.

## Non-response Handling

Emerson addresses non-response using a weighting system. The data is adjusted based on key demographic variables like age, gender, race, education, and party affiliation, making the sample more representative of the electorate. Weights are applied to compensate for under- or over-represented groups, ensuring balanced results.

## Questionnaire Design

The Emerson survey features a straightforward, efficient design focused on key issues and voter preferences. Strengths of the questionnaire include:

-   Simplicity: Clear, concise questions makes the survey easy for respondents to understand and to provide accurate answers.

-   Topical relevance: The poll zeroes in on vote preference and demographic splits, delivering timely insights during election cycles.

However, weaknesses include:

-   Lack of nuance: While the survey captures basic preferences, it does not delve deeply into voters' motivations or the factors driving their choices.

-   Mode limitations: Different survey modes can affect how respondents engage with the questions. This can potentially influence the depth and accuracy of voter responses.

## Conclusion

Emerson College Polling's mixed-mode methodology is effective in balancing cost with broad demographic reach. Its weighting techniques contribute to reliable and accurate results despite non-responses. While the use of multiple survey modes increases population representation, possible biases—such as lack of nuance and mode effects need to be taken into consideration when interpreting results.

In sum, this method is a strong predictor of election outcomes and continues to adapt to the evolving challenges of modern polling.

# Idealized Methodology

## Overview

This methodology outlines an election forecasting plan with a budget of \$100,000. Heavily inspired by TIPP Insights and Emerson College's hybrid approaches, the survey methodology balances cost-efficiency and data accuracy by using different modes of data collection and wide demographic reach.

This forecasting plan will predict the popular vote and Electoral College outcomes, emphasizing key swing states and underrepresented voter groups.

## Sampling Approach

This idealized methodology uses a multi-mode hybrid sampling approach. This allows the use of both probability-based and non-probability-based methods:

-   Probability-based methods:

    -   Live phone interviews (landline and cell): As used by TIPP, live phone interviews is a direct and personal method of collecting responses. Voter lists would be sourced from Aristotle to ensure we contact likely voters. This mode reaches older voters and voters with less internet connection.

    -   Interactive Voice Response (IVR): Similar to Emerson’s approach, IVR for landline users balances costs and ensures demographic representation, particularly among older or less tech-savvy populations. This method captures quick responses and is less expensive than live interviews.

-   Non-probability methods:

    -   MMS-to-web (mobile): Following Emerson’s MMS-to-web strategy, we would send multimedia messages to likely voters’ cell phones, encouraging them to complete an online survey. This method works well for younger, mobile-first voters, helping fill in gaps where traditional methods might miss.

    -   Online panel recruitment via CINT: To mirror TIPP’s online panel approach, a portion of the survey would be conducted via an opt-in panel of pre-screened voters from CINT, capturing tech-savvy respondents. CINT’s platform ensures that different voter groups, including hard-to-reach groups, are represented in the polling.

## Recruitment Strategy

By using Aristotle’s voter lists and CINT's platform as a sampling frame, the sample consists of likely voters based on registration status and past voting behavior. Quota sampling would be used to match the demographic profile of the U.S. electorate. This ensures that respondents in the sample reflect key demographic variables such as age, gender, race, and education level.

We would oversample key swing states like Pennsylvania, Wisconsin, and Florida. Key swing states are states where there is no overwhelming support for a particular candidate or political party. The outcome in these states often determine the overall result of the election, making them critical in securing a win in the Electoral College. Oversampling these states allows a more accurate forecasting in these battlegrounds, which are critical for predicting the Electoral College outcome.

## Data Validation

Given that the data is collected in a hybrid format, we would use duplicate detection methods via IP addresses or phone numbers to eliminate multiple responses from the same individual. Also, we would implement attention checks to identify respondents who are not fully engaged. Both duplicate respondents and unengaged respondents bias results, and should be removed from the data set.

Following TIPP’s iterative weighting process, the data will be weighted to ensure that the sample aligns with key demographics (race, gender, education) based on census data.

## Conclusion

By combining TIPP’s live phone/IVR hybrid and Emerson College’s mixed-mode survey methods, this idealized methodology captures diverse voter demographics. It uses both probability-based and non-probability-based sampling methods to forecast the 2024 U.S. Presidential Election.

A link to the survey can be found at: https://forms.gle/haGLaQBPQhrXE6Xa7

# Copy of the Survey

\includepdf[pages=-]{../other/surveyfiles/SurveyCopy.pdf}

# Model Diagnostics

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-dgplotone
#| 
knitr::include_graphics(here("other", "graphs", "diagnostic_plot_model3.png"))
```

The diagnostic plots for Model 3ni sgiven in [@fig-dgplotone]. We notice the following:

1.  **Residuals vs. Fitted Plot**: This plot checks for non-linearity and heteroscedasticity. The residuals are scattered around the horizontal axis with no clear pattern, indicating that the linearity assumption is reasonable. However, there is slight spread around the center, suggesting some mild heteroscedasticity.

2.  **Normal Q-Q Plot**: This plot assesses the normality of residuals. Most points align along the diagonal line, although there are some deviations at the tails. This suggests that the residuals are approximately normally distributed, with minor deviations in the extremes.

3.  **Scale-Location Plot**: This plot further checks for homoscedasticity. The residuals appear to be evenly spread across the fitted values, supporting the homoscedasticity assumption, although slight deviations are present in certain regions.

4.  **Residuals vs. Leverage Plot**: This plot identifies potential influential points. While most points have low leverage, a few points exhibit higher leverage, as indicated by their distance from the center. However, no points exceed Cook’s distance threshold, indicating no extreme outliers that would unduly influence the model.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-dgplottwo
#| 
knitr::include_graphics(here("other", "graphs", "diagnostic_plot_model6.png"))
```

The diagnostic plots for Model 6 is given in [@fig-dgplottwo]. We notice the following:

1. **Residuals vs Fitted**: This plot suggests a fairly random spread of residuals around zero, indicating that the linearity assumption holds reasonably well. However, there is slight clustering of points in the center, which could indicate minor heteroscedasticity but is not severe.

2. **Normal Q-Q Plot**: The Q-Q plot shows that the residuals generally follow a normal distribution, with only a few deviations at the tails. This suggests that the normality assumption is mostly met, though some minor deviations in the upper tail indicate possible outliers.

3. **Scale-Location (Spread-Location) Plot**: The residuals appear randomly dispersed with no clear pattern, supporting the homoscedasticity (constant variance) assumption. However, some observations near the upper range might indicate slight variance inconsistency, though it is minimal.

4. **Residuals vs Leverage**: This plot does not show any influential outliers with high leverage that might impact the model unduly. The Cook's distance lines show that no data points exceed these thresholds, indicating that no individual observation is disproportionately influencing the model.

Overall, the diagnostic plots for both model suggests that they meet the assumptions for linear regression fairly well, with minor deviations that are not expected to severely impact the model’s validity.

# Code Styling

The code in this paper was reviewed and formatted for consistency using lintr [@citelintr] and styler [@citestyler], ensuring readability and adherence to style standards.

# Reproducibility

To replicate the findings presented in this paper, users should execute the scripts available in the GitHub repository. Begin by running the 00-install_packages.R script, which installs all required packages for the analysis.

# Acknowledgments

We extend our gratitude to [@citeTS], which provided invaluable guidance in establishing a reproducible workflow and inspired many of the code structures used in this paper.

\newpage

# References

### 2024-10-31 paper/paper.qmd
Fixed the path to shapefile data.

@@
 ---
-title: Forecasting 2024 U.S. Election using Polls-of-Polls
-subtitle: Using a Linear Regression Model to Adjust Pollster Biases and Predict Kamala Harris Vote Percentage at 47.78%
+title: "Forecasting the 2024 U.S. Presidential Election: An Analysis for Battleground States"
+subtitle: "Trump Favored Nationally, While Harris Leads in 4 of 7 Battleground States"
 author: 
@@
 date-format: long
-abstract: "This paper uses a polls of polls approach to predict voter support for Kamala Harris in the 2024 U.S.
-election, aggregating multiple polls to reduce individual bias and improve accuracy.
-Through a regression model incorporating factors such as pollster, sample size, state, and recency, we predict her overall support to be 47.78%, just below the 50% threshold.
-Our analysis highlights the importance of poll quality and recency in understanding trends in public opinion.
-These findings are significant as they provide a more reliable estimate of electoral outcomes, helping to inform political strategies and public discourse."
+abstract: >
+  This study employs a polls-of-polls approach to predict support for Kamala Harris and Donald Trump in the 2024 U.S.
+Presidential Election, with a focus on key battleground states to forecast the likely winner of the electoral college.
+By aggregating multiple polls and applying a weighted linear regression model with predictors like pollster reliability, sample size, state, and recency, we estimate a higher national support for Trump relative to Harris.
+Our analysis also reveals a close competition in battleground states, with Trump holding a narrow lead in Arizona, Georgia, and North Carolina, while Harris leads in Michigan, Nevada, Pennsylvania, and Wisconsin.
+These findings underscore the value of aggregating polls over relying on individual surveys, offering a more comprehensive and robust forecast of electoral outcomes.
 format: 
   pdf:
-    documentclass: article
     number-sections: true
@@
       - \usepackage{pdfpages}
+      - '\KOMAoption{captions}{tableheading}'
+      - \usepackage[margin=1in]{geometry}
+
 bibliography: references.bib
+
 ---
@@
 library(sf)
+library(patchwork)
+library(ggfortify)
 
@@
 analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
-mapping_data <- st_read(here("data/03-mapping_data/US"))
-mean_support_data <- read_parquet(here("data/02-analysis_data/map_data.parquet"))
+us_states <- st_read(here("data/03-mapping_data/US"))
 ```
+
+\vspace{0.7cm}
 
@@
 While individual polls provide snapshots of public opinion, they are often subject to biases and methodological differences.
-To address these limitations, this paper uses a "polls of polls" approach to predict support for Kamala Harris, aggregating multiple polls to smooth out inconsistencies.
-By combining data from various polling organizations, our aim is to create a more accurate forecast of her voter support leading up to the 2024 U.S.
-election.
+In this paper, we aim to forecast voter support for Kamala Harris and Donald Trump by aggregating data from multiple polling sources, reducing individual poll biases, and improving overall prediction accuracy.
+Our analysis focuses on estimating the levels of support for each candidate not only nationally but also in key swing states that are likely to be decisive in determining the electoral college outcome.
 
-We employ a linear regression model to estimate Kamala Harris' support percentage, incorporating key variables such as pollster, sample size, state, and recency.
-To account for the varying reliability of different polls, we weight predictions using each pollster’s numeric grade.
-Our model also explores the role of recency, with recent polls showing a decline in support for Harris.
-The final weighted prediction for her support stands at 47.78%, indicating that her overall backing falls just below the 50% threshold, suggesting potential challenges in gaining majority support.
+The estimand of this study is the level of voter support for each candidate, Kamala Harris and Donald Trump, as reported across multiple polls.To estimate this, we developed a regression model incorporating variables such as pollster, sample size, state, and recency of the poll, with an emphasis on accurately capturing state-level dynamics.
+Our findings indicate that, on a national level, support between Kamala Harris and Donald Trump is closely balanced.
+Harris’s estimated national support stands close to 48%, while Trump’s support is slightly higher, around 49%.
+We found that Trump holds a marginal lead in battleground states like Arizona and Georgia, with a support margin slightly above 1%.
+In contrast, Harris shows narrow leads in Michigan, Nevada, Pennsylvania, and Wisconsin, with her support margin in Wisconsin reaching over 2%, making it her strongest battleground.
+North Carolina emerges as one of the tightest races, with Trump leading by only 0.26%, underscoring the competitive nature of these regions.
 
 This paper contributes to the field of election forecasting by emphasizing the importance of poll quality and recency in prediction models.
-The structure of the paper proceeds as follows: we first introduce the dataset and methodology, followed by model selection and results.
-We then discuss the implications of our findings, and conclude with limitations and potential directions for future research.
+By focusing on battleground states where the support margins are particularly close, our approach highlights the areas where campaign strategies and voter turnout efforts may have the greatest impact.
+This could provides insights for political strategists, media analysts, and the general public by offering a nuanced view of the electoral landscape.
 
-# Data
-
+This paper is organized as follows: In {#sec-data}, we clean the data, explore summary statistics, plot distributions of key variables, and examine relationships between variables.
+In {#sec-model}, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions.
+Finally, in {#sec-discussion}, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work.
+\vspace{0.7cm}
+# Data {#sec-data}
+\vspace{0.7cm}
 ## Measurement
@@
 
+\vspace{0.7cm}
+
 ## Data Cleaning
@@
 This cutoff was selected to retain mid to high-level pollsters for more reliable results.
-These steps were performed using functions from the `dpylr` package [@dplyr].
+These steps were performed using functions from the `Tidyverse` package [@tidy].
 Furthermore, dates were standardized and converted into a proper format for analysis using the `lubridate` package [@lubridate].
@@
 
-### Summary Statistics of Variables of Interest
+## Explorations of Variables of Interest
 
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-#| echo: false
+### Summary Statistics of Key Variables
 
-library(dplyr)
-library(modelsummary)
-library(tools)  
-library(kableExtra)
+\vspace{0.7cm}
 
-# Set options to use kableExtra for table formatting
-options(modelsummary_factory_default = 'kableExtra')
+For the purposes of analysis in this paper, only a subset of variables was selected.
+Two additional variable, national and state, was created using the existing 'state' and 'end date' variable.
+A short description of each variable of interest is given below.
 
-# Create the summary statistics table for the specified numerical variables
-summary_table <- analysis_data %>%
-  select(numeric_grade, pollscore, sample_size, num_support) %>%
-  # Remove underscores and capitalize the first letter of each word
-  rename_with(~ toTitleCase(gsub("_", " ", .))) %>%
-  datasummary_skim(
-    fun_numeric = list(
-      Mean = mean,
-      SD = sd,
-      Min. = min,
-      Median = median,
-      Max. = max
-    )
-  )
+-   *Pollster*: The name of the polling organization conducting the poll (e.g., YouGov, RMG Research). This variable helps adjust for poll-specific biases.
 
-print(summary_table)
+-   *Numeric Grade*: A numeric rating given to the pollster, representing the quality or reliability of the organization (e.g., 3.0), with higher grades indicating more reliable pollsters.
 
-#(summary_table)
+-   *Pollscore*: A score that reflects the reliability of each pollster, capturing their historical accuracy and any systematic bias. Negative values indicate better predictive accuracy.
 
-#kable(summary_table, caption = "Summary of Variables of Interest") %>% kable_styling(full_width = FALSE)
+-   *State*: The U.S. state where the poll was conducted, allowing for analysis of regional differences in candidate support.
 
-```
+-   *National*: A binary variable indicating whether the poll is national (1 for national polls, 0 for state polls).
 
-The summary statistics table presents key descriptive measures for the numeric variables of interest.
-The Numeric Grade variable has a mean of 2.3, with values ranging from a minimum of 1.1 to a maximum of 3.0.
-It has a standard deviation of 0.6, indicating relatively low variability.
-The Pollscore variable averages -0.5, with a standard deviation of 0.6 and a range between -1.5 and 1.7.
-The Sample Size has a wider range, with an average of 1962.5 and a standard deviation of 2608.0, indicating significant variation across polls, with values spanning from 382 to 12084.
-Lastly, Num Support, calculated as the number of respondents supporting Kamala Harris, shows a mean of 949.2, with values ranging from 172 to 6042 and a standard deviation of 1290.0, highlighting substantial differences across the polls.
+-   *End Date*: The date the poll ended, reflecting the currency of the data.
 
-### Distribution of Sample Size of Polls
+-   *Sample Size*: The total number of respondents in the poll.
+
+-   *Candidate Name*: The name of the candidate being polled (e.g., Kamala Harris or Donald Trump), identifying the focus of each poll result.
+
+-   *Pct (Percentage)*: The percentage of respondents in the poll who support the specified candidate.
+
+-   *Recency*: This variable measures how recent each poll is. It was calculated by subtracting the end date of the poll from the end date of the most recent poll.
+
+\newpage
 
@@
 #| echo: false
+#| label: tbl-sumstat
+#| tbl-cap: "Summary Statistics of Numerical Variable"
+analysis_data |> 
+  select(numeric_grade, pollscore, sample_size, pct, recency) |> 
+  datasummary_skim(histogram = FALSE)
+```
 
-library(ggplot2)
+\vspace{0.7cm}
+
+From [@tbl-sumstat] we see that the average poll in our dataset has a numeric grade of 2.2, suggesting that most polls are of moderate to high quality in terms of reliability.
+The mean pollscore of -0.5 indicates a general trend toward negative values, which suggests these polls are relatively accurate, as more negative values imply reduced bias.
+Additionally, the large average sample size of 1908 respondents across polls provides a stable basis for our model, reducing variability and enhancing the reliability of predictions for candidate support.
+
+\newpage
+
+### Variation of Poll Quality and Support for Candiates by Pollster
+
+\vspace{0.7cm}
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: tbl-topster
+#| tbl-cap: "Top 5 most frequent pollsters, with count of polls, average pollscore (lower scores indicate less bias), and average numeric grade (higher values indicate greater reliability)."
+# Get the top 10 pollsters by frequency along with their average pollscore and numeric grade
+top_pollsters <- analysis_data %>%
+  group_by(pollster) %>%           # Group by the 'pollster' variable
+  summarize(
+    Count = n(),                   # Count the number of occurrences
+    Average_Pollscore = round(mean(pollscore, na.rm = TRUE), 2),  # Average pollscore
+    Average_Numeric_Grade = round(mean(numeric_grade, na.rm = TRUE), 2),  # Average numeric grade
+    .groups = 'drop'               # Drop grouping structure for cleaner output
+  ) %>%
+  arrange(desc(Count)) %>%         # Arrange in descending order of count
+  slice_head(n = 5)                # Select the top 10 pollsters
+
+# Rename the columns to remove underscores and capitalize 'Pollster'
+colnames(top_pollsters) <- c("Pollster", "Count", "Average Pollscore", "Average Numeric Grade")
+
+kable(top_pollsters, format = "markdown")
+
+```
+
+[@tbl-topster] displays the top 5 most frequent pollsters in the dataset, with each pollster’s total poll count, average pollscore (measuring reliability and historical accuracy), and average numeric grade (indicating overall quality).
+We observe that the most frequent pollster, Morning Consult, has a moderate pollscore and a slightly above-average numeric grade, indicating it provides fairly reliable data.
+However, there is variability in pollscore and numeric grade across the top pollsters, reflecting differences in quality and potential biases among them, which the model needs to adjust for to achieve accurate predictions.
+
+```{r, fig.width = 12, fig.height = 6}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-ster
+#| fig-cap: 'Support distribution for Kamala Harris and Donald Trump by major pollsters, highlighting variability in reported support across organizations and the value of aggregating multiple polls for balanced insights.'
+
+# Filter the data for top 5 pollsters, if needed
+top_pollsters <- analysis_data %>%
+  count(pollster) %>%
+  top_n(4, n) %>%
+  pull(pollster)
+
+filtered_data <- analysis_data %>%
+  filter(pollster %in% top_pollsters)
+
+# Plot the density plots for each pollster with overlaid densities for each candidate
+ggplot(filtered_data, aes(x = pct, fill = candidate_name)) +
+  geom_density(alpha = 0.5, color = "black") +
+  scale_fill_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
+  facet_wrap(~ pollster, ncol = 1) +
+  labs(x = "Support (%)", y = "Density") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 0, hjust = 1),
+        axis.text.y = element_text(size = 10),
+        legend.title = element_blank(),
+        legend.position = "right",
+        panel.spacing = unit(0.5, "lines"))
+
+```
+
+[@fig-ster] shows the distribution of support for Kamala Harris and Donald Trump by pollster, highlighting variability across different polling organizations.
+Morning Consult, for example, demonstrates a wide range in reported support, with more spread in Trump’s support.
+In contrast, Siena/NYT shows less variation, with Harris consistently leading.
+This variability across pollsters emphasizes the importance of aggregating multiple polls to account for organization-specific biases and ensure a more balanced view of candidate support.
+
+### Sample Size of Polls
+
+\vspace{0.7cm}
+
+```{r, fig.width = 12, fig.height = 6}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-ss
+#| fig-cap: "Distribution of Sample Sizes Across Polls: Majority of polls have sample sizes under 5,000, with a few outliers at larger sizes."
 
@@
   geom_histogram(binwidth = 500, fill = "lightblue", color = "black", alpha = 0.7) +
-  labs(title = "Distribution of Sample Sizes",
+  labs(
        x = "Sample Size",
@@
 
-The histogram of sample sizes reveals a clear right-skewed distribution.
+[@fig-ss] reveals a clear right-skewed distribution.
 Most of the sample sizes are clustered between 0 and 3000, with a sharp peak around 1000-1500.
@@
 
-### Most Frequent Pollsters
+### Distribution of Numeric Grade and Pollscore of Polls
+
+```{r, fig.width = 12, fig.height = 6}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-reli
+#| fig-cap: "Distribution of Numeric Grade and Pollscore among polling organizations, highlighting variability in pollster reliability and potential bias across polls."
+# Histogram with density line for numeric_grade
+density_plot <- ggplot(analysis_data, aes(x = numeric_grade)) +
+  geom_histogram(aes(y = after_stat(density)), binwidth = 0.1, fill = "lightblue", color = "black", alpha = 0.6) +
+  geom_density(color = "black", size = 0.7) +
+  labs(x = "Numeric Grade", y = "Density") +
+  theme_minimal()
+
+# Histogram with density line for pollscore with more x-axis labels
+bar_plot <- ggplot(analysis_data, aes(x = pollscore)) +
+  geom_histogram(aes(y = after_stat(density)), binwidth = 0.2, fill = "lightblue", color = "black", alpha = 0.6) +
+  geom_density(color = "black", size = 0.7) +
+  labs(x = "Pollscore", y = "Density") +
+  scale_x_continuous(breaks = seq(-1.5, 1.5, by = 0.5)) +  # Adds more frequent x-axis labels
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Combine the plots side by side with a wider layout
+combined_plot <- density_plot + bar_plot + plot_layout(ncol = 2, widths = c(1.2, 1.2))
+
+# Display the combined plot
+combined_plot
+
+```
+
+[@fig-reli] displays the distribution of Numeric Grade (left) and Pollscore (right) across polling organizations.
+The Numeric Grade distribution shows that most pollsters are rated between 1.5 and 3, with peaks around 2.0 and 2.5, indicating a concentration of pollsters with moderate to high reliability scores.
+In contrast, the Pollscore distribution, where lower values indicate higher reliability, shows a range primarily between -1.5 and 0, with a notable peak around -0.5.
+This suggests that while many polls demonstrate relatively low bias, there is still variability in reliability across organizations.
+The distinction between these two metrics emphasizes the need to consider both quality (Numeric Grade) and potential systematic bias (Pollscore) when weighting polls in the model.
+
+\vspace{0.7cm}
+
+### Distribution of Polls by Poll Type and Candidate
+
+[@fig-pie] charts illustrate the distribution of polls between candidates and poll types.
+The left chart shows that the majority of polls are conducted at the state level, with a smaller portion being national.
+The right chart shows that polling is almost evenly split between Kamala Harris and Donald Trump.
+This distribution underscores the model’s balanced approach to capturing state-level nuances as well as broader national trends, providing a comprehensive view of candidate support across different contexts
+
+```{r, fig.width = 12, fig.height = 4}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-pie
+#| fig-cap: "Poll distribution by type (state vs. national) and candidate (Trump vs. Harris), showing a majority of state polls and near-equal coverage for each candidate."
+# Prepare data for National vs. State pie chart
+national_state_data <- analysis_data %>%
+  count(national) %>%
+  mutate(label = ifelse(national == 1, "National", "State"))
+
+# Pie chart for National vs. State
+national_state_pie <- ggplot(national_state_data, aes(x = "", y = n, fill = label)) +
+  geom_col(color = "darkgrey") +
+  coord_polar(theta = "y") +
+  scale_fill_manual(values = c("State" = "darkgrey", "National" = "white")) +
+  labs(fill = "Poll Type") +
+  theme_void() +
+  theme(legend.position = "bottom")
+
+# Prepare data for Kamala vs. Trump pie chart
+candidate_data <- analysis_data %>%
+  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
+  count(candidate_name)
+
+# Pie chart for Kamala vs. Trump
+candidate_pie <- ggplot(candidate_data, aes(x = "", y = n, fill = candidate_name)) +
+  geom_col(color = "white") +
+  coord_polar(theta = "y") +
+  scale_fill_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
+  labs( fill = "Candidate") +
+  theme_void() +
+  theme(legend.position = "bottom")
+
+# Combine the two pie charts side by side
+combined_plot <- national_state_pie + candidate_pie + plot_layout(ncol = 2)
+
+# Display the combined plot
+combined_plot
+
+```
+
+\newpage
+
+### Support Trend For Candidates
+
+```{r, fig.width = 12, fig.height = 6}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-trend
+#| fig-cap: "Figure showing support trends for Kamala Harris and Donald Trump over time.
+Trend lines indicate slight shifts in support as the election nears, with consistent polling frequency throughout the period."
+
+# Sample data creation (Replace this with your actual data)
+# Assuming your dataset `analysis_data` has columns: date, pct, and candidate_name
+
+# Filtering for only the necessary candidates if required
+filtered_data <- analysis_data %>%
+  filter(candidate_name %in% c("Kamala Harris", "Donald Trump"))
+
+# Plotting
+ggplot(filtered_data, aes(x = as.Date(end_date), y = pct, color = candidate_name)) +
+  geom_point(alpha = 0.3, size = 1.5) +
+  geom_smooth(method = "loess", se = TRUE, span = 0.2) +
+  scale_color_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
+  labs(x = "Date", y = "Support (%)", title = "Support Trends for Kamala Harris and Donald Trump") +
+  theme_minimal() +
+  theme(
+    legend.position = "right",
+    legend.title = element_blank(),
+    axis.text.x = element_text(angle = 0),
+    axis.title.x = element_text(margin = margin(t = 10)),
+    axis.title.y = element_text(margin = margin(r = 10))
+  )
+
+```
+
+[@fig-trend] shows the support trends for Kamala Harris and Donald Trump from August to October.
+Each point represents a poll result, color-coded by candidate, with the trend lines highlighting the overall changes in support over time.
+Kamala Harris's support remains relatively steady but shows slight fluctuations around mid-September, while Donald Trump's support appears to have a small upward trend toward October.
+The distribution of points is dense throughout, reflecting consistent polling activity, though some dates show more concentrated polling.
+This visualization indicates that while both candidates maintain stable support levels, minor shifts occur as the election approaches, underscoring the importance of tracking trends over time rather than relying on individual polls.
+
+\newpage
+
+### Relationship Between Sample Size and Suppoert for Candidates
+
+```{r, fig.width = 12, fig.height = 6}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-pct
+#| fig-cap: "Relationship between Sample Size and Support Percentage for Donald Trump and Kamala Harris, showing slight downward and upward trends in support with increasing sample size for Trump and Harris, respectively."
+
+# Filter data for both candidates and cap sample size at 12,000
+filtered_data <- analysis_data %>%
+  filter(sample_size <= 12000, candidate_name %in% c("Kamala Harris", "Donald Trump"))
+
+# Create the plot for both Kamala Harris and Donald Trump
+ggplot(filtered_data, aes(x = sample_size, y = pct, color = candidate_name)) +
+  geom_point(size = 0.5, alpha = 0.7) +  # Smaller dots with slight transparency
+  geom_smooth(aes(group = candidate_name), method = "lm", se = FALSE, color = "darkgrey", size = 0.8) +  # LOESS smoothing line
+
+  # Custom colors for candidate
+  scale_color_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
+
+  # Facet wrap by candidate name to get two plots
+  facet_wrap(~ candidate_name, scales = "free_y") +
+
+  # Minimal theme adjustments
+  theme_minimal() +
+  theme(
+    panel.grid.major = element_blank(),  # Remove major gridlines
+    panel.grid.minor = element_blank(),  # Remove minor gridlines
+    legend.position = "none",  # Remove legend as colors are self-explanatory
+    plot.title = element_text(hjust = 0.5),  # Center the title
+    axis.line = element_line(color = "black"),  # Add axis lines
+    panel.border = element_blank(),  # Remove panel border
+    axis.line.y.right = element_line(color = "black"),  # Add right spine
+    axis.line.x.top = element_blank()  # Remove top spine
+  ) +
+  labs(x = "Sample Size", y = "Support (%)")  # Title and axis labels
+
+```
+
+[@fig-pct] shows the relationship between sample size and support percentage for Donald Trump and Kamala Harris.
+For both candidates, the majority of polls have a sample size below 2,500, but there are some larger polls exceeding 10,000 respondents.
+In Trump’s chart, there’s a slight downward trend, suggesting that larger sample sizes may show marginally lower support.
+In contrast, Harris's chart shows a slight upward trend with larger sample sizes, indicating a marginal increase in support with larger poll samples.
+This highlights the variability in polling support depending on the sample size, emphasiszing the importance of including sampel size in our model.
+
+### Variability of Candidate support Across US States
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| inlcude: false
+
+# Create the dataset for mapping support
+support_by_state <- analysis_data %>%
+  # Filter for only Kamala Harris and Donald Trump
+  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
+  # Group by state and candidate
+  group_by(state, candidate_name) %>%
+  # Calculate the mean support percentage for each candidate in each state
+  summarize(mean_support = mean(pct, na.rm = TRUE)) %>%
+  # Spread the data so that each state has a column for each candidate's mean support
+  pivot_wider(names_from = candidate_name, values_from = mean_support) %>%
+  # Rename columns for clarity
+  rename(kamala_mean_support = `Kamala Harris`, trump_mean_support = `Donald Trump`) %>%
+  # Calculate total mean support for both candidates in each state
+  mutate(total_support = kamala_mean_support + trump_mean_support,
+         # Calculate proportion of support for each candidate out of the total
+         kamala_proportion = kamala_mean_support / total_support,
+         trump_proportion = trump_mean_support / total_support) %>%
+# Select relevant columns
+  select(state, kamala_proportion, trump_proportion)
+
+```
 
@@
 #| echo: false
-# Get the top 10 pollsters by frequency along with their average pollscore and numeric grade
-top_pollsters <- analysis_data %>%
-  group_by(pollster) %>%           # Group by the 'pollster' variable
-  summarize(
-    Count = n(),                   # Count the number of occurrences
-    Average_Pollscore = round(mean(pollscore, na.rm = TRUE), 2),  # Average pollscore
-    Average_Numeric_Grade = round(mean(numeric_grade, na.rm = TRUE), 2),  # Average numeric grade
-    .groups = 'drop'               # Drop grouping structure for cleaner output
-  ) %>%
-  arrange(desc(Count)) %>%         # Arrange in descending order of count
-  slice_head(n = 10)                # Select the top 10 pollsters
+#| include: false
 
-# Rename the columns to remove underscores and capitalize 'Pollster'
-colnames(top_pollsters) <- c("Pollster", "Count", "Average Pollscore", "Average Numeric Grade")
+map_data <- us_states %>%
+  left_join(support_by_state, by = c("NAME" = "state")) 
+```
 
-kable(top_pollsters, format = "markdown", caption = "Top 10 Pollsters by Frequency")
+```{r, fig.width = 12, fig.height = 6}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-map1
+#| fig-cap: "Map showing the proportion of support for Kamala Harris relative to Donald Trump across U.S.
+states.
+Blue indicates states where Harris has higher support, while red indicates states where Trump leads.
+Color intensity reflects the magnitude of support difference, with gray indicating insufficient polling data."
+
+# Filter out Alaska, Hawaii, and Puerto Rico using subset
+map_data_continental <- subset(map_data, !NAME %in% c("Alaska", "Hawaii", "Puerto Rico"))
+
+ggplot(map_data_continental) +
+  geom_sf(aes(fill = kamala_proportion * 100), color = "white", size = 0.2) +
+  geom_text(aes(label = STUSPS, geometry = geometry), 
+            stat = "sf_coordinates", color = "black", size = 3) +  # Adjust color and size as needed
+  scale_fill_distiller(palette = "RdBu", direction = 1, na.value = "grey", 
+                       name = "Harris's Mean Support in Polls (%)") +
+  theme_minimal() +
+  theme(
+    panel.grid = element_blank(),
+    axis.text = element_blank(),
+    axis.title = element_blank(),
+    axis.ticks = element_blank(),
+    aspect.ratio = 0.5  # Make the plot wider
+  )
 
@@
 
-This table lists the top 10 pollsters based on the frequency of polls they conducted.
-Morning Consult leads with 235 polls, followed by Siena/NYT and Redfield & Wilton Strategies with 94 and 88 polls, respectively.
-The table also shows each pollster's average poll score and average numeric grade.
-Most pollsters have negative poll scores, indicating a lower perceived reliability, except for Redfield & Wilton Strategies.
-However, the numeric grades vary, with Siena/NYT, YouGov, and Emerson receiving relatively high grades around 3, while others like TIPP and Redfield & Wilton Strategies are rated slightly lower.
+[@fig-map1] shows the average support for Kamala Harris as a proportion relative to both Kamala Harris and Donald Trump across the continental U.S.
+states.
+States shaded in blue indicate a higher proportion of support for Harris, while red shades represent stronger support for Trump, with color intensity reflecting the support margin.
+Gray-shaded states lack sufficient polling data.
 
-# Forecasting Election Outcome through Pooling Polls
+In key swing states, such as Pennsylvania, Michigan, and Arizona, we observe a balanced distribution of support, suggesting close competition in these battleground areas.
+Meanwhile, traditional Democratic strongholds, like California and New York, show a clear preference for Harris, with deeper blue shades.
+Conversely, traditionally Republican states, such as Texas and Tennessee, exhibit higher support for Trump, with intense red shades highlighting his advantage.
+This visualization emphasizes the varied regional dynamics of candidate support, with distinct patterns in both competitive and historically partisan states.
+
+# Forecasting Election Outcome through Pooling Polls {#sec-model}
 
@@
 
-## **Exploration of Independent Variables**
-
-### Support for Kamala Harris by Sample Size and Polling Level
-
-[@fig-sample] suggests that the relationship between support for Kamala Harris and sample size shows a slight positive trend, indicating a small increase in support as sample sizes grow.
-Both national (blue) and state (green) polls reflect this upward trend, although the effect is minimal.
-Since voting outcomes often depend on small margins, this subtle increase in support with larger sample sizes could still be relevant in predicting overall support.
-
-```{r}
-#| warning: false
-#| message: false
-#| echo: false
-#| label: fig-sample
-#| fig-cap: Sample Size vs. Support for Kamala Harris, Colored by Pollscope (National vs. State)
-
-# Custom colors for pollscope
-national_color <- "chartreuse3"
-state_color <- "steelblue"
-
-# Create the plot for Kamala Harris with the requested adjustments
-ggplot(analysis_data %>% filter(candidate_name == "Kamala Harris"), 
-       aes(x = sample_size, y = pct, color = factor(national))) +
-  geom_point(size = 0.5, alpha = 0.9) +  # Smaller dots with slight transparency
-  geom_smooth(aes(group = 1), method = "lm", se = FALSE, color = "azure4") +  # Linear model smoothing
-
-  # Custom colors for national (1 for National, 0 for State)
-  scale_color_manual(values = c("1" = national_color, "0" = state_color), 
-                     labels = c("National", "State")) +  # Pollscope categories with custom colors
-
-  # Same minimal styling
-  theme_minimal() +
-  theme(
-    panel.grid.major = element_blank(),  # Remove major gridlines
-    panel.grid.minor = element_blank(),  # Remove minor gridlines
-    legend.title = element_blank(),  # No title for the legend
-    legend.position = "bottom",  # Move legend to the bottom
-    plot.title = element_text(hjust = 0.5),  # Center the title
-    axis.line = element_line(color = "black"),  # Add axis lines
-    panel.border = element_blank(),  # Remove panel border
-    axis.line.y.right = element_line(color = "black"),  # Add right spine
-    axis.line.x.top = element_blank()  # Remove top spine
-  ) +
-  labs(x = "Sample Size", y = "Support (%)")  # No title, just axis labels
-
-```
-
-### Variation in Support for Kamala Harris Among Most Frequent Pollsters
-
-[@fig-top5pollsters] above shows the variation in support for Kamala Harris across the five most frequent pollsters.
-Notably, the median support levels differ between pollsters, indicating variability in central estimates of Harris’s support.
-For instance, Siena/NYT reports a higher median compared to Redfield & Wilton Strategies.
-Additionally, the differing ranges of support estimates highlight potential discrepancies in the methodologies, sample sizes, and biases employed by each pollster.
-This suggests that different pollster have different interpretations of candidate support, underlining the importance of aggregating data from multiple sources to avoid over-reliance on any single pollster.
-
-```{r}
-#| warning: false
-#| message: false
-#| echo: false
-#| label: fig-top5pollsters
-#| fig-cap: "Support for Kamala Harris across the top 5 pollsters by poll count."
-
-# Get top 5 pollsters by poll count
-top5_pollsters <- analysis_data %>%
-  filter(candidate_name == "Kamala Harris") %>%
-  count(pollster, sort = TRUE) %>%
-  top_n(5, n) %>%
-  pull(pollster)
-
-# Filter the data for only these top 5 pollsters
-top5_data <- analysis_data %>%
-  filter(pollster %in% top5_pollsters & candidate_name == "Kamala Harris")
-
-# Custom colors for boxplot fill
-boxplot_colors <- c("chartreuse3", "steelblue", "indianred", "purple", "darkorange")
-
-# Create the boxplots with thinner borders, smaller points, and rotated x-axis labels
-ggplot(top5_data, aes(x = pollster, y = pct, fill = pollster)) +
-  geom_boxplot(size = 0.2, outlier.size = 0.5) +  # Thinner box lines and smaller outliers
-  scale_fill_manual(values = boxplot_colors) +
-  theme_minimal() +  # Minimal theme
-  theme(
-    panel.grid.major = element_blank(),  # Remove major gridlines
-    panel.grid.minor = element_blank(),  # Remove minor gridlines
-    legend.position = "none",  # No legend needed for boxplots
-    axis.line = element_line(color = "black"),  # Add axis lines
-    plot.title = element_text(hjust = 0.5),  # Center title
-    axis.text.x = element_text(angle = 20, hjust = 1),  # Rotate x-axis labels for better readability
-    axis.line.y.right = element_line(color = "black"),  # Add right spine for clean look
-    axis.line.x.top = element_blank(),  # Remove top spine
-    panel.border = element_blank()  # Remove panel border
-  ) +
-  labs(x = "Pollster", y = "Support (%)")  # Labels
-
-```
-
-### Support for Kamala Harris Based on Recency of Polls and Pollster Quality
-
-In [@fig-enddate], we observe a slight increasing trend in support for Kamala Harris as the end date of polls progresses towards October.
-This may indicate a gradual improvement in her polling performance over time.
-The color gradient, reflecting the pollster's error and bias (Pollscore), reveals variations in bias: greener points represent more negative biases, while redder points indicate more positive biases.
-
-While the recency of the polls shows a stable trend in Harris's support, the wide distribution of bias (Pollscore) values suggests that pollster quality can vary substantially.
-These variations could influence the model's predictions if left unaddressed.
-Hence, including both recency and pollster quality as predictor variables in the model might be valuable to account for such variations.
-Further model diagnostics and validation will help decide if these variables are significant.
-
-```{r}
-#| warning: false
-#| message: false
-#| echo: false
-#| label: fig-enddate
-#| fig-cap: "Average Support for Kamala Harris in Different States estimated by Polls"
-
-# Define the color gradient based on pollscore
-ggplot(analysis_data %>% filter(candidate_name == "Kamala Harris"), aes(x = end_date, y = pct, color = pollscore)) +
-  geom_point(size = 0.8, alpha = 1) +  # Adjust point size and transparency
-  geom_smooth(method = "lm", se = FALSE, color = "azure4") +  # Add a smooth trend line for overall support
-  scale_color_gradient2(low = "blue", mid = "green", high = "red", midpoint = 0) +  # Color gradient
-  theme_minimal() +  # Minimal theme as usual
-  theme(
-    panel.grid.major = element_blank(),  # Remove major gridlines
-    panel.grid.minor = element_blank(),  # Remove minor gridlines
-    legend.position = "bottom",  # Legend at the bottom
-    axis.line = element_line(color = "black"),  # Add axis lines
-    panel.border = element_blank(),  # Remove panel border
-    axis.line.y.right = element_line(color = "black"),  # Add right spine for clean look
-    axis.line.x.top = element_blank()  # Remove top spine
-  ) +
-  labs(x = "Date Poll Ended",
-       y = "Support (%)",
-       color = "Error and Bias of Pollster (Low to High)")  # Updated label for color gradient
-
-```
-
-# Model
+## Model
 
@@
 
-We aim to model the percentage of support for Kamala Harris in each poll as a function of the pollster the sample size, the state, and the recency of the poll.
+We aim to model the percentage of support for Kamala Harris and Donald Trump in each poll as a function of the pollster the sample size, the state, and the recency of the poll.
 
@@
 
--   $y_i$​ is the percentage of support for Kamala Harris in poll iii,
+-   $y_i$​ is the percentage of support for candidate in poll iii,
 
@@
 
-We compared three models with different sets of predictors, gradually increasing the model complexity.
-Model 1 only included **pollster** and **sample size** as predictors, while Model 2 introduced **state** as an additional factor.
-Finally, Model 3 incorporated **recency** as another key predictor.
-Among these models, Model 3 demonstrated the highest $R^2$ value (0.731), indicating the best fit and ability to explain the variation in the data.
-Additionally, Model 3 had the lowest AIC (4786.815) and BIC (5203.787) values, suggesting that it provided the best balance between model complexity and predictive performance.
-Therefore, Model 3 was selected as the most appropriate model for smoothing out polling biases and differences, offering the most accurate prediction of support for Kamala Harris.
-
 ```{r}
+#| echo: false
 #| warning: false
 #| message: false
-#| echo: false
-# Load the models from RDS files
-model_1 <- readRDS(here("models/model_1.rds"))
-model_2 <- readRDS(here("models/model_2.rds"))
-model_3 <- readRDS(here("models/model_3.rds"))
-
-# Function to get model details
-get_model_details <- function(model, model_name, include_recency = FALSE) {
-  r2 <- summary(model)$r.squared
-  adj_r2 <- summary(model)$adj.r.squared
-  f_stat <- summary(model)$fstatistic[1]
-
-  # Extract p-values for Sample Size and Recency
-  sample_size_pvalue <- tidy(model) %>% filter(term == "sample_size") %>% pull(p.value)
-  sample_size_significance <- ifelse(sample_size_pvalue < 0.05, paste0("Significant (p = ", round(sample_size_pvalue, 4), ")"), paste0("Not significant (p = ", round(sample_size_pvalue, 4), ")"))
-
-  recency_significance <- "Not included"
-  if (include_recency) {
-    recency_pvalue <- tidy(model) %>% filter(term == "recency") %>% pull(p.value)
-    recency_significance <- ifelse(recency_pvalue < 0.05, paste0("Highly significant (p < 2e-16)"), paste0("Not significant (p = ", round(recency_pvalue, 4), ")"))
-  }
-
-  return(data.frame(
-    Model = model_name,
-    Variables_Included = ifelse(include_recency, "Pollster + Sample Size + State + Recency", "Pollster + Sample Size + State"),
-    R2 = round(r2, 3),
-    Adjusted_R2 = round(adj_r2, 3),
-    F_statistic = round(f_stat, 2),
-    Sample_Size_Significance = sample_size_significance,
-    Recency_Significance = recency_significance,
-    States_Significant = ifelse(model_name == "Model 1", "Not included", ifelse(model_name == "Model 2", "Several states", "More states"))
-  ))
-}
-
-# Generate the summaries for all models
-model_1_summary <- get_model_details(model_1, "Model 1")
-model_2_summary <- get_model_details(model_2, "Model 2")
-model_3_summary <- get_model_details(model_3, "Model 3", include_recency = TRUE)
-
-# Combine all summaries into one data frame
-all_summaries <- bind_rows(model_1_summary, model_2_summary, model_3_summary)
-
-# Print the table using kable for better formatting
-
-kable(all_summaries, format = "markdown",row.names = FALSE, col.names = c("Model", "Variables Included", "R²", "Adjusted R²", "F-statistic", "Sample Size Significance", "Recency Significance", "States Significant"))
+#| label: fig-causal-model
+#| fig-cap: "Directed Acyclic Graph illustrating factors influencing candidate support, with poll attributes (Pollster, Sample Size, State, Recency) as direct contributors and unobserved confounders representing potential biases in polling results."
+# Load and display the saved PNG
+knitr::include_graphics(here("other", "graphs", "polls_of_polls_dag.png"))
 
@@
 
+In our model, we aim to smooth out discrepancies and biases across various polling organizations using a polls-of-polls approach.
+Given the potential for individual pollsters to introduce systematic differences—due to variations in sampling methods, question phrasing, and historical leanings—our model includes a pollster variable to adjust for these organization-specific biases.
+This allows us to capture an aggregated view of public support that is less susceptible to the idiosyncrasies of any single poll.
+Furthermore, we incorporate sample size as a predictor, as polls with larger samples tend to yield more stable and reliable results, reducing random fluctuations caused by smaller samples.
+The state variable accounts for regional political differences, ensuring that the model captures varying levels of support across geographic and demographic groups, which is crucial for understanding the nuanced political landscape.
+Additionally, recency is included to prioritize more recent polls, as public opinion can shift rapidly in response to political events, and recent data is generally more reflective of current sentiment.
+By integrating these factors, our model seeks to produce a more stable and comprehensive measure of candidate support, reducing noise from individual poll discrepancies and focusing on a balanced, up-to-date aggregation of polling data.
+
+We opted for a linear model due to its capacity to quantify the marginal effects of each predictor (pollster, sample size, state, and recency) on candidate support in a straightforward manner.
+This structure is well-suited to our polls-of-polls approach, as it allows for the estimation of fixed effects that can control for systematic biases across pollsters, while accommodating the influence of sample size and recency as continuous variables.
+All modeling was conducted using the base R package [@citeR], specifically utilizing the lm() function from the stats package for linear regression analysis.
+
+## Model Results
+
 ```{r}
+#| echo: false
 #| warning: false
 #| message: false
-#| echo: false
+#| label: tbl-summary
+#| tbl-cap: "Model performance summary showing improved fit and accuracy as State and Recency are added, with R² increasing from 0.375 in Model 1 to 0.774 in Model 3, and RMSE decreasing from 3.053 to 1.836."
+# Load your models
+model1 <- readRDS(here("models/model1.rds"))
+model2 <- readRDS(here("models/model2.rds"))
+model3 <- readRDS(here("models/model3.rds"))
 
-# Create a data frame for AIC and BIC comparison
-aic_bic_table <- data.frame(
+# Extract summary statistics and add variables included in each model
+model_summary <- tibble::tibble(
   Model = c("Model 1", "Model 2", "Model 3"),
-  AIC = c(AIC(model_1), AIC(model_2), AIC(model_3)),
-  BIC = c(BIC(model_1), BIC(model_2), BIC(model_3))
+  Variables = c("Pollster, Sample Size", "Pollster, Sample Size, State", "Pollster, Sample Size, State, Recency"),
+  `R²` = c(summary(model1)$r.squared, summary(model2)$r.squared, summary(model3)$r.squared),
+  `Adjusted R²` = c(summary(model1)$adj.r.squared, summary(model2)$adj.r.squared, summary(model3)$adj.r.squared),
+  `AIC` = c(AIC(model1), AIC(model2), AIC(model3)),
+  `BIC` = c(BIC(model1), BIC(model2), BIC(model3)),
+  `RMSE` = c(sqrt(mean(residuals(model1)^2)), sqrt(mean(residuals(model2)^2)), sqrt(mean(residuals(model3)^2)))
 )
 
-# Print the table using kable for LaTeX format
-kable(aic_bic_table, format = "markdown", col.names = c("Model", "AIC", "BIC"), booktabs = TRUE)
+# Display the table using kable
+model_summary %>%
+  kable(caption = "Model Summary with Included Variables", digits = 3) %>%
+  kable_styling(full_width = FALSE)
 
@@
 
-## Model Results
+[@tbl-summary] summarizes the performance metrics for three models with progressively added variables.
+Model 1, which includes only Pollster and Sample Size, achieves an R² of 0.375, indicating that these variables alone explain about 37.5% of the variance in candidate support.
+Model 2 incorporates State as an additional predictor, resulting in a substantial improvement, with an R² of 0.720 and a reduction in both AIC and RMSE, showing better model fit and predictive accuracy.
+Model 3 further adds Recency, which increases the R² to 0.774 and decreases the RMSE to 1.836, indicating enhanced explanatory power and prediction accuracy.
+This progression highlights the benefit of adding contextual variables like State and Recency to better capture the complexities of polling data.
 
-The table below shows the model output, where only a few states and pollsters are displayed to keep the output concise, as the full model includes many variables.
-We included "State California," "State Texas," "Pollster Redfield & Wilton Strategies," and "Pollster Echelon Insights" to provide examples of how different states and pollsters influence support for Kamala Harris.
-
-The variable "Sample Size" has a positive coefficient, suggesting that polls with larger sample sizes are associated with slightly higher support percentages.
-"Recency" has a negative and highly significant coefficient, indicating that more recent polls tend to show lower support levels.
-Additionally, certain states like California and Texas show significant effects on the percentage of support, with California showing a notable increase.
-Similarly, different pollsters also display varying biases, as seen with the coefficients for "Pollster Redfield & Wilton Strategies" and "Pollster Echelon Insights." The model shows strong goodness-of-fit, with an $R^2$ of 0.731, indicating that 73.1% of the variability in support can be explained by the included variables.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-# Load necessary package
-library(stargazer)
-
-# Use stargazer to display only the specified variables and custom labels
-stargazer(model_3, 
-          type = "text",  # Can be "text", "html", or "latex"
-          covariate.labels = c("Sample Size", "Recency", "State California", "State Texas",
-                               "Pollster Redfield & Wilton Strategies", "Pollster Echelon Insights"),  # Renaming the variables
-          keep = c("sample_size", "recency", "stateCalifornia", "stateTexas", 
-                   "pollsterRedfield & Wilton Strategies", "pollsterEchelon Insights"),  # Keep only these variables
-          omit.stat = c("f", "ser"),  # Optionally omit F-statistics or other GOF stats
-          star.cutoffs = c(0.05, 0.01, 0.001),  # Custom significance stars
-          single.row = TRUE)  # Display coefficients on one row for readability
-
-```
-
-# Prediction
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# Create a new dataset called 'predict_data' with predicted values and weights
-predict_data <- analysis_data %>%
-  # Add the predicted values (fitted values from model_3)
-  mutate(fitted_values = fitted(model_3),
-         # Create weights based on numeric grade
-         weight = numeric_grade / sum(numeric_grade))
-
-# Calculate the weighted fitted values for Kamala Harris support
-weighted_support <- sum(predict_data$fitted_values * predict_data$weight)
-
-```
+## Prediction
 
@@
 $$
-w_i = \frac{\mathrm{numeric\_grade}_i}{\sum_{i=1}^{n} \mathrm{numeric\_grade}_i}
+w_i = \frac{\mathrm{numeric\_grade}_i \times (\mathrm{maxPollscore} - \mathrm{pollscore}_i)}{\sum_{i=1}^{n} \mathrm{numeric\_grade}_i \times (\mathrm{maxPollscore} - \mathrm{pollscore}_i)}
 $$
@@
 
-Using these weights, the overall weighted prediction of Kamala Harris' support is calculated by summing the weighted predicted values from our regression model:
+-   $pollscore_i$ is the is the pollscore of poll i which reflects the estimated bias of the poll (with more negative values indicating less bias)
+
+-   $maxPollscore$ is the maximum pollscore across all polls.
+
+The weight assigned to each poll combines both its quality (as represented by the numeric grade) and its level of bias (as indicated by the pollscore).
+Since a more negative pollscore reflects a lower level of bias, the formula uses the difference between the maximum pollscore and each poll’s specific pollscore.
+This approach gives more weight to polls that are both highly graded (indicating higher reliability and transparency) and less biased.By combining these two factors, the weighting system emphasizes polls that are reliable and minimally biased, ensuring that they have a stronger influence in the overall calculation.
+Additionally, the total of all weights is normalized to sum to one, so each poll’s weight is proportionate to its quality and relative lack of bias, resulting in a more balanced and accurate average of public support.
+
+Using these weights, the overall weighted prediction of candidate's support is calculated by summing the weighted predicted values from our regression model:
 
@@
 
-Applying this approach, our model predicts Kamala Harris' overall support to be **47.78%**.
-This prediction indicates that, after adjusting for biases in the polls, Harris' support is slightly below the 50% threshold.
-The weighted approach accounts for the differences in poll reliability and thus provides a more accurate estimate than using raw averages of the polls.
+### Comparing Overall Weighted Support Across All Polls
 
-This result suggests that Kamala Harris is facing some challenges in gaining majority support based on the current polling data.
-However, the model reflects the weighted average from multiple sources, smoothing out individual poll biases to offer a comprehensive view of her support landscape.
+Aggregating support across all polls and applying our weighted approach, we estimate the overall support for each candidate.
+The weights, calculated based on both the poll's numeric grade and poll score, ensure that higher-quality polls with less bias contribute more to our estimates.
+Based on this approach, Kamala Harris has an estimated overall weighted support of around 47.77%, while Donald Trump stands at around 46.31%.
+This aggregation provides a comprehensive view, taking into account the varied methodologies and sampling qualities of different polling organizations.
 
-# Discussion
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| include: true
 
-In this paper, we set out to predict Kamala Harris' support using a polls-of-polls approach.
-By aggregating multiple polls, we aimed to smooth out the inherent biases present in individual surveys and create a more accurate and reliable forecast.
-Our analysis was driven by a linear regression model that used key predictors, such as pollster, sample size, state, and recency of the polls.
-After obtaining predicted values from the model, we applied a weighting scheme based on the numeric grade of each pollster, reflecting the quality and reliability of the poll.
-By assigning higher weights to more reliable pollsters, we generated a final prediction for Kamala Harris' overall support, accounting for variations across different states and polling organizations.
+model6 <- readRDS(here("models/model6.rds"))
 
-One key insight from our analysis is the significance of poll recency.
-The negative and highly significant coefficient for recency indicates that **older** polls show lower support for Kamala Harris, while more recent polls tend to show higher support.
-This suggests that her popularity may be increasing over time, or that certain factors have shifted voter sentiment in her favor in more recent months.
-Inncorporating recency into the model also greatly improved its overall performance.
-Adding recency as a predictor significantly increased the model's explanatory power, raising the R² value and lowering both AIC and BIC scores compared to models that excluded it.
-This indicates that accounting for the timing of polls is essential to accurately predicting voter sentiment, as public opinion is highly dynamic and evolves rapidly in response to events.
-Including recency strengthens the model's ability to reflect these temporal shifts, leading to more reliable forecasts.
+# Step 1: Subset data for Kamala Harris, calculate weights, join with fitted values from model3, and calculate weighted_pct
+kamala_data <- analysis_data %>%
+  filter(candidate_name == "Kamala Harris") %>%
+  mutate(
+    weight = numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore) /
+             sum(numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore), na.rm = TRUE),
+    fitted_values = fitted(model3),  # Replace this with actual fitted values for Kamala
+    weighted_pct = fitted_values * weight  # Calculate weighted percentage for Kamala
+  )
 
-Our analysis also highlights the variability in support for Kamala Harris across different states and pollsters.
-For example, certain states like California and Texas exhibit distinct effects, with California significantly boosting Harris’ predicted support.
-Additionally, different pollsters show varying levels of bias, as seen with Redfield & Wilton Strategies and Echelon Insights, which have divergent effects on support levels.
-This underscores the geographic and methodological complexities in predicting election outcomes, demonstrating that voter preferences are not homogenous across regions or polling firms.
+# Calculate the overall weighted prediction for Kamala Harris
+kamala_weighted_support <- sum(kamala_data$weighted_pct, na.rm = TRUE)
 
-Despite the strengths of the polls-of-polls methodology, there are several limitations to our approach.
-First, our model assumes that the relationship between the predictors (pollster, sample size, state, recency) and support for Kamala Harris is linear, which may not always be the case.
-Non-linear trends or interactions between variables might exist that were not captured by the model.
-Additionally, while we weighted polls based on their numeric grade, this metric might not fully reflect the true accuracy or bias of each pollster.
-Moreover, we only included polls conducted after Harris' official candidacy announcement, potentially omitting relevant data from earlier in the election cycle that could provide a fuller picture.
+# Step 2: Subset data for Donald Trump, calculate weights, join with fitted values from model6, and calculate weighted_pct
+trump_data <- analysis_data %>%
+  filter(candidate_name == "Donald Trump") %>%
+  mutate(
+    weight = numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore) /
+             sum(numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore), na.rm = TRUE),
+    fitted_values = fitted(model6),  # Replace this with actual fitted values for Trump
+    weighted_pct = fitted_values * weight  # Calculate weighted percentage for Trump
+  )
 
-Future research could delve into more sophisticated methods to capture non-linear relationships and interactions between variables.
-For instance, incorporating machine learning algorithms like random forests or neural networks could allow for more flexible modeling of polling data.
-Another area of improvement could be refining the pollster weighting mechanism by integrating other factors, such as the historical performance of the pollster or specific characteristics of the electorate surveyed.
-Additionally, conducting more in-depth analyses of key battleground states could reveal more granular insights into voter behavior, especially in regions where small shifts in support could have outsized impacts on election outcomes.
+# Calculate the overall weighted prediction for Donald Trump
+trump_weighted_support <- sum(trump_data$weighted_pct, na.rm = TRUE)
 
-In conclusion, while our model provides valuable predictions of Kamala Harris' support, there is always room to enhance the predictive accuracy by refining methodologies and incorporating new data sources.
-The evolving nature of elections demands continuous adjustment of models to reflect changes in public opinion, candidate strategies, and electoral dynamics.
+```
+
+### State-Level Predictions
+
+In this section, we examine the predicted support for Kamala Harris and Donald Trump within each state, based on our weighted approach.
+By aggregating poll results within each state and applying weights that adjust for poll quality and bias, we can estimate candidate support with a more localized perspective.
+This allows us to identify state-by-state competition and highlight key battleground states where the margins are close.
+
+For each state, we compare the predicted support percentages and determine the projected winner based on which candidate has higher weighted support.
+Additionally, we calculate the margin of difference between the candidates, which can provide insights into how close each state’s race is.
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| include: false
+# Define the max pollscore across the dataset
+max_pollscore <- max(analysis_data$pollscore, na.rm = TRUE)
+
+# Step 1: Subset data for Kamala Harris, join fitted values, then calculate weights within each state
+kamala_data_a2 <- analysis_data %>%
+  filter(candidate_name == "Kamala Harris") %>%
+  mutate(fitted_values = fitted(model3)) %>%  # Add fitted values from model 3 for Kamala Harris
+  group_by(state) %>%
+  mutate(
+    weight = numeric_grade * (max_pollscore - pollscore) /
+             sum(numeric_grade * (max_pollscore - pollscore), na.rm = TRUE)
+  ) %>%
+  # Calculate the overall weighted prediction for each state
+  summarize(support_kamala = sum(fitted_values * weight, na.rm = TRUE))
+
+# Step 2: Repeat for Donald Trump with model6
+trump_data_a2 <- analysis_data %>%
+  filter(candidate_name == "Donald Trump") %>%
+  mutate(fitted_values = fitted(model6)) %>%  # Add fitted values from model 6 for Donald Trump
+  group_by(state) %>%
+  mutate(
+    weight = numeric_grade * (max_pollscore - pollscore) /
+             sum(numeric_grade * (max_pollscore - pollscore), na.rm = TRUE)
+  ) %>%
+  # Calculate the overall weighted prediction for each state
+  summarize(support_trump = sum(fitted_values * weight, na.rm = TRUE))
+
+# Step 3: Join the two datasets by state
+state_support_data <- left_join(kamala_data_a2, trump_data_a2, by = "state")
+
+# Add a new column 'winner' based on higher support values
+state_support_data <- state_support_data %>%
+  mutate(
+    winner = ifelse(support_kamala > support_trump, "Kamala Harris", "Donald Trump"),
+    margin = abs(support_kamala - support_trump)
+  )
+```
+
+```{r, fig.width = 12, fig.height = 6}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-map2
+#| fig-cap: "Predicted 2024 U.S.
+Presidential Election winner by state, with Kamala Harris-leaning states in blue, Donald Trump-leaning states in red, and states lacking sufficient polling data in gray.
+The map highlights traditional Democratic and Republican strongholds as well as key battleground states."
+
+# Filter out Alaska, Hawaii, and Puerto Rico from state_support_data
+map_data_continental <- subset(us_states, !NAME %in% c("Alaska", "Hawaii", "Puerto Rico"))
+
+# Merge with state_support_data to add winner info
+map_data_continental <- map_data_continental %>%
+  left_join(state_support_data,  by = c("NAME" = "state"))
+
+# Define the colors for each category
+color_scale <- c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred", "No Polling Data" = "grey")
+
+# Plotting the map with ggplot2
+ggplot(map_data_continental) +
+  geom_sf(aes(fill = winner), color = "white", size = 0.2) +
+  geom_text(aes(label = STUSPS, geometry = geometry), 
+            stat = "sf_coordinates", color = "black", size = 3) +  # Adjust text size as needed
+  scale_fill_manual(
+    values = color_scale, 
+    na.value = "grey",
+    name = "2024 Election Winner",
+    labels = c("Kamala Harris", "Donald Trump", "No Polling Data")
+  ) +
+  theme_minimal() +
+  theme(
+    panel.grid = element_blank(),
+    axis.text = element_blank(),
+    axis.title = element_blank(),
+    axis.ticks = element_blank(),
+    aspect.ratio = 0.5  # Adjust the aspect ratio to make the map wider
+  )
+
+```
+
+[@fig-map2] presents the predicted winner of the 2024 U.S.
+Presidential Election by state based on aggregated polling data, with red representing states projected to favor Kamala Harris and blue indicating those favoring Donald Trump.
+Notably, traditional Democratic strongholds in the Northeast and West Coast, such as California, New York, and Washington, show solid support for Harris.
+Conversely, traditional Republican states, including Texas, Florida, and much of the South, show strong support for Trump.
+
+Swing states, like Pennsylvania, Michigan, and Wisconsin, are predominantly leaning toward Harris, indicating a possible advantage for her in these key battlegrounds, although other typical swing states like Arizona and Georgia lean toward Trump.
+Gray-shaded states lack sufficient polling data to make a prediction, reflecting areas of data scarcity in the model's projections.
+This visualization highlights the geographical and political divides across the U.S., as well as the critical role swing states play in determining the election outcome.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-battleground-states
+#| tbl-cap: "Predicted support for Kamala Harris and Donald Trump across key battleground states, indicating the winner and the support margin (%) in each state."
+
+# Define battleground states based on provided list
+battleground_states <- c("Pennsylvania", "Michigan", "North Carolina", "Wisconsin", "Nevada", "Arizona", "Georgia")
+
+# Display the table of battleground states with adjusted column names
+battleground_table <- state_support_data %>%
+  filter(state %in% battleground_states) %>%
+  select(
+    State = state,
+    `Kamala Support (%)` = support_kamala,
+    `Trump Support (%)` = support_trump,
+    `Predicted Winner` = winner,
+    `Support Margin (%)` = margin
+  )
+
+# Display the table using kable
+knitr::kable(battleground_table)
+
+```
+
+USAFacts[@usafacts],FiveThirtyEight[@FiveThirtyEight] and many other polling websites are calling Arizona, Georgia, Michigan, Pennsylvania, Wisconsin, North Carolina, Nevada the 'swing states' of the 2024 presidential election.Polling in swing states is crucial for election predictions because these states, with their historically close voting margins, often determine the overall outcome by tipping the electoral balance toward one candidate.
+[@tbl-battleground-states] summarizes the predicted support percentages for Kamala Harris and Donald Trump in key battleground states, along with the predicted winner and the support margin.
+The data shows close margins in these states, with some like Arizona and Georgia leaning slightly towards Trump, while others such as Michigan and Pennsylvania favor Harris by narrow margins.
+The support margin column highlights the competitiveness of these states, with all margins below 2.2%, indicating highly contested races.
+
+# Discussion {#sec-discussion}
+
+In this paper, we set out to predict the support for both Kamala Harris and Donald Trump in the 2024 U.S.
+Presidential Election using a “polls-of-polls” approach.
+By aggregating multiple polls, we sought to mitigate the inherent biases present in individual surveys, creating a more accurate and balanced forecast across candidates.
+Our analysis was driven by a linear regression model using key predictors, such as pollster, sample size, state, and recency of the polls.
+After calculating predicted values from the model, we applied a weighting scheme based on each pollster’s numeric grade and pollscore, incorporating both reliability and bias.
+This weighted approach enabled us to produce an overall prediction of candidate support, reflecting variations across different states and polling organizations.
+
+One key insight from our analysis is the importance of poll recency.
+Including recency notably increased the model’s explanatory power, with improvements in R² and reductions in RMSE compared to models without it.
+This underscores the dynamic nature of public opinion, where recent events can influence voter perceptions and candidate support.
+
+Our state-level analysis highlights close competition observed in key battleground states, emphasizing their pivotal role in the 2024 election.
+Our model reveals that in states like Pennsylvania, Michigan, and Wisconsin, support levels for Harris and Trump are nearly tied, reflecting the intense contest for these critical electoral votes.
+The predictions show that even minor shifts in support within these swing states could decisively impact the overall election outcome.
+This analysis shows the importance of regional polling in capturing nuanced voter sentiment and the heightened influence that battleground states hold in shaping the final results.
+
+However, there are limitations to our approach.
+Our model assumes linear relationships between predictors and candidate support, which might not capture more complex or interaction-based trends.
+Additionally, while we weighted polls based on their numeric grades, these scores may not fully reflect each pollster’s accuracy, leaving room for improvement.
+Future research could explore non-linear methods, such as machine learning, to capture potential interactions between variables.
+Refining the weighting mechanism by considering pollster performance history or state-specific polling nuances could further enhance predictive accuracy.
+
+In conclusion, this paper contributes to election forecasting by providing a comprehensive, weighted estimate of candidate support across both national and state levels.
+While our predictions provide valuable insights into potential election outcomes, the evolving nature of voter sentiment and poll accuracy reflects the need for adaptive models that account for ongoing changes in public opinion.
 
@@
 # Copy of the Survey
+
 \includepdf[pages=-]{../other/surveyfiles/SurveyCopy.pdf}
+
+# Model Diagnostics
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-dgplotone
+#| 
+knitr::include_graphics(here("other", "graphs", "diagnostic_plot_model3.png"))
+```
+
+The diagnostic plots for Model 3ni sgiven in [@fig-dgplotone]. We notice the following:
+
+1.
+**Residuals vs.
+Fitted Plot**: This plot checks for non-linearity and heteroscedasticity.
+The residuals are scattered around the horizontal axis with no clear pattern, indicating that the linearity assumption is reasonable.
+However, there is slight spread around the center, suggesting some mild heteroscedasticity.
+
+2.
+**Normal Q-Q Plot**: This plot assesses the normality of residuals.
+Most points align along the diagonal line, although there are some deviations at the tails.
+This suggests that the residuals are approximately normally distributed, with minor deviations in the extremes.
+
+3.
+**Scale-Location Plot**: This plot further checks for homoscedasticity.
+The residuals appear to be evenly spread across the fitted values, supporting the homoscedasticity assumption, although slight deviations are present in certain regions.
+
+4.
+**Residuals vs.
+Leverage Plot**: This plot identifies potential influential points.
+While most points have low leverage, a few points exhibit higher leverage, as indicated by their distance from the center.
+However, no points exceed Cook’s distance threshold, indicating no extreme outliers that would unduly influence the model.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-dgplottwo
+#| 
+knitr::include_graphics(here("other", "graphs", "diagnostic_plot_model6.png"))
+```
+
+The diagnostic plots for Model 6 is given in [@fig-dgplottwo]. We notice the following:
+
+1.
+**Residuals vs Fitted**: This plot suggests a fairly random spread of residuals around zero, indicating that the linearity assumption holds reasonably well.
+However, there is slight clustering of points in the center, which could indicate minor heteroscedasticity but is not severe.
+
+2.
+**Normal Q-Q Plot**: The Q-Q plot shows that the residuals generally follow a normal distribution, with only a few deviations at the tails.
+This suggests that the normality assumption is mostly met, though some minor deviations in the upper tail indicate possible outliers.
+
+3.
+**Scale-Location (Spread-Location) Plot**: The residuals appear randomly dispersed with no clear pattern, supporting the homoscedasticity (constant variance) assumption.
+However, some observations near the upper range might indicate slight variance inconsistency, though it is minimal.
+
+4.
+**Residuals vs Leverage**: This plot does not show any influential outliers with high leverage that might impact the model unduly.
+The Cook's distance lines show that no data points exceed these thresholds, indicating that no individual observation is disproportionately influencing the model.
+
+Overall, the diagnostic plots for both model suggests that they meet the assumptions for linear regression fairly well, with minor deviations that are not expected to severely impact the model’s validity.
+
+# Code Styling
+
+The code in this paper was reviewed and formatted for consistency using lintr [@citelintr] and styler [@citestyler], ensuring readability and adherence to style standards.
+
+# Reproducibility
+
+To replicate the findings presented in this paper, users should execute the scripts available in the GitHub repository.
+Begin by running the 00-install_packages.R script, which installs all required packages for the analysis.
+
+# Acknowledgments
+
+We extend our gratitude to [@citeTS], which provided invaluable guidance in establishing a reproducible workflow and inspired many of the code structures used in this paper.
 

### 2024-11-01 paper/paper.qmd
re-work Appendix A

@@
 
-Emerson College Polling, known for its extensive online polling methods, conducted a survey from October 14 to 16, 2024, targeting 1,000 likely voters.
-The poll measured voter preferences between Kamala Harris and Donald Trump, revealing a near tie with Harris at 50% and Trump at 49%.
-
-## Population, Frame, and Sample
-
-The target population is likely voters in the U.S.
-elections.
-These voters are determined by the voters' likeliness to vote in the upcoming elections and the voter's voting history.
-Both of these factors are self-reported.
-
-The sampling frame is the voters that are on the Aristotle list and on the online panel provided by CINT.
-Aristotle is an online supplier of voter files and CINT is a software company that connects researchers, such as Emerson, to online panels.
-This approach helps ensure a representative sample by accurately reflecting the electorate.
-
-The sample size is the 1,000 likely voters chosen arbitrarily from the sampling frame.
-Likely voters are determined by a combination of voter history, registration status, and demographic data.
-These factors are self-reported.
+Emerson College Polling, also known as ECP, is a nationally-ranked polling center that aims to accurately reflect public opinion through a mixed-mode methodology [@about_ecp].
+ECP conducted a survey from October 14 to 16, 2024, targeting 1,000 likely voters in the 2024 U.S.
+Presidential Election.
+The poll measured voter preferences between Kamala Harris and Donald Trump, revealing a near tie with Harris at 50% and Trump at 49% [@ecp_poll].
+Appendix A takes a closer look at the methodology used by ECP for this poll.
+
+## Population, Frame, and Sample of the Poll
+
+To conduct a poll, it is important to first define the target population, the sampling frame and the sample.
+Defining these elements sets a clear direction for the research and assures a clear object of research [@akman].
+
+The target population is the whole group of interest the researchers would like to make conclusions on [@def_terms].
+For the poll conducted by ECP in October, the target population is likely voters in the U.S.
+elections [@ecp_poll].
+Choosing likely voters as the target population makes the poll results more accurate as not all registered voters end up voting in the Presidential Election.
+In fact, the last U.S.
+Presidential Election in 2020 only had 66.8% of registered voters participate in the elections, which is the highest voter turnout in the 21st century [@us_census].
+In ECP’s methodology, likely voters are determined by a combination of voter history, registration status, and demographic data.
+These factors are self-reported [@ecp_poll].
+
+The sampling frame is the part inside of the target population that has a chance to be sampled upon [@def_terms].
+In ECP’s methodology, the sampling frame is the voters that are on Aristotle’s database and on the online panel provided by CINT [@about_ecp].
+Aristotle is an online source of voter and consumer data that has serviced many large political campaigns, PACs and corporations both within the U.S and abroad [@aristotle].
+CINT is the largest global research marketplace that connects researchers, such as Emerson, to survey respondents [@cint].
+
+The sample of a study is the part of the sampling frame that is measured and is part of the data set [@def_terms].
+In the case of a poll like ECP’s, the sample is the respondents of the survey.
+These respondents are directly represented in the data set [@ecp_poll].
+The sample size is 1,000 likely voters chosen arbitrarily from the sampling frame.
 
@@
 
-Emerson College used a mixed-mode sampling approach for its polling. Specifically, three primary modes were used to collect data for its October 14-16, 2024 poll:
+Emerson College used a mixed-mode sampling approach for its polling [@about_ecp]. Specifically, three primary modes were used to collect data for its October 14-16, 2024 poll [@ecp_poll] :
 
@@
 
-The multi-mode sampling approach used in this poll has several advantages:
-
--   Wider demographic reach: Having a combination of data collection methods allows Emerson to gather responses from a broader demographic group.
-In this polling, MMS-to-web surveys mostly gather responses from younger voters and those who primarily use mobile devices.
-IVR target older and rural voters, who are more likely to respond via landlines.
-Online panels is more efficient in reaching tech-savvy individuals.
-
--   Cost-efficient: Automated technologies reduce the cost compared to live interviews. In a polling context, this allows for a larger sample size.
-
-However, the trade-offs include:
-
--   Coverage bias: IVR only reaches individuals with landlines, which may skew results towards older demographics.
-
--   Self-selection bias: Respondents in online panels tend to be more engaged, which can lead to over-representation of politically active individuals.
+The multi-mode sampling approach used by ECP in this poll has many advantages.
+By using multiple data collection methods, ECP reduces coverage bias [@mora].
+In a survey, coverage bias means there is some part of the target population that has zero chance of being part of the sample and will not be represented in the data [@def_cb].
+For example, MMS-to-web surveys tend to attract responses from younger voters and individuals who rely primarily on mobile devices, capturing a segment that may otherwise be underrepresented.
+Interactive voice response (IVR) calls target older and rural voters, who are more inclined to use landlines.
+Online panels gather responses from tech-savvy individuals who prefer online engagement over other methods.
+Multi-mode sampling is also more cost-effective [@mora].
+Automated technologies such as MMS-to-text web surveys and IVR reduce expenses compared to traditional live interviews.
+Lower costs allow the poll to reach a larger sample size, which also results in more accurate and representative conclusions [@large_ss].
+However, using a multi-mode sampling approach also introduces a trade-off.
+This sampling approach introduces a measurement error [@mora].
+Each method—whether IVR, web surveys, or online panels—brings its own unique biases.
+Combining these creates inconsistencies between respondent groups.
+This increases the overall margin of error by adding variability.
 
@@
 
-Emerson addresses non-response using a weighting system.
-The data is adjusted based on key demographic variables like age, gender, race, education, and party affiliation, making the sample more representative of the electorate.
-Weights are applied to compensate for under- or over-represented groups, ensuring balanced results.
+ECP addresses non-response using a weighting system.
+The data is adjusted based on demographic variables like age, gender, race, education, and party affiliation, making the sample more representative of the electorate [@about_ecp].
+Weights are applied to even out under- and over-represented groups, making the data more similar to the electorate [@weight].
 
@@
 
-The Emerson survey features a straightforward, efficient design focused on key issues and voter preferences. Strengths of the questionnaire include:
-
--   Simplicity: Clear, concise questions makes the survey easy for respondents to understand and to provide accurate answers.
-
--   Topical relevance: The poll zeroes in on vote preference and demographic splits, delivering timely insights during election cycles.
-
-However, weaknesses include:
-
--   Lack of nuance: While the survey captures basic preferences, it does not delve deeply into voters' motivations or the factors driving their choices.
-
--   Mode limitations: Different survey modes can affect how respondents engage with the questions. This can potentially influence the depth and accuracy of voter responses.
+The Emerson survey is designed with simplicity and efficiency.
+One of its strengths is its straightforward format, using clear and unambiguous questions.
+This simplicity makes the survey easy for respondents to understand.
+Another strength is its topical relevance.
+By focusing on vote preferences and demographic splits, the survey provides timely insights, especially valuable during election cycles.
+However, the survey has some limitations.
+One is its lack of nuance; while it captures basic voter preferences, it doesn’t explore the motivations or factors driving those choices in depth.
+Additionally, the various survey modes can affect respondent engagement with the questions, which may also influence the accuracy and depth of their answers.
 
@@
 Its weighting techniques contribute to reliable and accurate results despite non-responses.
-While the use of multiple survey modes increases population representation, possible biases—such as lack of nuance and mode effects need to be taken into consideration when interpreting results.
-
-In sum, this method is a strong predictor of election outcomes and continues to adapt to the evolving challenges of modern polling.
+While the use of multiple survey modes reduces coverage bias, other possible biases—such as lack of nuance and combined measurement errors must be considered during the analysis.
 

### 2024-11-01 paper/paper.tex
cleaning

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{siunitx}

  \newcolumntype{d}{S[
    input-open-uncertainty=,
    input-close-uncertainty=,
    parse-numbers = false,
    table-align-text-pre=false,
    table-align-text-post=false
  ]}

\KOMAoption{captions}{tableheading}
\usepackage{pdfpages}
\usepackage[margin=1in]{geometry}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Forecasting the 2024 U.S. Presidential Election: An Analysis for Battleground States},
  pdfauthor={Shamayla Durrin; Denise Chang; Krishna Kumar},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Forecasting the 2024 U.S. Presidential Election: An Analysis for
Battleground States\thanks{Code and data are available at:
\url{https://github.com/krishnak30/US_elections}.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Trump Favored Nationally, While Harris Leads in 4 of 7
Battleground States}
\author{Shamayla Durrin \and Denise Chang \and Krishna Kumar}
\date{November 1, 2024}

\begin{document}
\maketitle
\begin{abstract}
This study employs a polls-of-polls approach to predict support for
Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election,
with a focus on key battleground states to forecast the likely winner of
the electoral college. By aggregating multiple polls and applying a
weighted linear regression model with predictors like pollster
reliability, sample size, state, and recency, we estimate a higher
national support for Trump relative to Harris. Our analysis also reveals
a close competition in battleground states, with Trump holding a narrow
lead in Arizona, Georgia, and North Carolina, while Harris leads in
Michigan, Nevada, Pennsylvania, and Wisconsin. These findings underscore
the value of aggregating polls over relying on individual surveys,
offering a more comprehensive and robust forecast of electoral outcomes.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable]}{\end{tcolorbox}}\fi

\vspace{0.7cm}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

While individual polls provide snapshots of public opinion, they are
often subject to biases and methodological differences. In this paper,
we aim to forecast voter support for Kamala Harris and Donald Trump by
aggregating data from multiple polling sources, reducing individual poll
biases, and improving overall prediction accuracy. Our analysis focuses
on estimating the levels of support for each candidate not only
nationally but also in key swing states that are likely to be decisive
in determining the electoral college outcome.

The estimand of this study is the level of voter support for each
candidate, Kamala Harris and Donald Trump, as reported across multiple
polls.To estimate this, we developed a regression model incorporating
variables such as pollster, sample size, state, and recency of the poll,
with an emphasis on accurately capturing state-level dynamics. Our
findings indicate that, on a national level, support between Kamala
Harris and Donald Trump is closely balanced. Harris's estimated national
support stands close to 48\%, while Trump's support is slightly higher,
around 49\%. We found that Trump holds a marginal lead in battleground
states like Arizona and Georgia, with a support margin slightly above
1\%. In contrast, Harris shows narrow leads in Michigan, Nevada,
Pennsylvania, and Wisconsin, with her support margin in Wisconsin
reaching over 2\%, making it her strongest battleground. North Carolina
emerges as one of the tightest races, with Trump leading by only 0.26\%,
underscoring the competitive nature of these regions.

This paper contributes to the field of election forecasting by
emphasizing the importance of poll quality and recency in prediction
models. By focusing on battleground states where the support margins are
particularly close, our approach highlights the areas where campaign
strategies and voter turnout efforts may have the greatest impact. This
could provides insights for political strategists, media analysts, and
the general public by offering a nuanced view of the electoral
landscape.

This paper is organized as follows: In \{\#sec-data\}, we clean the
data, explore summary statistics, plot distributions of key variables,
and examine relationships between variables. In \{\#sec-model\}, we
discuss our forecasting approach, model selection, justification for the
chosen model, and the mechanism of deriving poll weights based on
pollster reliability, ultimately presenting our predictions. Finally, in
\{\#sec-discussion\}, we address the broader implications of our
findings, acknowledge limitations, and suggest directions for future
work. \vspace{0.7cm} \# Data \{\#sec-data\} \vspace{0.7cm} \#\#
Measurement

In the dataset of our analysis, the process of measurement begins by
capturing real-world public opinion through surveys. Polling
organizations transform this observed phenomenon into structured data by
recording support percentages, pollster details, and methodological
choices, thus turning public sentiment into quantifiable entries in the
dataset.

Public opinion polls are essential for understanding voter preferences
and electoral dynamics, providing data that informs policy decisions and
campaign strategies. Our data set captures public support for candidates
at specific times, along with information on methodologies, sample
sizes, and pollster ratings.

Polling organizations use various methods, such as App panels, IVR
(Interactive Voice Response), Live phone calls, Text-to-web systems etc.
These methods can influence the validity and reliability of results,
with larger sample sizes generally yielding more reliable estimates.
Some methods, like live phone interviews, may be seen as more
trustworthy than others.

The dataset also includes important measures of pollster performance,
such as numeric grades, poll scores, and transparency scores. These
metrics are calculated by considering bias, race difficulty, predictive
error, and transparency, allowing us to assess the accuracy of each
poll. The ``pollscore,'' for instance, reflects a pollster's past
performance and reliability, adjusted for potential biases.

\vspace{0.7cm}

\hypertarget{data-cleaning}{%
\subsection{Data Cleaning}\label{data-cleaning}}

The raw data for this project, sourced from FiveThirtyEight,
(FiveThirtyEight 2024) underwent a series of cleaning steps to prepare
it for analysis. Initially, duplicate rows were removed to ensure that
only unique observations remained, facilitated by the \texttt{janitor}
package (Firke 2023). A new binary variable, `national', was created to
indicate whether a poll was conducted at the national or state level.
Missing values in the `state' column were replaced with ``Not
Applicable,'' and numeric grades were evaluated to filter out
low-quality pollsters, keeping only those with a numeric grade above 1.
This cutoff was selected to retain mid to high-level pollsters for more
reliable results. These steps were performed using functions from the
\texttt{Tidyverse} package (Wickham et al. 2019). Furthermore, dates
were standardized and converted into a proper format for analysis using
the \texttt{lubridate} package (Grolemund and Wickham 2011). Polls
related to Kamala Harris were retained for further analysis, and
percentage support values were transformed into actual numbers of
supporters based on sample size. Additionally, pollster counts below
five were excluded to focus on more reliable data sources. Polls
regarding Kamala Harris were filtered to include only those conducted
after her official candidacy announcement on July 21, 2024, ensuring the
data reflects post-announcement public sentiment.The cleaned dataset was
saved in Parquet format for efficient storage and retrieval, using the
\texttt{arrow} package (Richardson et al. 2024).

\hypertarget{explorations-of-variables-of-interest}{%
\subsection{Explorations of Variables of
Interest}\label{explorations-of-variables-of-interest}}

\hypertarget{summary-statistics-of-key-variables}{%
\subsubsection{Summary Statistics of Key
Variables}\label{summary-statistics-of-key-variables}}

\vspace{0.7cm}

For the purposes of analysis in this paper, only a subset of variables
was selected. Two additional variable, national and state, was created
using the existing `state' and `end date' variable. A short description
of each variable of interest is given below.

\begin{itemize}
\item
  \emph{Pollster}: The name of the polling organization conducting the
  poll (e.g., YouGov, RMG Research). This variable helps adjust for
  poll-specific biases.
\item
  \emph{Numeric Grade}: A numeric rating given to the pollster,
  representing the quality or reliability of the organization (e.g.,
  3.0), with higher grades indicating more reliable pollsters.
\item
  \emph{Pollscore}: A score that reflects the reliability of each
  pollster, capturing their historical accuracy and any systematic bias.
  Negative values indicate better predictive accuracy.
\item
  \emph{State}: The U.S. state where the poll was conducted, allowing
  for analysis of regional differences in candidate support.
\item
  \emph{National}: A binary variable indicating whether the poll is
  national (1 for national polls, 0 for state polls).
\item
  \emph{End Date}: The date the poll ended, reflecting the currency of
  the data.
\item
  \emph{Sample Size}: The total number of respondents in the poll.
\item
  \emph{Candidate Name}: The name of the candidate being polled (e.g.,
  Kamala Harris or Donald Trump), identifying the focus of each poll
  result.
\item
  \emph{Pct (Percentage)}: The percentage of respondents in the poll who
  support the specified candidate.
\item
  \emph{Recency}: This variable measures how recent each poll is. It was
  calculated by subtracting the end date of the poll from the end date
  of the most recent poll.
\end{itemize}

\newpage

\hypertarget{tbl-sumstat}{}
\begin{table}
\caption{\label{tbl-sumstat}Summary Statistics of Numerical Variable }\tabularnewline

\centering
\begin{tabular}[t]{lrrrrrrr}
\toprule
  & Unique (\#) & Missing (\%) & Mean & SD & Min & Median & Max\\
\midrule
numeric\_grade & 21 & 0 & \num{2.2} & \num{0.6} & \num{1.0} & \num{2.1} & \num{3.0}\\
pollscore & 21 & 0 & \num{-0.5} & \num{0.6} & \num{-1.5} & \num{-0.4} & \num{1.7}\\
sample\_size & 594 & 0 & \num{1908.8} & \num{2602.0} & \num{147.0} & \num{999.0} & \num{20762.0}\\
pct & 216 & 0 & \num{46.9} & \num{4.0} & \num{25.0} & \num{47.0} & \num{70.0}\\
recency & 91 & 0 & \num{43.6} & \num{25.9} & \num{0.0} & \num{41.0} & \num{90.0}\\
\bottomrule
\end{tabular}
\end{table}

\vspace{0.7cm}

From Table~\ref{tbl-sumstat} we see that the average poll in our dataset
has a numeric grade of 2.2, suggesting that most polls are of moderate
to high quality in terms of reliability. The mean pollscore of -0.5
indicates a general trend toward negative values, which suggests these
polls are relatively accurate, as more negative values imply reduced
bias. Additionally, the large average sample size of 1908 respondents
across polls provides a stable basis for our model, reducing variability
and enhancing the reliability of predictions for candidate support.

\newpage

\hypertarget{variation-of-poll-quality-and-support-for-candiates-by-pollster}{%
\subsubsection{Variation of Poll Quality and Support for Candiates by
Pollster}\label{variation-of-poll-quality-and-support-for-candiates-by-pollster}}

\vspace{0.7cm}

\hypertarget{tbl-topster}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3867}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.0800}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2400}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2933}}@{}}
\caption{\label{tbl-topster}Top 5 most frequent pollsters, with count of
polls, average pollscore (lower scores indicate less bias), and average
numeric grade (higher values indicate greater
reliability).}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Pollster
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Count
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Average Pollscore
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Average Numeric Grade
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Pollster
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Count
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Average Pollscore
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Average Numeric Grade
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Morning Consult & 470 & -0.3 & 1.9 \\
Siena/NYT & 188 & -1.5 & 3.0 \\
Redfield \& Wilton Strategies & 176 & 0.4 & 1.8 \\
Emerson & 116 & -1.1 & 2.9 \\
YouGov & 113 & -1.1 & 3.0 \\
\end{longtable}

Table~\ref{tbl-topster} displays the top 5 most frequent pollsters in
the dataset, with each pollster's total poll count, average pollscore
(measuring reliability and historical accuracy), and average numeric
grade (indicating overall quality). We observe that the most frequent
pollster, Morning Consult, has a moderate pollscore and a slightly
above-average numeric grade, indicating it provides fairly reliable
data. However, there is variability in pollscore and numeric grade
across the top pollsters, reflecting differences in quality and
potential biases among them, which the model needs to adjust for to
achieve accurate predictions.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-ster-1.pdf}

}

\caption{\label{fig-ster}Support distribution for Kamala Harris and
Donald Trump by major pollsters, highlighting variability in reported
support across organizations and the value of aggregating multiple polls
for balanced insights.}

\end{figure}

Figure~\ref{fig-ster} shows the distribution of support for Kamala
Harris and Donald Trump by pollster, highlighting variability across
different polling organizations. Morning Consult, for example,
demonstrates a wide range in reported support, with more spread in
Trump's support. In contrast, Siena/NYT shows less variation, with
Harris consistently leading. This variability across pollsters
emphasizes the importance of aggregating multiple polls to account for
organization-specific biases and ensure a more balanced view of
candidate support.

\hypertarget{sample-size-of-polls}{%
\subsubsection{Sample Size of Polls}\label{sample-size-of-polls}}

\vspace{0.7cm}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-ss-1.pdf}

}

\caption{\label{fig-ss}Distribution of Sample Sizes Across Polls:
Majority of polls have sample sizes under 5,000, with a few outliers at
larger sizes.}

\end{figure}

Figure~\ref{fig-ss} reveals a clear right-skewed distribution. Most of
the sample sizes are clustered between 0 and 3000, with a sharp peak
around 1000-1500. This indicates that the majority of polls are
conducted with smaller sample sizes. As sample size increases, the
frequency significantly drops, with very few polls conducted with sample
sizes larger than 5000, though there are a few outliers with sizes
approaching 10,000 or more. This wide range in sample sizes might affect
the precision of estimates in different polls.

\hypertarget{distribution-of-numeric-grade-and-pollscore-of-polls}{%
\subsubsection{Distribution of Numeric Grade and Pollscore of
Polls}\label{distribution-of-numeric-grade-and-pollscore-of-polls}}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-reli-1.pdf}

}

\caption{\label{fig-reli}Distribution of Numeric Grade and Pollscore
among polling organizations, highlighting variability in pollster
reliability and potential bias across polls.}

\end{figure}

Figure~\ref{fig-reli} displays the distribution of Numeric Grade (left)
and Pollscore (right) across polling organizations. The Numeric Grade
distribution shows that most pollsters are rated between 1.5 and 3, with
peaks around 2.0 and 2.5, indicating a concentration of pollsters with
moderate to high reliability scores. In contrast, the Pollscore
distribution, where lower values indicate higher reliability, shows a
range primarily between -1.5 and 0, with a notable peak around -0.5.
This suggests that while many polls demonstrate relatively low bias,
there is still variability in reliability across organizations. The
distinction between these two metrics emphasizes the need to consider
both quality (Numeric Grade) and potential systematic bias (Pollscore)
when weighting polls in the model.

\vspace{0.7cm}

\hypertarget{distribution-of-polls-by-poll-type-and-candidate}{%
\subsubsection{Distribution of Polls by Poll Type and
Candidate}\label{distribution-of-polls-by-poll-type-and-candidate}}

Figure~\ref{fig-pie} charts illustrate the distribution of polls between
candidates and poll types. The left chart shows that the majority of
polls are conducted at the state level, with a smaller portion being
national. The right chart shows that polling is almost evenly split
between Kamala Harris and Donald Trump. This distribution underscores
the model's balanced approach to capturing state-level nuances as well
as broader national trends, providing a comprehensive view of candidate
support across different contexts

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-pie-1.pdf}

}

\caption{\label{fig-pie}Poll distribution by type (state vs.~national)
and candidate (Trump vs.~Harris), showing a majority of state polls and
near-equal coverage for each candidate.}

\end{figure}

\newpage

\hypertarget{support-trend-for-candidates}{%
\subsubsection{Support Trend For
Candidates}\label{support-trend-for-candidates}}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-trend-1.pdf}

}

\caption{\label{fig-trend}Figure showing support trends for Kamala
Harris and Donald Trump over time. Trend lines indicate slight shifts in
support as the election nears, with consistent polling frequency
throughout the period.}

\end{figure}

Figure~\ref{fig-trend} shows the support trends for Kamala Harris and
Donald Trump from August to October. Each point represents a poll
result, color-coded by candidate, with the trend lines highlighting the
overall changes in support over time. Kamala Harris's support remains
relatively steady but shows slight fluctuations around mid-September,
while Donald Trump's support appears to have a small upward trend toward
October. The distribution of points is dense throughout, reflecting
consistent polling activity, though some dates show more concentrated
polling. This visualization indicates that while both candidates
maintain stable support levels, minor shifts occur as the election
approaches, underscoring the importance of tracking trends over time
rather than relying on individual polls.

\newpage

\hypertarget{relationship-between-sample-size-and-suppoert-for-candidates}{%
\subsubsection{Relationship Between Sample Size and Suppoert for
Candidates}\label{relationship-between-sample-size-and-suppoert-for-candidates}}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-pct-1.pdf}

}

\caption{\label{fig-pct}Relationship between Sample Size and Support
Percentage for Donald Trump and Kamala Harris, showing slight downward
and upward trends in support with increasing sample size for Trump and
Harris, respectively.}

\end{figure}

Figure~\ref{fig-pct} shows the relationship between sample size and
support percentage for Donald Trump and Kamala Harris. For both
candidates, the majority of polls have a sample size below 2,500, but
there are some larger polls exceeding 10,000 respondents. In Trump's
chart, there's a slight downward trend, suggesting that larger sample
sizes may show marginally lower support. In contrast, Harris's chart
shows a slight upward trend with larger sample sizes, indicating a
marginal increase in support with larger poll samples. This highlights
the variability in polling support depending on the sample size,
emphasiszing the importance of including sampel size in our model.

\hypertarget{variability-of-candidate-support-across-us-states}{%
\subsubsection{Variability of Candidate support Across US
States}\label{variability-of-candidate-support-across-us-states}}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-map1-1.pdf}

}

\caption{\label{fig-map1}Map showing the proportion of support for
Kamala Harris relative to Donald Trump across U.S. states. Blue
indicates states where Harris has higher support, while red indicates
states where Trump leads. Color intensity reflects the magnitude of
support difference, with gray indicating insufficient polling data.}

\end{figure}

Figure~\ref{fig-map1} shows the average support for Kamala Harris as a
proportion relative to both Kamala Harris and Donald Trump across the
continental U.S. states. States shaded in blue indicate a higher
proportion of support for Harris, while red shades represent stronger
support for Trump, with color intensity reflecting the support margin.
Gray-shaded states lack sufficient polling data.

In key swing states, such as Pennsylvania, Michigan, and Arizona, we
observe a balanced distribution of support, suggesting close competition
in these battleground areas. Meanwhile, traditional Democratic
strongholds, like California and New York, show a clear preference for
Harris, with deeper blue shades. Conversely, traditionally Republican
states, such as Texas and Tennessee, exhibit higher support for Trump,
with intense red shades highlighting his advantage. This visualization
emphasizes the varied regional dynamics of candidate support, with
distinct patterns in both competitive and historically partisan states.

\hypertarget{sec-model}{%
\section{Forecasting Election Outcome through Pooling
Polls}\label{sec-model}}

\hypertarget{forecasting-approach}{%
\subsection{Forecasting Approach}\label{forecasting-approach}}

The polls of polls methodology is widely used in election prediction as
it aggregates multiple polls to provide a more reliable estimate of
voter support, rather than relying on any single poll. The goal is to
reduce errors and biases present in individual polls by using a weighted
average of many different polls.

In our approach, we will employ linear modeling of voter support
percentage (pct) on pollster and other independent variables such as
sample size, poll recency, and poll scope (state vs.~national). This
will allow us to smooth out the inherent noise, biases, and variability
across different pollsters. Once we obtain the predicted values from our
model, we will weight these predictions based on the numeric grade
(quality score) of each pollster to calculate an overall national
estimate of the outcome. Additionally, we will separately compute
estimates for key battleground states, where voter behavior can be more
volatile and pivotal in deciding the final outcome of the election. This
approach helps us capture both national trends and critical state-level
dynamics.

\hypertarget{model}{%
\subsection{Model}\label{model}}

In this section, we aim to address the inherent biases and differences
present in various polling data to arrive at a robust prediction model.
The core challenge lies in selecting a model with an optimal balance
between complexity and fit, ensuring it accurately captures the dynamics
of polling data while avoiding overfitting. To this end, we carefully
evaluated different model specifications to determine the most
appropriate one for our forecasting purpose.

Given that variables like numeric grade and pollscore are perfectly
collinear with pollster, they were excluded from the regression analysis
to avoid multicollinearity issues. These variables, however, remain
integral to our weighting strategy, where they will be used to adjust
for differences in polling accuracy and reliability. Instead, we focus
on key features such as pollster, sample size, state, and recency,
gradually adding complexity to the model.

By systematically comparing model specifications that incorporate these
variables, we aim to select the model with the right balance between
predictive accuracy and generalizability, ultimately providing the best
possible forecast.

\hypertarget{model-set-up}{%
\subsection{Model Set Up}\label{model-set-up}}

We aim to model the percentage of support for Kamala Harris and Donald
Trump in each poll as a function of the pollster the sample size, the
state, and the recency of the poll.

\[
y_i = \alpha + \beta_1 \cdot \mathrm{pollster}_i + \beta_2 \cdot \mathrm{sample\_size}_i + \beta_3 \cdot \mathrm{recency}_i +  \beta_4 \cdot \mathrm{state}_i + \epsilon_i
\]

Where

\begin{itemize}
\item
  \(y_i\)\hspace{0pt} is the percentage of support for candidate in poll
  iii,
\item
  \(α\) is the intercept,
\item
  \(β_1\) captures the effect of the polling organization,
\item
  \(β_2​\) captures the effect of the sample size,
\item
  \(β_3\)\hspace{0pt} captures the effect of recency (how recent the
  poll is),
\item
  \(β_4\) capture the effects of the different states
\item
  \hspace{0pt} \(\epsilon_i\) represents the error term, assumed to
  follow a normal distribution with mean 0.
\end{itemize}

\hypertarget{model-justification}{%
\subsection{Model Justification}\label{model-justification}}

\begin{figure}

{\centering \includegraphics{../other/graphs/polls_of_polls_dag.png}

}

\caption{\label{fig-causal-model}Directed Acyclic Graph illustrating
factors influencing candidate support, with poll attributes (Pollster,
Sample Size, State, Recency) as direct contributors and unobserved
confounders representing potential biases in polling results.}

\end{figure}

In our model, we aim to smooth out discrepancies and biases across
various polling organizations using a polls-of-polls approach. Given the
potential for individual pollsters to introduce systematic
differences---due to variations in sampling methods, question phrasing,
and historical leanings---our model includes a pollster variable to
adjust for these organization-specific biases. This allows us to capture
an aggregated view of public support that is less susceptible to the
idiosyncrasies of any single poll. Furthermore, we incorporate sample
size as a predictor, as polls with larger samples tend to yield more
stable and reliable results, reducing random fluctuations caused by
smaller samples. The state variable accounts for regional political
differences, ensuring that the model captures varying levels of support
across geographic and demographic groups, which is crucial for
understanding the nuanced political landscape. Additionally, recency is
included to prioritize more recent polls, as public opinion can shift
rapidly in response to political events, and recent data is generally
more reflective of current sentiment. By integrating these factors, our
model seeks to produce a more stable and comprehensive measure of
candidate support, reducing noise from individual poll discrepancies and
focusing on a balanced, up-to-date aggregation of polling data.

We opted for a linear model due to its capacity to quantify the marginal
effects of each predictor (pollster, sample size, state, and recency) on
candidate support in a straightforward manner. This structure is
well-suited to our polls-of-polls approach, as it allows for the
estimation of fixed effects that can control for systematic biases
across pollsters, while accommodating the influence of sample size and
recency as continuous variables. All modeling was conducted using the
base R package (R Core Team 2024), specifically utilizing the lm()
function from the stats package for linear regression analysis.

\hypertarget{model-results}{%
\subsection{Model Results}\label{model-results}}

\hypertarget{tbl-summary}{}
\begin{longtable}[t]{llrrrrr}
\caption{\label{tbl-summary}Model performance summary showing improved fit and accuracy as State and
Recency are added, with R² increasing from 0.375 in Model 1 to 0.774 in
Model 3, and RMSE decreasing from 3.053 to 1.836. }Model Summary with Included Variables}\\
\toprule
Model & Variables & R² & Adjusted R² & AIC & BIC & RMSE\\
\midrule
Model 1 & Pollster, Sample Size & 0.375 & 0.320 & 6498.418 & 7026.156 & 3.053\\
Model 2 & Pollster, Sample Size, State & 0.720 & 0.685 & 5574.796 & 6292.110 & 2.043\\
Model 3 & Pollster, Sample Size, State, Recency & 0.774 & 0.746 & 5311.836 & 6034.274 & 1.836\\
\bottomrule
\end{longtable}

Table~\ref{tbl-summary} summarizes the performance metrics for three
models with progressively added variables. Model 1, which includes only
Pollster and Sample Size, achieves an R² of 0.375, indicating that these
variables alone explain about 37.5\% of the variance in candidate
support. Model 2 incorporates State as an additional predictor,
resulting in a substantial improvement, with an R² of 0.720 and a
reduction in both AIC and RMSE, showing better model fit and predictive
accuracy. Model 3 further adds Recency, which increases the R² to 0.774
and decreases the RMSE to 1.836, indicating enhanced explanatory power
and prediction accuracy. This progression highlights the benefit of
adding contextual variables like State and Recency to better capture the
complexities of polling data.

\hypertarget{prediction}{%
\subsection{Prediction}\label{prediction}}

To predict Kamala Harris' overall support, we used a weighted average
approach based on the quality of each poll. The weights are calculated
using each poll's \texttt{numeric\_grade}, which reflects the
reliability and transparency of the polling methodology.

We define the weight for each pollster \(w_i\) as follows:

\[
w_i = \frac{\mathrm{numeric\_grade}_i \times (\mathrm{maxPollscore} - \mathrm{pollscore}_i)}{\sum_{i=1}^{n} \mathrm{numeric\_grade}_i \times (\mathrm{maxPollscore} - \mathrm{pollscore}_i)}
\]

where:

\begin{itemize}
\item
  \(w_i\)\hspace{0pt} represents the weight assigned to poll i,
\item
  \(numericgrade_i\) is the numeric grade of poll i, and
\item
  \(n\) is the total number of polls used in the analysis.
\item
  \(pollscore_i\) is the is the pollscore of poll i which reflects the
  estimated bias of the poll (with more negative values indicating less
  bias)
\item
  \(maxPollscore\) is the maximum pollscore across all polls.
\end{itemize}

The weight assigned to each poll combines both its quality (as
represented by the numeric grade) and its level of bias (as indicated by
the pollscore). Since a more negative pollscore reflects a lower level
of bias, the formula uses the difference between the maximum pollscore
and each poll's specific pollscore. This approach gives more weight to
polls that are both highly graded (indicating higher reliability and
transparency) and less biased.By combining these two factors, the
weighting system emphasizes polls that are reliable and minimally
biased, ensuring that they have a stronger influence in the overall
calculation. Additionally, the total of all weights is normalized to sum
to one, so each poll's weight is proportionate to its quality and
relative lack of bias, resulting in a more balanced and accurate average
of public support.

Using these weights, the overall weighted prediction of candidate's
support is calculated by summing the weighted predicted values from our
regression model:

\[
\text{Overall Weighted Support} = \sum_{i=1}^{n} w_i \cdot \hat{y}_i
\]

\begin{itemize}
\tightlist
\item
  \(\hat{y}_i\)\hspace{0pt} is the predicted percentage of support for
  Kamala Harris from poll i.
\end{itemize}

\hypertarget{comparing-overall-weighted-support-across-all-polls}{%
\subsubsection{Comparing Overall Weighted Support Across All
Polls}\label{comparing-overall-weighted-support-across-all-polls}}

Aggregating support across all polls and applying our weighted approach,
we estimate the overall support for each candidate. The weights,
calculated based on both the poll's numeric grade and poll score, ensure
that higher-quality polls with less bias contribute more to our
estimates. Based on this approach, Kamala Harris has an estimated
overall weighted support of around 47.77\%, while Donald Trump stands at
around 46.31\%. This aggregation provides a comprehensive view, taking
into account the varied methodologies and sampling qualities of
different polling organizations.

\hypertarget{state-level-predictions}{%
\subsubsection{State-Level Predictions}\label{state-level-predictions}}

In this section, we examine the predicted support for Kamala Harris and
Donald Trump within each state, based on our weighted approach. By
aggregating poll results within each state and applying weights that
adjust for poll quality and bias, we can estimate candidate support with
a more localized perspective. This allows us to identify state-by-state
competition and highlight key battleground states where the margins are
close.

For each state, we compare the predicted support percentages and
determine the projected winner based on which candidate has higher
weighted support. Additionally, we calculate the margin of difference
between the candidates, which can provide insights into how close each
state's race is.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-map2-1.pdf}

}

\caption{\label{fig-map2}Predicted 2024 U.S. Presidential Election
winner by state, with Kamala Harris-leaning states in blue, Donald
Trump-leaning states in red, and states lacking sufficient polling data
in gray. The map highlights traditional Democratic and Republican
strongholds as well as key battleground states.}

\end{figure}

Figure~\ref{fig-map2} presents the predicted winner of the 2024 U.S.
Presidential Election by state based on aggregated polling data, with
red representing states projected to favor Kamala Harris and blue
indicating those favoring Donald Trump. Notably, traditional Democratic
strongholds in the Northeast and West Coast, such as California, New
York, and Washington, show solid support for Harris. Conversely,
traditional Republican states, including Texas, Florida, and much of the
South, show strong support for Trump.

Swing states, like Pennsylvania, Michigan, and Wisconsin, are
predominantly leaning toward Harris, indicating a possible advantage for
her in these key battlegrounds, although other typical swing states like
Arizona and Georgia lean toward Trump. Gray-shaded states lack
sufficient polling data to make a prediction, reflecting areas of data
scarcity in the model's projections. This visualization highlights the
geographical and political divides across the U.S., as well as the
critical role swing states play in determining the election outcome.

\hypertarget{tbl-battleground-states}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1705}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2159}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2045}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1932}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2159}}@{}}
\caption{\label{tbl-battleground-states}Predicted support for Kamala
Harris and Donald Trump across key battleground states, indicating the
winner and the support margin (\%) in each state.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
State
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Kamala Support (\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Trump Support (\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Predicted Winner
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Support Margin (\%)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
State
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Kamala Support (\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Trump Support (\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Predicted Winner
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Support Margin (\%)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Arizona & 46.89009 & 48.21778 & Donald Trump & 1.3276921 \\
Georgia & 47.09997 & 48.21968 & Donald Trump & 1.1197167 \\
Michigan & 47.44921 & 46.34289 & Kamala Harris & 1.1063207 \\
Nevada & 47.48644 & 46.47506 & Kamala Harris & 1.0113774 \\
North Carolina & 47.60615 & 47.87192 & Donald Trump & 0.2657668 \\
Pennsylvania & 48.01476 & 46.96981 & Kamala Harris & 1.0449520 \\
Wisconsin & 48.67186 & 46.50662 & Kamala Harris & 2.1652329 \\
\end{longtable}

USAFacts(USA Facts 2024),FiveThirtyEight(FiveThirtyEight 2024) and many
other polling websites are calling Arizona, Georgia, Michigan,
Pennsylvania, Wisconsin, North Carolina, Nevada the `swing states' of
the 2024 presidential election.Polling in swing states is crucial for
election predictions because these states, with their historically close
voting margins, often determine the overall outcome by tipping the
electoral balance toward one candidate.
Table~\ref{tbl-battleground-states} summarizes the predicted support
percentages for Kamala Harris and Donald Trump in key battleground
states, along with the predicted winner and the support margin. The data
shows close margins in these states, with some like Arizona and Georgia
leaning slightly towards Trump, while others such as Michigan and
Pennsylvania favor Harris by narrow margins. The support margin column
highlights the competitiveness of these states, with all margins below
2.2\%, indicating highly contested races.

\hypertarget{sec-discussion}{%
\section{Discussion}\label{sec-discussion}}

In this paper, we set out to predict the support for both Kamala Harris
and Donald Trump in the 2024 U.S. Presidential Election using a
``polls-of-polls'' approach. By aggregating multiple polls, we sought to
mitigate the inherent biases present in individual surveys, creating a
more accurate and balanced forecast across candidates. Our analysis was
driven by a linear regression model using key predictors, such as
pollster, sample size, state, and recency of the polls. After
calculating predicted values from the model, we applied a weighting
scheme based on each pollster's numeric grade and pollscore,
incorporating both reliability and bias. This weighted approach enabled
us to produce an overall prediction of candidate support, reflecting
variations across different states and polling organizations.

One key insight from our analysis is the importance of poll recency.
Including recency notably increased the model's explanatory power, with
improvements in R² and reductions in RMSE compared to models without it.
This underscores the dynamic nature of public opinion, where recent
events can influence voter perceptions and candidate support.

Our state-level analysis highlights close competition observed in key
battleground states, emphasizing their pivotal role in the 2024
election. Our model reveals that in states like Pennsylvania, Michigan,
and Wisconsin, support levels for Harris and Trump are nearly tied,
reflecting the intense contest for these critical electoral votes. The
predictions show that even minor shifts in support within these swing
states could decisively impact the overall election outcome. This
analysis shows the importance of regional polling in capturing nuanced
voter sentiment and the heightened influence that battleground states
hold in shaping the final results.

However, there are limitations to our approach. Our model assumes linear
relationships between predictors and candidate support, which might not
capture more complex or interaction-based trends. Additionally, while we
weighted polls based on their numeric grades, these scores may not fully
reflect each pollster's accuracy, leaving room for improvement. Future
research could explore non-linear methods, such as machine learning, to
capture potential interactions between variables. Refining the weighting
mechanism by considering pollster performance history or state-specific
polling nuances could further enhance predictive accuracy.

In conclusion, this paper contributes to election forecasting by
providing a comprehensive, weighted estimate of candidate support across
both national and state levels. While our predictions provide valuable
insights into potential election outcomes, the evolving nature of voter
sentiment and poll accuracy reflects the need for adaptive models that
account for ongoing changes in public opinion.

\newpage

\appendix

\hypertarget{appendix}{%
\section*{Appendix}\label{appendix}}
\addcontentsline{toc}{section}{Appendix}

\hypertarget{emerson-college-polling-methodology-october-14-16-2024}{%
\section{Emerson College Polling Methodology (October 14-16,
2024)}\label{emerson-college-polling-methodology-october-14-16-2024}}

\hypertarget{overview}{%
\subsection{Overview}\label{overview}}

Emerson College Polling, also known as ECP, is a nationally-ranked
polling center that aims to accurately reflect public opinion through a
mixed-mode methodology (Emerson College Polling 2024a). ECP conducted a
survey from October 14 to 16, 2024, targeting 1,000 likely voters in the
2024 U.S. Presidential Election. The poll measured voter preferences
between Kamala Harris and Donald Trump, revealing a near tie with Harris
at 50\% and Trump at 49\% (Emerson College Polling 2024b). Appendix A
takes a closer look at the methodology used by ECP for this poll.

\hypertarget{population-frame-and-sample-of-the-poll}{%
\subsection{Population, Frame, and Sample of the
Poll}\label{population-frame-and-sample-of-the-poll}}

To conduct a poll, it is important to first define the target
population, the sampling frame and the sample. Defining these elements
sets a clear direction for the research and assures a clear object of
research (Akman 2023).

The target population is the whole group of interest the researchers
would like to make conclusions on (University of Massachusetts 2022).
For the poll conducted by ECP in October, the target population is
likely voters in the U.S. elections (Emerson College Polling 2024b).
Choosing likely voters as the target population makes the poll results
more accurate as not all registered voters end up voting in the
Presidential Election. In fact, the last U.S. Presidential Election in
2020 only had 66.8\% of registered voters participate in the elections,
which is the highest voter turnout in the 21st century (U.S. Census
Bureau 2021). In ECP's methodology, likely voters are determined by a
combination of voter history, registration status, and demographic data.
These factors are self-reported (Emerson College Polling 2024b).

The sampling frame is the part inside of the target population that has
a chance to be sampled upon (University of Massachusetts 2022). In ECP's
methodology, the sampling frame is the voters that are on Aristotle's
database and on the online panel provided by CINT (Emerson College
Polling 2024a). Aristotle is an online source of voter and consumer data
that has serviced many large political campaigns, PACs and corporations
both within the U.S and abroad (Aristotle 2024). CINT is the largest
global research marketplace that connects researchers, such as Emerson,
to survey respondents (CINT 2024).

The sample of a study is the part of the sampling frame that is measured
and is part of the data set (University of Massachusetts 2022). In the
case of a poll like ECP's, the sample is the respondents of the survey.
These respondents are directly represented in the data set (Emerson
College Polling 2024b). The sample size is 1,000 likely voters chosen
arbitrarily from the sampling frame.

\hypertarget{sample-recruitment-and-approach}{%
\subsection{Sample Recruitment and
Approach}\label{sample-recruitment-and-approach}}

Emerson College used a mixed-mode sampling approach for its polling
(Emerson College Polling 2024a). Specifically, three primary modes were
used to collect data for its October 14-16, 2024 poll (Emerson College
Polling 2024b) :

\begin{itemize}
\item
  MMS-to-web text surveys: Respondents were contacted via text messages
  sent to cell phones. These messages directed participants to complete
  the survey online. This method was conducted using voter lists
  provided by Aristotle.
\item
  Interactive Voice Response (IVR): Landlines were targeted using
  automated phone calls where respondents could answer questions using
  their keypads. The contact information for this was also sourced from
  Aristotle's voter lists.
\item
  Online Panel from CINT: Emerson utilized a pre-screened, opt-in online
  panel of voters provided by CINT.
\end{itemize}

\hypertarget{advantages-and-trade-offs-in-mixed-mode-sampling}{%
\subsection{Advantages and Trade-offs in Mixed-Mode
Sampling}\label{advantages-and-trade-offs-in-mixed-mode-sampling}}

The multi-mode sampling approach used by ECP in this poll has many
advantages. By using multiple data collection methods, ECP reduces
coverage bias (Mora 2011). In a survey, coverage bias means there is
some part of the target population that has zero chance of being part of
the sample and will not be represented in the data (Cummins 2021). For
example, MMS-to-web surveys tend to attract responses from younger
voters and individuals who rely primarily on mobile devices, capturing a
segment that may otherwise be underrepresented. Interactive voice
response (IVR) calls target older and rural voters, who are more
inclined to use landlines. Online panels gather responses from
tech-savvy individuals who prefer online engagement over other methods.
Multi-mode sampling is also more cost-effective (Mora 2011). Automated
technologies such as MMS-to-text web surveys and IVR reduce expenses
compared to traditional live interviews. Lower costs allow the poll to
reach a larger sample size, which also results in more accurate and
representative conclusions (Charlesworth Author Services 2022). However,
using a multi-mode sampling approach also introduces a trade-off. This
sampling approach introduces a measurement error (Mora 2011). Each
method---whether IVR, web surveys, or online panels---brings its own
unique biases. Combining these creates inconsistencies between
respondent groups. This increases the overall margin of error by adding
variability.

\hypertarget{non-response-handling}{%
\subsection{Non-response Handling}\label{non-response-handling}}

ECP addresses non-response using a weighting system. The data is
adjusted based on demographic variables like age, gender, race,
education, and party affiliation, making the sample more representative
of the electorate (Emerson College Polling 2024a). Weights are applied
to even out under- and over-represented groups, making the data more
similar to the electorate (Elliott 2020).

\hypertarget{questionnaire-design}{%
\subsection{Questionnaire Design}\label{questionnaire-design}}

The Emerson survey is designed with simplicity and efficiency. One of
its strengths is its straightforward format, using clear and unambiguous
questions. This simplicity makes the survey easy for respondents to
understand. Another strength is its topical relevance. By focusing on
vote preferences and demographic splits, the survey provides timely
insights, especially valuable during election cycles. However, the
survey has some limitations. One is its lack of nuance; while it
captures basic voter preferences, it doesn't explore the motivations or
factors driving those choices in depth. Additionally, the various survey
modes can affect respondent engagement with the questions, which may
also influence the accuracy and depth of their answers.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

Emerson College Polling's mixed-mode methodology is effective in
balancing cost with broad demographic reach. Its weighting techniques
contribute to reliable and accurate results despite non-responses. While
the use of multiple survey modes reduces coverage bias, other possible
biases---such as lack of nuance and combined measurement errors must be
considered during the analysis.

\hypertarget{idealized-methodology}{%
\section{Idealized Methodology}\label{idealized-methodology}}

\hypertarget{overview-1}{%
\subsection{Overview}\label{overview-1}}

This methodology outlines an election forecasting plan with a budget of
\$100,000. Heavily inspired by TIPP Insights and Emerson College's
hybrid approaches, the survey methodology balances cost-efficiency and
data accuracy by using different modes of data collection and wide
demographic reach.

This forecasting plan will predict the popular vote and Electoral
College outcomes, emphasizing key swing states and underrepresented
voter groups.

\hypertarget{sampling-approach}{%
\subsection{Sampling Approach}\label{sampling-approach}}

This idealized methodology uses a multi-mode hybrid sampling approach.
This allows the use of both probability-based and non-probability-based
methods:

\begin{itemize}
\item
  Probability-based methods:

  \begin{itemize}
  \item
    Live phone interviews (landline and cell): As used by TIPP, live
    phone interviews is a direct and personal method of collecting
    responses. Voter lists would be sourced from Aristotle to ensure we
    contact likely voters. This mode reaches older voters and voters
    with less internet connection.
  \item
    Interactive Voice Response (IVR): Similar to Emerson's approach, IVR
    for landline users balances costs and ensures demographic
    representation, particularly among older or less tech-savvy
    populations. This method captures quick responses and is less
    expensive than live interviews.
  \end{itemize}
\item
  Non-probability methods:

  \begin{itemize}
  \item
    MMS-to-web (mobile): Following Emerson's MMS-to-web strategy, we
    would send multimedia messages to likely voters' cell phones,
    encouraging them to complete an online survey. This method works
    well for younger, mobile-first voters, helping fill in gaps where
    traditional methods might miss.
  \item
    Online panel recruitment via CINT: To mirror TIPP's online panel
    approach, a portion of the survey would be conducted via an opt-in
    panel of pre-screened voters from CINT, capturing tech-savvy
    respondents. CINT's platform ensures that different voter groups,
    including hard-to-reach groups, are represented in the polling.
  \end{itemize}
\end{itemize}

\hypertarget{recruitment-strategy}{%
\subsection{Recruitment Strategy}\label{recruitment-strategy}}

By using Aristotle's voter lists and CINT's platform as a sampling
frame, the sample consists of likely voters based on registration status
and past voting behavior. Quota sampling would be used to match the
demographic profile of the U.S. electorate. This ensures that
respondents in the sample reflect key demographic variables such as age,
gender, race, and education level.

We would oversample key swing states like Pennsylvania, Wisconsin, and
Florida. Key swing states are states where there is no overwhelming
support for a particular candidate or political party. The outcome in
these states often determine the overall result of the election, making
them critical in securing a win in the Electoral College. Oversampling
these states allows a more accurate forecasting in these battlegrounds,
which are critical for predicting the Electoral College outcome.

\hypertarget{data-validation}{%
\subsection{Data Validation}\label{data-validation}}

Given that the data is collected in a hybrid format, we would use
duplicate detection methods via IP addresses or phone numbers to
eliminate multiple responses from the same individual. Also, we would
implement attention checks to identify respondents who are not fully
engaged. Both duplicate respondents and unengaged respondents bias
results, and should be removed from the data set.

Following TIPP's iterative weighting process, the data will be weighted
to ensure that the sample aligns with key demographics (race, gender,
education) based on census data.

\hypertarget{conclusion-1}{%
\subsection{Conclusion}\label{conclusion-1}}

By combining TIPP's live phone/IVR hybrid and Emerson College's
mixed-mode survey methods, this idealized methodology captures diverse
voter demographics. It uses both probability-based and
non-probability-based sampling methods to forecast the 2024 U.S.
Presidential Election.

A link to the survey can be found at:
https://forms.gle/haGLaQBPQhrXE6Xa7

\hypertarget{copy-of-the-survey}{%
\section{Copy of the Survey}\label{copy-of-the-survey}}

\includepdf[pages=-]{../other/surveyfiles/SurveyCopy.pdf}

\hypertarget{model-diagnostics}{%
\section{Model Diagnostics}\label{model-diagnostics}}

\begin{figure}

{\centering \includegraphics{../other/graphs/diagnostic_plot_model3.png}

}

\caption{\label{fig-dgplotone}\textbf{?(caption)}}

\end{figure}

The diagnostic plots for Model 3ni sgiven in Figure~\ref{fig-dgplotone}.
We notice the following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Residuals vs.~Fitted Plot}: This plot checks for non-linearity
  and heteroscedasticity. The residuals are scattered around the
  horizontal axis with no clear pattern, indicating that the linearity
  assumption is reasonable. However, there is slight spread around the
  center, suggesting some mild heteroscedasticity.
\item
  \textbf{Normal Q-Q Plot}: This plot assesses the normality of
  residuals. Most points align along the diagonal line, although there
  are some deviations at the tails. This suggests that the residuals are
  approximately normally distributed, with minor deviations in the
  extremes.
\item
  \textbf{Scale-Location Plot}: This plot further checks for
  homoscedasticity. The residuals appear to be evenly spread across the
  fitted values, supporting the homoscedasticity assumption, although
  slight deviations are present in certain regions.
\item
  \textbf{Residuals vs.~Leverage Plot}: This plot identifies potential
  influential points. While most points have low leverage, a few points
  exhibit higher leverage, as indicated by their distance from the
  center. However, no points exceed Cook's distance threshold,
  indicating no extreme outliers that would unduly influence the model.
\end{enumerate}

\begin{figure}

{\centering \includegraphics{../other/graphs/diagnostic_plot_model6.png}

}

\caption{\label{fig-dgplottwo}\textbf{?(caption)}}

\end{figure}

The diagnostic plots for Model 6 is given in Figure~\ref{fig-dgplottwo}.
We notice the following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Residuals vs Fitted}: This plot suggests a fairly random
  spread of residuals around zero, indicating that the linearity
  assumption holds reasonably well. However, there is slight clustering
  of points in the center, which could indicate minor heteroscedasticity
  but is not severe.
\item
  \textbf{Normal Q-Q Plot}: The Q-Q plot shows that the residuals
  generally follow a normal distribution, with only a few deviations at
  the tails. This suggests that the normality assumption is mostly met,
  though some minor deviations in the upper tail indicate possible
  outliers.
\item
  \textbf{Scale-Location (Spread-Location) Plot}: The residuals appear
  randomly dispersed with no clear pattern, supporting the
  homoscedasticity (constant variance) assumption. However, some
  observations near the upper range might indicate slight variance
  inconsistency, though it is minimal.
\item
  \textbf{Residuals vs Leverage}: This plot does not show any
  influential outliers with high leverage that might impact the model
  unduly. The Cook's distance lines show that no data points exceed
  these thresholds, indicating that no individual observation is
  disproportionately influencing the model.
\end{enumerate}

Overall, the diagnostic plots for both model suggests that they meet the
assumptions for linear regression fairly well, with minor deviations
that are not expected to severely impact the model's validity.

\hypertarget{code-styling}{%
\section{Code Styling}\label{code-styling}}

The code in this paper was reviewed and formatted for consistency using
lintr (Hester et al. 2024) and styler (Müller and Walthert 2024),
ensuring readability and adherence to style standards.

\hypertarget{reproducibility}{%
\section{Reproducibility}\label{reproducibility}}

To replicate the findings presented in this paper, users should execute
the scripts available in the GitHub repository. Begin by running the
00-install\_packages.R script, which installs all required packages for
the analysis.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

We extend our gratitude to (Alexander 2023), which provided invaluable
guidance in establishing a reproducible workflow and inspired many of
the code structures used in this paper.

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-akman}{}}%
Akman, Sena. 2023. \emph{What Is Target Population: Definition \&
Examples}. \url{https://forms.app/en/blog/target-population}.

\leavevmode\vadjust pre{\hypertarget{ref-citeTS}{}}%
Alexander, Rohan. 2023. {``Telling Stories with Data.''} Chapman;
Hall/CRC. \url{https://tellingstorieswithdata.com/}.

\leavevmode\vadjust pre{\hypertarget{ref-aristotle}{}}%
Aristotle. 2024. \emph{Data}. \url{https://www.aristotle.com/data/}.

\leavevmode\vadjust pre{\hypertarget{ref-large_ss}{}}%
Charlesworth Author Services. 2022. \emph{The Importance of Having Large
Sample Sizes for Your Research}.
\url{https://www.cwauthors.com/article/importance-of-having-large-sample-sizes-for-research}.

\leavevmode\vadjust pre{\hypertarget{ref-cint}{}}%
CINT. 2024. \emph{Cint}. \url{https://www.cint.com/}.

\leavevmode\vadjust pre{\hypertarget{ref-def_cb}{}}%
Cummins, Emily. 2021. \emph{Coverage Bias: Definition \& Examples}.
\url{https://study.com/academy/lesson/coverage-bias-definition-examples.html}.

\leavevmode\vadjust pre{\hypertarget{ref-weight}{}}%
Elliott, Roxana. 2020. \emph{Weighting Survey Data: Methods and
Advantages}.
\url{https://www.geopoll.com/blog/weighting-survey-data-raking-cell-weighting/}.

\leavevmode\vadjust pre{\hypertarget{ref-about_ecp}{}}%
Emerson College Polling. 2024a. \emph{About Us}.
\url{https://emersoncollegepolling.com/about/}.

\leavevmode\vadjust pre{\hypertarget{ref-ecp_poll}{}}%
---------. 2024b. \emph{October 2024 Tracking National Poll: Harris 49}.
\url{https://emersoncollegepolling.com/october-2024-tracking-national-poll-harris-49-trump-48/}.

\leavevmode\vadjust pre{\hypertarget{ref-janitor}{}}%
Firke, Sam. 2023. \emph{Janitor: Simple Tools for Examining and Cleaning
Dirty Data}. \url{https://CRAN.R-project.org/package=janitor}.

\leavevmode\vadjust pre{\hypertarget{ref-FiveThirtyEight}{}}%
FiveThirtyEight. 2024. {``{FiveThirtyEight U.S. Election Polls}.''}
\url{https://projects.fivethirtyeight.com/polls/}.

\leavevmode\vadjust pre{\hypertarget{ref-lubridate}{}}%
Grolemund, Garrett, and Hadley Wickham. 2011. {``Dates and Times Made
Easy with {lubridate}.''} \emph{Journal of Statistical Software} 40 (3):
1--25. \url{https://www.jstatsoft.org/v40/i03/}.

\leavevmode\vadjust pre{\hypertarget{ref-citelintr}{}}%
Hester, Jim, Florent Angly, Russ Hyde, Michael Chirico, Kun Ren,
Alexander Rosenstock, and Indrajeet Patil. 2024. \emph{Lintr: A 'Linter'
for r Code}. \url{https://CRAN.R-project.org/package=lintr}.

\leavevmode\vadjust pre{\hypertarget{ref-mora}{}}%
Mora, Michaela. 2011. \emph{Understanding the Pros and Cons of
Mixed-Mode Research}.
\url{https://www.relevantinsights.com/articles/pros-and-cons-of-mixed-mode-research/}.

\leavevmode\vadjust pre{\hypertarget{ref-citestyler}{}}%
Müller, Kirill, and Lorenz Walthert. 2024. \emph{Styler: Non-Invasive
Pretty Printing of r Code}.
\url{https://CRAN.R-project.org/package=styler}.

\leavevmode\vadjust pre{\hypertarget{ref-citeR}{}}%
R Core Team. 2024. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-arrow}{}}%
Richardson, Neal, Ian Cook, Nic Crane, Dewey Dunnington, Romain
François, Jonathan Keane, Dragoș Moldovan-Grünfeld, Jeroen Ooms, Jacob
Wujciak-Jens, and Apache Arrow. 2024. \emph{Arrow: Integration to
'Apache' 'Arrow'}. \url{https://CRAN.R-project.org/package=arrow}.

\leavevmode\vadjust pre{\hypertarget{ref-def_terms}{}}%
University of Massachusetts. 2022. \emph{Populations and Samples}.
\url{https://people.umass.edu/~biep540w/pdf/sampling.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-us_census}{}}%
U.S. Census Bureau. 2021. \emph{2020 Presidential Election Voting and
Registration Tables Now Available}.
\url{https://www.census.gov/newsroom/press-releases/2021/2020-presidential-election-voting-and-registration-tables-now-available.html}.

\leavevmode\vadjust pre{\hypertarget{ref-usafacts}{}}%
USA Facts. 2024. {``{What Are the Current Swing States and How Have They
Changed Over Time?}''}
\url{https://usafacts.org/articles/what-are-the-current-swing-states-and-how-have-they-changed-over-time/}.

\leavevmode\vadjust pre{\hypertarget{ref-tidy}{}}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\end{CSLReferences}

\end{document}

### 2024-11-01 paper/paper.qmd
cleaning files

@@
 bibliography: references.bib
-
 ---
@@
 
-This paper is organized as follows: In {#sec-data}, we clean the data, explore summary statistics, plot distributions of key variables, and examine relationships between variables.
-In {#sec-model}, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions.
-Finally, in {#sec-discussion}, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work.
-\vspace{0.7cm}
-# Data {#sec-data}
-\vspace{0.7cm}
-## Measurement
+This paper is organized as follows: In \ref{sec-data}, we clean the data, explore summary statistics, plot distributions of key variables, and examine relationships between variables.
+In \ref{sec-model}, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions.
+Finally, in \ref{sec-discussion}, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work.
+\vspace{0.7cm} \# Data {#sec-data} \vspace{0.7cm} \## Measurement
 
@@
 This methodology outlines an election forecasting plan with a budget of \$100,000.
-Heavily inspired by TIPP Insights and Emerson College's hybrid approaches, the survey methodology balances cost-efficiency and data accuracy by using different modes of data collection and wide demographic reach.
-
-This forecasting plan will predict the popular vote and Electoral College outcomes, emphasizing key swing states and underrepresented voter groups.
+This survey is heavily inspired by TIPP Insights and Emerson College Polling (ECP), which are two respected polling institutions.
+TIPP Insights is known for accurately predicting major U.S.
+elections, including presidential races, by combining traditional live phone interviews with probability-based sampling [@tipp_accuracy].
+Their history of reliable forecasts has made TIPP a trusted name in polling, that specializes in capturing voter sentiment and representing diverse demographics.
+Similarly, Emerson College is known for its mixed-mode polling strategy, which combines Interactive Voice Response (IVR) with online polling to reach both older, landline users and younger, mobile-first voters [@about_ecp].
+
+By incorporating techniques from both of these institutions, our methodology aims to achieve high accuracy and broad demographic reach, ensuring a balanced, cost-effective approach to election forecasting.
 
@@
 
-This idealized methodology uses a multi-mode hybrid sampling approach. This allows the use of both probability-based and non-probability-based methods:
-
--   Probability-based methods:
-
-    -   Live phone interviews (landline and cell): As used by TIPP, live phone interviews is a direct and personal method of collecting responses.
-Voter lists would be sourced from Aristotle to ensure we contact likely voters.
-This mode reaches older voters and voters with less internet connection.
-
-    -   Interactive Voice Response (IVR): Similar to Emerson’s approach, IVR for landline users balances costs and ensures demographic representation, particularly among older or less tech-savvy populations.
-This method captures quick responses and is less expensive than live interviews.
-
--   Non-probability methods:
-
-    -   MMS-to-web (mobile): Following Emerson’s MMS-to-web strategy, we would send multimedia messages to likely voters’ cell phones, encouraging them to complete an online survey.
-This method works well for younger, mobile-first voters, helping fill in gaps where traditional methods might miss.
-
-    -   Online panel recruitment via CINT: To mirror TIPP’s online panel approach, a portion of the survey would be conducted via an opt-in panel of pre-screened voters from CINT, capturing tech-savvy respondents.
-CINT’s platform ensures that different voter groups, including hard-to-reach groups, are represented in the polling.
+This idealized methodology employs a multi-mode hybrid sampling approach, utilizing both probability-based and non-probability-based methods.
+Combining both methods provides both accuracy and inclusivity in polling.
+Overall, this method creates a more complete view of voter preferences [@baker].
+
+Probability-based methods include live phone interviews (landline and cell), as used by TIPP.
+Live interviews are a direct, personal method of collecting responses, with voter lists sourced from Aristotle [@tipp_methodology].
+This method is effective in reaching older voters and voters with less internet connectivity.
+Interactive Voice Response (IVR) is also used, following Emerson’s strategy for landline users [@about_ecp], to balance costs and ensure representation of less tech-savvy populations.
+
+For non-probability methods, this approach includes MMS-to-web surveys for mobile users, inspired by Emerson’s MMS-to-web strategy [@about_ecp].
+Likely voters receive multimedia messages on their cell phones encouraging them to complete an online survey, which helps reach younger, mobile-first voters.
+Additionally, online panel recruitment is done through CINT.
+This follows TIPP's online panel strategy, where a part of the survey is delivered to a pre-screened panel [@tipp_methodology].
+
+The combination of these sampling methods and collection methods ensures a more accurate representation of voter groups, including harder-to-reach populations.
 
@@
 
-By using Aristotle’s voter lists and CINT's platform as a sampling frame, the sample consists of likely voters based on registration status and past voting behavior.
+By using Aristotle’s voter lists and CINT's platform as a sampling frame, the sample consists of likely voters.
+Likely voters are determined based on registration status and past voting behavior [@ecp_poll].
 Quota sampling would be used to match the demographic profile of the U.S.
 electorate.
-This ensures that respondents in the sample reflect key demographic variables such as age, gender, race, and education level.
-
-We would oversample key swing states like Pennsylvania, Wisconsin, and Florida.
-Key swing states are states where there is no overwhelming support for a particular candidate or political party.
+Quota sampling is a sampling method that relies on a non-random selection of units based on predetermined proportions [@def_quota].
+This ensures that respondents in the sample reflect demographic variables such as age, gender, race, and education level.
+
+We would oversample key swing states like Pennsylvania, Arizona, and Georgia [@swing_state].
+Key swing states are states where there is no overwhelming support for a particular candidate or political party [@barker].
 The outcome in these states often determine the overall result of the election, making them critical in securing a win in the Electoral College.
@@
 
-Following TIPP’s iterative weighting process, the data will be weighted to ensure that the sample aligns with key demographics (race, gender, education) based on census data.
+The data will be weighted to ensure that the sample aligns with key demographics (race, gender, education) based on census data.
+Weights are applied to even out under- and over-represented groups, making the data more similar to the electorate [@weight].
 

### 2024-11-01 paper/paper.tex
cleaning files

@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, breakable, boxrule=0pt, sharp corners, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 
-This paper is organized as follows: In \{\#sec-data\}, we clean the
+This paper is organized as follows: In \ref{sec-data}, we clean the
 data, explore summary statistics, plot distributions of key variables,
-and examine relationships between variables. In \{\#sec-model\}, we
+and examine relationships between variables. In \ref{sec-model}, we
 discuss our forecasting approach, model selection, justification for the
@@
 pollster reliability, ultimately presenting our predictions. Finally, in
-\{\#sec-discussion\}, we address the broader implications of our
+\ref{sec-discussion}, we address the broader implications of our
 findings, acknowledge limitations, and suggest directions for future
@@
 This methodology outlines an election forecasting plan with a budget of
-\$100,000. Heavily inspired by TIPP Insights and Emerson College's
-hybrid approaches, the survey methodology balances cost-efficiency and
-data accuracy by using different modes of data collection and wide
-demographic reach.
-
-This forecasting plan will predict the popular vote and Electoral
-College outcomes, emphasizing key swing states and underrepresented
-voter groups.
+\$100,000. This survey is heavily inspired by TIPP Insights and Emerson
+College Polling (ECP), which are two respected polling institutions.
+TIPP Insights is known for accurately predicting major U.S. elections,
+including presidential races, by combining traditional live phone
+interviews with probability-based sampling (Rakich 2023). Their history
+of reliable forecasts has made TIPP a trusted name in polling, that
+specializes in capturing voter sentiment and representing diverse
+demographics. Similarly, Emerson College is known for its mixed-mode
+polling strategy, which combines Interactive Voice Response (IVR) with
+online polling to reach both older, landline users and younger,
+mobile-first voters (Emerson College Polling 2024a).
+
+By incorporating techniques from both of these institutions, our
+methodology aims to achieve high accuracy and broad demographic reach,
+ensuring a balanced, cost-effective approach to election forecasting.
 
@@
 
-This idealized methodology uses a multi-mode hybrid sampling approach.
-This allows the use of both probability-based and non-probability-based
-methods:
-
-\begin{itemize}
-\item
-  Probability-based methods:
-
-  \begin{itemize}
-  \item
-    Live phone interviews (landline and cell): As used by TIPP, live
-    phone interviews is a direct and personal method of collecting
-    responses. Voter lists would be sourced from Aristotle to ensure we
-    contact likely voters. This mode reaches older voters and voters
-    with less internet connection.
-  \item
-    Interactive Voice Response (IVR): Similar to Emerson's approach, IVR
-    for landline users balances costs and ensures demographic
-    representation, particularly among older or less tech-savvy
-    populations. This method captures quick responses and is less
-    expensive than live interviews.
-  \end{itemize}
-\item
-  Non-probability methods:
-
-  \begin{itemize}
-  \item
-    MMS-to-web (mobile): Following Emerson's MMS-to-web strategy, we
-    would send multimedia messages to likely voters' cell phones,
-    encouraging them to complete an online survey. This method works
-    well for younger, mobile-first voters, helping fill in gaps where
-    traditional methods might miss.
-  \item
-    Online panel recruitment via CINT: To mirror TIPP's online panel
-    approach, a portion of the survey would be conducted via an opt-in
-    panel of pre-screened voters from CINT, capturing tech-savvy
-    respondents. CINT's platform ensures that different voter groups,
-    including hard-to-reach groups, are represented in the polling.
-  \end{itemize}
-\end{itemize}
+This idealized methodology employs a multi-mode hybrid sampling
+approach, utilizing both probability-based and non-probability-based
+methods. Combining both methods provides both accuracy and inclusivity
+in polling. Overall, this method creates a more complete view of voter
+preferences (Baker et al. 2013).
+
+Probability-based methods include live phone interviews (landline and
+cell), as used by TIPP. Live interviews are a direct, personal method of
+collecting responses, with voter lists sourced from Aristotle (TIPP
+Staff 2023). This method is effective in reaching older voters and
+voters with less internet connectivity. Interactive Voice Response (IVR)
+is also used, following Emerson's strategy for landline users (Emerson
+College Polling 2024a), to balance costs and ensure representation of
+less tech-savvy populations.
+
+For non-probability methods, this approach includes MMS-to-web surveys
+for mobile users, inspired by Emerson's MMS-to-web strategy (Emerson
+College Polling 2024a). Likely voters receive multimedia messages on
+their cell phones encouraging them to complete an online survey, which
+helps reach younger, mobile-first voters. Additionally, online panel
+recruitment is done through CINT. This follows TIPP's online panel
+strategy, where a part of the survey is delivered to a pre-screened
+panel (TIPP Staff 2023).
+
+The combination of these sampling methods and collection methods ensures
+a more accurate representation of voter groups, including
+harder-to-reach populations.
 
@@
 By using Aristotle's voter lists and CINT's platform as a sampling
-frame, the sample consists of likely voters based on registration status
-and past voting behavior. Quota sampling would be used to match the
-demographic profile of the U.S. electorate. This ensures that
-respondents in the sample reflect key demographic variables such as age,
+frame, the sample consists of likely voters. Likely voters are
+determined based on registration status and past voting behavior
+(Emerson College Polling 2024b). Quota sampling would be used to match
+the demographic profile of the U.S. electorate. Quota sampling is a
+sampling method that relies on a non-random selection of units based on
+predetermined proportions (Nikolopoulou 2022). This ensures that
+respondents in the sample reflect demographic variables such as age,
 gender, race, and education level.
 
-We would oversample key swing states like Pennsylvania, Wisconsin, and
-Florida. Key swing states are states where there is no overwhelming
-support for a particular candidate or political party. The outcome in
-these states often determine the overall result of the election, making
-them critical in securing a win in the Electoral College. Oversampling
-these states allows a more accurate forecasting in these battlegrounds,
-which are critical for predicting the Electoral College outcome.
+We would oversample key swing states like Pennsylvania, Arizona, and
+Georgia (FitzGerald 2024). Key swing states are states where there is no
+overwhelming support for a particular candidate or political party
+(Barket 2020). The outcome in these states often determine the overall
+result of the election, making them critical in securing a win in the
+Electoral College. Oversampling these states allows a more accurate
+forecasting in these battlegrounds, which are critical for predicting
+the Electoral College outcome.
 
@@
 
-Following TIPP's iterative weighting process, the data will be weighted
-to ensure that the sample aligns with key demographics (race, gender,
-education) based on census data.
+The data will be weighted to ensure that the sample aligns with key
+demographics (race, gender, education) based on census data. Weights are
+applied to even out under- and over-represented groups, making the data
+more similar to the electorate (Elliott 2020).
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-baker}{}}%
+Baker, Reg, Michael Brick, Nancy Bates, Mike Battaglia, Mick Couper,
+Jill Dever, Krista Gile, and Roger Tourangeau. 2013. \emph{REPORT OF THE
+AAPOR TASK FORCE ON NONPROBABILITY SAMPLING}.
+\url{https://aapor.org/wp-content/uploads/2022/11/NPS_TF_Report_Final_7_revised_FNL_6_22_13-2.pdf}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-barker}{}}%
+Barket, Elaine. 2020. \emph{Recommendation for Key Management: Part 1 --
+General}.
+\url{https://nvlpubs.nist.gov/nistpubs/SpecialPublications/NIST.SP.800-57pt1r5.pdf}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-large_ss}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-swing_state}{}}%
+FitzGerald, James. 2024. \emph{The Seven Swing States Set to Decide the
+2024 US Election}. \url{https://www.bbc.com/news/articles/c511pyn3xw3o}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-FiveThirtyEight}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-def_quota}{}}%
+Nikolopoulou, Kassiani. 2022. \emph{What Is Quota Sampling? \textbar{}
+Definition \& Examples}.
+\url{https://www.scribbr.com/methodology/quota-sampling/}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-citeR}{}}%
@@
 \url{https://www.R-project.org/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-tipp_accuracy}{}}%
+Rakich, Nathaniel. 2023. \emph{The Polls Were Historically Accurate in
+2022}.
+\url{https://fivethirtyeight.com/features/2022-election-polling-accuracy/}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-tipp_methodology}{}}%
+TIPP Staff. 2023. \emph{TIPP Indexes Methodology}.
+\url{https://tippinsights.com/tipp-indexes-methodology/}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-def_terms}{}}%

### 2024-11-01 paper/paper.qmd
include measurement of variables and references

@@
 Finally, in \ref{sec-discussion}, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work.
-\vspace{0.7cm} \# Data {#sec-data} \vspace{0.7cm} \## Measurement
-
-In the dataset of our analysis, the process of measurement begins by capturing real-world public opinion through surveys.
-Polling organizations transform this observed phenomenon into structured data by recording support percentages, pollster details, and methodological choices, thus turning public sentiment into quantifiable entries in the dataset.
-
-Public opinion polls are essential for understanding voter preferences and electoral dynamics, providing data that informs policy decisions and campaign strategies.
-Our data set captures public support for candidates at specific times, along with information on methodologies, sample sizes, and pollster ratings.
-
-Polling organizations use various methods, such as App panels, IVR (Interactive Voice Response), Live phone calls, Text-to-web systems etc.
-These methods can influence the validity and reliability of results, with larger sample sizes generally yielding more reliable estimates.
-Some methods, like live phone interviews, may be seen as more trustworthy than others.
-
-The dataset also includes important measures of pollster performance, such as numeric grades, poll scores, and transparency scores.
-These metrics are calculated by considering bias, race difficulty, predictive error, and transparency, allowing us to assess the accuracy of each poll.
-The "pollscore," for instance, reflects a pollster’s past performance and reliability, adjusted for potential biases.
+\vspace{0.7cm}
+
+# Data {#sec-data}
 
@@
 
-### Summary Statistics of Key Variables
+### Summary Statistics of Key Variables and Measurement Method
 
@@
 Two additional variable, national and state, was created using the existing 'state' and 'end date' variable.
-A short description of each variable of interest is given below.
-
--   *Pollster*: The name of the polling organization conducting the poll (e.g., YouGov, RMG Research). This variable helps adjust for poll-specific biases.
-
--   *Numeric Grade*: A numeric rating given to the pollster, representing the quality or reliability of the organization (e.g., 3.0), with higher grades indicating more reliable pollsters.
-
--   *Pollscore*: A score that reflects the reliability of each pollster, capturing their historical accuracy and any systematic bias. Negative values indicate better predictive accuracy.
-
--   *State*: The U.S. state where the poll was conducted, allowing for analysis of regional differences in candidate support.
+A short description of each variable of interest and their respective measurement methods is given below.
+
+-   *Pollster*: The name of the polling organization conducting the poll (e.g., YouGov, RMG Research).
+This variable helps adjust for poll-specific biases.
+Every pollster that publicly published a scientific poll about the 2024 U.S.
+Presidential Election is included in the data set.
+The value measured for this variable is the name that appears in the scientific poll [@538_methodology].
+
+-   *Numeric Grade*: A numeric rating given to the pollster, representing the accuracy and methodological transparency of the organization (e.g., 3.0), with higher grades indicating more reliable pollsters.
+The value of this variable is calculated by the 538 team.
+All national and state-level polls that were conducted in 1998 or later are considered to determine and to assign a numeric grade to each pollster [@538_morris].
+
+-   *Pollscore*: A score that reflects the reliability of each pollster, capturing their historical accuracy and biases.
+Negative values indicate better predictive accuracy, and rewards pollsters who are accurate and precise.
+This variable is calculated by aggregating predictive error and predictive bias together [@538_morris].
+Predictive error is the difference between expectations and reality [@def_pred_error] and predictive bias is the systematic bias introduced in the methodology [@def_pred_bias].
+
+-   *State*: The U.S.
+state where the poll was conducted, allowing for analysis of regional differences in candidate support.
+The value measured for this variable is the primary state that appears in the scientific poll's methodology [@538_methodology].
 
@@
 
--   *End Date*: The date the poll ended, reflecting the currency of the data.
-
--   *Sample Size*: The total number of respondents in the poll.
+-   *End Date*: The date the poll ended, reflecting the currency of the data. This variable is taken from the publication, and is the last date for which the survey was active [@538_methodology].
+
+-   *Sample Size*: The total number of respondents in the poll. This is measure by the number of completed surveys [@538_methodology].
 
@@
 
--   *Recency*: This variable measures how recent each poll is. It was calculated by subtracting the end date of the poll from the end date of the most recent poll.
+-   *Recency*: This variable measures how recent each poll is. This variable is calculated by subtracting the end date of the poll from the end date of the most recent poll.
 

### 2024-11-01 paper/paper.qmd
remove redundate reference. fixed spacing errors.

@@
 Finally, in \ref{sec-discussion}, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work.
+
 \vspace{0.7cm}
@@
 This cutoff was selected to retain mid to high-level pollsters for more reliable results.
-These steps were performed using functions from the `Tidyverse` package [@tidy].
-Furthermore, dates were standardized and converted into a proper format for analysis using the `lubridate` package [@lubridate].
+These steps were performed using functions from the `tidyverse` package [@tidy].
+Dates were also standardized and converted into a proper format for analysis using the same package.
 Polls related to Kamala Harris were retained for further analysis, and percentage support values were transformed into actual numbers of supporters based on sample size.
@@
 
-For the purposes of analysis in this paper, only a subset of variables was selected.
-Two additional variable, national and state, was created using the existing 'state' and 'end date' variable.
+For the purposes this analysis, a subset of variables was selected from the raw data set.
+Two additional variables, 'national' and 'state', were created using the existing 'state' and 'end date' variable.
 A short description of each variable of interest and their respective measurement methods is given below.
@@
 
-From [@tbl-sumstat] we see that the average poll in our dataset has a numeric grade of 2.2, suggesting that most polls are of moderate to high quality in terms of reliability.
-The mean pollscore of -0.5 indicates a general trend toward negative values, which suggests these polls are relatively accurate, as more negative values imply reduced bias.
-Additionally, the large average sample size of 1908 respondents across polls provides a stable basis for our model, reducing variability and enhancing the reliability of predictions for candidate support.
+@tbl-sumstat shows that the average poll in our dataset has a numeric grade of 2.2.
+This suggests most polls are of moderate to high quality in terms of reliability.
+The mean pollscore of -0.5 suggests these polls are relatively accurate, as negative values imply reduced bias.
+Additionally, the large average sample size of 1908 respondents across polls provides a stable basis for our model, reducing variability and ensuring reliable predictions for candidate support.
 
@@
 
-[@tbl-topster] displays the top 5 most frequent pollsters in the dataset, with each pollster’s total poll count, average pollscore (measuring reliability and historical accuracy), and average numeric grade (indicating overall quality).
+@tbl-topster displays the top 5 most frequent pollsters in the dataset, with each pollster’s total poll count, average pollscore (measuring reliability and historical accuracy), and average numeric grade (indicating overall quality).
 We observe that the most frequent pollster, Morning Consult, has a moderate pollscore and a slightly above-average numeric grade, indicating it provides fairly reliable data.
@@
 
-[@fig-ster] shows the distribution of support for Kamala Harris and Donald Trump by pollster, highlighting variability across different polling organizations.
+@fig-ster shows the distribution of support for Kamala Harris and Donald Trump by pollster, highlighting variability across different polling organizations.
 Morning Consult, for example, demonstrates a wide range in reported support, with more spread in Trump’s support.
@@
 
-[@fig-ss] reveals a clear right-skewed distribution.
-Most of the sample sizes are clustered between 0 and 3000, with a sharp peak around 1000-1500.
-This indicates that the majority of polls are conducted with smaller sample sizes.
+@fig-ss shows a clear right-skewed distribution.
+Most of the sample sizes are clustered between 0 and 3000 respondents, with a sharp peak around 1000-1500 respondents.
+This indicates that the majority of polls have smaller sample sizes.
 As sample size increases, the frequency significantly drops, with very few polls conducted with sample sizes larger than 5000, though there are a few outliers with sizes approaching 10,000 or more.
-This wide range in sample sizes might affect the precision of estimates in different polls.
+This wide range in sample sizes can affect the precision of estimates across different polls.
 
@@
 
-[@fig-reli] displays the distribution of Numeric Grade (left) and Pollscore (right) across polling organizations.
+@fig-reli shows the distribution of Numeric Grade (left) and Pollscore (right) across polling organizations.
 The Numeric Grade distribution shows that most pollsters are rated between 1.5 and 3, with peaks around 2.0 and 2.5, indicating a concentration of pollsters with moderate to high reliability scores.
@@
 
-[@fig-pie] charts illustrate the distribution of polls between candidates and poll types.
+@fig-pie illustrates the distribution of polls between candidates and poll types.
 The left chart shows that the majority of polls are conducted at the state level, with a smaller portion being national.
@@
 
-[@fig-trend] shows the support trends for Kamala Harris and Donald Trump from August to October.
+@fig-trend shows the support trends for Kamala Harris and Donald Trump from August to October.
 Each point represents a poll result, color-coded by candidate, with the trend lines highlighting the overall changes in support over time.
@@
 
-[@fig-pct] shows the relationship between sample size and support percentage for Donald Trump and Kamala Harris.
+@fig-pct shows the relationship between sample size and support percentage for Donald Trump and Kamala Harris.
 For both candidates, the majority of polls have a sample size below 2,500, but there are some larger polls exceeding 10,000 respondents.
@@
 
-[@fig-map1] shows the average support for Kamala Harris as a proportion relative to both Kamala Harris and Donald Trump across the continental U.S.
+@fig-map1 shows the average support for Kamala Harris as a proportion relative to both Kamala Harris and Donald Trump across the continental U.S.
 states.
@@
 
--   $y_i$​ is the percentage of support for candidate in poll iii,
+-   $y_i$ is the percentage of support for candidate in poll i,
 
@@
 
--   $β_2​$ captures the effect of the sample size,
-
--   $β_3$​ captures the effect of recency (how recent the poll is),
+-   $β_2$ captures the effect of the sample size,
+
+-   $β_3$ captures the effect of recency (how recent the poll is),
 
@@
 
--   ​ $\epsilon_i$ represents the error term, assumed to follow a normal distribution with mean 0.
+-   $\epsilon_i$ represents the error term, assumed to follow a normal distribution with mean 0.
 
@@
 
-[@tbl-summary] summarizes the performance metrics for three models with progressively added variables.
+@tbl-summary summarizes the performance metrics for three models with progressively added variables.
 Model 1, which includes only Pollster and Sample Size, achieves an R² of 0.375, indicating that these variables alone explain about 37.5% of the variance in candidate support.
@@
 
--   $w_i$​ represents the weight assigned to poll i,
+-   $w_i$ represents the weight assigned to poll i,
 
@@
 
--   $\hat{y}_i$​ is the predicted percentage of support for Kamala Harris from poll i.
+-   $\hat{y}_i$ is the predicted percentage of support for Kamala Harris from poll i.
 
@@
 
-[@fig-map2] presents the predicted winner of the 2024 U.S.
+@fig-map2 presents the predicted winner of the 2024 U.S.
 Presidential Election by state based on aggregated polling data, with red representing states projected to favor Kamala Harris and blue indicating those favoring Donald Trump.
@@
 USAFacts[@usafacts],FiveThirtyEight[@FiveThirtyEight] and many other polling websites are calling Arizona, Georgia, Michigan, Pennsylvania, Wisconsin, North Carolina, Nevada the 'swing states' of the 2024 presidential election.Polling in swing states is crucial for election predictions because these states, with their historically close voting margins, often determine the overall outcome by tipping the electoral balance toward one candidate.
-[@tbl-battleground-states] summarizes the predicted support percentages for Kamala Harris and Donald Trump in key battleground states, along with the predicted winner and the support margin.
+@tbl-battleground-states summarizes the predicted support percentages for Kamala Harris and Donald Trump in key battleground states, along with the predicted winner and the support margin.
 The data shows close margins in these states, with some like Arizona and Georgia leaning slightly towards Trump, while others such as Michigan and Pennsylvania favor Harris by narrow margins.

### 2024-11-01 paper/paper.tex
commit intermediate files

@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, breakable, boxrule=0pt, sharp corners, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 findings, acknowledge limitations, and suggest directions for future
-work. \vspace{0.7cm} \# Data \{\#sec-data\} \vspace{0.7cm} \#\#
-Measurement
-
-In the dataset of our analysis, the process of measurement begins by
-capturing real-world public opinion through surveys. Polling
-organizations transform this observed phenomenon into structured data by
-recording support percentages, pollster details, and methodological
-choices, thus turning public sentiment into quantifiable entries in the
-dataset.
-
-Public opinion polls are essential for understanding voter preferences
-and electoral dynamics, providing data that informs policy decisions and
-campaign strategies. Our data set captures public support for candidates
-at specific times, along with information on methodologies, sample
-sizes, and pollster ratings.
-
-Polling organizations use various methods, such as App panels, IVR
-(Interactive Voice Response), Live phone calls, Text-to-web systems etc.
-These methods can influence the validity and reliability of results,
-with larger sample sizes generally yielding more reliable estimates.
-Some methods, like live phone interviews, may be seen as more
-trustworthy than others.
-
-The dataset also includes important measures of pollster performance,
-such as numeric grades, poll scores, and transparency scores. These
-metrics are calculated by considering bias, race difficulty, predictive
-error, and transparency, allowing us to assess the accuracy of each
-poll. The ``pollscore,'' for instance, reflects a pollster's past
-performance and reliability, adjusted for potential biases.
+work.
+
+\vspace{0.7cm}
+
+\hypertarget{sec-data}{%
+\section{Data}\label{sec-data}}
 
@@
 reliable results. These steps were performed using functions from the
-\texttt{Tidyverse} package (Wickham et al. 2019). Furthermore, dates
-were standardized and converted into a proper format for analysis using
-the \texttt{lubridate} package (Grolemund and Wickham 2011). Polls
-related to Kamala Harris were retained for further analysis, and
-percentage support values were transformed into actual numbers of
-supporters based on sample size. Additionally, pollster counts below
-five were excluded to focus on more reliable data sources. Polls
-regarding Kamala Harris were filtered to include only those conducted
-after her official candidacy announcement on July 21, 2024, ensuring the
-data reflects post-announcement public sentiment.The cleaned dataset was
-saved in Parquet format for efficient storage and retrieval, using the
-\texttt{arrow} package (Richardson et al. 2024).
+\texttt{tidyverse} package (Wickham et al. 2019). Dates were also
+standardized and converted into a proper format for analysis using the
+same package. Polls related to Kamala Harris were retained for further
+analysis, and percentage support values were transformed into actual
+numbers of supporters based on sample size. Additionally, pollster
+counts below five were excluded to focus on more reliable data sources.
+Polls regarding Kamala Harris were filtered to include only those
+conducted after her official candidacy announcement on July 21, 2024,
+ensuring the data reflects post-announcement public sentiment.The
+cleaned dataset was saved in Parquet format for efficient storage and
+retrieval, using the \texttt{arrow} package (Richardson et al. 2024).
 
@@
 
-\hypertarget{summary-statistics-of-key-variables}{%
-\subsubsection{Summary Statistics of Key
-Variables}\label{summary-statistics-of-key-variables}}
+\hypertarget{summary-statistics-of-key-variables-and-measurement-method}{%
+\subsubsection{Summary Statistics of Key Variables and Measurement
+Method}\label{summary-statistics-of-key-variables-and-measurement-method}}
 
@@
 
-For the purposes of analysis in this paper, only a subset of variables
-was selected. Two additional variable, national and state, was created
-using the existing `state' and `end date' variable. A short description
-of each variable of interest is given below.
+For the purposes this analysis, a subset of variables was selected from
+the raw data set. Two additional variables, `national' and `state', were
+created using the existing `state' and `end date' variable. A short
+description of each variable of interest and their respective
+measurement methods is given below.
 
@@
   poll (e.g., YouGov, RMG Research). This variable helps adjust for
-  poll-specific biases.
+  poll-specific biases. Every pollster that publicly published a
+  scientific poll about the 2024 U.S. Presidential Election is included
+  in the data set. The value measured for this variable is the name that
+  appears in the scientific poll (Morris 2024b).
 \item
   \emph{Numeric Grade}: A numeric rating given to the pollster,
-  representing the quality or reliability of the organization (e.g.,
-  3.0), with higher grades indicating more reliable pollsters.
+  representing the accuracy and methodological transparency of the
+  organization (e.g., 3.0), with higher grades indicating more reliable
+  pollsters. The value of this variable is calculated by the 538 team.
+  All national and state-level polls that were conducted in 1998 or
+  later are considered to determine and to assign a numeric grade to
+  each pollster (Morris 2024a).
 \item
   \emph{Pollscore}: A score that reflects the reliability of each
-  pollster, capturing their historical accuracy and any systematic bias.
-  Negative values indicate better predictive accuracy.
+  pollster, capturing their historical accuracy and biases. Negative
+  values indicate better predictive accuracy, and rewards pollsters who
+  are accurate and precise. This variable is calculated by aggregating
+  predictive error and predictive bias together (Morris 2024a).
+  Predictive error is the difference between expectations and reality
+  (Renewable Energy 2012) and predictive bias is the systematic bias
+  introduced in the methodology (Aguinis and Culpepper 2024).
 \item
   \emph{State}: The U.S. state where the poll was conducted, allowing
-  for analysis of regional differences in candidate support.
+  for analysis of regional differences in candidate support. The value
+  measured for this variable is the primary state that appears in the
+  scientific poll's methodology (Morris 2024b).
 \item
@@
   \emph{End Date}: The date the poll ended, reflecting the currency of
-  the data.
-\item
-  \emph{Sample Size}: The total number of respondents in the poll.
+  the data. This variable is taken from the publication, and is the last
+  date for which the survey was active (Morris 2024b).
+\item
+  \emph{Sample Size}: The total number of respondents in the poll. This
+  is measure by the number of completed surveys (Morris 2024b).
 \item
@@
 \item
-  \emph{Recency}: This variable measures how recent each poll is. It was
-  calculated by subtracting the end date of the poll from the end date
-  of the most recent poll.
+  \emph{Recency}: This variable measures how recent each poll is. This
+  variable is calculated by subtracting the end date of the poll from
+  the end date of the most recent poll.
 \end{itemize}
@@
 
-From Table~\ref{tbl-sumstat} we see that the average poll in our dataset
-has a numeric grade of 2.2, suggesting that most polls are of moderate
-to high quality in terms of reliability. The mean pollscore of -0.5
-indicates a general trend toward negative values, which suggests these
-polls are relatively accurate, as more negative values imply reduced
+Table~\ref{tbl-sumstat} shows that the average poll in our dataset has a
+numeric grade of 2.2. This suggests most polls are of moderate to high
+quality in terms of reliability. The mean pollscore of -0.5 suggests
+these polls are relatively accurate, as negative values imply reduced
 bias. Additionally, the large average sample size of 1908 respondents
 across polls provides a stable basis for our model, reducing variability
-and enhancing the reliability of predictions for candidate support.
+and ensuring reliable predictions for candidate support.
 
@@
 
-Figure~\ref{fig-ss} reveals a clear right-skewed distribution. Most of
-the sample sizes are clustered between 0 and 3000, with a sharp peak
-around 1000-1500. This indicates that the majority of polls are
-conducted with smaller sample sizes. As sample size increases, the
-frequency significantly drops, with very few polls conducted with sample
-sizes larger than 5000, though there are a few outliers with sizes
-approaching 10,000 or more. This wide range in sample sizes might affect
-the precision of estimates in different polls.
+Figure~\ref{fig-ss} shows a clear right-skewed distribution. Most of the
+sample sizes are clustered between 0 and 3000 respondents, with a sharp
+peak around 1000-1500 respondents. This indicates that the majority of
+polls have smaller sample sizes. As sample size increases, the frequency
+significantly drops, with very few polls conducted with sample sizes
+larger than 5000, though there are a few outliers with sizes approaching
+10,000 or more. This wide range in sample sizes can affect the precision
+of estimates across different polls.
 
@@
 
-Figure~\ref{fig-reli} displays the distribution of Numeric Grade (left)
-and Pollscore (right) across polling organizations. The Numeric Grade
+Figure~\ref{fig-reli} shows the distribution of Numeric Grade (left) and
+Pollscore (right) across polling organizations. The Numeric Grade
 distribution shows that most pollsters are rated between 1.5 and 3, with
@@
 
-Figure~\ref{fig-pie} charts illustrate the distribution of polls between
+Figure~\ref{fig-pie} illustrates the distribution of polls between
 candidates and poll types. The left chart shows that the majority of
@@
 \item
-  \(y_i\)\hspace{0pt} is the percentage of support for candidate in poll
-  iii,
+  \(y_i\) is the percentage of support for candidate in poll i,
 \item
@@
 \item
-  \(β_2​\) captures the effect of the sample size,
-\item
-  \(β_3\)\hspace{0pt} captures the effect of recency (how recent the
-  poll is),
+  \(β_2\) captures the effect of the sample size,
+\item
+  \(β_3\) captures the effect of recency (how recent the poll is),
 \item
@@
 \item
-  \hspace{0pt} \(\epsilon_i\) represents the error term, assumed to
-  follow a normal distribution with mean 0.
+  \(\epsilon_i\) represents the error term, assumed to follow a normal
+  distribution with mean 0.
 \end{itemize}
@@
 \item
-  \(w_i\)\hspace{0pt} represents the weight assigned to poll i,
+  \(w_i\) represents the weight assigned to poll i,
 \item
@@
 \item
-  \(\hat{y}_i\)\hspace{0pt} is the predicted percentage of support for
-  Kamala Harris from poll i.
+  \(\hat{y}_i\) is the predicted percentage of support for Kamala Harris
+  from poll i.
 \end{itemize}
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-def_pred_bias}{}}%
+Aguinis, Herman, and Steven Culpepper. 2024. \emph{Improving Our
+Understanding of Predictive Bias in Testing}.
+\url{https://psycnet.apa.org/fulltext/2024-15734-001.html}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-akman}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lubridate}{}}%
-Grolemund, Garrett, and Hadley Wickham. 2011. {``Dates and Times Made
-Easy with {lubridate}.''} \emph{Journal of Statistical Software} 40 (3):
-1--25. \url{https://www.jstatsoft.org/v40/i03/}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-citelintr}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-538_morris}{}}%
+Morris, Elliott. 2024a. \emph{How 538's Pollster Ratings Work}.
+\url{https://abcnews.go.com/538/538s-pollster-ratings-work/story?id=105398138}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-538_methodology}{}}%
+---------. 2024b. \emph{Trump Leads in Swing-State Polls and Is Tied
+with Biden Nationally}.
+\url{https://abcnews.go.com/538/trump-leads-swing-state-polls-tied-biden-nationally/story?id=109506070}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-citestyler}{}}%
@@
 \url{https://fivethirtyeight.com/features/2022-election-polling-accuracy/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-def_pred_error}{}}%
+Renewable Energy. 2012. \emph{Prediction Error}.
+\url{https://www.sciencedirect.com/topics/engineering/prediction-error}.
 

### 2024-11-01 paper/paper.tex
commit intermediary files

@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi
 

### 2024-11-01 paper/paper.qmd
Resolved cross referencing issue in the intro.
correct typos and remove whitespace

@@
 #| label: fig-dgplotone
-#| 
+
 knitr::include_graphics(here("other", "graphs", "diagnostic_plot_model3.png"))
@@
 #| label: fig-dgplottwo
-#| 
+
 knitr::include_graphics(here("other", "graphs", "diagnostic_plot_model6.png"))

### 2024-11-01 scripts/00-simulate_data.R
fixed prereqs and coding typose

@@ -4,8 +4,7 @@
 # Date: October 19 2024
 # Contact: krishna.kumar@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-
+# Pre-requisites: 00-install_packages.R
 
 #### Workspace setup ####
 
@@ -71,7 +70,6 @@ candidate_name <- sample(c("Donald Trump", "Kamala Harris", "other"), n, replace
 
 pct <- round(runif(n, min = 30, max = 70), 1)
 
-
 # Combining all variables into a tibble #
 
 simulated_data <- tibble(
@@ -88,7 +86,5 @@ simulated_data <- tibble(
   candidate_name = candidate_name,
   pct = pct
 )
-
-head(simulated_data, 20)
 
 write.csv(simulated_data, file = "data/00-simulated_data/simulated_data.csv")

### 2024-11-01 scripts/01-test_simulated_data.R

@@ -5,8 +5,8 @@
 # Contact: krishna.kumar@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
+  # - 00-install_packages.R
+  # - 00-simulate_data.R 
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-11-01 scripts/04-create_graph.R

@@ -4,7 +4,7 @@
 # Date: 30-10-2024
 # Contact: shamayla.islam@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Libraries DiagrammeR, DiagrammeRsvg, and rsvg must be installed.
+# Pre-requisites: 00-install_packages.R
 # Additional Notes: This chunk saves the DAG as a PNG file to the specified directory for use in the main report.
 
 # Load necessary libraries

### 2024-11-01 scripts/03-clean_data.R
updated

@@ -89,6 +89,32 @@ cleaned_data <- cleaned_data %>%
 cleaned_data <- cleaned_data %>%
   filter(end_date >= as.Date("2024-07-21"))
 
+# Calculate recency of every poll and make a column 
+
+reference_date <- max(cleaned_data$end_date, na.rm = TRUE)  # Most recent poll date
+
+analysis_data <- cleaned_data %>%
+  mutate(recency = as.numeric(reference_date - end_date))
+
+
 #### Save data ####
 
 write_parquet(cleaned_data, "data/02-analysis_data/analysis_data.parquet")
+
+#### Creating Data For Mapping ####
+
+deduped_data <- raw_data %>%
+  distinct()  # Remove duplicate rows
+
+# Step 2: Filter for Kamala Harris and non-empty states
+harris_data <- deduped_data %>%
+  filter(candidate_name == "Kamala Harris" & !is.na(state))
+
+# Step 3: Group by state and calculate mean support for Kamala Harris
+mapping_data <- harris_data %>%
+  group_by(state) %>%
+  summarise(mean_support_kamala = mean(pct, na.rm = TRUE))
+
+# Step 4: Save the aggregated data to a parquet file
+write_parquet(mapping_data, "data/02-analysis_data/mapping_data.parquet")
+

### 2024-11-01 scripts/06-model_data.R

@@ -1,19 +1,22 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models the percentage of support for Kamala Harris (pct) based on various factors such as pollster, poll type (state vs. national), 
+#          sample size, and other relevant variables.
+# Author: Shamayla Durrin
+# Date: 19 October 2024
+# Contact: shamayla.islam@mil.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Requires the cleaned dataset of polls, including variables like pollster, state, sample size, and pct for Kamala Harris.
+# Any other information needed: Ensure the necessary packages (e.g., tidyverse, broom, and modeling packages) are installed before running the model.
 
 
 #### Workspace setup ####
+
 library(tidyverse)
 library(rstanarm)
+library(arrow)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
 ### Model data ####
 first_model <-
@@ -31,7 +34,7 @@ first_model <-
 #### Save model ####
 saveRDS(
   first_model,
-  file = "models/first_model.rds"
+  file = "models/first_m\del.rds"
 )
 
 

### 2024-11-01 paper/paper.qmd
fix title ans telegraph appendices

@@
 ---
-title: "Forecasting the 2024 U.S. Presidential Election: An Analysis for Battleground States"
+title: "Forecasting the 2024 U.S. Presidential Election: An Analysis of Battleground States"
 subtitle: "Trump Favored Nationally, While Harris Leads in 4 of 7 Battleground States"
@@
 library(here)
-library(ggplot2)
 library(kableExtra)
@@
 library(RColorBrewer)
-library(stringr)
 library(modelsummary)
@@
 
-This paper is organized as follows: In \ref{sec-data}, we clean the data, explore summary statistics, plot distributions of key variables, and examine relationships between variables.
-In \ref{sec-model}, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions.
-Finally, in \ref{sec-discussion}, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work.
+This paper is organized as follows: In Section @sec-data, we clean the data, explore summary statistics, plot distributions of key variables, and examine relationships between variables.
+In Section @sec-model, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions.
+In Section @sec-discussion, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work.
+[Appendix -@sec-pollster-meth] contains a closer look at the methodology used by Emerson College Polling for its poll conducted from October 14 t0 16, 2024.
+[Appendix -@sec-idealized-meth] contains a discussion about an idealized methodology that we would use to forecast the 2024 U.S.
+Presidential Election if we had a budget of  \$100,000.
 
@@
 
-# Emerson College Polling Methodology (October 14-16, 2024)
+# Emerson College Polling Methodology (October 14-16, 2024) {#sec-pollster-meth}
 
@@
 
-# Idealized Methodology
+# Idealized Methodology {#sec-idealized-meth}
 

### 2024-11-01 paper/paper.qmd
adjust decimals in figure 4. Fix captions.

@@
 Polls related to Kamala Harris were retained for further analysis, and percentage support values were transformed into actual numbers of supporters based on sample size.
-Additionally, pollster counts below five were excluded to focus on more reliable data sources.
+Pollster counts below five were excluded to focus on more reliable data sources.
 Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the `arrow` package [@arrow].
@@
 ### Summary Statistics of Key Variables and Measurement Method
-
-\vspace{0.7cm}
 
@@
 #| label: fig-trend
-#| fig-cap: "Figure showing support trends for Kamala Harris and Donald Trump over time.
-Trend lines indicate slight shifts in support as the election nears, with consistent polling frequency throughout the period."
+#| fig-cap: "Support trends for Kamala Harris and Donald Trump over time. Trend lines indicate slight shifts in support as the election nears, with consistent polling frequency throughout the period."
 
@@
 #| label: fig-map1
-#| fig-cap: "Map showing the proportion of support for Kamala Harris relative to Donald Trump across U.S.
+#| fig-cap: "Proportion of support for Kamala Harris relative to Donald Trump across U.S.
 states.
@@
 #| label: fig-causal-model
-#| fig-cap: "Directed Acyclic Graph illustrating factors influencing candidate support, with poll attributes (Pollster, Sample Size, State, Recency) as direct contributors and unobserved confounders representing potential biases in polling results."
+#| fig-cap: "Factors influencing candidate support, with poll attributes (Pollster, Sample Size, State, Recency) as direct contributors and unobserved confounders representing potential biases in polling results."
 # Load and display the saved PNG
@@
     `Support Margin (%)` = margin
+  ) %>%
+  mutate(
+    `Kamala Support (%)` = round(`Kamala Support (%)`, 2),
+    `Trump Support (%)` = round(`Trump Support (%)`, 2),
+    `Support Margin (%)` = round(`Support Margin (%)`, 2)
   )

### 2024-11-01 paper/paper.qmd
remove figure title for figure 5

@@
 
-\vspace{0.7cm}
-
 ### Distribution of Polls by Poll Type and Candidate
@@
   scale_color_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
-  labs(x = "Date", y = "Support (%)", title = "Support Trends for Kamala Harris and Donald Trump") +
+  labs(x = "Date", y = "Support (%)") +
   theme_minimal() +

### 2024-11-01 paper/paper.qmd
fixed a little prose

@@
 
-\vspace{0.7cm}
-
 # Introduction
@@
 While individual polls provide snapshots of public opinion, they are often subject to biases and methodological differences.
-In this paper, we aim to forecast voter support for Kamala Harris and Donald Trump by aggregating data from multiple polling sources, reducing individual poll biases, and improving overall prediction accuracy.
-Our analysis focuses on estimating the levels of support for each candidate not only nationally but also in key swing states that are likely to be decisive in determining the electoral college outcome.
+In this paper, we aim to forecast voter support for Kamala Harris and Donald Trump by using data from multiple polling sources, reducing individual poll biases, and improving overall prediction accuracy.
+Our analysis focuses on estimating the levels of support for each candidate nationally while focusing on key swing states, which are likely to be decisive in determining the electoral college outcome.
 
@@
 
-\vspace{0.7cm}
-
 # Data {#sec-data}
-
-\vspace{0.7cm}
 
@@
 
-For the purposes this analysis, a subset of variables was selected from the raw data set.
+For this analysis, a subset of variables was selected from the raw data set.
 Two additional variables, 'national' and 'state', were created using the existing 'state' and 'end date' variable.
@@
 
-@tbl-sumstat shows that the average poll in our dataset has a numeric grade of 2.2.
+@tbl-sumstat shows the average poll in our data set has a numeric grade of 2.2.
 This suggests most polls are of moderate to high quality in terms of reliability.
-The mean pollscore of -0.5 suggests these polls are relatively accurate, as negative values imply reduced bias.
-Additionally, the large average sample size of 1908 respondents across polls provides a stable basis for our model, reducing variability and ensuring reliable predictions for candidate support.
+The mean pollscore of -0.5 suggests these polls are accurate, as negative values imply reduced bias.
+Also, the large average sample size of 1908 respondents across polls reduces variability and ensures reliable predictions for the candidate support model.
 
@@
 ### Variation of Poll Quality and Support for Candiates by Pollster
-
-\vspace{0.7cm}
 
@@
 
-@tbl-topster displays the top 5 most frequent pollsters in the dataset, with each pollster’s total poll count, average pollscore (measuring reliability and historical accuracy), and average numeric grade (indicating overall quality).
-We observe that the most frequent pollster, Morning Consult, has a moderate pollscore and a slightly above-average numeric grade, indicating it provides fairly reliable data.
-However, there is variability in pollscore and numeric grade across the top pollsters, reflecting differences in quality and potential biases among them, which the model needs to adjust for to achieve accurate predictions.
+@tbl-topster shows the top 5 most frequent pollsters in the dataset, with each pollster’s total poll count, average pollscore (measuring reliability and historical accuracy), and average numeric grade (indicating overall quality).
+The most frequent pollster, Morning Consult, has a moderate pollscore and a slightly above-average numeric grade, suggesting the data provided is reliable.
+However, there is variability in pollscore and numeric grade across the top pollsters, showing differences in quality and potential biases among them.
 
@@
 In contrast, Siena/NYT shows less variation, with Harris consistently leading.
-This variability across pollsters emphasizes the importance of aggregating multiple polls to account for organization-specific biases and ensure a more balanced view of candidate support.
+This variability across pollsters shows the importance of aggregating multiple polls to account for organization-specific biases and ensure a more balanced view of candidate support.
 
@@
 This indicates that the majority of polls have smaller sample sizes.
-As sample size increases, the frequency significantly drops, with very few polls conducted with sample sizes larger than 5000, though there are a few outliers with sizes approaching 10,000 or more.
+As sample size increases, the frequency decreases, with few polls conducted with sample sizes larger than 5000, though there are a few outliers with sizes approaching 10,000 or more.
 This wide range in sample sizes can affect the precision of estimates across different polls.
@@
 @fig-reli shows the distribution of Numeric Grade (left) and Pollscore (right) across polling organizations.
-The Numeric Grade distribution shows that most pollsters are rated between 1.5 and 3, with peaks around 2.0 and 2.5, indicating a concentration of pollsters with moderate to high reliability scores.
+The Numeric Grade distribution shows that most pollsters are rated between 1.5 and 3, with peaks around 2.0 and 2.5, suggesting a concentration of pollsters with moderate to high reliability scores.
 In contrast, the Pollscore distribution, where lower values indicate higher reliability, shows a range primarily between -1.5 and 0, with a notable peak around -0.5.
@@
 In this paper, we set out to predict the support for both Kamala Harris and Donald Trump in the 2024 U.S.
-Presidential Election using a “polls-of-polls” approach.
+Presidential Election using a “polls-of-polls” approach ADD REFERENCE HERE.
 By aggregating multiple polls, we sought to mitigate the inherent biases present in individual surveys, creating a more accurate and balanced forecast across candidates.
@@
 
-In conclusion, this paper contributes to election forecasting by providing a comprehensive, weighted estimate of candidate support across both national and state levels.
-While our predictions provide valuable insights into potential election outcomes, the evolving nature of voter sentiment and poll accuracy reflects the need for adaptive models that account for ongoing changes in public opinion.
+In conclusion, this paper contributes to election forecasting by providing a weighted estimate of candidate support across both national and state levels.The evolving nature of voter sentiment and poll accuracy suggests the need for adaptive models that takes in consideration ongoing changes in public opinion.
 

### 2024-11-01 scripts/03-test_analysis_data.R
rename scripts

@@ -1,0 +1,10 @@
+#### Preamble ####
+# Purpose: Tests the structure and validity of the analysis data set
+# Author: Krishna Kumar
+# Date: November 1 2024
+# Contact: krishna.kumar@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# - 00-install_packages.R
+# - 02-clean_data.R
+

### 2024-11-01 scripts/00-simulate_data.R
use styler on scripts

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Simulates a data set of polling results for forecasting the upcoming US elections. 
+# Purpose: Simulates a data set of polling results for forecasting the upcoming US elections.
 # Author: Krishna Kumar
 # Date: October 19 2024
 # Contact: krishna.kumar@mail.utoronto.ca
@@ -34,25 +34,74 @@ pollscore <- round(runif(n, min = -1, max = 1), 2)
 
 # simulating methodology for three assumed types #
 
-methodology <- sample(c('Online', 'In-Person', 'Hybrid'), n, replace = TRUE)
+methodology <-
+  sample(c('Online', 'In-Person', 'Hybrid'), n, replace = TRUE)
 
 # simulating states #
 
-states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", 
-            "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", 
-            "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", 
-            "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", 
-            "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", 
-            "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", 
-            "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming")
+states <-
+  c(
+    "Alabama",
+    "Alaska",
+    "Arizona",
+    "Arkansas",
+    "California",
+    "Colorado",
+    "Connecticut",
+    "Delaware",
+    "Florida",
+    "Georgia",
+    "Hawaii",
+    "Idaho",
+    "Illinois",
+    "Indiana",
+    "Iowa",
+    "Kansas",
+    "Kentucky",
+    "Louisiana",
+    "Maine",
+    "Maryland",
+    "Massachusetts",
+    "Michigan",
+    "Minnesota",
+    "Mississippi",
+    "Missouri",
+    "Montana",
+    "Nebraska",
+    "Nevada",
+    "New Hampshire",
+    "New Jersey",
+    "New Mexico",
+    "New York",
+    "North Carolina",
+    "North Dakota",
[…]

### 2024-11-01 scripts/01-test_simulated_data.R

@@ -4,14 +4,15 @@
 # Date: October 19 2024
 # Contact: krishna.kumar@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
-  # - 00-install_packages.R
-  # - 00-simulate_data.R 
+# Pre-requisites:
+# - 00-install_packages.R
+# - 00-simulate_data.R
 
 #### Workspace setup ####
 library(tidyverse)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+analysis_data <-
+  read_csv("data/00-simulated_data/simulated_data.csv")
 
 # Test if the data was successfully loaded
 if (exists("analysis_data")) {
@@ -45,9 +46,17 @@ if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
 }
 
 # Check if the 'state' column contains only valid Australian state names
-valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", 
-                  "Western Australia", "Tasmania", "Northern Territory", 
-                  "Australian Capital Territory")
+valid_states <-
+  c(
+    "New South Wales",
+    "Victoria",
+    "Queensland",
+    "South Australia",
+    "Western Australia",
+    "Tasmania",
+    "Northern Territory",
+    "Australian Capital Territory"
+  )
 
 if (all(analysis_data$state %in% valid_states)) {
   message("Test Passed: The 'state' column contains only valid Australian state names.")
@@ -56,7 +65,8 @@ if (all(analysis_data$state %in% valid_states)) {
 }
 
 # Check if the 'party' column contains only valid party names
-valid_parties <- c("Labor", "Liberal", "Greens", "National", "Other")
+valid_parties <-
+  c("Labor", "Liberal", "Greens", "National", "Other")
 
 if (all(analysis_data$party %in% valid_parties)) {
   message("Test Passed: The 'party' column contains only valid party names.")
@@ -72,7 +82,8 @@ if (all(!is.na(analysis_data))) {
 }
 
 # Check if there are no empty strings in 'division', 'state', and 'party' columns
-if (all(analysis_data$division != "" & analysis_data$state != "" & analysis_data$party != "")) {
+if (all(analysis_data$division != "" &
+        analysis_data$state != "" & analysis_data$party != "")) {
   message("Test Passed: There are no empty strings in 'division', 'state', or 'party'.")
 } else {
   stop("Test Failed: There are empty strings in one or more columns.")

### 2024-11-01 scripts/02-clean_data.R

@@ -17,55 +17,59 @@ raw_data <- read_csv(here("data/01-raw_data/raw_data.csv"))
 
 #### Data Cleaning ####
 
-# Remove duplicate rows if any 
-
+# Remove duplicate rows if any
 cleaned_data <- raw_data %>% distinct()
 
 # Create a new column 'National' based on whether the 'state' column is empty or not
-
 cleaned_data <- cleaned_data %>%
   mutate(national = ifelse(is.na(state), 1, 0))
 
 # Replace missing values (NA) in the 'state' column with Not Applicable
-
 cleaned_data <- cleaned_data %>%
   mutate(state = ifelse(is.na(state), "Not Applicable", state))
 
 # Keep only the relevant columns
 cleaned_data <- cleaned_data %>%
-  select(poll_id, pollster, numeric_grade, pollscore, state, national, end_date, sample_size,candidate_name, pct)
+  select(
+    poll_id,
+    pollster,
+    numeric_grade,
+    pollscore,
+    state,
+    national,
+    end_date,
+    sample_size,
+    candidate_name,
+    pct
+  )
 
 # Remove any rows that have missing values in any column
 cleaned_data <- cleaned_data %>% drop_na()
 
-# Look at the range of numeric grade to select a cutoff 
-
+# Look at the range of numeric grade to select a cutoff
 numeric_grade_range <- range(cleaned_data$numeric_grade)
 
-cutoffgrade <- 1 # this cutoff was selected  to retain mid to high quality pollster 
+cutoffgrade <-
+  1 # this cutoff was selected  to retain mid to high quality pollster
 
-# Remove pollster that has low numeric grade 
-
-cleaned_data <- cleaned_data %>% filter(numeric_grade >= cutoffgrade)
+# Remove pollster that has low numeric grade
+cleaned_data <-
+  cleaned_data %>% filter(numeric_grade >= cutoffgrade)
 
 
 # Ensure the start_date and end_date columns are in date format
 cleaned_data <- cleaned_data %>%
-  mutate(
-    end_date = parse_date_time(end_date, orders = c("mdy", "ymd"))
-  )
+  mutate(end_date = parse_date_time(end_date, orders = c("mdy", "ymd")))
 
 # Apply clean_names to clean the column names
 cleaned_data <- cleaned_data %>%
   clean_names()
 
 # Ensure that end_date is a Date object
-
[…]

### 2024-11-01 scripts/04-model_data.R

@@ -7,19 +7,16 @@
 # Pre-requisites: Ensure that the necessary libraries (tidyverse, arrow, etc.) are installed and data is available in the required format.
 # Any other information needed? Data should include variables for support, pollster, sample size, state, and recency.
 
-
 #### Workspace setup ####
-
 library(tidyverse)
 library(here)
 library(arrow)
 
 #### Read data ####
-analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
+analysis_data <-
+  read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
 
-
-# Filter all the observations regarding Harris. Model 1,2,3 will be about Harris. 
-
+# Filter all the observations regarding Harris. Model 1,2,3 will be about Harris.
 harris_data <- analysis_data %>%
   filter(candidate_name %in% c("Kamala Harris"))
 
@@ -29,15 +26,16 @@ model1 <- lm(pct ~ pollster + sample_size, data = harris_data)
 saveRDS(model1, here("models", "model1.rds"))
 
 # Model 2: Adding state as a predictor
-model2 <- lm(pct ~ pollster + sample_size + state, data = harris_data)
+model2 <-
+  lm(pct ~ pollster + sample_size + state, data = harris_data)
 saveRDS(model2, here("models", "model2.rds"))
 
 # Model 3: Adding recency as a predictor
-model3 <- lm(pct ~ pollster + sample_size + state + recency, data = harris_data)
+model3 <-
+  lm(pct ~ pollster + sample_size + state + recency, data = harris_data)
 saveRDS(model3, here("models", "model3.rds"))
 
-# Filter all the observations regarding Trump. Model 4,5,6 will be about Trump. 
-
+# Filter all the observations regarding Trump. Model 4,5,6 will be about Trump.
 trump_data <- analysis_data %>%
   filter(candidate_name %in% c("Donald Trump"))
 
@@ -47,10 +45,11 @@ model4 <- lm(pct ~ pollster + sample_size, data = trump_data)
 saveRDS(model4, here("models", "model4.rds"))
 
 # Model 5: Adding state as a predictor
-model5 <- lm(pct ~ pollster + sample_size + state, data = trump_data)
+model5 <-
+  lm(pct ~ pollster + sample_size + state, data = trump_data)
 saveRDS(model5, here("models", "model5.rds"))
 
 # Model 6: Adding recency as a predictor
[…]

### 2024-11-01 scripts/05-create_graph.R

@@ -7,7 +7,7 @@
 # Pre-requisites: 00-install_packages.R
 # Additional Notes: This chunk saves the DAG as a PNG file to the specified directory for use in the main report.
 
-# Load necessary libraries
+#### Workspace setup ####
 library(DiagrammeR)
 library(DiagrammeRsvg)
 library(rsvg)
@@ -47,20 +47,30 @@ export_svg(graph) |>
   charToRaw() |>
   rsvg::rsvg_png("other/graphs/polls_of_polls_dag.png", width = 1000)
 
-
-
 # Define the file paths using the here() function
-output_path_model3 <- here("other", "graphs", "diagnostic_plot_model3.png")
-output_path_model6 <- here("other", "graphs", "diagnostic_plot_model6.png")
+output_path_model3 <-
+  here("other", "graphs", "diagnostic_plot_model3.png")
+output_path_model6 <-
+  here("other", "graphs", "diagnostic_plot_model6.png")
 
 # Save diagnostic plot for model3
 png(output_path_model3, width = 800, height = 600)
-autoplot(model3, label.size = 3, smooth.colour = "blue", smooth.linetype = "dashed") + theme_classic()
+autoplot(
+  model3,
+  label.size = 3,
+  smooth.colour = "blue",
+  smooth.linetype = "dashed"
+) + theme_classic()
 dev.off()
 
 # Save diagnostic plot for model6
 png(output_path_model6, width = 800, height = 600)
-autoplot(model6, label.size = 3, smooth.colour = "blue", smooth.linetype = "dashed") + theme_classic()
+autoplot(
+  model6,
+  label.size = 3,
+  smooth.colour = "blue",
+  smooth.linetype = "dashed"
+) + theme_classic()
 dev.off()
 
 

### 2024-11-01 paper/paper.qmd
fix typo in table 4

@@
 
-USAFacts[@usafacts],FiveThirtyEight[@FiveThirtyEight] and many other polling websites are calling Arizona, Georgia, Michigan, Pennsylvania, Wisconsin, North Carolina, Nevada the 'swing states' of the 2024 presidential election.Polling in swing states is crucial for election predictions because these states, with their historically close voting margins, often determine the overall outcome by tipping the electoral balance toward one candidate.
+USAFacts [@usafacts], FiveThirtyEight [@FiveThirtyEight] and many other polling websites are calling Arizona, Georgia, Michigan, Pennsylvania, Wisconsin, North Carolina, Nevada the 'swing states' of the 2024 presidential election.
+Swing states are crucial for election predictions because these states, with their historically close voting margins, often determine the overall outcome by tipping the electoral balance toward one candidate.
 @tbl-battleground-states summarizes the predicted support percentages for Kamala Harris and Donald Trump in key battleground states, along with the predicted winner and the support margin.
-The data shows close margins in these states, with some like Arizona and Georgia leaning slightly towards Trump, while others such as Michigan and Pennsylvania favor Harris by narrow margins.
-The support margin column highlights the competitiveness of these states, with all margins below 2.2%, indicating highly contested races.
+The data shows close margins in these states, with some like Arizona and Georgia leaning towards Trump, while others such as Michigan and Pennsylvania favor Harris by narrow margins.
+The support margin column highlights the competitiveness of these states, with all margins below 2.2%.
 

### 2024-11-02 paper/paper.qmd

--- 2024-11-01 paper/paper.qmd
@@
 ---
-title: "Forecasting the 2024 U.S. Presidential Election: An Analysis of Battleground States"
+title: "Forecasting the 2024 U.S. Presidential Election: An Analysis for Battleground States"
 subtitle: "Trump Favored Nationally, While Harris Leads in 4 of 7 Battleground States"
@@
 
-This paper is organized as follows: In Section @sec-data, we clean the data, explore summary statistics, plot distributions of key variables, and examine relationships between variables.
-In Section @sec-model, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions.
-In Section @sec-discussion, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work.
-[Appendix -@sec-pollster-meth] contains a closer look at the methodology used by Emerson College Polling for its poll conducted from October 14 t0 16, 2024.
-[Appendix -@sec-idealized-meth] contains a discussion about an idealized methodology that we would use to forecast the 2024 U.S.
-Presidential Election if we had a budget of  \$100,000.
+This paper is organized as follows: In \ref{sec-data}, we clean the data, explore summary statistics, plot distributions of key variables, and examine relationships between variables.
+In \ref{sec-model}, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions.
+Finally, in \ref{sec-discussion}, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work.
 
@@
 Polls related to Kamala Harris were retained for further analysis, and percentage support values were transformed into actual numbers of supporters based on sample size.
-Pollster counts below five were excluded to focus on more reliable data sources.
+Additionally, pollster counts below five were excluded to focus on more reliable data sources.
 Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the `arrow` package [@arrow].
@@
 ### Summary Statistics of Key Variables and Measurement Method
+
+\vspace{0.7cm}
 
@@
 
+\vspace{0.7cm}
+
 ### Distribution of Polls by Poll Type and Candidate
@@
 #| label: fig-trend
-#| fig-cap: "Support trends for Kamala Harris and Donald Trump over time. Trend lines indicate slight shifts in support as the election nears, with consistent polling frequency throughout the period."
+#| fig-cap: "Figure showing support trends for Kamala Harris and Donald Trump over time.
+Trend lines indicate slight shifts in support as the election nears, with consistent polling frequency throughout the period."
 
@@
   scale_color_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
-  labs(x = "Date", y = "Support (%)") +
+  labs(x = "Date", y = "Support (%)", title = "Support Trends for Kamala Harris and Donald Trump") +
   theme_minimal() +
@@
 #| label: fig-map1
-#| fig-cap: "Proportion of support for Kamala Harris relative to Donald Trump across U.S.
+#| fig-cap: "Map showing the proportion of support for Kamala Harris relative to Donald Trump across U.S.
 states.
@@
 #| label: fig-causal-model
-#| fig-cap: "Factors influencing candidate support, with poll attributes (Pollster, Sample Size, State, Recency) as direct contributors and unobserved confounders representing potential biases in polling results."
+#| fig-cap: "Directed Acyclic Graph illustrating factors influencing candidate support, with poll attributes (Pollster, Sample Size, State, Recency) as direct contributors and unobserved confounders representing potential biases in polling results."
 # Load and display the saved PNG
@@
     `Support Margin (%)` = margin
-  ) %>%
-  mutate(
-    `Kamala Support (%)` = round(`Kamala Support (%)`, 2),
-    `Trump Support (%)` = round(`Trump Support (%)`, 2),
-    `Support Margin (%)` = round(`Support Margin (%)`, 2)
   )
@@
 
-# Emerson College Polling Methodology (October 14-16, 2024) {#sec-pollster-meth}
+# Emerson College Polling Methodology (October 14-16, 2024)
 
@@
 
-# Idealized Methodology {#sec-idealized-meth}
+# Idealized Methodology
 

### 2024-11-04 scripts/00-install_packages.R
restore scripts from earlier version

@@ -25,3 +25,5 @@ install.packages("modelsummary")
 install.packages("broom")
 install.packages("sf")
 install.packages("patchwork")
+install.packages("testthat")
+install.packages("readr")

### 2024-11-04 scripts/00-simulate_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Simulates a data set of polling results for forecasting the upcoming US elections. 
+# Purpose: Simulates a data set of polling results for forecasting the upcoming US elections.
 # Author: Krishna Kumar
 # Date: October 19 2024
 # Contact: krishna.kumar@mail.utoronto.ca
@@ -21,57 +21,98 @@ n <- 100
 poll_id <- sample(1:1000, n, replace = TRUE)
 
 # simulating pollster ID #
-
 pollster_id <- sample(1:100, n, replace = TRUE)
 
 # simulating numeric grade #
-
 numeric_grade <- round(runif(n, min = 0.0, max = 10.0), 2)
 
 # simulating poll score #
-
 pollscore <- round(runif(n, min = -1, max = 1), 2)
 
 # simulating methodology for three assumed types #
-
-methodology <- sample(c('Online', 'In-Person', 'Hybrid'), n, replace = TRUE)
+methodology <-
+  sample(c('Online', 'In-Person', 'Hybrid'), n, replace = TRUE)
 
 # simulating states #
-
-states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", 
-            "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", 
-            "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", 
-            "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", 
-            "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", 
-            "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", 
-            "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming")
+states <-
+  c(
+    "Alabama",
+    "Alaska",
+    "Arizona",
+    "Arkansas",
+    "California",
+    "Colorado",
+    "Connecticut",
+    "Delaware",
+    "Florida",
+    "Georgia",
+    "Hawaii",
+    "Idaho",
+    "Illinois",
+    "Indiana",
+    "Iowa",
+    "Kansas",
+    "Kentucky",
+    "Louisiana",
+    "Maine",
+    "Maryland",
[…]

### 2024-11-04 scripts/01-test_simulated_data.R

@@ -10,21 +10,14 @@
 
 #### Workspace setup ####
 
-#install.packages("arrow")
-#install.packages("testthat")
-
 library(tidyverse)
 library(testthat)
 library(arrow)
 library(readr)
 
-<<<<<<< HEAD
 # loading the data #
-simulated_data <- read_csv("data/00-simulated_data/simulated_data.csv")
-=======
-analysis_data <-
+simulated_data <-
   read_csv("data/00-simulated_data/simulated_data.csv")
->>>>>>> 2e22a206dc9788f5708daba86f1e4ee71eec1e75
 
 # Checking data load success
 if (exists("simulated_data") && is.data.frame(simulated_data)) {
@@ -41,13 +34,59 @@ if (all(!is.na(simulated_data))) {
 }
 
 # Checking for valid state names
-valid_states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", 
-                  "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", 
-                  "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", 
-                  "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", 
-                  "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", 
-                  "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", 
-                  "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming")
+valid_states <-
+  c(
+    "Alabama",
+    "Alaska",
+    "Arizona",
+    "Arkansas",
+    "California",
+    "Colorado",
+    "Connecticut",
+    "Delaware",
+    "Florida",
+    "Georgia",
+    "Hawaii",
+    "Idaho",
+    "Illinois",
+    "Indiana",
+    "Iowa",
+    "Kansas",
+    "Kentucky",
+    "Louisiana",
+    "Maine",
+    "Maryland",
+    "Massachusetts",
+    "Michigan",
+    "Minnesota",
+    "Mississippi",
+    "Missouri",
+    "Montana",
+    "Nebraska",
+    "Nevada",
+    "New Hampshire",
[…]

### 2024-11-04 scripts/02-clean_data.R

@@ -12,109 +12,84 @@
 library(tidyverse)
 library(janitor)
 library(arrow)
-library(lubridate)
-raw_data <- read_csv("data/01-raw_data/raw_data.csv")
+
+raw_data <- read_csv(here("data/01-raw_data/raw_data.csv"))
 
 #### Data Cleaning ####
 
-# Remove duplicate rows if any 
-
+# Remove duplicate rows if any
 cleaned_data <- raw_data %>% distinct()
 
 # Create a new column 'National' based on whether the 'state' column is empty or not
-
 cleaned_data <- cleaned_data %>%
   mutate(national = ifelse(is.na(state), 1, 0))
 
 # Replace missing values (NA) in the 'state' column with Not Applicable
-
 cleaned_data <- cleaned_data %>%
   mutate(state = ifelse(is.na(state), "Not Applicable", state))
 
 # Keep only the relevant columns
 cleaned_data <- cleaned_data %>%
-  select(poll_id, pollster_id, pollster, numeric_grade, pollscore, state, start_date, national, end_date, sample_size, population, candidate_name, pct)
+  select(
+    poll_id,
+    pollster,
+    numeric_grade,
+    pollscore,
+    state,
+    national,
+    end_date,
+    sample_size,
+    candidate_name,
+    pct
+  )
 
 # Remove any rows that have missing values in any column
 cleaned_data <- cleaned_data %>% drop_na()
 
-# Look at the range of numeric grade to select a cutoff 
-
+# Look at the range of numeric grade to select a cutoff
 numeric_grade_range <- range(cleaned_data$numeric_grade)
 
-cutoffgrade <- 1 # this cutoff was selected  to retain mid to high quality pollster 
+cutoffgrade <-
+  1 # this cutoff was selected  to retain mid to high quality pollster
 
-# Remove pollster that has low numeric grade 
-
-cleaned_data <- cleaned_data %>% filter(numeric_grade >= cutoffgrade)
-
-
-#Filter the dataset to keep only rows with Kamala Harris or Donald Trump
-
-cleaned_data <- cleaned_data %>%
-  filter(candidate_name %in% c("Kamala Harris"))
+# Remove pollster that has low numeric grade
+cleaned_data <-
+  cleaned_data %>% filter(numeric_grade >= cutoffgrade)
 
 # Ensure the start_date and end_date columns are in date format
 cleaned_data <- cleaned_data %>%
-  mutate(
[…]

### 2024-11-04 scripts/03-test_analysis_data.R

@@ -1,24 +1,23 @@
 #### Preamble ####
-<<<<<<< HEAD
 # Purpose: Tests the structure and validity of the analysis polling results data set
 # Author: Krishna Kumar
 # Date: October 19 2024
 # Contact: krishna.kumar@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
-  # - 00-install_packages.R
-  # - 00-simulate_data.R 
+# Pre-requisites:
+# - 00-install_packages.R
+# - 00-simulate_data.R
 
 #### Workspace setup ####
-
-#install.packages("here")
+library(tidyverse)
+library(testthat)
+library(arrow)
+library(readr)
 library(here)
 
 # Load analysis data
-analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
-
-head(analysis_data)
-unique(analysis_data$state)
+analysis_data <-
+  read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
 
 # Check if data is loaded and is a data frame
 test_that("Data load and structure", {
@@ -32,14 +31,45 @@ test_that("No missing values in the dataset", {
 })
 
 # Define valid states
-valid_states <- c("Not Applicable", "Washington", "Oregon", "Florida", "Arizona", 
-                  "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania",  
-                  "Wisconsin", "Maryland", "Ohio", "Texas", "California",    
-                  "Minnesota", "New Mexico", "Montana", "Nebraska", "Utah",          
-                  "Massachusetts", "New Hampshire", "Iowa", "Nebraska CD-2", "New York",      
-                  "Virginia", "Connecticut", "Rhode Island", "Colorado", "Missouri",      
-                  "Indiana", "Alaska", "Vermont", "Maine", "Maine CD-1",    
-                  "Maine CD-2")
+valid_states <-
+  c(
+    "Not Applicable",
+    "Washington",
+    "Oregon",
+    "Florida",
+    "Arizona",
+    "Georgia",
+    "Michigan",
+    "Nevada",
+    "North Carolina",
+    "Pennsylvania",
+    "Wisconsin",
+    "Maryland",
+    "Ohio",
+    "Texas",
+    "California",
+    "Minnesota",
+    "New Mexico",
+    "Montana",
+    "Nebraska",
+    "Utah",
+    "Massachusetts",
[…]

### 2024-11-04 scripts/04-model_data.R

@@ -1,40 +1,57 @@
 #### Preamble ####
-# Purpose: Models the percentage of support for Kamala Harris (pct) based on various factors such as pollster, poll type (state vs. national), 
-#          sample size, and other relevant variables.
+# Purpose: Models the percentage of support for Kamala Harris in each poll as a function of various predictors.
 # Author: Shamayla Durrin
-# Date: 19 October 2024
-# Contact: shamayla.islam@mil.utoronto.ca
+# Date: 30-10-2024
+# Contact: shamayla.islam@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Requires the cleaned dataset of polls, including variables like pollster, state, sample size, and pct for Kamala Harris.
-# Any other information needed: Ensure the necessary packages (e.g., tidyverse, broom, and modeling packages) are installed before running the model.
-
+# Pre-requisites:
+# - 00-install_packages.R
+# - 02-clean_data.R
+# Any other information needed? Data should include variables for support, pollster, sample size, state, and recency.
 
 #### Workspace setup ####
-
 library(tidyverse)
-library(rstanarm)
+library(here)
 library(arrow)
 
 #### Read data ####
-analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
+analysis_data <-
+  read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# Filter all the observations regarding Harris. Model 1,2,3 will be about Harris.
+harris_data <- analysis_data %>%
+  filter(candidate_name %in% c("Kamala Harris"))
 
+# Model 1: Only pollster and sample size as predictors
+model1 <- lm(pct ~ pollster + sample_size, data = harris_data)
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_m\del.rds"
-)
+saveRDS(model1, here("models", "model1.rds"))
 
+# Model 2: Adding state as a predictor
+model2 <-
[…]

### 2024-11-04 scripts/05-create_graph.R

@@ -7,7 +7,7 @@
 # Pre-requisites: 00-install_packages.R
 # Additional Notes: This chunk saves the DAG as a PNG file to the specified directory for use in the main report.
 
-# Load necessary libraries
+#### Workspace setup ####
 library(DiagrammeR)
 library(DiagrammeRsvg)
 library(rsvg)
@@ -47,20 +47,28 @@ export_svg(graph) |>
   charToRaw() |>
   rsvg::rsvg_png("other/graphs/polls_of_polls_dag.png", width = 1000)
 
-
-
 # Define the file paths using the here() function
-output_path_model3 <- here("other", "graphs", "diagnostic_plot_model3.png")
-output_path_model6 <- here("other", "graphs", "diagnostic_plot_model6.png")
+output_path_model3 <-
+  here("other", "graphs", "diagnostic_plot_model3.png")
+output_path_model6 <-
+  here("other", "graphs", "diagnostic_plot_model6.png")
 
 # Save diagnostic plot for model3
 png(output_path_model3, width = 800, height = 600)
-autoplot(model3, label.size = 3, smooth.colour = "blue", smooth.linetype = "dashed") + theme_classic()
+autoplot(
+  model3,
+  label.size = 3,
+  smooth.colour = "blue",
+  smooth.linetype = "dashed"
+) + theme_classic()
 dev.off()
 
 # Save diagnostic plot for model6
 png(output_path_model6, width = 800, height = 600)
-autoplot(model6, label.size = 3, smooth.colour = "blue", smooth.linetype = "dashed") + theme_classic()
+autoplot(
+  model6,
+  label.size = 3,
+  smooth.colour = "blue",
+  smooth.linetype = "dashed"
+) + theme_classic()
 dev.off()
-
-

### 2024-11-04 paper/paper.qmd
remove taboo words

--- 2024-11-01 paper/paper.qmd
@@
 abstract: >
-  This study employs a polls-of-polls approach to predict support for Kamala Harris and Donald Trump in the 2024 U.S.
+  This study employs a poll-of-polls approach to predict support for Kamala Harris and Donald Trump in the 2024 U.S.
 Presidential Election, with a focus on key battleground states to forecast the likely winner of the electoral college.
 By aggregating multiple polls and applying a weighted linear regression model with predictors like pollster reliability, sample size, state, and recency, we estimate a higher national support for Trump relative to Harris.
-Our analysis also reveals a close competition in battleground states, with Trump holding a narrow lead in Arizona, Georgia, and North Carolina, while Harris leads in Michigan, Nevada, Pennsylvania, and Wisconsin.
-These findings underscore the value of aggregating polls over relying on individual surveys, offering a more comprehensive and robust forecast of electoral outcomes.
+Our analysis also shows a close competition in battleground states, with Trump holding a narrow lead in Arizona, Georgia, and North Carolina, while Harris leads in Michigan, Nevada, Pennsylvania, and Wisconsin.
+These findings underscore the value of aggregating polls over relying on individual surveys, offering a robust forecast of electoral outcomes.
 format: 
@@
 By focusing on battleground states where the support margins are particularly close, our approach highlights the areas where campaign strategies and voter turnout efforts may have the greatest impact.
-This could provides insights for political strategists, media analysts, and the general public by offering a nuanced view of the electoral landscape.
+This analyasis is useful for political strategists, media analysts, and the general public by offering a view of the electoral landscape.
 
@@
 
-## Explorations of Variables of Interest
+## Variables of Interest
 
@@
 #| label: fig-ster
-#| fig-cap: 'Support distribution for Kamala Harris and Donald Trump by major pollsters, highlighting variability in reported support across organizations and the value of aggregating multiple polls for balanced insights.'
+#| fig-cap: 'Support distribution for Kamala Harris and Donald Trump by major pollsters, highlighting variability in reported support across organizations and the value of aggregating multiple polls.'
 
@@
 The right chart shows that polling is almost evenly split between Kamala Harris and Donald Trump.
-This distribution underscores the model’s balanced approach to capturing state-level nuances as well as broader national trends, providing a comprehensive view of candidate support across different contexts
+This distribution underscores the model’s balanced approach to capturing state-level differences as well as broader national trends, providing a view of candidate support across different contexts
 
@@
 # Filter out Alaska, Hawaii, and Puerto Rico using subset
-map_data_continental <- subset(map_data, !NAME %in% c("Alaska", "Hawaii", "Puerto Rico"))
+map_data_continental <-
+  subset(map_data,!NAME %in% c("Alaska", "Hawaii", "Puerto Rico"))
 
 ggplot(map_data_continental) +
-  geom_sf(aes(fill = kamala_proportion * 100), color = "white", size = 0.2) +
-  geom_text(aes(label = STUSPS, geometry = geometry), 
-            stat = "sf_coordinates", color = "black", size = 3) +  # Adjust color and size as needed
-  scale_fill_distiller(palette = "RdBu", direction = 1, na.value = "grey", 
-                       name = "Harris's Mean Support in Polls (%)") +
+  geom_sf(aes(fill = kamala_proportion * 100),
+          color = "white",
+          size = 0.2) +
+  geom_text(
+    aes(label = STUSPS, geometry = geometry),
+    stat = "sf_coordinates",
+    color = "black",
+    size = 3
+  ) +  # Adjust color and size as needed
+  scale_fill_distiller(
+    palette = "RdBu",
+    direction = 1,
+    na.value = "grey",
+    name = "Harris's Mean Support in Polls (%)"
+  ) +
   theme_minimal() +
@@
   )
-
 ```
@@
 Additionally, we will separately compute estimates for key battleground states, where voter behavior can be more volatile and pivotal in deciding the final outcome of the election.
-This approach helps us capture both national trends and critical state-level dynamics.
+This approach helps us capture both national trends and state-level dynamics.
 
@@
 #| fig-cap: "Factors influencing candidate support, with poll attributes (Pollster, Sample Size, State, Recency) as direct contributors and unobserved confounders representing potential biases in polling results."
+
 # Load and display the saved PNG
@@
 Furthermore, we incorporate sample size as a predictor, as polls with larger samples tend to yield more stable and reliable results, reducing random fluctuations caused by smaller samples.
-The state variable accounts for regional political differences, ensuring that the model captures varying levels of support across geographic and demographic groups, which is crucial for understanding the nuanced political landscape.
-Additionally, recency is included to prioritize more recent polls, as public opinion can shift rapidly in response to political events, and recent data is generally more reflective of current sentiment.
-By integrating these factors, our model seeks to produce a more stable and comprehensive measure of candidate support, reducing noise from individual poll discrepancies and focusing on a balanced, up-to-date aggregation of polling data.
+The state variable accounts for regional political differences, ensuring that the model captures varying levels of support across geographic and demographic groups, which is important for understanding the political landscape.
+Additionally, recency is included to prioritize more recent polls, as public opinion can shift in response to political events, and recent data is generally more reflective of current sentiment.
+By integrating these factors, our model seeks to produce a more stable measure of candidate support, reducing noise from individual poll discrepancies and focusing on a balanced, up-to-date aggregation of polling data.
 
@@
 #| tbl-cap: "Model performance summary showing improved fit and accuracy as State and Recency are added, with R² increasing from 0.375 in Model 1 to 0.774 in Model 3, and RMSE decreasing from 3.053 to 1.836."
+
 # Load your models
@@
   Model = c("Model 1", "Model 2", "Model 3"),
-  Variables = c("Pollster, Sample Size", "Pollster, Sample Size, State", "Pollster, Sample Size, State, Recency"),
-  `R²` = c(summary(model1)$r.squared, summary(model2)$r.squared, summary(model3)$r.squared),
-  `Adjusted R²` = c(summary(model1)$adj.r.squared, summary(model2)$adj.r.squared, summary(model3)$adj.r.squared),
+  Variables = c(
+    "Pollster, Sample Size",
+    "Pollster, Sample Size, State",
+    "Pollster, Sample Size, State, Recency"
+  ),
+  `R²` = c(
+    summary(model1)$r.squared,
+    summary(model2)$r.squared,
+    summary(model3)$r.squared
+  ),
+  `Adjusted R²` = c(
+    summary(model1)$adj.r.squared,
+    summary(model2)$adj.r.squared,
+    summary(model3)$adj.r.squared
+  ),
   `AIC` = c(AIC(model1), AIC(model2), AIC(model3)),
   `BIC` = c(BIC(model1), BIC(model2), BIC(model3)),
-  `RMSE` = c(sqrt(mean(residuals(model1)^2)), sqrt(mean(residuals(model2)^2)), sqrt(mean(residuals(model3)^2)))
+  `RMSE` = c(sqrt(mean(
+    residuals(model1) ^ 2
+  )), sqrt(mean(
+    residuals(model2) ^ 2
+  )), sqrt(mean(
+    residuals(model3) ^ 2
+  )))
 )
@@
 Based on this approach, Kamala Harris has an estimated overall weighted support of around 47.77%, while Donald Trump stands at around 46.31%.
-This aggregation provides a comprehensive view, taking into account the varied methodologies and sampling qualities of different polling organizations.
+This aggregation takes into account the varied methodologies and sampling qualities of different polling organizations.
 
@@
     weight = numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore) /
-             sum(numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore), na.rm = TRUE),
-    fitted_values = fitted(model3),  # Replace this with actual fitted values for Kamala
+      sum(numeric_grade * (
+        max(analysis_data$pollscore, na.rm = TRUE) - pollscore
+      ), na.rm = TRUE),
+    fitted_values = fitted(model3),
+    # Replace this with actual fitted values for Kamala
     weighted_pct = fitted_values * weight  # Calculate weighted percentage for Kamala
@@
 # Calculate the overall weighted prediction for Kamala Harris
-kamala_weighted_support <- sum(kamala_data$weighted_pct, na.rm = TRUE)
+kamala_weighted_support <-
+  sum(kamala_data$weighted_pct, na.rm = TRUE)
 
@@
     weight = numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore) /
-             sum(numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore), na.rm = TRUE),
-    fitted_values = fitted(model6),  # Replace this with actual fitted values for Trump
+      sum(numeric_grade * (
+        max(analysis_data$pollscore, na.rm = TRUE) - pollscore
+      ), na.rm = TRUE),
+    fitted_values = fitted(model6),
+    # Replace this with actual fitted values for Trump
     weighted_pct = fitted_values * weight  # Calculate weighted percentage for Trump
@@
 trump_weighted_support <- sum(trump_data$weighted_pct, na.rm = TRUE)
-
 ```
@@
 For each state, we compare the predicted support percentages and determine the projected winner based on which candidate has higher weighted support.
-Additionally, we calculate the margin of difference between the candidates, which can provide insights into how close each state’s race is.
+We also calculate the margin of difference between the candidates, which shows how close each state’s race is.
 
@@
   group_by(state) %>%
-  mutate(
-    weight = numeric_grade * (max_pollscore - pollscore) /
-             sum(numeric_grade * (max_pollscore - pollscore), na.rm = TRUE)
-  ) %>%
+  mutate(weight = numeric_grade * (max_pollscore - pollscore) /
+           sum(numeric_grade * (max_pollscore - pollscore), na.rm = TRUE)) %>%
   # Calculate the overall weighted prediction for each state
@@
 # Step 3: Join the two datasets by state
-state_support_data <- left_join(kamala_data_a2, trump_data_a2, by = "state")
+state_support_data <-
+  left_join(kamala_data_a2, trump_data_a2, by = "state")
 
@@
 # Filter out Alaska, Hawaii, and Puerto Rico from state_support_data
-map_data_continental <- subset(us_states, !NAME %in% c("Alaska", "Hawaii", "Puerto Rico"))
+map_data_continental <-
+  subset(us_states,!NAME %in% c("Alaska", "Hawaii", "Puerto Rico"))
 
@@
 # Define the colors for each category
-color_scale <- c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred", "No Polling Data" = "grey")
+color_scale <-
+  c(
+    "Kamala Harris" = "steelblue",
+    "Donald Trump" = "indianred",
+    "No Polling Data" = "grey"
+  )
 
@@
   geom_sf(aes(fill = winner), color = "white", size = 0.2) +
-  geom_text(aes(label = STUSPS, geometry = geometry), 
-            stat = "sf_coordinates", color = "black", size = 3) +  # Adjust text size as needed
+  geom_text(
+    aes(label = STUSPS, geometry = geometry),
+    stat = "sf_coordinates",
+    color = "black",
+    size = 3
+  ) +  # Adjust text size as needed
   scale_fill_manual(
@@
   ) 
-
 ```
@@
 Gray-shaded states lack sufficient polling data to make a prediction, reflecting areas of data scarcity in the model's projections.
-This visualization highlights the geographical and political divides across the U.S., as well as the critical role swing states play in determining the election outcome.
+This visualization highlights the geographical and political divides across the U.S., as well as the importance of swing states play in election outcome.
 
@@
 # Define battleground states based on provided list
-battleground_states <- c("Pennsylvania", "Michigan", "North Carolina", "Wisconsin", "Nevada", "Arizona", "Georgia")
+battleground_states <-
+  c(
+    "Pennsylvania",
+    "Michigan",
+    "North Carolina",
+    "Wisconsin",
+    "Nevada",
+    "Arizona",
+    "Georgia"
+  )
 
@@
 knitr::kable(battleground_table)
-
 ```
@@
 USAFacts [@usafacts], FiveThirtyEight [@FiveThirtyEight] and many other polling websites are calling Arizona, Georgia, Michigan, Pennsylvania, Wisconsin, North Carolina, Nevada the 'swing states' of the 2024 presidential election.
-Swing states are crucial for election predictions because these states, with their historically close voting margins, often determine the overall outcome by tipping the electoral balance toward one candidate.
+Swing states are important for election predictions because these states, with their historically close voting margins, often determine the overall outcome by tipping the electoral balance toward one candidate.
 @tbl-battleground-states summarizes the predicted support percentages for Kamala Harris and Donald Trump in key battleground states, along with the predicted winner and the support margin.
@@
 Our state-level analysis highlights close competition observed in key battleground states, emphasizing their pivotal role in the 2024 election.
-Our model reveals that in states like Pennsylvania, Michigan, and Wisconsin, support levels for Harris and Trump are nearly tied, reflecting the intense contest for these critical electoral votes.
+Our model shows that in states like Pennsylvania, Michigan, and Wisconsin, support levels for Harris and Trump are nearly tied, reflecting the intense contest for these electoral votes.
 The predictions show that even minor shifts in support within these swing states could decisively impact the overall election outcome.
-This analysis shows the importance of regional polling in capturing nuanced voter sentiment and the heightened influence that battleground states hold in shaping the final results.
+This analysis shows the importance of regional polling in capturing voter sentiment and the heightened influence that battleground states hold in shaping the final results.
 
@@
 Future research could explore non-linear methods, such as machine learning, to capture potential interactions between variables.
-Refining the weighting mechanism by considering pollster performance history or state-specific polling nuances could further enhance predictive accuracy.
+Refining the weighting mechanism by considering pollster performance history or state-specific polling differences could further enhance predictive accuracy.
 
@@
 Another strength is its topical relevance.
-By focusing on vote preferences and demographic splits, the survey provides timely insights, especially valuable during election cycles.
+By focusing on vote preferences and demographic splits, the survey provides valuable information during election cycles.
 However, the survey has some limitations.
-One is its lack of nuance; while it captures basic voter preferences, it doesn’t explore the motivations or factors driving those choices in depth.
+One is the lack of deeper understanding; while it captures basic voter preferences, it doesn’t explore the motivations or factors driving those choices in depth.
 Additionally, the various survey modes can affect respondent engagement with the questions, which may also influence the accuracy and depth of their answers.
@@
 Its weighting techniques contribute to reliable and accurate results despite non-responses.
-While the use of multiple survey modes reduces coverage bias, other possible biases—such as lack of nuance and combined measurement errors must be considered during the analysis.
+While the use of multiple survey modes reduces coverage bias, other possible biases—such as lack of deeper understanding and combined measurement errors must be considered during the analysis.
 
@@
 Key swing states are states where there is no overwhelming support for a particular candidate or political party [@barker].
-The outcome in these states often determine the overall result of the election, making them critical in securing a win in the Electoral College.
-Oversampling these states allows a more accurate forecasting in these battlegrounds, which are critical for predicting the Electoral College outcome.
+The outcome in these states often determine the overall result of the election, making them key in securing a win in the Electoral College.
+Oversampling these states allows a more accurate forecasting in these battlegrounds, which is important for predicting the Electoral College outcome.
 

### 2024-11-04 paper/paper.qmd
separate discussion in sub-categories

@@
 
-In this paper, we set out to predict the support for both Kamala Harris and Donald Trump in the 2024 U.S.
-Presidential Election using a “polls-of-polls” approach ADD REFERENCE HERE.
-By aggregating multiple polls, we sought to mitigate the inherent biases present in individual surveys, creating a more accurate and balanced forecast across candidates.
-Our analysis was driven by a linear regression model using key predictors, such as pollster, sample size, state, and recency of the polls.
-After calculating predicted values from the model, we applied a weighting scheme based on each pollster’s numeric grade and pollscore, incorporating both reliability and bias.
-This weighted approach enabled us to produce an overall prediction of candidate support, reflecting variations across different states and polling organizations.
-
-One key insight from our analysis is the importance of poll recency.
+## Key Findings
+
+This paper predicts the support for Kamala Harris and Donald Trump in the 2024 U.S.
+Presidential Election using a “poll-of-polls” approach.
+By aggregating multiple polls, we reduce the influence of individual surveys' biases and create a more accurate and balanced forecast across candidates.
+Our model uses predictors, such as pollster, sample size, state, and recency of the polls.
+We applied a weighting scheme based on each pollster’s numeric grade and pollscore which takes in consideration reliability and bias.
+This weighted approach produces a prediction of candidate support that reflects variations across different states and polling organizations.
+
+One main finding from our analysis is the importance of poll recency.
 Including recency notably increased the model’s explanatory power, with improvements in R² and reductions in RMSE compared to models without it.
@@
 
-However, there are limitations to our approach.
+## Weaknesses and Future Directions
+
 Our model assumes linear relationships between predictors and candidate support, which might not capture more complex or interaction-based trends.

### 2024-11-04 paper/paper.qmd
render pdf
elaborate discussion and add references

@@
 model_summary %>%
-  kable(caption = "Model Summary with Included Variables", digits = 3) %>%
-  kable_styling(full_width = FALSE)
+  kable(caption = "Model Summary with Included Variables", digits = 3)
 
@@
 
+## Polling Uncertainty in 2024 Election
+
+Polling provides a general sense of public opinion, but it shouldn’t be viewed as a guaranteed predictor of election outcomes.
+Past U.S.
+elections show why polls can be misleading.
+For example, in 2016, polls showed Hillary Clinton in the lead.
+According to national polls, Clinton would have won popular vote by about 3.2% [@durand].
+However, Donald Trump won battleground states, surprising many who relied on those predictions [@silver].
+Trump won electoral vote with 306 votes, while Clinton had 232 [@durand].
+
+In 2020, 93% of national polls overestimated support for President Joe Biden [@kennedy].
+Biden's lead was especially overestimated in states like Florida and Wisconsin, where Trump performed better than expected [@edwards].
+These examples highlight the limitations of polls, especially when they can't capture late shifts in voter preferences, changes in turnout, and the difficulty of reaching a representative sample [@pew].
+While polls can show trends, they are not always reliable for predicting what will happen on election day.
+
+## Trump and Harris' Path to Victory
+
+Both Donald Trump and Kamala Harris have potential paths to victory in the 2024 election.
+For Trump, one factor is the “shy voter” effect, where some supporters may be hesitant to share their views with pollsters due to social pressures.
+This can lead to polls underestimating his actual support [@woodie].
+Additionally, Trump’s base includes groups that are harder to reach through standard polling, such as rural voters and individuals who avoid mainstream media [@brown].
+These voters may be underrepresented in polls but can play an important role in election outcomes.
+
+On the other hand, Kamala Harris could secure a win if her campaign mobilizes specific demographics, including young people, women, and minority voters.
+As the first Black and South Asian woman on a presidential ticket, Harris’s candidacy may energize these groups and boost turnout among historically Democratic-leaning voters [@pew].
+The growing diversity and youth of the electorate could also benefit her, especially in urban areas and battleground states [@blow].
+With strong efforts in voter registration and mobilization, Harris could turn polling support into actual votes, providing her with a strong chance at victory if turnout aligns with her campaign’s goals.
+
 ## Weaknesses and Future Directions

### 2024-11-04 paper/paper.qmd
fix appendix and type out survey

@@
 These findings underscore the value of aggregating polls over relying on individual surveys, offering a robust forecast of electoral outcomes.
-format: 
-  pdf:
+format: pdf
 number-sections: true
-    header-includes:
-      - \usepackage{pdfpages}
-      - '\KOMAoption{captions}{tableheading}'
-      - \usepackage[margin=1in]{geometry}
-
 bibliography: references.bib
@@
 
-This paper is organized as follows: In Section @sec-data, we clean the data, explore summary statistics, plot distributions of key variables, and examine relationships between variables.
+This paper is organized as follows: In Section @sec-data, we show summary statistics, plot distributions of key variables, and examine relationships between variables.
 In Section @sec-model, we discuss our forecasting approach, model selection, justification for the chosen model, and the mechanism of deriving poll weights based on pollster reliability, ultimately presenting our predictions.
 In Section @sec-discussion, we address the broader implications of our findings, acknowledge limitations, and suggest directions for future work.
+[Appendix -@sec-data-model] contains details of the data cleaning process and model diagnostics.
 [Appendix -@sec-pollster-meth] contains a closer look at the methodology used by Emerson College Polling for its poll conducted from October 14 t0 16, 2024.
@@
 Presidential Election if we had a budget of  \$100,000.
+[Appendix -@sec-copy] contains a copy of a survey that would be used in the idealized methodology.
+[Appendix -@sec-styling] contains specifications on code styling.
+[Appendix -@sec-reproducibility] contains notes on reproducibility and [Appendix -@acknowledgment] acknowledges important ressources consulted for the production of this paper.
 
@@
 
+# Data Cleaning and Modeling {#sec-data-model}
+
+## Data Cleaning
+
+The raw data for this project, sourced from FiveThirtyEight, [@FiveThirtyEight] underwent a series of cleaning steps to prepare it for analysis.
+Initially, duplicate rows were removed to ensure that only unique observations remained, facilitated by the `janitor` package [@janitor].
+A new binary variable, 'national', was created to indicate whether a poll was conducted at the national or state level.
+Missing values in the 'state' column were replaced with "Not Applicable," and numeric grades were evaluated to filter out low-quality pollsters, keeping only those with a numeric grade above 1.
+This cutoff was selected to retain mid to high-level pollsters for more reliable results.
+These steps were performed using functions from the `tidyverse` package [@tidy].
+Dates were also standardized and converted into a proper format for analysis using the same package.
+Polls related to Kamala Harris were retained for further analysis, and percentage support values were transformed into actual numbers of supporters based on sample size.
+Pollster counts below five were excluded to focus on more reliable data sources.
+Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the `arrow` package [@arrow].
+
+## Model Diagnostics
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-dgplotone
+
+knitr::include_graphics(here("other", "graphs", "diagnostic_plot_model3.png"))
+```
+
+The diagnostic plots for Model 3 is given in [@fig-dgplotone]. We notice the following:
+
+1.
+**Residuals vs.
+Fitted Plot**: This plot checks for non-linearity and heteroscedasticity.
+The residuals are scattered around the horizontal axis with no clear pattern, indicating that the linearity assumption is reasonable.
+However, there is slight spread around the center, suggesting some mild heteroscedasticity.
+
+2.
+**Normal Q-Q Plot**: This plot assesses the normality of residuals.
+Most points align along the diagonal line, although there are some deviations at the tails.
+This suggests that the residuals are approximately normally distributed, with minor deviations in the extremes.
+
+3.
+**Scale-Location Plot**: This plot further checks for homoscedasticity.
+The residuals appear to be evenly spread across the fitted values, supporting the homoscedasticity assumption, although slight deviations are present in certain regions.
+
+4.
+**Residuals vs.
+Leverage Plot**: This plot identifies potential influential points.
+While most points have low leverage, a few points exhibit higher leverage, as indicated by their distance from the center.
+However, no points exceed Cook’s distance threshold, indicating no extreme outliers that would unduly influence the model.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-dgplottwo
+
+knitr::include_graphics(here("other", "graphs", "diagnostic_plot_model6.png"))
+```
+
+The diagnostic plots for Model 6 is given in [@fig-dgplottwo]. We notice the following:
+
+1.
+**Residuals vs Fitted**: This plot suggests a fairly random spread of residuals around zero, indicating that the linearity assumption holds reasonably well.
+However, there is slight clustering of points in the center, which could indicate minor heteroscedasticity but is not severe.
+
+2.
+**Normal Q-Q Plot**: The Q-Q plot shows that the residuals generally follow a normal distribution, with only a few deviations at the tails.
+This suggests that the normality assumption is mostly met, though some minor deviations in the upper tail indicate possible outliers.
+
+3.
+**Scale-Location (Spread-Location) Plot**: The residuals appear randomly dispersed with no clear pattern, supporting the homoscedasticity (constant variance) assumption.
+However, some observations near the upper range might indicate slight variance inconsistency, though it is minimal.
+
+4.
+**Residuals vs Leverage**: This plot does not show any influential outliers with high leverage that might impact the model unduly.
+The Cook's distance lines show that no data points exceed these thresholds, indicating that no individual observation is disproportionately influencing the model.
+
+Overall, the diagnostic plots for both model suggests that they meet the assumptions for linear regression fairly well, with minor deviations that are not expected to severely impact the model’s validity.
+
 # Emerson College Polling Methodology (October 14-16, 2024) {#sec-pollster-meth}
@@
 
-# Copy of the Survey
-
-\includepdf[pages=-]{../other/surveyfiles/SurveyCopy.pdf}
-
-# Model Diagnostics
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-dgplotone
-
-knitr::include_graphics(here("other", "graphs", "diagnostic_plot_model3.png"))
-```
-
-The diagnostic plots for Model 3ni sgiven in [@fig-dgplotone]. We notice the following:
-
-1.
-**Residuals vs.
-Fitted Plot**: This plot checks for non-linearity and heteroscedasticity.
-The residuals are scattered around the horizontal axis with no clear pattern, indicating that the linearity assumption is reasonable.
-However, there is slight spread around the center, suggesting some mild heteroscedasticity.
-
-2.
-**Normal Q-Q Plot**: This plot assesses the normality of residuals.
-Most points align along the diagonal line, although there are some deviations at the tails.
-This suggests that the residuals are approximately normally distributed, with minor deviations in the extremes.
-
-3.
-**Scale-Location Plot**: This plot further checks for homoscedasticity.
-The residuals appear to be evenly spread across the fitted values, supporting the homoscedasticity assumption, although slight deviations are present in certain regions.
-
-4.
-**Residuals vs.
-Leverage Plot**: This plot identifies potential influential points.
-While most points have low leverage, a few points exhibit higher leverage, as indicated by their distance from the center.
-However, no points exceed Cook’s distance threshold, indicating no extreme outliers that would unduly influence the model.
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-dgplottwo
-
-knitr::include_graphics(here("other", "graphs", "diagnostic_plot_model6.png"))
-```
-
-The diagnostic plots for Model 6 is given in [@fig-dgplottwo]. We notice the following:
-
-1.
-**Residuals vs Fitted**: This plot suggests a fairly random spread of residuals around zero, indicating that the linearity assumption holds reasonably well.
-However, there is slight clustering of points in the center, which could indicate minor heteroscedasticity but is not severe.
-
-2.
-**Normal Q-Q Plot**: The Q-Q plot shows that the residuals generally follow a normal distribution, with only a few deviations at the tails.
-This suggests that the normality assumption is mostly met, though some minor deviations in the upper tail indicate possible outliers.
-
-3.
-**Scale-Location (Spread-Location) Plot**: The residuals appear randomly dispersed with no clear pattern, supporting the homoscedasticity (constant variance) assumption.
-However, some observations near the upper range might indicate slight variance inconsistency, though it is minimal.
-
-4.
-**Residuals vs Leverage**: This plot does not show any influential outliers with high leverage that might impact the model unduly.
-The Cook's distance lines show that no data points exceed these thresholds, indicating that no individual observation is disproportionately influencing the model.
-
-Overall, the diagnostic plots for both model suggests that they meet the assumptions for linear regression fairly well, with minor deviations that are not expected to severely impact the model’s validity.
-
-# Code Styling
+# Copy of the Survey {#sec-copy}
+
+\begingroup
+\small
+**2024 U.S. Presidential Election Survey**
+
+**Introduction:**
+Thank you for your interest in the 2024 U.S.
+Presidential Election Survey.
+The 2024 U.S.
+Presidential Election is set to take place on November 5, 2024.
+This election will decide the next President of the United States, a position that holds significant influence over national and global policies.
+
+Your participation in this survey is entirely voluntary, and you can choose to withdraw at any time without any consequences.
+The information collected will be used solely for research purposes to gain insights into voter preferences and trends as we approach this important election.
+All responses will remain confidential and will be reported only in aggregate form.
+
+**Contact Information:**
+
+- Denise Chang, University of Toronto. Email: dede.chang@mail.utoronto.ca
+
+- Shamayla Durrin, University of Toronto. Email: shamayla.islam@mail.utoronto.ca
+
+- Krishna Kumar, University of Toronto. Email: krishna.kumar@mail.utoronto.ca
+
+---
+
+**Section 1: Voting Intention**
+
+1. Consent to Participate
+By clicking the button below, I consent to participate in this survey, and I agree that my responses may be used for research purposes, including analysis and reporting on voter sentiment related to the 2024 U.S.
+Presidential Election.
+I understand that my participation is voluntary and that I can withdraw at any time without any consequences.
+My responses will be kept confidential and will be used only in aggregate form.
+   - I agree
+
+2. How likely are you to vote in the 2024 U.S. Presidential Election?
+   - Not likely at all
+   - Not very likely
+   - Somewhat likely
+   - Extremely likely
+   - Prefer not to say
+
+3. If the 2024 U.S. Presidential Election were held today, who would you vote for?
+   - Kamala Harris (Democratic Party)
+   - Donald Trump (Republican Party)
+   - Undecided
+   - Prefer not to say
+
+4. **If you answered "Undecided" in the previous question, who are you leaning towards?**
+   - Kamala Harris (Democratic Party)
+   - Donald Trump (Republican Party)
+   - Prefer not to say
+
+---
+
+**Section 2: Key Issue Priorities**
+
+5. What are the three most important issues influencing your vote in this election? (Select up to 3)
+   - Economy
+   - Immigration
+   - Healthcare
+   - National Security
+   - Education
+   - Climate Change
+
+6. **How has the state of the U.S. economy influenced your voting choice?
+   - Not at all
+   - Somewhat
+   - Greatly
+   - Unsure
+
+---
+
+**Section 3: Voting History**
+
+7. Did you vote in the 2020 U.S. Presidential Election?
+   - Yes
+   - No
+   - Prefer not to say
+
+8. If you voted in 2020, for whom did you vote?
+   - Joe Biden (Democratic Party)
+   - Donald Trump (Republican Party)
+   - Prefer not to say
+
+---
+
+**Section 4: Voter Engagement**
+
+9. How motivated are you to vote in the 2024 election compared to past elections?
+   - Less motivated
+   - About the same
+   - More motivated
+   - Unsure
+
+10. Which of the following describes your voting method?
+    - I plan to vote in person on Election Day
+    - I plan to vote early in person
+    - I plan to vote by mail
+    - I have already voted (via early voting or mail-in)
+    - I do not plan to vote
+
+---
+
+**Section 5: Demographic Questions**
+
+11. What is your age range? (years)
+    - 18-24
+    - 25-34
+    - 35-44
+    - 45-54
+    - 55-64
+    - 65 or older
+    - Prefer not to say
+
+12. What gender do you identify as?
+    - Male
+    - Female
+    - Non-binary
+    - Prefer not to say
+
+13. Which of the following best describes your race or ethnicity?
+    - White or Caucasian
+    - Black or African American
+    - Hispanic or Latino of any race
+    - Asian
+    - Other or multiple races
+
+14. What is the highest level of education you have completed?
+    - Less than high school
+    - High school diploma or GED
+    - Some college
+    - No degree
+    - Associate degree
+    - Bachelor’s degree
+    - Graduate or professional degree
+    - Prefer not to say
+
+15. What is your current employment status?
+    - Employed full-time
+    - Employed part-time
+    - Self-employed
+    - Unemployed
+    - Student
+    - Retired
+    - Prefer not to say
+
+16. Do you consider yourself a supporter of any of the following political parties?
+    - Democratic Party
+    - Republican Party
+    - Independent
+    - Prefer not to say
+
+---
+
+**End Message**:
+
+Thank you for completing this survey! Your responses are valuable in understanding voter opinions and trends in the lead-up to the 2024 U.S. Presidential Election.
+
+\endgroup
+
+# Code Styling {#sec-styling}
 
@@
 
-# Reproducibility
+# Reproducibility {#sec-reproducibility}
 
@@
 
-# Acknowledgments
+# Acknowledgments {#sec-acknowledgement}
 

### 2024-11-04 paper/paper.qmd

@@
   This study employs a poll-of-polls approach to predict support for Kamala Harris and Donald Trump in the 2024 U.S.
-Presidential Election, with a focus on key battleground states to forecast the likely winner of the electoral college.
-By aggregating multiple polls and applying a weighted linear regression model with predictors like pollster reliability, sample size, state, and recency, we estimate a higher national support for Trump relative to Harris.
-Our analysis also shows a close competition in battleground states, with Trump holding a narrow lead in Arizona, Georgia, and North Carolina, while Harris leads in Michigan, Nevada, Pennsylvania, and Wisconsin.
-These findings underscore the value of aggregating polls over relying on individual surveys, offering a robust forecast of electoral outcomes.
+Presidential Election, focusing on key battleground states to forecast the likely winner of the electoral college.
+By aggregating multiple polls and applying a weighted linear regression model with predictors like pollster reliability, sample size, state, and recency, we estimate higher national support for Trump than Harris.
+Our analysis also shows a close competition in battleground states, with Trump holding a slight lead in Arizona, Georgia, and North Carolina, while Harris leads in Michigan, Nevada, Pennsylvania, and Wisconsin.
+These findings show the value of aggregating polls over relying on individual surveys, offering a robust forecast of electoral outcomes.
 format: pdf
@@
 In this paper, we aim to forecast voter support for Kamala Harris and Donald Trump by using data from multiple polling sources, reducing individual poll biases, and improving overall prediction accuracy.
-Our analysis focuses on estimating the levels of support for each candidate nationally while focusing on key swing states, which are likely to be decisive in determining the electoral college outcome.
-
-The estimand of this study is the level of voter support for each candidate, Kamala Harris and Donald Trump, as reported across multiple polls.To estimate this, we developed a regression model incorporating variables such as pollster, sample size, state, and recency of the poll, with an emphasis on accurately capturing state-level dynamics.
+Our analysis estimates the levels of support for each candidate nationally while focusing on swing states, which are likely to be decisive in determining the Electoral College outcome.
+
+The estimand of this study is the level of voter support for each candidate, Kamala Harris and Donald Trump, as reported across multiple polls.
+To estimate this, we developed a regression model incorporating variables such as pollster, sample size, state, and recency of the poll, with an emphasis on accurately capturing state-level dynamics.
 Our findings indicate that, on a national level, support between Kamala Harris and Donald Trump is closely balanced.
-Harris’s estimated national support stands close to 48%, while Trump’s support is slightly higher, around 49%.
-We found that Trump holds a marginal lead in battleground states like Arizona and Georgia, with a support margin slightly above 1%.
-In contrast, Harris shows narrow leads in Michigan, Nevada, Pennsylvania, and Wisconsin, with her support margin in Wisconsin reaching over 2%, making it her strongest battleground.
-North Carolina emerges as one of the tightest races, with Trump leading by only 0.26%, underscoring the competitive nature of these regions.
-
-This paper contributes to the field of election forecasting by emphasizing the importance of poll quality and recency in prediction models.
-By focusing on battleground states where the support margins are particularly close, our approach highlights the areas where campaign strategies and voter turnout efforts may have the greatest impact.
-This analyasis is useful for political strategists, media analysts, and the general public by offering a view of the electoral landscape.
+Harris’s estimated national support is around 48%, while Trump’s support is slightly higher, around 49%.
+We found that Trump holds a lead in battleground states like Arizona and Georgia, with a support margin of over 1%.
+In contrast, Harris shows small leads in Michigan, Nevada, Pennsylvania, and Wisconsin, with her support margin in Wisconsin over 2%, making it her strongest battleground.
+North Carolina is one of the tightest races, with Trump leading by only 0.26%.
+This analysis is useful for political strategists, media analysts, and the general public by offering a view of the electoral landscape.
 
@@
 [Appendix -@sec-data-model] contains details of the data cleaning process and model diagnostics.
-[Appendix -@sec-pollster-meth] contains a closer look at the methodology used by Emerson College Polling for its poll conducted from October 14 t0 16, 2024.
+[Appendix -@sec-pollster-meth] contains a closer look at the methodology used by Emerson College Polling for its poll conducted from October 14 to 16, 2024.
 [Appendix -@sec-idealized-meth] contains a discussion about an idealized methodology that we would use to forecast the 2024 U.S.
@@
 [Appendix -@sec-styling] contains specifications on code styling.
-[Appendix -@sec-reproducibility] contains notes on reproducibility and [Appendix -@acknowledgment] acknowledges important ressources consulted for the production of this paper.
+[Appendix -@sec-reproducibility] contains notes on reproducibility and [Appendix -@acknowledgment] acknowledges important resources consulted for the production of this paper.
 
@@
 
-## Data Cleaning
-
-The raw data for this project, sourced from FiveThirtyEight, [@FiveThirtyEight] underwent a series of cleaning steps to prepare it for analysis.
-Initially, duplicate rows were removed to ensure that only unique observations remained, facilitated by the `janitor` package [@janitor].
-A new binary variable, 'national', was created to indicate whether a poll was conducted at the national or state level.
-Missing values in the 'state' column were replaced with "Not Applicable," and numeric grades were evaluated to filter out low-quality pollsters, keeping only those with a numeric grade above 1.
-This cutoff was selected to retain mid to high-level pollsters for more reliable results.
-These steps were performed using functions from the `tidyverse` package [@tidy].
-Dates were also standardized and converted into a proper format for analysis using the same package.
-Polls related to Kamala Harris were retained for further analysis, and percentage support values were transformed into actual numbers of supporters based on sample size.
-Pollster counts below five were excluded to focus on more reliable data sources.
-Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the `arrow` package [@arrow].
-
 ## Variables of Interest
@@
 
+We measure voters' intentions by surveying a sample of individuals at various points leading up to the election, asking about their current opinions on candidates, issues, or likelihood to vote.
+Each response is recorded with a timestamp, allowing us to analyze changes over time (recency) and connect shifting opinions to projected behaviors.
+We then aggregate and analyze these responses to predict patterns, identifying trends that could indicate likely voting actions on Election Day.
+
 For this analysis, a subset of variables was selected from the raw data set.
-Two additional variables, 'national' and 'state', were created using the existing 'state' and 'end date' variable.
+Two additional variables, 'national' and 'state', were created using the existing 'state' and 'end date' variables.
 A short description of each variable of interest and their respective measurement methods is given below.
@@
 The value of this variable is calculated by the 538 team.
-All national and state-level polls that were conducted in 1998 or later are considered to determine and to assign a numeric grade to each pollster [@538_morris].
+All national and state-level polls that were conducted in 1998 or later are considered to determine and assign a numeric grade to each pollster [@538_morris].
 
@@
 
--   *End Date*: The date the poll ended, reflecting the currency of the data. This variable is taken from the publication, and is the last date for which the survey was active [@538_methodology].
-
--   *Sample Size*: The total number of respondents in the poll. This is measure by the number of completed surveys [@538_methodology].
+-   *End Date*: The date the poll ended, reflecting the currency of the data. This variable is taken from the publication and is the last date for which the survey was active [@538_methodology].
+
+-   *Sample Size*: The total number of respondents in the poll. This is measured by the number of completed surveys [@538_methodology].
 
@@
 This suggests most polls are of moderate to high quality in terms of reliability.
-The mean pollscore of -0.5 suggests these polls are accurate, as negative values imply reduced bias.
+The mean poll score of -0.5 suggests these polls are accurate, as negative values imply reduced bias.
 Also, the large average sample size of 1908 respondents across polls reduces variability and ensures reliable predictions for the candidate support model.
@@
 
-@tbl-topster shows the top 5 most frequent pollsters in the dataset, with each pollster’s total poll count, average pollscore (measuring reliability and historical accuracy), and average numeric grade (indicating overall quality).
-The most frequent pollster, Morning Consult, has a moderate pollscore and a slightly above-average numeric grade, suggesting the data provided is reliable.
-However, there is variability in pollscore and numeric grade across the top pollsters, showing differences in quality and potential biases among them.
+@tbl-topster shows the top 5 most frequent pollsters in the dataset, with each pollster’s total poll count, average poll score (measuring reliability and historical accuracy), and average numeric grade (indicating overall quality).
+The most frequent pollster, Morning Consult, has an average poll score and a slightly above-average numeric grade, suggesting the data provided is reliable.
+However, there is variability in scores and numeric grades across the top pollsters, showing differences in quality and potential biases among them.
 
@@
 
-@fig-ster shows the distribution of support for Kamala Harris and Donald Trump by pollster, highlighting variability across different polling organizations.
-Morning Consult, for example, demonstrates a wide range in reported support, with more spread in Trump’s support.
+@fig-ster shows the distribution of support for Kamala Harris and Donald Trump by pollsters, highlighting variability across different polling organizations.
+Morning Consult, for example, demonstrates a wide range of reported support, with more spread in Trump’s support.
 In contrast, Siena/NYT shows less variation, with Harris consistently leading.
@@
 @fig-reli shows the distribution of Numeric Grade (left) and Pollscore (right) across polling organizations.
-The Numeric Grade distribution shows that most pollsters are rated between 1.5 and 3, with peaks around 2.0 and 2.5, suggesting a concentration of pollsters with moderate to high reliability scores.
+The numerical grade distribution shows that most pollsters are rated between 1.5 and 3, with peaks around 2.0 and 2.5, suggesting a concentration of pollsters with moderate to high reliability scores.
 In contrast, the Pollscore distribution, where lower values indicate higher reliability, shows a range primarily between -1.5 and 0, with a notable peak around -0.5.
 This suggests that while many polls demonstrate relatively low bias, there is still variability in reliability across organizations.
-The distinction between these two metrics emphasizes the need to consider both quality (Numeric Grade) and potential systematic bias (Pollscore) when weighting polls in the model.
+The distinction between these two metrics emphasizes the need to consider both quality (Numeric Grade) and potential systematic bias (Pollscore) when weighing polls in the model.
 
@@
 The right chart shows that polling is almost evenly split between Kamala Harris and Donald Trump.
-This distribution underscores the model’s balanced approach to capturing state-level differences as well as broader national trends, providing a view of candidate support across different contexts
+This distribution underscores the model’s balanced approach to capturing state-level differences as well as broader national trends, providing a view of candidate support across different contexts.
 
@@
 @fig-trend shows the support trends for Kamala Harris and Donald Trump from August to October.
-Each point represents a poll result, color-coded by candidate, with the trend lines highlighting the overall changes in support over time.
+Each point represents a poll result, color-coded by candidates, with the trend lines highlighting the overall changes in support over time.
 Kamala Harris's support remains relatively steady but shows slight fluctuations around mid-September, while Donald Trump's support appears to have a small upward trend toward October.
@@
 In contrast, Harris's chart shows a slight upward trend with larger sample sizes, indicating a marginal increase in support with larger poll samples.
-This highlights the variability in polling support depending on the sample size, emphasiszing the importance of including sampel size in our model.
+This highlights the variability in polling support depending on the sample size, emphasizing the importance of including sample size in our model.
 
@@
 
-In key swing states, such as Pennsylvania, Michigan, and Arizona, we observe a balanced distribution of support, suggesting close competition in these battleground areas.
-Meanwhile, traditional Democratic strongholds, like California and New York, show a clear preference for Harris, with deeper blue shades.
+In swing states, such as Pennsylvania, Michigan, and Arizona, there is a balanced distribution of support, suggesting close competition in these battleground areas.
+Meanwhile, traditional Democratic strongholds, like California and New York, show a preference for Harris, with deeper blue shades.
 Conversely, traditionally Republican states, such as Texas and Tennessee, exhibit higher support for Trump, with intense red shades highlighting his advantage.
@@
 Once we obtain the predicted values from our model, we will weight these predictions based on the numeric grade (quality score) of each pollster to calculate an overall national estimate of the outcome.
-Additionally, we will separately compute estimates for key battleground states, where voter behavior can be more volatile and pivotal in deciding the final outcome of the election.
+Additionally, we will separately compute estimates for key battleground states, where voter behavior can be more volatile and pivotal in deciding the outcome of the election.
 This approach helps us capture both national trends and state-level dynamics.
@@
 
-Given that variables like numeric grade and pollscore are perfectly collinear with pollster, they were excluded from the regression analysis to avoid multicollinearity issues.
+Given that variables like numeric grade and poll score are perfectly collinear with the pollster, they were excluded from the regression analysis to avoid multicollinearity issues.
 These variables, however, remain integral to our weighting strategy, where they will be used to adjust for differences in polling accuracy and reliability.
@@
 
-By systematically comparing model specifications that incorporate these variables, we aim to select the model with the right balance between predictive accuracy and generalizability, ultimately providing the best possible forecast.
+By comparing model specifications that incorporate these variables, we aim to select the model with the right balance between predictive accuracy and generalizability, ultimately providing the best possible forecast.
 
@@
 
-We opted for a linear model due to its capacity to quantify the marginal effects of each predictor (pollster, sample size, state, and recency) on candidate support in a straightforward manner.
+We opted for a linear model due to its capacity to quantify the marginal effects of each predictor (pollster, sample size, state, and recency) on candidate support.
 This structure is well-suited to our polls-of-polls approach, as it allows for the estimation of fixed effects that can control for systematic biases across pollsters, while accommodating the influence of sample size and recency as continuous variables.
@@
 
-The weight assigned to each poll combines both its quality (as represented by the numeric grade) and its level of bias (as indicated by the pollscore).
-Since a more negative pollscore reflects a lower level of bias, the formula uses the difference between the maximum pollscore and each poll’s specific pollscore.
-This approach gives more weight to polls that are both highly graded (indicating higher reliability and transparency) and less biased.By combining these two factors, the weighting system emphasizes polls that are reliable and minimally biased, ensuring that they have a stronger influence in the overall calculation.
+The weight assigned to each poll combines both its quality (as represented by the numeric grade) and its level of bias (as indicated by the poll score).
+Since a more negative poll score reflects a lower level of bias, the formula uses the difference between the maximum poll score and each poll’s specific poll score.
+This approach gives more weight to polls that are both highly graded (indicating higher reliability and transparency) and less biased.
+By combining these two factors, the weighting system emphasizes polls that are reliable and minimally biased, ensuring that they have a stronger influence on the overall calculation.
 Additionally, the total of all weights is normalized to sum to one, so each poll’s weight is proportionate to its quality and relative lack of bias, resulting in a more balanced and accurate average of public support.
@@
 Gray-shaded states lack sufficient polling data to make a prediction, reflecting areas of data scarcity in the model's projections.
-This visualization highlights the geographical and political divides across the U.S., as well as the importance of swing states play in election outcome.
+This visualization highlights the geographical and political divides across the U.S., as well as the importance of swing states in election outcomes.
 
@@
 
-USAFacts [@usafacts], FiveThirtyEight [@FiveThirtyEight] and many other polling websites are calling Arizona, Georgia, Michigan, Pennsylvania, Wisconsin, North Carolina, Nevada the 'swing states' of the 2024 presidential election.
+Many polling websites such as USAFacts [@usafacts] and FiveThirtyEight [@FiveThirtyEight] are calling Arizona, Georgia, Michigan, Pennsylvania, Wisconsin, North Carolina and Nevada the 'swing states' of the 2024 presidential election.
 Swing states are important for election predictions because these states, with their historically close voting margins, often determine the overall outcome by tipping the electoral balance toward one candidate.
@@
 Our model uses predictors, such as pollster, sample size, state, and recency of the polls.
-We applied a weighting scheme based on each pollster’s numeric grade and pollscore which takes in consideration reliability and bias.
+We applied a weighting scheme based on each pollster’s numeric grade and poll score which takes into consideration reliability and bias.
 This weighted approach produces a prediction of candidate support that reflects variations across different states and polling organizations.
@@
 For example, in 2016, polls showed Hillary Clinton in the lead.
-According to national polls, Clinton would have won popular vote by about 3.2% [@durand].
+According to national polls, Clinton would have won the popular vote by about 3.2% [@durand].
 However, Donald Trump won battleground states, surprising many who relied on those predictions [@silver].
-Trump won electoral vote with 306 votes, while Clinton had 232 [@durand].
+Trump won the electoral vote with 306 votes, while Clinton had 232 [@durand].
 
@@
 While polls can show trends, they are not always reliable for predicting what will happen on election day.
-
 ## Trump and Harris' Path to Victory
@@
 
-In conclusion, this paper contributes to election forecasting by providing a weighted estimate of candidate support across both national and state levels.The evolving nature of voter sentiment and poll accuracy suggests the need for adaptive models that takes in consideration ongoing changes in public opinion.
+In conclusion, this paper contributes to election forecasting by providing a weighted estimate of candidate support across both national and state levels.
+The evolving nature of voter sentiment and poll accuracy suggests the need for adaptive models that take into consideration ongoing changes in public opinion.
 
@@
 The raw data for this project, sourced from FiveThirtyEight, [@FiveThirtyEight] underwent a series of cleaning steps to prepare it for analysis.
-Initially, duplicate rows were removed to ensure that only unique observations remained, facilitated by the `janitor` package [@janitor].
+Initially, duplicate rows were removed to ensure that unique observations remained, facilitated by the `janitor` package [@janitor].
 A new binary variable, 'national', was created to indicate whether a poll was conducted at the national or state level.
@@
 Pollster counts below five were excluded to focus on more reliable data sources.
-Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the `arrow` package [@arrow].
+Polls regarding Kamala Harris were filtered to include only those conducted after her official candidacy announcement on July 21, 2024, ensuring the data reflects post-announcement public sentiment.
+The cleaned dataset was saved in Parquet format for efficient storage and retrieval, using the `arrow` package [@arrow].
 
@@
 
-The diagnostic plots for Model 3 is given in [@fig-dgplotone]. We notice the following:
+The diagnostic plots for Model 3 are given in [@fig-dgplotone]. We notice the following:
 
@@
 The residuals are scattered around the horizontal axis with no clear pattern, indicating that the linearity assumption is reasonable.
-However, there is slight spread around the center, suggesting some mild heteroscedasticity.
+However, there is a slight spread around the center, suggesting some mild heteroscedasticity.
 
@@
 
-The diagnostic plots for Model 6 is given in [@fig-dgplottwo]. We notice the following:
+The diagnostic plots for Model 6 are given in [@fig-dgplottwo]. We notice the following:
 
@@
 **Residuals vs Fitted**: This plot suggests a fairly random spread of residuals around zero, indicating that the linearity assumption holds reasonably well.
-However, there is slight clustering of points in the center, which could indicate minor heteroscedasticity but is not severe.
+However, there is a slight clustering of points in the center, which could indicate minor heteroscedasticity but is not severe.
 
@@
 
-Overall, the diagnostic plots for both model suggests that they meet the assumptions for linear regression fairly well, with minor deviations that are not expected to severely impact the model’s validity.
+Overall, the diagnostic plots for both models suggest that they meet the assumptions for linear regression fairly well, with minor deviations that are not expected to severely impact the model’s validity.
 
@@
 Choosing likely voters as the target population makes the poll results more accurate as not all registered voters end up voting in the Presidential Election.
-In fact, the last U.S.
+The last U.S.
 Presidential Election in 2020 only had 66.8% of registered voters participate in the elections, which is the highest voter turnout in the 21st century [@us_census].
@@
 In ECP’s methodology, the sampling frame is the voters that are on Aristotle’s database and on the online panel provided by CINT [@about_ecp].
-Aristotle is an online source of voter and consumer data that has serviced many large political campaigns, PACs and corporations both within the U.S and abroad [@aristotle].
+Aristotle is an online source of voter and consumer data that has serviced many large political campaigns, PACs and corporations both within the U.S.
+and abroad [@aristotle].
 CINT is the largest global research marketplace that connects researchers, such as Emerson, to survey respondents [@cint].
@@
 The Emerson survey is designed with simplicity and efficiency.
-One of its strengths is its straightforward format, using clear and unambiguous questions.
+One of its strengths is its straightforward format, using unambiguous questions.
 This simplicity makes the survey easy for respondents to understand.

### 2024-11-04 paper/paper.qmd
Added model summary.

@@
 
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| include: false
+
+model6 <- readRDS(here("models/model6.rds"))
+
+# Subset the data for each candidate
+harris_data <- analysis_data %>%
+  filter(candidate_name == "Kamala Harris")
+
+trump_data <- analysis_data %>%
+  filter(candidate_name == "Donald Trump")
+
+# Create model summary data for Harris
+modelsumharris <- harris_data %>%
+  mutate(
+    fit = predict(model3, newdata = ., interval = "confidence"),
+    lower = fit[, "lwr"],
+    upper = fit[, "upr"],
+    fitted = fit[, "fit"]
+  ) %>%
+  select(end_date, fitted, lower, upper, state)  # Include state column
+
+# Create model summary data for Trump
+modelsumtrump <- trump_data %>%
+  mutate(
+    fit = predict(model6, newdata = ., interval = "confidence"),
+    lower = fit[, "lwr"],
+    upper = fit[, "upr"],
+    fitted = fit[, "fit"]
+  ) %>%
+  select(end_date, fitted, lower, upper, state)  # Include state column
+
+# Combine the datasets for plotting
+modelsumharris$candidate_name <- "Kamala Harris"
+modelsumtrump$candidate_name <- "Donald Trump"
+combined_data <- bind_rows(modelsumharris, modelsumtrump)
+```
+```{r}
+
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-modsum1
+#| eval: true
+#| fig-cap: 'Model summaries for Kamala Harris and Donald Trump at the national level using polling data.'
+
+ggplot(combined_data, aes(x = as.Date(end_date), y = fitted, color = candidate_name, fill = candidate_name)) +
+  geom_smooth(method = "loess", se = TRUE, span = 0.2) +
+  scale_color_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
+  scale_fill_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
+  labs(x = "Date", y = "Support (%)") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+    legend.title = element_blank(),
+    axis.text.x = element_text(angle = 0),
+    axis.title.x = element_text(margin = margin(t = 10)),
+    axis.title.y = element_text(margin = margin(r = 10))
+  ) +
+  facet_wrap(~ candidate_name, scales = "free_y")
+
+```
+```{r}
+
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-modsum2
+#| eval: true 
+#| fig-cap: 'Model summary of Donald Trump’s support trends across key battleground states.'
+# Define the target states
+target_states <- c("Arizona", "Georgia", "Michigan", "Pennsylvania", "Wisconsin", "North Carolina", "Nevada")
+
+# Filter combined_data for Donald Trump, exclude national data, and limit to target states
+trump_state_data <- combined_data %>%
+  filter(candidate_name == "Donald Trump", state %in% target_states)
+
+# Plotting
+ggplot(trump_state_data, aes(x = as.Date(end_date), y = fitted, fill = candidate_name)) +
+  geom_smooth(method = "loess", se = TRUE, span = 0.2, color = "indianred") +
+  scale_fill_manual(values = c("Donald Trump" = "indianred")) +
+  labs(x = "Date", y = "Support (%)",
+       title = "Support trends for Donald Trump across different states") +
+  theme_minimal() +
+  theme(
+    legend.position = "bottom",
+    legend.title = element_blank(),
+    axis.text.x = element_text(angle = 0),
+    axis.title.x = element_text(margin = margin(t = 10)),
+    axis.title.y = element_text(margin = margin(r = 10))
+  ) +
+  facet_wrap(~ state, scales = "free_y")  # Separate subplots for each state
+
+```
+```{r}
+
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-modsum1
+#| fig-cap: 'Model summary of Kamala Harris’s support trends across key battleground states.'
+# Define the target states
+target_states <- c("Arizona", "Georgia", "Michigan", "Pennsylvania", "Wisconsin", "North Carolina", "Nevada")
+
+# Filter combined_data for Kamala Harris and limit to target states
+harris_state_data <- combined_data %>%
+  filter(candidate_name == "Kamala Harris", state %in% target_states)
+
+# Plotting
+ggplot(harris_state_data, aes(x = as.Date(end_date), y = fitted, fill = candidate_name)) +
+  geom_smooth(method = "loess", se = TRUE, span = 0.2, color = "steelblue") +
+  scale_fill_manual(values = c("Kamala Harris" = "steelblue")) +
+  labs(x = "Date", y = "Support (%)",
+       title = "Support trends for Kamala Harris across different states") +
+  theme_minimal() +
+  theme(
+    legend.position = "bottom",
+    legend.title = element_blank(),
+    axis.text.x = element_text(angle = 0),
+    axis.title.x = element_text(margin = margin(t = 10)),
+    axis.title.y = element_text(margin = margin(r = 10))
+  ) +
+  facet_wrap(~ state, scales = "free_y")  # Separate subplots for each state
+
+```
+
 ## Prediction

### 2024-11-04 paper/paper.qmd
fix decimal

--- 2024-11-04 paper/paper.qmd
@@
   ),
-  `AIC` = c(AIC(model1), AIC(model2), AIC(model3)),
-  `BIC` = c(BIC(model1), BIC(model2), BIC(model3)),
+  `AIC` = round(c(AIC(model1), AIC(model2), AIC(model3)), 0),
+  `BIC` = round(c(BIC(model1), BIC(model2), BIC(model3)), 0),
   `RMSE` = c(sqrt(mean(

### 2024-11-04 paper/paper.qmd
render pdf

@@
 
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-modsum1
+#| fig-cap: 'Model summaries for Kamala Harris and Donald Trump at the national level using polling data.'
+
+# Load and display the saved PNG
+knitr::include_graphics(here("other", "graphs", "sum1.png"))
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-modsum2
+#| fig-cap: 'Model summary of Donald Trump’s support trends across key battleground states.'
+
+# Load and display the saved PNG
+knitr::include_graphics(here("other", "graphs", "sum2.png"))
+```
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-modsum3
+#| fig-cap: 'Model summary of Kamala Harris’s support trends across key battleground states.'
+
+# Load and display the saved PNG
+knitr::include_graphics(here("other", "graphs", "sum3.png"))
+```
+
 # Emerson College Polling Methodology (October 14-16, 2024) {#sec-pollster-meth}

### 2024-11-04 paper/paper.qmd
fix rendering

@@
 #| label: fig-dgplotone
+#| #| fig-cap: 'Diagnostic plots for Model 3'
 
@@
 #| label: fig-dgplottwo
+#| fig-cap: 'Diagnostic plots for Model 6'
 
